### 2022-02-03 README.Rmd
Initial commit.

---
output: github_document
bibliography: inst/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r packages, echo = FALSE, include=FALSE}
# packages
library(magrittr)
library(stringr)
library(purrr)
```

```{r metadata-manuscript, echo = FALSE, include=FALSE}
# Manuscript metadata
manuscript1_yaml <- rmarkdown::yaml_front_matter("analysis/paper/001-paper-main.Rmd")
manuscript2_yaml <- rmarkdown::yaml_front_matter("analysis/paper/001-reply-main.Rmd")

## info to display

# manuscript1
manuscript1 <- list()
manuscript1$authors <-
  manuscript1_yaml$author %>%
  purrr::map_chr(function(x) paste0(x$given_name, " ", x$surname)) %>%
  knitr::combine_words()

manuscript1$year <- 
  paste0("(",  eval(parse(text = stringr::str_extract(manuscript1_yaml$date, pattern = "^\\d+"))), "):")

manuscript1$title <-
  manuscript1_yaml$title %>%
  paste0(".")

manuscript1$journal <-
  eval(parse(text = manuscript1_yaml$journal %>% stringr::str_remove("^`r ") %>% stringr::str_remove("`$"))) %>%
  names()

manuscript1$doi <-
  manuscript1_yaml$doi %>%
  paste0(".")
```

# hklmirs

This repository contains the data and code for our submitted manuscript:

> `r paste(manuscript1$authors, "(submitted)", manuscript1$title)`

### How to cite

Please cite this compendium as:

> Henning Teickner and Klaus-Holger Knorr, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for "`r manuscript1_yaml$title`"_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown source documents
    needed to reproduce the manuscript, including figures and tables. The main 
    script is [001-paper-main.Rmd](analysis/paper/001-paper-main.Rmd). Additional scripts are:
        - [002-paper-m-original-models.Rmd](analysis/paper/002-paper-m-original-models.Rmd): Computes the original models used in @Hodgkins.2018 and models with the same model structure, but as Bayesian models.  
        - [003-paper-m-gaussian-beta.Rmd](analysis/paper/003-paper-m-gaussian-beta.Rmd): Computes models assuming a Beta distribution for holocellulose and Klason lignin contents and compares them to the original models.
        - [004-paper-m-reduce-underfitting.Rmd](analysis/paper/004-paper-m-reduce-underfitting.Rmd): Extents the Beta regression models by including additional variables (additional peaks) or using a different approach (using measured spectral intensities of binned spectra instead of extracted peaks), and validates these models using LOO-CV.
        - [005-paper-m-minerals.Rmd](analysis/paper/005-paper-m-minerals.Rmd): Uses the models from `003-paper-m-gaussian-beta.Rmd` to test how accurate a model for holocellulose content is which is also calibrated on training samples with higher mineral contents.
        - [006-paper-m-prediction-domain.Rmd](analysis/paper/006-paper-m-prediction-domain.Rmd): Analyzes the prediction domain [@Wadoux.2021] of the original models and the modified models and identifie under which conditions models extrapolate for peat and vegetation smaples from @Hodgkins.2018.
        - [007-paper-m-prediction-differences.Rmd](analysis/paper/007-paper-m-prediction-differences.Rmd): Compares predictions for the training data and the peat and vegetation data from @Hodgkins.2018 for the original models from @Hodgkins.2018 and the modified models from `004-paper-m-reduce-underfitting.Rmd`.
        - [008-paper-supplementary.Rmd](analysis/paper/008-paper-supplementary.Rmd): Computes supplementary analyses and figures.
  - [:file\_folder: data](/analysis/data): Data used in the analysis. Note that raw data is not stored in [:file\_folder: raw_data](/analysis/data/raw_data) (empty folder), but in [:file\_folder: /inst/extdata](/inst/extdata). [:file\_folder: derived_data](/analysis/data/derived_data) contains derived data computed from the scripts. The raw data are derived from @Hodgkins.2018.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations (empty folder).

The other folders in this directory follow the standard naming scheme and function of folders in R packages. There are the following directories and files:

- `README.md`/`README.Rmd`: Readme for the compendium.  
- `DESCRIPTION`: The R package DESCRIPTION file for the compendium.  
- `NAMESPACE`: The R package NAMESPACE file for the compendium.  
- `LICENSE.md`: Details on the license for the code in the compendium.  
- `CONTRIBUTING.md` and `CONDUCT.md`: Files with information on how to contribute to the compendium.  
- `Dockerfile`: Dockerfile to build a Docker image for the compendium.  
- `.Rbuildignore`, `.gitignore`, `.dockerignore`: Files to ignore during R package building, to ignore by Git, and to ignore while building a Docker image, respectively.  
- `renv.lock`: renv lock file (Lists all R package dependencies and versions and can be used to restore the R package library using renv). `renv.lock` was created by running     `renv::snapshot()` in the R package directory and it uses the information included in the `DESCRIPTION` file.  
- `.Rprofile`: Code to run upon opening the R-project.  
- `R`, `man`, `inst`, `data-raw`, `data`, `src`: Default folders for making the R package run.

  - Folder `inst/extdata`: Folder with the raw data used for the analyses. All files in this folder are derived from @Hodgkins.2018.

## How to run in your broswer or download and run locally

You can download the compendium as a zip from from this URL: <https://doi.org/> ---todo

Or you can install this compendium as an R package, hklmirs, from GitHub with:

```{r installation, eval=FALSE}
remotes::install_github("henningte/hklmirs")
```

## How to use

#### Reproduce the analyses

To reproduce the analyses for the paper, open the Rstudio project included in this research compendium and run the Rmarkdown script in `analysis/paper/001-paper-main.rmd`.

Running the whole script takes about two hours and occupies additional disk space of ~2 Gb.

Alternatively, the [Dockerfile](Dockerfile) can be used to build a Docker image from which all analyses can be reproduced. The Dockerfile ensures that all required dependencies are installed (e.g. specific R packages; this is managed using the R package [renv](https://rstudio.github.io/renv/articles/renv.html)).

The [Dockerfile](Dockerfile) provides instructions how to build a Docker image from the Dockerfile and how to run the image in a Docker container. It occupies disk space of ~5 Gb.

When the Docker image runs in a container, go to `localhost:8787` in your Browser. You will find an RStudio interface where you can log in with username `rstudio` and password `hkl`. Here you can find the Rmarkdown scripts (`hklmirs/analysis/paper/001-paper-main.rmd`) as described above.

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse. See the sources section for licenses for data derived from external sources and how to give credit to the original author(s) and the source.

### Sources

All files in `inst/extdata` are derived from @Hodgkins.2018. These data are licensed under the [CC-BY 4.0](http://creativecommons.org/licenses/by/4.0/) license (see https://www.nature.com/articles/s41467-018-06050-2#rightslink).

The format of this research compendium is inspired by @Marwick.2018 and was created with rrtools [@Marwick.2019]. The Rmarkdown template for the main article is from the rticles package [@Allaire.2020].

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### References

### 2022-02-03 analysis/paper/001-paper-main.Rmd
Initial commit.

---
title: "Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra"
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'soil')`"
author:
  - given_name: Henning
    surname: Teickner
    email: henning.teickner@uni-muenster.de
    affiliation: 1
  - given_name: Klaus-Holger
    surname: Knorr
    email: kh.knorr@uni-muenster.de
    affiliation: 1
# If you have more than one corresponding author, add them manually using the following structure (note the commas):
# Two authors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# Three authors or more: Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry (j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)
# If the following line is uncommented, the "corresponding: true" above are ignored
correspongdingauthors: Henning Teickner (henning.teickner@uni-muenster.de)
# If an author is deceased, please mark the respective author name(s) with a dagger '†' and add a further affiliation; put the decease date in the 'address' field", see 'Nikolaus Copernicus' in template.
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: ILÖK, Ecohydrology and Biogeochemistry Group, University of Münster, Heisenbergstr. 2, 48149 Münster, Germany
abstract:  |
    To understand global soil organic matter (SOM) chemistry and its dynamics, we need tools to efficiently quantify SOM properties, for example prediction models using mid infrared spectra. All the advantages of such models rely on their critical validation. Recently, \cite{Hodgkins.2018} developed models to predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major compounds of organic matter in general. The models can be used to understand large-scale SOM gradients and have been used in various studies.  
    A research gap to fill is that these models have not been validated in detail yet. What are their limitations and how can we improve them? This study provides a validation with the aim to identify concrete steps to improve the models. As a first step, we provide several improvements using the original training data.  
    The major limitation we identified is that the original training data are not representative for a range of peat samples which causes both biased estimates and extrapolation uncertainty under the original models. In addition, the original models can in parctice produce unrealistic predictions (negative values or values $>100$ mass-%). Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid unrealistic predictions. Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples.  
    A key step to improve the models will be to collect training data that is representative for SOM formed under various conditions. This study opens directions to develop operational models to predict SOM holocellulose and Klason lignin contents from mid infrared spectra.
running:
  title: Improving Models to Predict Holocellulose and Klason Lignin Contents
  author: Teickner et al.
# This section is mandatory even if you declare that no competing interests are present.
competinginterests: |
  The authors declare that they have no conflict of interest.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section.
#copyrightstatement: |
#  The author's copyright for this publication is transferred to institution/company. 
### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #code: |
  #  use this to add a statement when having only software code available
  #data: |
  #  use this to add a statement when having only data sets available
  #sample: |
  #  use this section when having geoscientific samples available
  codedata: |
    Data and code to reproduce our analyzes are available via [---todo: add link].
#videosupplement: |
#  use this section when having video supplements available
authorcontribution: |
  HT (Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization); KH (Resources, Writing - Review & Editing, Supervision, Funding acquisition)
#disclaimer: |
#  We like Copernicus.
acknowledgements: |
  We like to thank @Hodgkins.2018 and @DelaCruz.2016 for provision of data and code to reproduce and modify their analyses, and @Allaire.2020 for provision of an Rmarkdown template. We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions.
#appendix: |
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_book:
    base_format: rticles::copernicus_article
    keep_tex: true
    citation_package: natbib
bibliography: references
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", 
                      fig.pos = "H")
```

```{r packages, echo=FALSE, include=FALSE}
# packages
library(ggplot2)
library(ir)
library(irpeat)
library(caret)
library(magrittr)
library(dplyr)
library(rstanarm)
library(brms)
library(loo)
library(projpred)
library(bayesplot)
library(patchwork)
library(ggrepel)
library(scales)
library(bib2df)
library(kableExtra)
library(purrr)
library(stringr)
library(tidyr)
```

```{r options-plot, echo=FALSE}
# plot theme
ggplot2::theme_set(ggplot2::theme_classic() + 
                     ggplot2::theme(strip.background = ggplot2::element_rect(fill = "lightgrey", colour = NA)))

# table print options
options(knitr.table.format = "latex")
```

\introduction[Introduction]

Understanding soil organic matter chemistry and how it changes is important to understand future global carbon dynamics. The chemistry of soil organic matter (SOM) controls how fast it can be decomposed [@Bengtsson.2018; @Shipley.2021]. To understand and predict these processes, we therefore need to measure SOM chemistry on a global scale. A challenge is that soils develop under diverse and changing environmental conditions which also affect SOM quality [@Scharlemann.2014; @Lehmann.2015]. As a consequence spatially and temporally resolved measurements on a global scale are needed. For this, we need methods to measure SOM chemistry efficiently.  
Mid infrared spectra (MIRS)-based models to predict SOM properties are a promising high-throughput approach which can replace more labor intensive or costly measurements [@ViscarraRossel.2006]. For example, MIRS have been used to predict elemental contents which otherwise are measured using e.g. combustion and gas chromatographic analysis of resulting gases. All the advantages of MIRS-based models rely on the accuracy of the computed models. This makes model validation a crucial step during model development.  
OM derived from plants and SOM is often characterized by step-wise chemical fractionation into holocellulose (acid soluble) --- a proxy for polysaccharides --- and lignin (acid insoluble) --- a proxy for aromatics --- [@DelaCruz.2016; @Elle.2019] and these variables are often important indicators for the chemical quality of OM in studies and models analyzing decomposition of SOM [@Leifeld.2012; @Biester.2014; @Worrall.2017; @Hodgkins.2018; @Bengtsson.2018; @Agren.1996a; @Bauer.2004; @Shipley.2021]. If these fractions can be predicted from MIRS, it may be possible to understand decomposition across larger scales and at higher spatial resolution.  
Several models to predict holocellulose and lignin contents in different OM types have been developed (some based on near infrared spectra) [summarized for lignin by @Elle.2019; for holocellulose see e.g. @Peltre.2011; @Sun.2011]. Most of these models consider only wood, material from few species, or specific vegetation organs [@Elle.2019]. Another problem is that most of these models cannot be easily reproduced and used since the raw data has not been published.  
Recently, @Hodgkins.2018 developed models to predict Klason lignin and holocellulose contents from attenuated total reflectance MIRS. The models have several advantages over existing approaches. The training data comprises several different OM types, such as paper products (cardboard, office paper, magazines, newspaper), leaves from diverse species (trees and graminoids), and wood samples [@DelaCruz.2016; @Hodgkins.2018]. Moreover, both raw data and model code are publicly available [@Hodgkins.2018; @Teickner.2020a]. For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020], most recently in @Verbeke.2022, and for future developments, e.g. by including additional data.   
A problem is that neither @Hodgkins.2018, nor a later study using the models provided a thorough model validation. To compute the models, the authors developed a procedure to extract peaks which are indicative for specific OM fractions, such as holocellulose and Klason lignin (Fig. \@ref(fig:hodgkins-peaks-sample)), from the MIRS. The peaks are baseline corrected, their maximum height computed and divided by the sum of absorbance values in the spectra. These values are used to calibrate the prediction models. As model validation, only the linearity between the target variables and predictors was assessed (supplementary material to @Hodgkins.2018), but not for example how predicted values match measured values. Moreover, no information has been provided if the training data is representative for SOM, particularly the peat samples analyzed in the study. Similarly, it is unclear if pre-selecting peaks may reduce the predictive accuracy of the models. Many of these concerns have also been raised by an anonymous reviewer of the paper (supplementary information to @Hodgkins.2018). Thus, an important research gap to fill is to validate the models and to provide concrete strategies for further improvements.

```{r hodgkins-peaks-sample, echo=FALSE, out.width="70%", fig.height=2.5, fig.width=6, fig.cap="Sample spectrum with the peaks and troughs, their heights, and their baselines (green) as detected by the script developed by \\cite{Hodgkins.2018}. The plot was created with the data and code from \\cite{Hodgkins.2018} as implemented in ir and irpeat, respectively."}
seed <- 1
temp <- 
  ir::ir_sample_data[36, ] %>% 
  ir::ir_bc(method = "rubberband") %>%
  ir::ir_normalise(method = "area") %>%
  irpeat:::irp_content_klh_hodgkins_prepare() %>%
  hklmirs::hkl_hodgkins_ggplot()
temp[[1]] +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

Our goals are to identify limitations of the original models and to give concrete recommendations for improvements. Moreover, we use the original data from @Hodgkins.2018 to provide improvements where possible. To this end, we conducted an exploratory analysis using the peat and peat-forming vegetation data provided by @Hodgkins.2018. Our exploratory analysis was guided by the following research questions:

1. _Is the normal distribution reasonable to predict holocellulose and Klason lignin contents for peat samples?_ Klason lignin and holocellulose contents as actually measured for the training data set in @Hodgkins.2018 cannot be negative or larger than 100 mass-%. For this reason, predictions and prediction uncertainties of the models should be in the interval [0, 100] mass-%. Yet, the original models assume a normal distribution which does not entail any lower or upper bound for predicted values. This can potentially result in unrealistic predictions. However, if no unrealistic predictions occur for "representative" peat samples, using a model with normal distribution can be justifiable in practice due the central limit theorem. In contrast, if unrealistic predictions occur, it is more reasonable to use a distribution which assumptions are in accordance with knowledge on the data generating process. For example, the beta distribution assumes a lower and upper bound which can be mapped to the interval [0, 100] mass-%. For this reason, we were interested if and under which conditions the original models may produce unrealistic predictions for the peat and vegetation samples.

2. _Can the predictive accuracy of the models be improved?_ The original models were parameterized with manually selected variables (normalized peak heights from the extraction procedure) [@Hodgkins.2018]: the peak `carb` (~1090 cm$^{-1}$) for holocellulose and the sum of the peaks `arom15` (~1515 cm$^{-1}$) and `arom16` (~1650 cm$^{-1}$) (`arom15arom16`) for Klason lignin. Whilst there is both a strong causal rationale behind this decision and it may support the robustness of the models [@Hodgkins.2018], it is also under risk of underfitting --- excluding informative variables and thus missing an opportunity to improve predictive accuracy. For this reason, we were interested if including more variables extracted from the MIRS results in models with a better predictive accuracy.

3. _Is the prediction domain of the training data representative for peat samples?_ If a model extrapolates outside the range of predictor variable values in the training data set, it is unclear if the predictions are valid, for instance because it cannot be evaluated if there is bias. An advantage of the independent reference sample set used by @Hodgkins.2018 is that the model is applicable on various OM types and e.g. no problem-specific correlations are used by the models to make predictions [@Hodgkins.2018]. For example, in peat, holocellulose content is controlled by decomposition, and other structures also controlled by decomposition might therefore provide information on holocellulose contents [@Leifeld.2012; @Biester.2014]. But this does not have to be the case for other samples, e.g. peat forming vegetation. An advantage is therefore the generality of the training data set @Hodgkins.2018 used. However, since it does not include peat samples, it is questionable if the prediction domain formed by the training samples covers peat samples and peat forming vegetation in general, and this can result in overlooking prediction failures due to extrapolation. For this reason, we were interested if the training data (its predictions domain, @Wadoux.2021) is representative for peat and peat forming vegetation.

4. _Do predictions of the improved models differ and if so, why?_ Computing modified models enables us to compare their predictions to those of the original models using the same training data, and to the actually measured contents. Model comparison often reveals detailed insights into the mechanics of a model and may provide hints to which model makes more correct predictions. We therefore were interested in analyzing differences between the models computed, what factors may cause differences in predictions, and if there are indications which models make more correct predictions.

5. _Can holocellulose contents also be predicted for samples with mineral interferences?_ During model calibration, @Hodgkins.2018 omitted some samples. For prediction of holocellulose, old magazine samples were omitted because the `carb` peak suffered from mineral interference from the samples' clay coatings [@Hodgkins.2018]. This decision is sensible, but since there are many organic soils with mineral admixtures --- for example due to volcanic ash, in base layers, or due to cryoturbation) [e.g. @Broder.2012; @Bockheim.2007; @Koven.2009; @Loisel.2014] --- we think that it is useful to have a model to predict holocellulose contents also for mineral rich samples. We therefore wanted to investigate if a suitable model can be computed by including the previously omitted old magazine samples during calibration.

In investigating these issues, our main goals are to provide a concrete plan for how to improve the original models. Moreover, we want to analyze under which conditions it may not be appropriate to use the original or modified models. Where possible using the original data, we want to provide improved models which we hope can be further improved in the future. In addition, this study provides general guidance for pitfalls during validation of spectral prediction models. With this, we want to contribute to the development of models to predict SOM holocellulose and Klason lignin contents which are important to provide diverse data to fit SOM decomposition models and to understand how environmental change affects global soil carbon dynamics.

# Methods

```{r opts, echo=FALSE}
# define global parameter values
k_threshold <- 0.7
seed <- 1
iter <- 4000
warmup <- iter %/% 2
chains <- 4
```

```{r data-import, echo=FALSE}
### load and preprocess the data

## training data

# get training data as ir object
d <- 
  ir::ir_sample_data %>%
  ir::ir_bc(method = "rubberband") %>%
  ir::ir_normalise(method = "area") %>%
  dplyr::mutate(
    hol = units::drop_units(holocellulose),
    kl = units::drop_units(klason_lignin)
  )

# flatten
d_flat <- irpeat:::irp_content_klh_hodgkins_prepare(d)

## data for peat samples
d_peat <- 
  list("../../inst/extdata/sample_mir.csv") %>%
  ir::ir_import_csv(sample_id = "from_colnames") %>%
  dplyr::left_join(
    read.csv("../../inst/extdata/sample_data.csv", header = TRUE), 
    by = "sample_id"
  ) %>%
  dplyr::arrange(core_label, sample_depth_lower)

# preprocessing
d_peat <-
  d_peat %>%
  ir::ir_bc(method = "rubberband") %>%
  ir::ir_normalise(method = "area")

## data for vegetation samples
d_veg <- 
  list("../../inst/extdata/Hodgkins_vegetation_mirs.csv") %>%
  ir::ir_import_csv(sample_id = "from_colnames") %>%
  dplyr::left_join(
    read.csv("../../inst/extdata/Hodgkins_vegetation_sample_info.csv", header = TRUE) %>%
      dplyr::rename(sample_id = "sample_name"), 
    by = "sample_id"
  )

# preprocessing
d_veg <-
  d_veg %>%
  ir::ir_bc(method = "rubberband") %>%
  ir::ir_normalise(method = "area")
```

Our aim is to provide a validation of the original models of @Hodgkins.2018 and concrete suggestions for improvements. To this end, we conducted a series of statistical analyzes to query each of our research questions listed in the introduction. During this, we also computed improved models using the same training data that was used to compute the original models. Since we used these improved models to uncover and analyze the limitations of the original models, we first describe steps how we improved the original models and then how we analyzed the limitations.  
We use the data and models collected and developed by @Hodgkins.2018. The training data used to calibrate the original models consist of various organic materials and comprise measured values for holocellulose and Klason lignin [@DelaCruz.2016]. The data are available via @Hodgkins.2018. In addition, we used a data set consisting of MIRS for `r length(unique(d_peat$core_label))` peat cores (`r nrow(d_peat)` samples in total) and `r nrow(d_veg)` vegetation samples collected at the same sites which are also provided by @Hodgkins.2018. We used these peat and vegetation data to evaluate the performance of the original and improved models for peat and peat forming vegetation in general. We assume that the peat samples are representative for a wide range of SOM because they cover a large latitudinal range and a wide range of degrees of decomposition, and were formed by different vegetation [@Hodgkins.2018].  
We computed all models as Bayesian models. This allowed us to consider parameter uncertainty in predictions and to compute models with good predictive performance [@Raftery.2003; @Piironen.2017]. For this reason, we also recomputed the original models as Bayesian models with weakly informative priors to facilitate model comparison using the same metrics and tools. Whenever we refer to "original model", we mean these Bayesian translations, and otherwise refer to "original non-Bayesian models". To show that we did not introduce any of the identified weaknesses using the Bayesian approach, we compared the predictions of the Bayesian models to the original non-Bayesian models (supplementary Fig. 1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).  
All analyzes were performed in R (`r R.Version()$version.string %>% stringr::str_remove("^R ") %>% stringr::str_remove("\\)") %>% stringr::str_replace(" \\(", ", ")`) [@RCoreTeam.2020]. Spectra were preprocessed using ir (`r packageVersion("ir")`) [@Teickner.2020]. The original script used by @Hodgkins.2018 to extract peaks (\url{https://github.com/shodgkins/FTIRbaselines}) as implemented in irpeat (`r packageVersion("irpeat")`) [@Teickner.2020a] was used to recompute the original models and extract peaks from MIRS. Bayesian models were computed using rstanarm (`r packageVersion("rstanarm")`) [@Goodrich.2020] and brms (`r packageVersion("brms")`) [@Burkner.2017; @Burkner.2018] which are interfaces to Stan (`r rstan::stan_version()`) [@StanDevelopmentTeam.2020; @StanDevelopmentTeam.2021]. All predictor variables were $z$-transformed. MCMC sampling was validated in addition using bayesplot (`r packageVersion("bayesplot")`) [@Gabry.2020]. The out-of-sample predictive performance of the models was estimated using the expected log predictive density (ELPD) estimated using Pareto smoothed importance sampling-leave-one-out cross-validation (PSIS-LOO) [@Vehtari.2017] using the loo package (`r packageVersion("loo")`) [@Vehtari.2019]. PSIS-LOO ELPD is an universal measure to compare the predictive performance of models and $-2 \cdot \text{PSIS-LOO ELPD}$ is the deviance [@Vehtari.2017].

```{r m-original-models-child, child="002-paper-m-original-models.Rmd", cache=TRUE, include=FALSE}
# get results from child document
```

## Normal versus beta regression models

Above, we mentioned that assuming a model with normal distribution can result in unrealistic predictions outside the interval [0, 100] mass-%. We were interested if and under which conditions unrealistic predictions can occur. Here, we differentiate two ways in which a prediction can be unrealistic: It can be an unrealistic point estimate (median predictions), and the prediction interval can cover values outside [0, 100] mass-%, even though the median prediction can be in [0, 100] mass-%.  
For our analysis, we first computed a sequence of normalized `carb` and `arom15arom16` peak heights such that the original model predict values completely covering the interval [0, 100] mass-% for holocellulose and Klason lignin, respectively. For these simulated peak heights, we computed median model predictions and 90% prediction intervals using the original models ("original Gaussian models"). Finally, we identified unrealistic predictions in terms of the point estimates and 90% prediction intervals, and the `carb` and `arom15arom16` peak heights under which these occur. We used these values as thresholds to decide if the models make unrealistic predictions in practice. To relate these values to real data, we computed median predictions and 90% prediction intervals also for the peat and vegetation samples and identified unrealistic predictions as for the simulation analysis.  
After identifying conditions under which unrealistic predictions occur, we compared this to the behavior of a model with assumptions in line with the data generating process. Beta regression models assume all values of the dependent variable to be in (0,100) mass-% which is a reasonable assumption given how the actual data is generated. Therefore, we repeated the previous analyses, but now using beta regression models. We used the same model structure and priors for intercept and slopes as the original models ("original beta models").  
To facilitate practical comparison between the modeling approaches, we compared the predictions of all models in the form of depth profiles of the holocellulose and Klason lignin contents. Altogether, this allowed us to identify unrealistic predictions of the original models, and to analyze how the improved models (beta regression models) perform in comparison.

```{r m-gaussian-beta-child, child="003-paper-m-gaussian-beta.Rmd", cache=TRUE, include=FALSE}
# get results from child document
```

## Reducing underfitting with more variables

The original approach of @Hodgkins.2018 was to include only manually selected variables which may result in underfitting and consequently relatively low predictive accuracy. We wanted to investigate if the predictive accuracy can be improved by including more variables in the models. Two approaches with complementary advantages and disadvantages were tested:

1. We used all peak heights and trough heights returned by the peak identification algorithm of @Hodgkins.2018 (Fig. \@ref(fig:hodgkins-peaks-sample)). The advantage is that this makes the prediction model probably more robust against calibration transfer issues in comparison to the second approach (a hypothesis that remains to be tested). The disadvantage is that additional information contained within the spectra cannot be fully exploited if these are not contained within the extracted peak heights. To avoid overfitting, we computed models using priors implying different amount of regularization for model coefficients (standard deviation of 1 and 0.5, respectively) in addition to the default weakly informative prior which we derived from rstanarm (`r packageVersion("rstanarm")`) (standard deviation of 2.5) [@Goodrich.2020].

2. We used all variables in the spectra and Bayesian regularization [@Piironen.2020], similar to [@Teickner.2022]. The advantage is that this approach can more fully exploit the information contained within the spectra because it is not restricted to specific peaks, resulting potentially in a better predictive performance. The disadvantage is that this approach may be more prone to calibration transfer issues. To reduce redundancy in the variables and impact of noise, we binned the spectra, testing bin widths of 10, 20, 30, 50, and 100 cm$^{-1}$. As regularizing priors, we used the regularized horseshoe prior with a parameterization as described in @Piironen.2017b, assuming a number of relevant variables of eight.

Out-of-sample model performance was estimated with PSIS-LOO ELPD as described above [@Vehtari.2017; @Vehtari.2019]. If a model has a larger PSIS-LOO ELPD than another model fitted on the same data, it has a larger (leave-one-out cross-validation) predictive accuracy, so models with larger PSIS-LOO ELPD are preferred [@Vehtari.2017]. To compare models, one typically computes the posterior distribution of the difference in PSIS-LOO ELPD between the model with the largest PSIS-LOO ELPD and the other models, such that one can evaluate the probability of a certain predictive advantage of the best model in comparison to the other models [@Vehtari.2017; @Vehtari.2019].  
To further validate and interpret the models using binned spectra, we visually identified the most important bins (largest absolute median coefficients $\ge0.2$) and linked these to molecular structures and causal relations that most probably result in the correlation with holocellulose and Klason lignin, respectively.

```{r m-reduce-underfitting-child, child="004-paper-m-reduce-underfitting.Rmd", cache=TRUE, include=FALSE}
# get results from child document
```

## Assessing the prediction domain of the training data

Is the training data used to compute the models representative for the spectral properties of peat and peat forming vegetation? To answer this question, we needed to compare the spectral properties of the training samples to those of the peat and vegetation samples [@Wadoux.2021]: If for a sample the value of a spectral variable included in a model exceeds the range of values for the same variable in the training data, the model extrapolates if applied on the sample [@Wadoux.2021]. If extrapolation occurs, it is unclear if the predictions are valid estimates for holocellulose and Klason lignin contents.  
For the original models, the prediction domain is simply the range of area normalized heights of the `carb` and `arom15arom16` peak, respectively [@Hodgkins.2018]. Therefore, we could directly compare the area normalized heights of the `carb` and `arom15arom16` peak of the peat and vegetation samples to the range of area normalized heights of the same peaks for the training data.  
For the improved models using binned spectra, the predictor variables represent a multivariate prediction domain. We therefore created plots with which we could compare the value ranges for each bin for the training data with the respective values in the peat and vegetation spectra. This allowed us to identify for which spectral variables extrapolation is an issue. Finally, by identifying the most important variables in the improved models, we could qualitatively summarize how large the risk of this extrapolation is.  
We did not additionally investigate the prediction domain for the models using all extracted peaks and troughs since these models had no predictive advantage in comparison to the models using binned spectra.

## Analyzing differences in predictions between the original models and the improved models 
We analyzed how predictions of the improved models (best model for each of the two approaches) differ from the original models. We found considerable differences, even within the spectral range of the training data and therefore analyzed what factors probably cause these differences.  
In a first step, we were interested if the original or improved models are biased. @Hodgkins.2018 provide no plot of residuals against predicted values or measured values against predicted values. Both plots are commonly used to detect any biases in a model fit [e.g. @Pineiro.2008]. We therefore created such plots for the original model and the improved model and compared indication of bias between both.  
In a second step, we compared how predictions of both models differ in practice. To this end, we compared predictions of both for the peat and vegetation data from @Hodgkins.2018, similarly to how we compared the Gaussian and beta regression models. Since we found considerable differences, we conducted a more targeted exploratory analysis to identify factors probably causing these differences.

## Predicting holocellulose contents in samples with mineral admixtures

In the introduction we mentioned that the original @Hodgkins.2018 model for holocellulose excluded training samples with admixtures of clay minerals (old magazine samples) since these interfere with the `carb` peak. By including more variables into the model than the `carb` peak (as described in the previous section), we tried to compute models that can also describe the holocellulose content in the training data samples with high clay content. For this, we fitted the best models for holocellulose from each of the two approaches described in the previous section again (binned spectra: `r loo_comparison_hol$bin_width[[1]]` cm$^{-1}$ bin width; extracted peaks: priors with standard deviation =`r loo_comparison_hol$prior_scale[loo_comparison_hol$variables == "peaks"][[1]]`), but this time including the four old magazine samples previously left out. We interpreted the coefficients of these binned models similarly to the versions not fitted with mineral rich samples.

```{r m-minerals-child, child="005-paper-m-minerals.Rmd", cache=TRUE, include=FALSE}
# get results from child document
```

```{r m-prediction-domain-child, child="006-paper-m-prediction-domain.Rmd", cache=TRUE, include=FALSE}
# get results from child document
```

```{r m-prediction-differences-child, child="007-paper-m-prediction-differences.Rmd", cache=TRUE, include=FALSE}
# get results from child document
```

# Results and discussion

## Is the normal distribution reasonable to predict holocellulose and Klason lignin contents for peat samples?

The original model for holocellulose produced unrealistic predictions: It had a negative point prediction for one strongly decomposed sample from a tropical peat core. Moreover, for `r d_peat_pred %>% dplyr::filter(variable_y == "Holocellulose" & model == "Gaussian, Bayesian") %>% dplyr::mutate(negative = lwr < 0) %>% dplyr::summarise(negative = sum(negative)/length(negative)) %>% dplyr::pull(negative) %>% magrittr::multiply_by(100) %>% round(0)`% of the peat samples in the data from @Hodgkins.2018, the lower 90% prediction interval covered negative values (supplementary Fig. 3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->). No point prediction or 90% prediction interval was $>100$ mass-%. The original model for Klason lignin did not produce unrealistic predictions for the peat samples (supplementary Fig. 3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->). Both models did not produce unrealistic predictions for the vegetation samples.   
Figure \@ref(fig:res-p-gaussian-beta-range) shows predicted medians and prediction intervals for both models across a potential range of the spectral predictor variables alongside the range covered by the peat and vegetation samples. The model with normal distribution produces negative lower prediction interval limits whenever the predicted median is lower than around `r d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose" & variable_limit == "fit_lwr") %>% dplyr::pull(fit) %>% magrittr::multiply_by(100) %>% round(0)` mass-% and upper prediction interval limits $>100$ mass-% whenever the predicted median is larger than around `r d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose" & variable_limit == "fit_upr") %>% dplyr::pull(fit) %>% magrittr::multiply_by(100) %>% round(0)` mass-%. For Klason lignin these values are around `r d_range_pred_limits %>% dplyr::filter(variable_y == "Klason lignin" & variable_limit == "fit_lwr") %>% dplyr::pull(fit) %>% magrittr::multiply_by(100) %>% round(0)` and `r d_range_pred_limits %>% dplyr::filter(variable_y == "Klason lignin" & variable_limit == "fit_upr") %>% dplyr::pull(fit) %>% magrittr::multiply_by(100) %>% round(0)` mass-%, respectively.  
The beta regression models produce realistic predictions across the complete range of the spectral predictor variables. In addition, the beta regression model has smaller prediction uncertainties for extreme values (Fig. \@ref(fig:res-p-gaussian-beta-range)). A consequence of this are larger predicted holocellulose contents with narrower prediction intervals for several (decomposed) peat samples under the beta regression model in comparison to the original normal model (Fig. \@ref(fig:res-p-gaussian-beta-range) and supplementary Fig. 3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->).  
Thus, whereas our results indicate little differences in choosing a model distribution for Klason lignin, for holocellulose contents of peat samples it is crucial to use a beta regression model to avoid unrealistic predictions. Nevertheless, it is generally advisable to use a beta regression model for mass contents. It may not be known in advance how large holocellulose or Klason lignin contents are for a given sample. And even for Klason lignin, contents may be low, e.g. due to high contents of minerals.

```{r res-p-gaussian-beta-range, warning=FALSE, echo=FALSE, out.width="70%", fig.height=3.5, fig.width=6.5, fig.cap="Predicted Holocellulose (a) and Klason lignin (b) contents using either the original Gaussian models or a beta regression model, versus predicitons of the original Gaussian model covering the entire interval [0, 100] mass-\\%. Coloured lines within shaded regions are median predictions. Shaded areas and boundaries are 90\\% prediction intervals. Coloured points are median point predictions for the peat (round points) and vegetation (crosses) samples from \\cite{Hodgkins.2018}. Colors differentiate predictions of the original Gaussian model (blue) and the model with beta distribution (red). White filled points indicate the fitted values for which the 90\\% prediction interval of the Gaussian model contains unrealistic values. Dashed grey lines represent the limit for realistic point predictions."}
p_gaussian_beta_range
```

## Can the predictive accuracy of the models be improved?

Our analysis shows that both strategies to include more variables result on average in more accurate predictions. If uncertainties for the estimated predictive accuracy are considered, the advantage is clear only for the models for holocellulose using binned spectra (Table \@ref(tab:res-tab-loo-comparison-res)), but not for Klason lignin. Consequently, using the data available, we could improve the model for holocellulose, but our analysis remains inconclusive with respect to Klason lignin.  
Is one of the strategies to include more variables advantageous? The on average overall best models, according to their PSIS-LOO ELPD, were those using binned spectra with bin widths of `r loo_comparison_hol$bin_width[[1]]` and `r loo_comparison_kl$bin_width[[1]]` for holocellulose and Klason lignin, respectively (Table \@ref(tab:res-tab-loo-comparison-res)). For Klason lignin, comparatively large prediction uncertainties result in the practical equivalence of both strategies. For holocellulose, using binned spectra (except bin width = 20) is more clearly better than using the extracted peaks. Summarized, this indicates that using binned spectra results in an improved predictive accuracy over the second approach to use all extracted peaks for holocellulose, whereas our validation remains inconclusive with respect to Klason lignin.  
To interpret the improved models using binned spectra, we plotted the median coefficients for the best models using binned spectra (Fig. \@ref(fig:res-p-coefficients)). From this plot, we identified the bins with the largest absolute median coefficients ($\ge0.2$, marked with points and labeled). The reason for the (on average) improved performance is that the improved models can use information contained in the complete spectra which is not contained in the manually selected peaks for the original models:  
Similarly to the original model, a variable near the `carb` peak is important to predict the holocellulose content (bin at ~950 cm$^{-1}$) since it is related to cellulose [@Cocozza.2003; @Stuart.2004; @Artz.2008], but two peaks related to aromatic structures are also important (~1200 cm$^{-1}$: C-O stretching in phenols and aromatic ethers [@Stuart.2004; @Kubo.2005]. ~1250 cm$^{-1}$: aromatic in plane C-H bending) [@Stuart.2004] (Fig. \@ref(fig:res-p-coefficients), first panel). This indicates that aromatics provide information on the holocellulose content in the training samples. A plausible explanation is that holocellulose and Klason lignin together are the major mass fractions in many OM types and both are controlled by the same processes, but often in different directions (decomposition, resource allocation trade-off during plant growth, selective removal during processing) [@Biester.2014; @Chen.2016] (supporting Fig. 4<!---\@ref(fig:res-sup-kl-hol)--->).   
For Klason lignin, the improved model does not use bins near to the `arom15` and `arom16` peaks, in contrast to the original model. Instead bins with large absolute coefficients are located in the fingerprint region and probably related to aromatic in plane C-H bending (~1270 and ~1160 cm$^{-1}$) [@Stuart.2004] (Fig. \@ref(fig:res-p-coefficients), last panel). We have no explanation why one coefficient is negative and the other positive. In section "What causes differences in predictions?", we provide a mechanistic interpretation for why the selected variables are important.

```{r res-tab-loo-comparison-res, echo=FALSE}
loo_comparison_res[[1]] <- 
  loo_comparison_res[[1]] %>%
  stringr::str_replace("\\\\toprule", "\\\\tophline") %>%
  stringr::str_replace("\\\\midrule", "\\\\middlehline") %>%
  stringr::str_replace("\\\\bottomrule", "\\\\bottomhline") %>%
  stringr::str_replace_all("\\\\addlinespace.+\\n", "") %>%
  stringr::str_remove_all("\\\\hspace\\{1em\\}") %>%
  stringr::str_replace_all("m1\\.", "\\\\hspace\\{1em\\}m1\\.") %>%
  stringr::str_replace_all("m2\\.", "\\\\hspace\\{1em\\}m2\\.")
loo_comparison_res
```

\clearpage

```{r res-p-coefficients, echo=FALSE, out.width="90%", fig.height=3, fig.width=6.5, fig.cap='Coefficients of the best models using binned spectra for Holocellulose (first two panels, model trained without and with samples containing minerals, respectively) and Klason lignin (last panel), plotted against the average wavenumber of the bins. Points are median coefficient estimates, error bars are 90\\% posterior intervals. Points with an absolute median coefficient $\\ge 0.2$ are labeled with the respective average wavenumber value. The grey shaded region is a reference spectrum.'}
p_coefficients
```

## Is the prediction domain of the training data representative for peat samples?

The training data set does not cover the range of the spectral variables relevant for predicting peat and peat forming vegetation holocellulose contents (Fig. \@ref(fig:res-p-prediction-domain) A and B). For the original models, the training data `carb` and `arom15arom16` peak heights cover the range of the vegetation samples. However, few peat samples have larger `arom15arom16` and many peat samples have smaller `carb` peak heights than covered by the training data. This suggests that the original models extrapolate to predict holocellulose contents in peat samples with small holocellulose contents and Klason lignin contents in peat samples with large Klason lignin contents.  
The same is true for the improved models using binned spectra (Fig. \@ref(fig:res-p-prediction-domain) C and D). Even though it is not possible to establish a direct relation between extrapolation for individual spectral variables and predicted contents, it is visible that for several bins, peat samples --- and to a lesser extent --- vegetation samples have mostly larger standardized predictor values than covered by the training data set. This is also true for bins with the largest absolute median coefficients (including the model for holocellulose fitted with mineral-rich samples; see section "Can holocellulose contents also be predicted for samples with mineral interferences?" below).  
Under what conditions does extrapolation occur? Samples with small `carb` peak, or for the binned spectra, large absorbance at ~`r m_pars %>% dplyr::filter(abs(median) >= 0.02 & variable_y == "Holocellulose (with minerals)") %>% dplyr::pull(wn) %>% sort() %>% knitr::combine_words()` cm$^{-1}$, are outside the prediction domain for holocellulose (Fig. \@ref(fig:res-p-prediction-domain)). Since the `carb` peak is small and absorbance at the selected wavenumbers is typically large in decomposed samples [@Cocozza.2003; @Tfaily.2014; @Hodgkins.2018] (and the respective coefficients are negative), this indicates that for holocellulose, extrapolation occurs mainly for more decomposed peat. 
This is also true for Klason lignin since extrapolation occurs for samples with large `arom15arom16` peak, or for the binned spectra, large absorbance at ~`r m_pars %>% dplyr::filter(abs(median) >= 0.02 & variable_y == "Klason lignin") %>% dplyr::pull(wn) %>% sort() %>% knitr::combine_words()` cm$^{-1}$, all of which are larger for more decomposed peat [@Cocozza.2003; @Tfaily.2014; @Hodgkins.2018].  
Overall, this indicates that the prediction domain formed by the training data does not cover peat --- particularly decomposed peat --- and partly peat forming vegetation samples. This is probably also true for non-peat SOM. Therefore the models' prediction can represent extrapolations in practice.

```{r res-p-prediction-domain, echo=FALSE, out.width="40%", fig.height=7, fig.width=4, fig.cap='Prediction domain for the original models (a, b), and the improved models using binned spectra (c, d). The prediction domain is the range of values covered by the predictor variables. The original models use only one predictor variable and therefore the predicition domain can be shown as spread of the training samples across the x axis (standardized height of the \\texttt{carb} and \\texttt{arom15arom16} peak, respectively) of a histogram (a, b; blue bars). Extrapolation occurs if other samples (peat, vegetation) exceed this range. Counts are number of samples in the respective datasets. The improved models include bins across the complete spectra (c, d). Therefore, the range in standardized predictor variable values (y axis) covered by the training data for each bin is shown as orange shaded area for the dataset with a bin width of 50 (c) and 10 cm$^{-1}$ (d), respectively, as used by the best improved models for holocellulose and Klason lignin. Extrapolation occurs if other samples (blue shaded regions) exceed these areas. Dashed lines indicate the most important variables for the holocellulose (c) and Klason lignin model (d), respectively (compare with Fig. \\ref{fig:res-p-coefficients})'}
p_prediction_domain
```

## How do predictions of the original and improved models differ?

There were considerable differences in predictions of the original and improved models using binned spectra for both Klason lignin and holocellulose for the peat samples (Fig. \@ref(fig:res-p-pred-diff-all1) and supplementary Fig. 5 <!---\@ref(fig:res-p-gaussian-beta-depth-profile-best)--->). For holocellulose, the original model tends to predict larger contents for peat, especially for samples with larger holocellulose contents (Fig. \@ref(fig:res-p-pred-diff-all1)). Similar, albeit less pronounced patterns are visible for the vegetation and training data. For Klason lignin, the improved model predicts up to 25 mass-% larger Klason lignin contents for some samples, but for others up to 10 mass-% smaller contents than the original model. This happens particularly in a region where the original model predicts Klason lignin contents between 25 and 35 mass-% (Fig. \@ref(fig:res-p-pred-diff-all1)).  
This is surprising for two reasons: First, for the training data, both models make quite similar predictions (Fig. \@ref(fig:res-p-pred-diff-all1)). The peat samples have `aro15arom16` values similar to the training data (Fig. \@ref(fig:res-p-prediction-domain)), but nevertheless both models produce contrasting predictions. The same is true for holocellulose for a relatively large fraction of peat samples (Fig. \@ref(fig:res-p-prediction-domain)). Second, both models have comparatively small bias within the range of the training data relevant to peat samples (supplementary Fig. 6 <!---\@ref(fig:res-p-y-yhat-best)--->). Thus, predictions are different even within the range covered by the training data where the models actually make similar predictions and these differences are not due to misfit of either model to the training data. This indicates that the spectra of the training data are not entirely representative for the peat samples, even if they are within the prediction domain (compare also with the previous section). What causes these differences?

```{r res-p-pred-diff-all1, echo=FALSE, out.width="80%", fig.height=6, fig.width=6, fig.cap='Predicted values of the original model versus predicted values of the original model (first column), and the best improved models using extracted peaks (second column) and binned spectra (last column), respectively, for holocellulose (a) and Klason lignin (b), respectively. Colours differentiate the training, peat, and vegetation data from \\cite{Hodgkins.2018}.'}
p_pred_diff_all1
```

## What causes differences in predictions?

We hypothesize that both the original and the improved models are not unbiased for samples with other spectral properties, even if these differences occur in variables not directly included in the models. For holocellulose, we could not find indications which model is better. For Klason lignin, we provide mechanistic evidence for why the prediction of the improved models probably are more correct. Therefore, a key result of our analysis is that additional training and validation data is required to compute models to predict SOM holocellulose and Klason lignin contents from MIRS.

### Holocellulose

In comparison to the training data, spectra for peat typically have larger absorption values at around 1200 cm$^{-1}$ and between (~1200 to 1500 cm$^{-1}$), but a smaller absorption in the region of the "OH-peak" (~3400 cm$^{-1}$) (supplementary Fig. 7 <!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).  
These spectral differences are directly related to the differences in predicted values of the original and improved model (supplementary Fig. 8 <!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->): The smaller the OH-peak is, the larger are predicted values of the original model in comparison to the improved model using binned spectra, especially for larger `carb` peak heights. Similarly, the larger the trough at 1200 cm$^{-1}$ or the absorbance at 1550 cm$^{-1}$ is, the larger are predictions of the original model in comparison to the improved model using binned spectra. This indicates that both the delineation of the `carb` peak and predictions using binned spectra are potentially sensitive to these differences and that this causes the observed differences in the predictions even in the range of the training data where both models yield similar predictions for the training samples.  
What causes this sensitivity? Peak heights --- for the original model --- or spectral variables (bins) --- for the improved models using binned spectra --- are normalized by the sum of absorbance values in the complete spectra [@Hodgkins.2018]. This is necessary to make spectra comparable, but also renders inferences sensitive to absorbance of other peaks not directly considered or with low influence in the original and improved models, respectively. Thus, if the OH-peak is smaller, the normalized height of the `carb` peak is larger than for the same spectrum with a larger OH-peak. Likewise, for the improved model using binned spectra a smaller OH-peak results in larger absorbances around ~1200 to 1600 cm$^{-1}$ where the most influential variables are located (supplementary Fig. 7 <!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)---> and 2<!---\@ref(fig:res-sup-spectra-sim-p))--->).  
Why do peat samples have such different spectral properties? We hypothesize the following mechanistic explanation for the differences: Decomposition of peat results in distinct changes in the absorbance at specific wavenumbers, e.g. the "OH-peak" and the fingerprint region [e.g. @Cocozza.2003]. For example, decomposition of phenols and disruption of tissue structures with hydrogen bonds result in a smaller absorbance around the OH-peak [@Kubo.2005; @Schellekens.2015]. The training data contain only undecomposed or industrially processed OM [@DelaCruz.2016; @Hodgkins.2018] which do not reflect such decomposition changes and therefore generally have a larger OH peak (middle row in supplementary Fig. 7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->). For decomposed OM, the same `carb` peak height or absorbance at specific wavenumbers therefore represents different holocellulose contents than for the training data.    
A consequence therefore is that it is questionable if the models (both original and improved) can be applied to peat and other SOM samples in general.

### Klason lignin

For Klason lignin, whereas the original model is unbiased across all samples (supplementary Fig. 6<!---\@ref(fig:res-p-y-yhat-best)--->), it is not within two classes of training samples. Samples of these classes differ in the relative contribution of the `arom15` and `arom16` peaks to `arom15arom16` (the sum of their heights) (Fig. \@ref(fig:res-p-pred-diff-train1-p2)). The first class (class 1) has high contributions of the `arom15` peak ($>30$%) and typically smaller `arom15arom16` values; the second class (class 2) a smaller contribution of the `arom15` peak and larger `arom15arom16` values. Samples of both classes can have large or small actual Klason lignin contents (Fig. \@ref(fig:res-p-pred-diff-train1-p2)). What we observed is that the model makes biased predictions for both classes: For small `arom15arom16` values, it underestimates contents, but for larger values, it increasingly overestimates Klason lignin contents (supplementary Fig. 9<!---\@ref(fig:res-p-pred-diff-kl-residuals)--->). For the improved model, this bias is much smaller (Fig. \@ref(fig:res-p-pred-diff-train1-p2) and supplementary Fig. 9<!---\@ref(fig:res-p-pred-diff-kl-residuals)--->).  
We suggest that it is this bias which causes the difference in predictions of the original and improved models. Conditional on the relative contribution of the `arom15` peak to `arom15arom16`, peat samples have smaller and larger `arom15arom16` values than the training data (supplementary Fig. 10<!---\@ref(fig:res-p-pred-diff-kl-histogram1)--->). If one extrapolates the biased predictions of the original model for the training data classes, this means larger under- and overestimation for the peat samples than for the training data using the original model (Fig. \@ref(fig:res-p-pred-diff-train1-p2)). This is supported by measured peat Klason lignin contents: Peat from temperate peatlands was reported to have Klason lignin contents as high as ~70 mass-% and ~57 mass-% on average [@Hayes.2015]. Even though this was peat from extraction sites which probably is more decomposed [@Hayes.2015], this indicates that peat Klason lignin contents can be much larger, also in temperate regions, than predictions of the original model for wood rich tropical peat [@Hodgkins.2018]. In contrast, predictions under the improved model most probably reflect more correct Klason lignin estimates because the bias is much smaller and predicted values are larger for deeper peat samples. Therefore, this bias very likely is the reason why predictions of the original model differ from that of the improved model. Moreover, our results indicate that the improved model makes more correct predictions.  
What causes this bias? We suggest that `arom15arom16` is a poor indicator for Klason lignin content for samples with varying contents of proteins. Protein C-N and C=O stretching and N-H bending cause strong absorbance around 1560 and 1650 cm$^{-1}$ [@Stuart.2004] and therefore it is not only aromatics which contribute to the `arom16` peak. Larger `arom16` peak heights may be indicative for Klason lignin, but can also be due to high protein contents.  
This is what distinguishes samples of class 1 and 2 (supplementary Fig. 11<!---\@ref(fig:res-p-pred-diff-train2)--->): Samples of class 1 are wood samples. Wood typically has smaller nitrogen contents, but larger Klason lignin contents [@Cowling.1966; @Aerts.1999; @DelaCruz.2016] which explains the large contribution of the `arom15` peak, indicative for aromatic skeletal vibrations (but not for proteins) [@Stuart.2004], to `arom15arom16`. Conversely, samples of class 2 are leaf and needle samples which may contain both varying contents of Klason lignin and proteins [@Aerts.1999; @Reich.2004; @DelaCruz.2016]. Leaves and needles typically have larger protein contents than wood [@Cowling.1966; @Aerts.1999; @Reich.2004] and therefore a smaller contribution of the `arom15` peak to `arom15arom16`. This is why the original model predicts high Klason lignin contents for samples of class 2, even though some of these actually have low Klason lignin contents (Fig. \@ref(fig:res-p-pred-diff-train1-p2)). The model balances this, but in doing so it introduces the observed bias. As a consequence, `arom15arom16` alone is a poor predictor for Klason lignin content of OM.  
The improved model using binned spectra gives more weight to variables related to aromatic skeletal structures and not in a region where proteins cause large absorbances [@Stuart.2004] (Fig. \@ref(fig:res-p-coefficients)). We suggest that these variables are better predictors for Klason lignin because they interfere less with other molecular structures, such as proteins [@Stuart.2004]. Overall, this mechanistic explanation provides additional evidence that the improved model using binned spectra probably makes more correct predictions.

```{r res-p-pred-diff-train1-p2, echo=FALSE, out.width="90%", fig.height=4, fig.width=7, fig.cap='Measured training data Klason lignin contents versus fitted values [mass-\\%] of the original model (first column), the improved model using extracted peaks (second column), and the improved model using binned spectra (third column). Points are scaled according to the relative contribution of the \\texttt{arom15} peak to \\texttt{arom15arom16}. The color gradient represents \\texttt{arom15arom16} values. The diagonal line represents values where measured and fitted values are identical. If points are below the line, the model overestimates contents.'}
p_pred_diff_train1_p2
```

## Can holocellulose contents also be predicted for samples with mineral interferences?

Our analysis shows that models with similar fit to the training data can be computed also if mineral-rich samples are included. We see this as proof-of-concept that holocellulose contents can also be predicted from MIRS for SOM samples with mineral admixtures.  
If the clay rich old magazine samples are included, the model using binned spectra had the best average predictive performance (ELPD on average `r round(abs(loo_comparison_mineral[2, 1]), 2)` larger). The model using extracted peaks had a worse, but similar performance ($\Delta\text{ELPD}= `r round(loo_comparison_mineral[2, 1], 2)`$, $\Delta\text{SE}= `r round(loo_comparison_mineral[2, 2], 2)`$). Moreover, the models trained on mineral rich samples had a similar fit to the remaining samples as our improved models not trained with the mineral rich samples (supplementary Fig. 12<!---\@ref(fig:res-p-y-yhat-minerals)--->). In comparison to this, the original model considerably overestimated the holocellulose content for these samples, as observed by @Hodgkins.2018. This was also the case for the improved models not trained with the mineral rich samples (supplementary Fig. 12<!---\@ref(fig:res-p-y-yhat-minerals)--->). We see this as proof-of-concept that it is possible to predict holocellulose contents from MIRS even for samples with high mineral contents. Moreover, there is no trade-off in predictive accuracy if mineral rich samples are included, suggesting that general purpose models can be developed.    
How do coefficients for the model trained with old magazine samples differ from the improved model trained without old magazine samples? According to Fig. \@ref(fig:res-p-coefficients) (middle panel), the "mineral-rich model" down-weights bins near the `carb` peak, in contrast to the improved model not trained on old magazine samples, and instead has a larger coefficient for an additional peak related to aromatic C=C stretching and amide N-H bending and C-N stretching (~1550 cm$^{-1}$) [@Stuart.2004] which corresponds to the `arom15` peak. This provides further evidence that it is possible to infer the holocellulose content via aromatics.

## What can we learn in general from this study?

Our validation analysis provides general lessons for validating models using spectral data. What can we learn from the model validation? First, even though @Hodgkins.2018 had a sensible rationale in including only selected peaks into their model based upon causal knowledge, this strategy has critical weaknesses. If predictive accuracy is the main goal, pre-selecting variables triggers underfitting and bias and therefore is ineffective. Instead, it is more effective to include a larger set of variables in a model and to use regularization to avoid overfitting and find relations between predictor variables and the target variable [@Piironen.2017].  
Second, a linear relation between the target variable to predict and a predictor variable is no sufficient validation of a spectral prediction model if the training data is not representative for the samples on which the model is applied. Most importantly, due to normalization, predictions can even be sensitive to variables not included in the model. Therefore, to assess if training data (the prediction domain) is representative, whole spectra have to be compared.  
Third, our analysis also shows that re-evaluating existing models with their original training data can be an effective way to improve the models. Here, it was important that one modification of the original model often addressed multiple limitations at once due to interdependences between the limitations: Improving the models' predictive performance did not only address underfitting, but reduced also the bias in the models (Fig. \@ref(fig:res-p-pred-diff-train1-p2)). Analyzing the prediction domain did not only show that the models extrapolate, but we could use this knowledge to analyze the practical impact of the bias. Lastly, including additional variables not only improved the predictive performance, but allowed us also to compute a model that probably is suitable for samples containing minerals. We therefore suggest that continuous re-evaluation of existing models can be an effective way to develop better models and should be part of a general model development workflow [@Gelman.2020].

## What are weaknesses of our analyses and opportunities for future improvements?

There are several problems we could not solve. The two most important problem is that also for the improved models, it is unclear if predictions are correct outside the prediction domain of the training data and therefore if the models are applicable to SOM in general.    
Further issues are: (1) The overall accuracy of the models probably can be improved with more training samples. (2) It is unclear how robust the original models and our improved models --- especially models using binned spectra --- are in terms of calibration transfer. We assume the binning and estimation procedure is relatively robust since bin widths are comparatively broad and all variables were standardized, meaning that only between sample variability in intensities are relevant. However, this remains to be tested. (3) Even though we computed a model that could fit and predict mineral rich samples, it is unclear if this holds in general. (4) Some of our models had computational issues (holocellulose: `r hklmirs::format_numbers(sum(d_models$n_divergent_transitions[d_models$variable_y == "Holocellulose"] > 0))` model with maximally `r hklmirs::format_numbers(max(d_models$n_divergent_transitions[d_models$variable_y == "Holocellulose"]))` divergent transitions, Klason lignin: `r hklmirs::format_numbers(sum(d_models$n_divergent_transitions[d_models$variable_y == "Klason lignin"] > 0))` models with maximally `r hklmirs::format_numbers(max(d_models$n_divergent_transitions[d_models$variable_y == "Klason lignin"]))` divergent transitions) that we could not resolve.    
Our analysis thus opens concrete and promising directions to further improve the models: We need training and validation data that includes SOM, particularly highly decomposed peat. Such data make it possible to analyze the impact of the bias and to compute models with less bias, higher representativity for SOM and peat, and potentially larger predictive accuracy. In addition to this, it is likely that an operative  model for prediction of holocellulose contents in SOM samples with mineral admixtures can be developed by including such samples with more diverse minerals. Ideally, such improvements would be performed across multiple labs with an archive of sample materials such that calibration transfer of the models between different mid infrared spectrometers can be further explored.  
To support such developments, we implemented the best models using binned spectra and extracted peaks and also the model using binned spectra that could predict mineral-rich samples into the R package irpeat [@Teickner.2020a].

\conclusions[Conclusions]

Our aim was to validate the original models of @Hodgkins.2018 to predict OM holocellulose and Klason lignin contents, identify weaknesses, provide a concrete strategy how to tackle these problems, and provide first improvements using the same data.  
The main weakness of the original models is the underlying training data: It is not in general representative for SOM, peat, and peat forming vegetation. This results in biased predictions for holocellulose and Klason lignin for SOM, such as peat. Manual variable selection favored this bias because it excluded information critical for better model fits. Finally, the original models can produce unrealistic (smaller than 0 or larger than 100 mass-%) predictions and prediction intervals, also in practice.   
Even though it was impossible to address the key problem of unrepresentative training data using the original data, we could address some of these issues, provide improved models, and develop a concrete strategy for future improvements. The improved models have less bias, avoid unrealistic predictions, and use information from the complete spectra and thus have a better predictive accuracy. Moreover, we provide a proof-of-concept that it is possible to predict holocellulose contents also for OM with mineral admixtures.  
Our analysis thus opens concrete and promising directions to further improve the models: A major opportunity is to collect training and validation data representative for SOM, such that the improved models can be extended and thoroughly validated. In a next step, potential calibration transfer issues can be addressed.  
Improved models to predict SOM holocellulose and Klason lignin contents can be of large importance in the long-run because they allow cost-efficient high-throughput analyses of SOM. Detailed understanding of SOM chemistry across large scales, as well as the processes that result in changes in SOM chemistry, are only possible if such fast and effective tools are available.

# Acknowldegements {.unnumbered}

```{r supporting-info, echo=FALSE, include=FALSE, eval = TRUE}
rmarkdown::render("008-paper-supplementary.Rmd", clean = FALSE)
```

```{r reply, echo=FALSE, include=FALSE, eval = TRUE}
rmarkdown::render("001-reply-main.Rmd", clean = FALSE)
```

```{r references-preparation, echo=FALSE, eval=FALSE}
# get bibkeys
bibkeys <- bib2df::bib2df("references.bib")

# get texts to search for bibkey
si_text <- 
  readLines("008-paper-supplementary.Rmd") %>%
  paste(collapse = " ")

main_text <- 
  readLines("001-paper-main.Rmd") %>%
  paste(collapse = " ")

bibkeys <-
  bibkeys %>%
  dplyr::mutate(
    refs_only_si = 
      purrr::map_lgl(BIBTEXKEY, function(x) {
        stringr::str_detect(si_text, x) & !stringr::str_detect(main_text, x)
      })
  )

\nocite{`r paste(bibkeys$BIBTEXKEY[!bibkeys$refs_only_si], collapse = ", ")`}
\bibliography{references}

\nocitesi{`r paste(bibkeys$BIBTEXKEY[bibkeys$refs_only_si], collapse = ", ")`}
\bibliographystylesi{apacite}
\bibliographysi{references}
```

### 2022-02-03 analysis/paper/002-paper-m-original-models.Rmd
Initial commit.

```{r m-original-models-prep}
## extract the peak heights
d_res_path <- "../data/derived_data/d_peaks"
dir.create(d_res_path)

# extract scaled normalized peak heights from MIRS
res <- 
  hklmirs::hkl_get_normalized_peak_heights(
    x = d_flat,
    export = d_res_path,
    verbose = FALSE,
    make_plots = FALSE,
    do_scale = TRUE
  )

# extract information on scaling for prediction for new datasets
res_scale <- 
  tibble::tibble(
    name = colnames(res),
    center = purrr::map_dbl(res, attr, which = "scaled:center"),
    scale = purrr::map_dbl(res, attr, which = "scaled:scale")
  )

# add peak info to d
d <- dplyr::bind_cols(d, res %>% purrr::map_df(as.numeric))
```

```{r m-original-models-models}
# compute the regression models
m1.1 <- lm(hol ~ carb, 
         data = d %>% dplyr::filter(sample_type != "old magazines"))
m2.1 <- lm(kl ~ arom15arom16, 
         data = d %>% dplyr::filter(sample_type != "office paper"))
```

```{r m-original-models-or-bayesian}
## compute the original regression models as Bayesian models

# define priors
rstanarm_prior <- rstanarm::normal(0, 2.5, autoscale = TRUE)
rstanarm_prior_aux <- rstanarm::exponential(1, autoscale = TRUE)

# compute models
m1.2 <- rstanarm::stan_glm(hol ~ carb, 
                          data = d %>% dplyr::filter(sample_type != "old magazines"),
                          family = gaussian,
                          chains = chains,
                          iter = iter,
                          warmup = warmup,
                          seed = seed,
                          prior = rstanarm_prior,
                          prior_intercept = rstanarm_prior,
                          prior_aux = rstanarm_prior_aux)
m2.2 <- rstanarm::stan_glm(kl ~ arom15arom16, 
                          data = d %>% dplyr::filter(sample_type != "office paper"),
                          family = gaussian,
                          chains = chains,
                          iter = iter,
                          warmup = warmup,
                          seed = seed,
                          prior = rstanarm_prior,
                          prior_intercept = rstanarm_prior,
                          prior_aux = rstanarm_prior_aux)

# compute loos
m1.2_loo <- loo(m1.2, 
                k_threshold = k_threshold)
m2.2_loo <- loo(m2.2, 
                k_threshold = k_threshold)
```

### 2022-02-03 analysis/paper/003-paper-m-gaussian-beta.Rmd
Initial commit.

```{r m-gaussian-beta-1}
## beta regression models

# define priors
brms_beta_priors <-
  c(
    brms::prior_string("normal(0, 2.5)", class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

# compute models
m1.3 <- brms::brm(hol ~ carb, 
                  data = d %>% dplyr::filter(sample_type != "old magazines"), 
                  family = Beta(link = "logit", link_phi = "log"),
                  prior = brms_beta_priors,
                  chains = chains,
                  seed = seed,
                  iter = iter,
                  warmup = warmup)

m2.3 <- brms::brm(kl ~ arom15arom16, 
                  data = d %>% dplyr::filter(sample_type != "office paper"), 
                  family = Beta(link = "logit", link_phi = "log"),
                  prior = brms_beta_priors,
                  chains = chains,
                  seed = seed,
                  iter = iter,
                  warmup = warmup)

# compute loos
m1.3_loo <- loo(m1.3, k_threshold = k_threshold)
m2.3_loo <- loo(m2.3, k_threshold = k_threshold)

# ---todo: brms and rstanarm yield very different uncertainty estimates. brms has smaller prediction intervals. The reason seems to be the different priors for phi (normal(0, 2.5) for rstanarm; gamma(0.01, 0.01) for brms).
```

```{r m-gaussian-beta-2}
## predictions for peat samples

# predict Klason lignin and holocellulose contents using the original model
d_peat_res_path <- "../data/derived_data/d_peat_peaks"
dir.create(d_peat_res_path)

d_peat <-
  d_peat %>%
  irpeat::irp_content_klh_hodgkins(export = d_peat_res_path, 
                                   verbose = FALSE, 
                                   make_plots = FALSE)

# import the extracted peaks and compute arom15arom16
d_peat_res <- read.csv(paste0(d_peat_res_path, "/Heights_Norm.Corr.csv"), 
                       header = TRUE, 
                       row.names = 1) %>%
  dplyr::mutate(arom15arom16 = arom15 + arom16) %>%
  purrr::map2_df(seq_len(nrow(res_scale)), function(x, i) {
    scale(x, 
          center = res_scale$center[[i]], 
          scale = res_scale$scale[[i]])
  })

# compute predictions
d_peat_pred <- 
  dplyr::bind_rows(
    m1.1 %>%
      predict(newdata = d_peat_res %>% dplyr::mutate(carb = as.numeric(carb)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Gaussian, Bayesian",
        variable_y = "Holocellulose"
      ),
    m2.1 %>%
      predict(newdata = d_peat_res %>% dplyr::mutate(arom15arom16 = as.numeric(arom15arom16)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Klason lignin"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Gaussian, Bayesian",
        variable_y = "Klason lignin"
      ),
    m1.3 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Beta",
        variable_y = "Holocellulose"
      ),
    m2.3 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Beta",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:6, function(x) {
      tibble::tibble(
        sample_id = d_peat$sample_id, 
        core_label = d_peat$core_label,
        sampling_longitude = d_peat$sampling_longitude, 
        sampling_latitude = d_peat$sampling_latitude, 
        sampling_altitude = d_peat$sampling_altitude, 
        sample_depth_lower = d_peat$sample_depth_lower, 
        sample_depth_upper = d_peat$sample_depth_upper
      )
    })
  )

# add fitted values from Gaussian model as separate variable
d_peat_pred <-
  dplyr::left_join(d_peat_pred,
                   d_peat_pred %>% 
                     dplyr::filter(model == "Gaussian, Bayesian") %>%
                     dplyr::select(variable_y, sample_id, fit) %>%
                     dplyr::rename(fit_gaussian = "fit"),
                   by = c("variable_y", "sample_id"))
```

```{r m-gaussian-beta-3}
## predictions for vegetation samples

## create data sets for predictions
d_veg_res_path <- "../data/derived_data/d_veg_peaks"
dir.create(d_veg_res_path)
d_veg_res <-
  irpeat::irp_content_klh_hodgkins(d_veg,
                                   export = d_veg_res_path,
                                   verbose = FALSE,
                                   make_plots = FALSE)

# import the extracted features and compute missing variables
d_veg_res <- read.csv(paste0(d_veg_res_path, "/Heights_Norm.Corr.csv"), 
                       header = TRUE, 
                       row.names = 1) %>%
  dplyr::mutate(arom15arom16 = arom15 + arom16) %>%
  purrr::map2_df(seq_len(nrow(res_scale)), function(x, i) {
    scale(x, 
          center = res_scale$center[[i]], 
          scale = res_scale$scale[[i]])
  })

# compute predictions
d_veg_pred <- 
  dplyr::bind_rows(
    m1.1 %>%
      predict(newdata = d_veg_res %>% dplyr::mutate(carb = as.numeric(carb)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Gaussian, Bayesian",
        variable_y = "Holocellulose"
      ),
    m2.1 %>%
      predict(newdata = d_veg_res %>% dplyr::mutate(arom15arom16 = as.numeric(arom15arom16)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Klason lignin"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Gaussian, Bayesian",
        variable_y = "Klason lignin"
      ),
    m1.3 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Beta",
        variable_y = "Holocellulose"
      ),
    m2.3 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Beta",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:6, function(x) {
      tibble::tibble(
        sample_id = d_veg$sample_id, 
        site_name = d_veg$site_name,
        species_name = d_veg$species_name,
        plant_part = d_veg$plant_part
      )
    })
  )

# add fitted values from Gaussian model as separate variable
d_veg_pred <-
  dplyr::left_join(d_veg_pred,
                   d_veg_pred %>% 
                     dplyr::filter(model == "Gaussian, Bayesian") %>%
                     dplyr::select(variable_y, sample_id, fit) %>%
                     dplyr::rename(fit_gaussian = "fit"),
                   by = c("variable_y", "sample_id"))
```

```{r m-gaussian-beta-4}
### plot of depth profiles for the peat core data: comparison of beta regression and Gaussian regression models

p1 <-
  d_peat_pred %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

p2 <-
  d_peat_pred %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

# combine plots
p_gaussian_beta_depth_profile <-
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r m-gaussian-beta-5}
### prediction across the whole possible range from 0 to 100 mass-%: comparison of beta regression and Gaussian regression models

# prediction range
d_range_res <-
  tibble::tibble(
    variable_y = c("holocellulose", "Klason lignin"),
    from = rep(-13, 2),
    to = from * (-1)
  )

# sequence of values across the prediction range
d_range_res <- 
  tibble::tibble(
    carb = 
      seq(from = d_range_res$from[d_range_res$variable_y == "holocellulose"], 
          to = d_range_res$to[d_range_res$variable_y == "holocellulose"], 
          length.out = 2 * 1e2),
    arom15arom16 = 
      seq(from = d_range_res$from[d_range_res$variable_y == "Klason lignin"], 
          to = d_range_res$to[d_range_res$variable_y == "Klason lignin"], 
          length.out = 2 * 1e2)
  )

# predictions
d_range_pred <- 
  dplyr::bind_rows(
    m1.1 %>%
      predict(newdata = d_range_res %>% dplyr::mutate(carb = as.numeric(carb)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_range_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        iter = seq_len(nrow(.)),
        model = "Gaussian, Bayesian",
        variable_y = "Holocellulose"
      ),
    m2.1 %>%
      predict(newdata = d_range_res %>% dplyr::mutate(arom15arom16 = as.numeric(arom15arom16)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Klason lignin"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_range_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        iter = seq_len(nrow(.)),
        model = "Gaussian, Bayesian",
        variable_y = "Klason lignin"
      ),
    m1.3 %>%
      brms::posterior_predict(newdata = d_range_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        iter = seq_len(nrow(.)),
        model = "Beta",
        variable_y = "Holocellulose"
      ),
    m2.3 %>%
      brms::posterior_predict(newdata = d_range_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        iter = seq_len(nrow(.)),
        model = "Beta",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:6, function(x) {
      d_range_res
    })
  )

# add fitted values from Gaussian model as separate variable
d_range_pred <-
  dplyr::left_join(d_range_pred,
                   d_range_pred %>% 
                     dplyr::filter(model == "Gaussian, Bayesian") %>%
                     dplyr::select(variable_y, iter, fit) %>%
                     dplyr::rename(fit_gaussian = "fit"),
                   by = c("variable_y", "iter"))

# detect fitted values where prediction intervals produce unrealistic predictions for the Gaussian models
d_range_pred_limits <-
  d_range_pred %>%
  dplyr::filter(model == "Gaussian, Bayesian") %>%
  dplyr::group_by(model, variable_y) %>%
  dplyr::summarise(index_fit_lwr = which(lwr > 0)[[1]],
                   index_fit_upr = which(upr > 1)[[1]],
                   fit_lwr = fit[index_fit_lwr],
                   fit_upr = fit[index_fit_upr],
                   lwr_lwr = lwr[index_fit_lwr],
                   upr_upr = upr[index_fit_upr],
                   fit_gaussian_lwr = fit_gaussian[index_fit_lwr],
                   fit_gaussian_upr = fit_gaussian[index_fit_upr],
                   .groups = "keep")

d_range_pred_limits <-
  dplyr::bind_cols(
    d_range_pred_limits %>%
      tidyr::pivot_longer(cols = dplyr::all_of(c("fit_lwr", "fit_upr")), 
                          values_to = "fit", 
                          names_to = "variable_limit"),
    d_range_pred_limits %>%
      tidyr::pivot_longer(cols = dplyr::all_of(c("fit_gaussian_lwr", "fit_gaussian_upr")), 
                          values_to = "fit_gaussian", 
                          names_to = "variable_gaussian_limit") %>%
      dplyr::ungroup() %>%
      dplyr::select(fit_gaussian)
  )
```

```{r m-gaussian-beta-6}
# plot
p1 <-
  d_range_pred %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
  geom_hline(yintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_vline(xintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, colour = NA) +
  geom_path() +
  geom_point(data = d_peat_pred %>% dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit)) +
  geom_point(data = d_veg_pred %>% dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 4) +
  geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 21, fill = "white", colour = "black") +
  labs(title = "Holocellulose", 
       x = expression(Predicted~values~"(Gaussian model)"),
       y = expression(Predicted~values)) +
  coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100)

p2 <-
  d_range_pred %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
  geom_hline(yintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_vline(xintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, colour = NA) +
  geom_path() +
  geom_point(data = d_peat_pred %>% dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit)) +
  geom_point(data = d_veg_pred %>% dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 4) +
  geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Klason lignin"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 21, fill = "white", colour = "black") +
  labs(title = "Klason lignin", 
       x = expression(Predicted~values~"(Gaussian model)"),
       y = expression(Predicted~values)) +
  coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100)

# combine plots
p_gaussian_beta_range <-
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

### 2022-02-03 analysis/paper/004-paper-m-reduce-underfitting.Rmd
Initial commit.

```{r m-reduce-underfitting-1}
### reduce undrfitting by including additional variables provided by the algorithm from Hodgkins et al. 2018 into the models

## Holocellulose

# same as m1.3, but with additional predictors
m1.5 <- 
  brms::brm(hol ~ carb + arom15 + arom16 + acids + aliph28 + aliph29 + trough16 + trough28, 
            data = d %>% dplyr::filter(sample_type != "old magazines"), 
            family = Beta(link = "logit", link_phi = "log"),
            prior = brms_beta_priors,
            chains = chains,
            seed = seed,
            iter = iter,
            warmup = warmup)

# same as m1.5, but more regularizing priors
brms_beta_priors1 <-
  c(
    brms::prior_string("normal(0, 0.5)", class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m1.6 <- 
  update(m1.5, prior = brms_beta_priors1,
         seed = seed)

# compute loos
m1.5_loo <- loo(m1.5, reloo = TRUE)
m1.6_loo <- loo(m1.6, reloo = TRUE)

## Klason lignin

# Same as m2.3, but with additional predictors
m2.5 <- 
  brms::brm(kl ~ carb + arom15 + arom16 + acids + aliph28 + aliph29 + trough16 + trough28, 
            data = d %>% dplyr::filter(sample_type != "office paper"), 
            family = Beta(link = "logit", link_phi = "log"),
            prior = brms_beta_priors,
            chains = chains,
            seed = seed,
            iter = iter,
            warmup = warmup)

# same as m2.5, but more regularizing priors
m2.6 <- update(m2.5, prior = brms_beta_priors1,
               seed = seed)

# compute loos
m2.5_loo <- loo(m2.5, reloo = TRUE)
m2.6_loo <- loo(m2.6, reloo = TRUE)
```

```{r m-reduce-underfitting-2}
### reduce underfitting by using the complete initial spectrum binned with varying bin widths ranging from 10 to 100 and using regularizing priors

# define bin widths
bin_widths <- c(10, 20, 30, 50, 100)

# create binned datasets
d_underfitting_flat <-
  purrr::map(bin_widths, function(x) {
    d %>%
      ir::ir_bin(width = x) %>%
      ir::ir_flatten()
  })

d_underfitting <- 
  purrr::map(d_underfitting_flat, function(x) {
    d %>%
      dplyr::select(hol, kl) %>%
      dplyr::bind_cols(
        x %>%
          dplyr::select(-1) %>%
          t() %>%
          as.data.frame() %>%
          scale(center = TRUE, scale = TRUE) %>%
          as.data.frame()
      )
  })

# define regularized horsheshoe prior
prior_hs_tau0 <-
  purrr::map_dbl(d_underfitting, function(x) {
    n <- nrow(x) - 4L # in each case, four observations are removed
    p <- ncol(x) - 2L # two target variables have to be subtracted
    p0 <- 8L # prior guess for the number of relevant variables
    p0/(p-p0) * 1/sqrt(n) # tau0
  })
```

```{r m-reduce-underfitting-3}
## Holocellulose

# fit models
brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[1]], ", df_global = 1, autoscale = TRUE)"),
                       class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m1.7 <- 
  brms::brm(
    hol ~ ., 
    data = 
      d_underfitting[[1]] %>% 
      dplyr::select(-kl) %>% 
      dplyr::filter(d$sample_type != "old magazines"),
    family = Beta(link = "logit", link_phi = "log"),
    prior = brms_beta_priors2,
    cores = chains,
    chains = chains,
    warmup = warmup,
    iter = iter,
    seed = seed ,
    control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 6 divergent transitions, low bulk ESS; with adapt_delta = 0.99: There were 183 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10, low bulk ESS
  )

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[2]], ", df_global = 1, autoscale = TRUE)"),
                       class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m1.8 <- 
  brms::brm(
    hol ~ ., 
    data = 
      d_underfitting[[2]] %>% 
      dplyr::select(-kl) %>% 
      dplyr::filter(d$sample_type != "old magazines"),
    family = Beta(link = "logit", link_phi = "log"),
    prior = brms_beta_priors2,
    cores = chains,
    chains = chains,
    warmup = warmup,
    iter = iter,
    seed = seed,
    control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 5 divergent transitions; with adapt_delta = 0.99: There were 3994 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10
  )

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[3]], ", df_global = 1, autoscale = TRUE)"),
                       class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m1.9 <- 
  brms::brm(
    hol ~ ., 
    data = 
      d_underfitting[[3]] %>% 
      dplyr::select(-kl) %>% 
      dplyr::filter(d$sample_type != "old magazines"),
    family = Beta(link = "logit", link_phi = "log"),
    prior = brms_beta_priors2,
    cores = chains,
    chains = chains,
    warmup = warmup,
    iter = iter,
    seed = seed,
    control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 11 divergent transitions; adapt_delta = 0.99: There were 2000 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10
  )

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[4]], ", df_global = 1, autoscale = TRUE)"),
                       class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m1.10 <- 
  brms::brm(
    hol ~ ., 
    data = 
      d_underfitting[[4]] %>% 
      dplyr::select(-kl) %>% 
      dplyr::filter(d$sample_type != "old magazines"),
    family = Beta(link = "logit", link_phi = "log"),
    prior = brms_beta_priors2,
    cores = chains,
    chains = chains,
    warmup = warmup,
    iter = iter,
    seed = seed,
    control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 8 divergent transitions; adapt_delta = 0.99: There were 1840 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10.
  )

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[5]], ", df_global = 1, autoscale = TRUE)"),
                       class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m1.11 <- 
  brms::brm(
    hol ~ ., 
    data = 
      d_underfitting[[5]] %>% 
      dplyr::select(-kl) %>% 
      dplyr::filter(d$sample_type != "old magazines"),
    family = Beta(link = "logit", link_phi = "log"),
    prior = brms_beta_priors2,
    cores = chains,
    chains = chains,
    warmup = warmup,
    iter = iter,
    seed = seed,
    control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 17 divergent transitions; adapt_delta = 0.99: no sampling issues
  ) 
```

```{r m-reduce-underfitting-4}
## Klason lignin

# fit models
brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[1]], ", df_global = 1, autoscale = TRUE)"),
                                             class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m2.7 <- 
  brms::brm(
    kl ~ ., 
                  data = 
                    d_underfitting[[1]] %>% 
                    dplyr::select(-hol) %>% 
                    dplyr::filter(d$sample_type != "office paper"),
                  family = Beta(link = "logit", link_phi = "log"),
                  prior = brms_beta_priors2,
                  cores = chains,
                  chains = chains,
                  warmup = warmup,
                  iter = iter,
                  seed = seed,
                  control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 49 divergent transitions; adapt_delta = 0.99: There were 7136 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. Low bulk ESS
    )

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[2]], ", df_global = 1, autoscale = TRUE)"),
                                             class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m2.8 <- 
  brms::brm(
  kl ~ ., 
                  data = 
                    d_underfitting[[2]] %>% 
                    dplyr::select(-hol) %>% 
                    dplyr::filter(d$sample_type != "office paper"),
                  family = Beta(link = "logit", link_phi = "log"),
                  prior = brms_beta_priors2,
                  cores = chains,
                  chains = chains,
                  warmup = warmup,
                  iter = iter,
                  seed = seed,
                  control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 136 divergent transitions and low tail ESS; adapt_delta = 0.99: 3 divergent transitions. There were 2000 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. 
  )

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[3]], ", df_global = 1, autoscale = TRUE)"),
                                             class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m2.9 <- brms::brm(kl ~ ., 
                  data = 
                    d_underfitting[[3]] %>% 
                    dplyr::select(-hol) %>% 
                    dplyr::filter(d$sample_type != "office paper"),
                  family = Beta(link = "logit", link_phi = "log"),
                  prior = brms_beta_priors2,
                  cores = chains,
                  chains = chains,
                  warmup = warmup,
                  iter = iter,
                  seed = seed,
                  control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 452 divergent transitions and low tail ESS; adapt_delta = 0.99: 5 divergent transitions. There were 1631 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10.
                  )

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[4]], ", df_global = 1, autoscale = TRUE)"),
                                             class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m2.10 <- brms::brm(kl ~ ., 
                   data = 
                     d_underfitting[[4]] %>% 
                     dplyr::select(-hol) %>% 
                     dplyr::filter(d$sample_type != "office paper"),
                   family = Beta(link = "logit", link_phi = "log"),
                   prior = brms_beta_priors2,
                   cores = chains,
                   chains = chains,
                   warmup = warmup,
                   iter = iter,
                   seed = seed,
                  control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 267 divergent transitions, low tail ESS; adapt_delta = 0.99: 7 divergent transitions 
                  )

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[5]], ", df_global = 1, autoscale = TRUE)"),
                                             class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m2.11 <- brms::brm(kl ~ ., 
                   data = 
                     d_underfitting[[5]] %>% 
                     dplyr::select(-hol) %>% 
                     dplyr::filter(d$sample_type != "office paper"),
                   family = Beta(link = "logit", link_phi = "log"),
                   prior = brms_beta_priors2,
                   cores = chains,
                   chains = chains,
                   warmup = warmup,
                   iter = iter,
                   seed = seed,
                  control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 81 divergent transitions; adapt_delta = 0.99: no sampling issues
                  ) 
```

```{r m-reduce-underfitting-5-1}
# compute loos
m1.7_loo <- loo(m1.7, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-5-2}
m1.8_loo <- loo(m1.8, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-5-3}
m1.9_loo <- loo(m1.9, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-5-4}
m1.10_loo <- loo(m1.10, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-5-5}
m1.11_loo <- loo(m1.11, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-6-1}
# compute loos
m2.7_loo <- loo(m2.7, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-6-2}
m2.8_loo <- loo(m2.8, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-6-3}
m2.9_loo <- loo(m2.9, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-6-4}
m2.10_loo <- loo(m2.10, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-6-5}
m2.11_loo <- loo(m2.11, reloo = TRUE, cores = chains)
```

```{r m-reduce-underfitting-7}
### create the tables summarizing the predictive performance comparison

# table with loo results
loo_comparison <-
  tibble::tibble(
    variably_y = c("Holocellulose", "Klason lignin"),
    loo_comparison = 
      list(
        loo::loo_compare(list(m1.2_loo, m1.3_loo, m1.5_loo, m1.6_loo, m1.7_loo, m1.8_loo, m1.9_loo, m1.10_loo, m1.11_loo)),
        loo::loo_compare(list(m2.2_loo, m2.3_loo, m2.5_loo, m2.6_loo, m2.7_loo, m2.8_loo, m2.9_loo, m2.10_loo, m2.11_loo))
      )
  )

## create table for printing

# holocellulose
loo_comparison_hol <-
  loo_comparison %>%
  dplyr::filter(variably_y == "Holocellulose") %>%
  dplyr::pull(loo_comparison) %>%
  as.data.frame() %>%
  dplyr::select(1:3) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), round, 1)) %>%
  dplyr::mutate(
    y_variable = "Holocellulose",
    model_code = rownames(.),
    bin_width = c(50, 100, 30, 10, rep("-", 4), 20),
    family = c(rep("Beta", 7), "Gaussian", "Beta"),
    original_model = c(rep(FALSE, 6), TRUE, TRUE, FALSE),
    prior_scale = c(rep("-", 4), 0.5, 2.5, 2.5, 2.5, "-"),
    variables = ifelse(bin_width == "-", "peaks", "bins")
  )

# Klason lignin
loo_comparison_kl <-
  loo_comparison %>%
  dplyr::filter(variably_y == "Klason lignin") %>%
  dplyr::pull(loo_comparison) %>%
  as.data.frame() %>%
  dplyr::select(1:3) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), round, 1)) %>%
  dplyr::mutate(
    y_variable = "Klason lignin",
    model_code = rownames(.),
    bin_width = c(10, rep("-", 4), 30, 20, 100, 50),
    family = c(rep("Beta", 3), "Gaussian", rep("Beta", 5)),
    original_model = c(rep(FALSE, 3), TRUE, TRUE, rep(FALSE, 4)),
    prior_scale = c("-", 0.5, 2.5, 2.5, 2.5, rep("-", 4)),
    variables = ifelse(bin_width == "-", "peaks", "bins")
  )

## combine
loo_comparison_res_colindex <- 
  c("model_code", "original_model", "family", "variables", "prior_scale", "bin_width", "elpd_loo", "elpd_diff", "se_diff")

loo_comparison_res <-
  dplyr::bind_rows(
    loo_comparison_hol %>%
      dplyr::select(dplyr::all_of(loo_comparison_res_colindex)),
    loo_comparison_kl %>%
      dplyr::select(dplyr::all_of(loo_comparison_res_colindex))
  ) %>%
  tibble::remove_rownames() %>%
  kableExtra::kable(booktabs = TRUE, 
                    col.names = c("Model", "Original model?", "Distribution", "Predictors", "Prior scale", "Bin width", "ELPD", "$\\Delta$ELPD", "$\\Delta$SE"), 
                    escape = FALSE,
                    caption = 'Overview on the relative predictive performance of the models for holocellulose and Klason lignin content as measured using PSIS-LOO ELPD. For each variable, the model with the best average predictive performance (largest ELPD) is at the top and the other models follow in descending order. Models ending in ".2" and ".3" are the models with the original model structure as developed by \\cite{Hodgkins.2018} (column "Original model?"). "Distribution" is the distribution assumed for the target variable. "Predictors" indicates if models were fitted with peaks extracted using the procedure of \\cite{Hodgkins.2018} ("peaks") or using binned spectra ("bins"). "Prior" scale indicates the standard deviation for Gaussian coefficients (numeric values), or that flat priors were assumed ("flat") or regularized horseshoe priors were used ("-"). "Bin width" is the width of bins (in wavenumber units). "ELPD" is the PSIS-LOO expected log-predictive density, $\\Delta$ELPD the difference in ELPD relative to the best average model, and $\\Delta$SE the standard error in the average $\\Delta$ELPD. \\label{tab:res-tab-loo-comparison-res}') %>%
  kableExtra::kable_styling(latex_options = c("hold_position"), position = "center") %>%
  kableExtra::pack_rows("Holocellulose", 1, 9) %>%
  kableExtra::pack_rows("Klason lignin", 10, 8)
```

```{r m-reduce-underfitting-8}
### compute predictions for peat profiles comparing the "best" model and th original Beta regression model

## create data sets for predictions

# bin width: 50
d_peat_bin50_scale <-
  d_underfitting_flat[[4]] %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(center = TRUE, scale = TRUE)

d_peat_bin50_scale <-
  tibble::tibble(
    center = attr(d_peat_bin50_scale, "scaled:center"),
    scale = attr(d_peat_bin50_scale, "scaled:scale")
  )

d_peat_bin50_flat <-
  d_peat %>%
  ir::ir_bin(width = 50) %>%
  ir::ir_flatten()

d_peat_bin50 <- 
  d_peat_bin50_flat %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(
    center = d_peat_bin50_scale$center, 
    scale = d_peat_bin50_scale$scale
  ) %>%
  as.data.frame()

# bin width: 10
d_peat_bin10_scale <-
  d_underfitting_flat[[1]] %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(center = TRUE, scale = TRUE)

d_peat_bin10_scale <-
  tibble::tibble(
    center = attr(d_peat_bin10_scale, "scaled:center"),
    scale = attr(d_peat_bin10_scale, "scaled:scale")
  )

d_peat_bin10_flat <-
  d_peat %>%
  ir::ir_bin(width = 10) %>%
  ir::ir_flatten()

d_peat_bin10 <- 
  d_peat_bin10_flat %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(
    center = d_peat_bin10_scale$center, 
    scale = d_peat_bin10_scale$scale
  ) %>%
  as.data.frame()

# predictions
d_peat_pred_best <- 
  dplyr::bind_rows(
    m1.10 %>%
      rstanarm::posterior_predict(newdata = d_peat_bin50) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Best binned spectra",
        variable_y = "Holocellulose"
      ),
    m1.6 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Best all peaks",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Original",
        variable_y = "Holocellulose"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Original",
        variable_y = "Klason lignin"
      ),
    m2.7 %>%
      rstanarm::posterior_predict(newdata = d_peat_bin10) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Best binned spectra",
        variable_y = "Klason lignin"
      ),
    m2.6 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Best all peaks",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:6, function(x) {
      tibble::tibble(
        sample_id = d_peat$sample_id, 
        core_label = d_peat$core_label,
        sampling_longitude = d_peat$sampling_longitude, 
        sampling_latitude = d_peat$sampling_latitude, 
        sampling_altitude = d_peat$sampling_altitude, 
        sample_depth_lower = d_peat$sample_depth_lower, 
        sample_depth_upper = d_peat$sample_depth_upper
      )
    })
  )

# summarize across depth values to reproduce key plots in Hodgkins et al. 2018
d_peat_pred_best_summary <-
  d_peat_pred_best %>%
  dplyr::filter(sample_depth_lower <= 50) %>%
  dplyr::group_by(variable_y, model, core_label) %>%
  dplyr::summarise(fit = mean(fit),
                   lwr = mean(lwr),
                   upr = mean(upr),
                   sampling_latitude = mean(sampling_latitude))

# data frame to compare fitted with measured values across all models
d_fitted_comparison <-
  dplyr::bind_rows(
    tibble::tibble(
      y = 
        d %>% 
        dplyr::filter(sample_type != "old magazines") %>% 
        dplyr::pull(hol) %>%
        rep(4),
      yhat = 
        list(m1.2, m1.5, m1.10) %>%
        purrr::map(function(x) {
          apply(posterior_predict(x), 2, median)
        }) %>%
        unlist() %>%
        c(m1.1 %>% predict(type = "response")),
      model = rep(c("Original", "Best all peaks", "Best binned spectra", "Original non-Bayesian"), 
                  each = d %>% dplyr::filter(sample_type != "old magazines") %>% nrow()),
      y_variable = "Holocellulose"
    ),
    tibble::tibble(
      y = 
        d %>% 
        dplyr::filter(sample_type != "office paper") %>% 
        dplyr::pull(kl) %>%
        rep(4),
      yhat = 
        list(m2.2, m2.6, m2.7) %>%
        purrr::map(function(x) {
          apply(posterior_predict(x), 2, median)
        }) %>%
        unlist() %>%
        c(m2.1 %>% predict(type = "response")),
      model = rep(c("Original", "Best all peaks", "Best binned spectra", "Original non-Bayesian"), 
                  each = d %>% dplyr::filter(sample_type != "office paper") %>% nrow()),
      y_variable = "Klason lignin"
    )
  )
```

```{r m-reduce-underfitting-9}
## plot depth profiles
p1 <-
  d_peat_pred_best %>%
  dplyr::filter(variable_y == "Holocellulose") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

p2 <-
  d_peat_pred_best %>%
  dplyr::filter(variable_y == "Klason lignin") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

# combine plots
p_gaussian_beta_depth_profile_best <-
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

### Reproduce plots in Hodgkins et al. 2018

## contents vs latitude across sites
p1 <-
  d_peat_pred_best_summary %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Best all peaks") %>%
  ggplot(aes(y = fit, x = sampling_latitude, colour = model)) +
  geom_errorbar(aes(ymin = lwr, ymax = upr)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x) +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"["*g/g[sample]*"]"),
       x = expression("Latitude ["*degree*N*"]")) +
  guides(colour = guide_legend(title = "Model"))

p2 <-
  d_peat_pred_best_summary %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Best all peaks") %>%
  ggplot(aes(y = fit, x = sampling_latitude, colour = model)) +
  geom_errorbar(aes(ymin = lwr, ymax = upr)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x) +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"["*g/g[sample]*"]"),
       x = expression("Latitude ["*degree*N*"]")) +
  guides(colour = guide_legend(title = "Model"))

# combine
p_sites_latitude <- 
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

## depth profiles for two latitude classes
p1 <- 
  d_peat_pred_best %>%
    dplyr::arrange(sample_depth_lower) %>%
    dplyr::filter(variable_y == "Holocellulose" & model != "Best all peaks") %>%
    ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
               x = sample_depth_lower, 
               colour = sampling_latitude >45, 
               fill = sampling_latitude >45)) +
    geom_hline(yintercept = c(0.09, 0.14, 0.21, 0.43), colour = "grey", linetype = 2) +
    geom_smooth(method = "loess", formula = y ~ x) +
    coord_flip() +
    labs(title = "Holocellulose", 
         y = expression(Holocellulose~content~"["*g/g[sample]*"]"),
         x = "Lower layer boundary depth [cm]") +
    facet_wrap(~ model, 
               scales = "free_y", 
               ncol = 1) +
    scale_x_reverse() + 
  guides(colour = guide_legend(title = expression(Latitude>45~degree*N)),
         fill = guide_legend(title = expression(Latitude>45~degree*N)))

p2 <- 
  d_peat_pred_best %>%
    dplyr::arrange(sample_depth_lower) %>%
    dplyr::filter(variable_y == "Klason lignin" & model != "Best all peaks") %>%
    ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
               x = sample_depth_lower, 
               colour = sampling_latitude >45, 
               fill = sampling_latitude >45)) +
    geom_hline(yintercept = c(0.22, 0.34, 0.38, 0.53), colour = "grey", linetype = 2) +
    geom_smooth(method = "loess", formula = y ~ x) +
    coord_flip() +
    labs(title = "Klason lignin", 
         y = expression(Klason~lignin~content~"["*g/g[sample]*"]"),
         x = "Lower layer boundary depth [cm]") +
    facet_wrap(~ model, 
               scales = "free_y", 
               ncol = 1) +
    scale_x_reverse() + 
  guides(colour = guide_legend(title = expression(Latitude>45~degree*N)),
         fill = guide_legend(title = expression(Latitude>45~degree*N)))

# combine
p_depth_profiles_latitude <- 
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

### plot fitted vs measured values
p1 <-
  d_fitted_comparison %>%
  dplyr::filter(y_variable == "Holocellulose" & model != "Original non-Bayesian") %>%
  dplyr::mutate(model = factor(model, levels = c("Original", "Best binned spectra", "Best all peaks"))) %>%
ggplot(aes(y = y, x = yhat)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  geom_point() +
  geom_smooth(method = "loess", formula = y ~ x, colour = "grey50") +
  labs(title = "Holocellulose", 
       x = "Fitted",
       y = "Measured") +
  coord_fixed(ratio = 1) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  facet_wrap( ~ model)

p2 <-
  d_fitted_comparison %>%
  dplyr::filter(y_variable == "Klason lignin" & model != "Original non-Bayesian") %>%
  dplyr::mutate(model = factor(model, levels = c("Original", "Best binned spectra", "Best all peaks"))) %>%
ggplot(aes(y = y, x = yhat)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  geom_point() +
  geom_smooth(method = "loess", formula = y ~ x, colour = "grey50") +
  labs(title = "Klason lignin", 
       x = "Fitted",
       y = "Measured") +
  coord_fixed(ratio = 1) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  facet_wrap( ~ model)

# combine plots
p_y_yhat_best <-
  p1 / p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r m-reduce-underfitting-10}
### compute predictions for vegetation samples comparing the "best" model and th original Gaussian regression model

# bin width: 50
d_veg_bin50_flat <-
  d_veg %>%
  ir::ir_bin(width = 50) %>%
  ir::ir_flatten()

d_veg_bin50 <- 
  d_veg_bin50_flat %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(center = d_peat_bin50_scale$center, 
        scale = d_peat_bin50_scale$scale) %>%
  as.data.frame()

# bin width: 10
d_veg_bin10_flat <-
  d_veg %>%
  ir::ir_bin(width = 10) %>%
  ir::ir_flatten()

d_veg_bin10 <- 
  d_veg_bin10_flat %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(center = d_peat_bin10_scale$center, 
        scale = d_peat_bin10_scale$scale) %>%
  as.data.frame()

# predictions
d_veg_pred_best <- 
  dplyr::bind_rows(
    m1.10 %>%
      rstanarm::posterior_predict(newdata = d_veg_bin50) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Best binned spectra",
        variable_y = "Holocellulose"
      ),
    m1.6 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Best all peaks",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Original",
        variable_y = "Holocellulose"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Original",
        variable_y = "Klason lignin"
      ),
    m2.7 %>%
      rstanarm::posterior_predict(newdata = d_veg_bin10) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Best binned spectra",
        variable_y = "Klason lignin"
      ),
    m2.6 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Best all peaks",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:6, function(x) {
      tibble::tibble(
        sample_id = d_veg$sample_id, 
        site_name = d_veg$site_name,
        species_name = d_veg$species_name,
        plant_part = d_veg$plant_part
      )
    })
  )

# plot comparing vegetation predictions
p_veg_species <-
  d_veg_pred_best %>%
  dplyr::filter(model != "Best all peaks") %>%
  dplyr::mutate(
    species_name_numeric = as.factor(species_name) %>% as.integer() %>% magrittr::add(ifelse(model == "Original", 0.2, -0.2))
  ) %>%
  ggplot(aes(x = species_name_numeric, y = fit, ymin = lwr, ymax = upr, colour = model)) +
  # geom_errorbar(position = position_jitterdodge(0.9, jitter.height = 0, jitter.width = 0.4, seed = 1), width = 0) +
  geom_ribbon(aes(group = sample_id), fill = "grey", alpha = 0.4, colour = NA) +
  geom_path(aes(group = sample_id), colour = "grey") +
  geom_point() +
  coord_flip() +
  geom_vline(xintercept = seq(from = 1.5, by = 1, length.out = length(unique(d_veg_pred_best$species_name)) - 1), colour = "grey", linetype = 2) +
  labs(x = "Species",
       y = expression("Predicted content [mass-%]")) +
  facet_grid(~ variable_y, scales = "free") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(title = "Model", ncol = 1)) + 
  scale_x_continuous(
    breaks = seq(from = 1,  by = 1, length.out = length(unique(d_veg_pred_best$species_name))),
    labels = levels(as.factor(d_veg_pred_best$species_name))
  ) +
  scale_y_continuous(labels = function(x) x * 100)

# classify sites so that they match the site-vegetation classification in Hodgkins et al. 2018
d_peat_pred_best <-
  d_peat_pred_best %>%
  dplyr::left_join(
    tibble::tibble(
      core_label = unique(d_peat_pred_best$core_label),
      barchart_label = c("Boreal fens", "Stordalen", "NC Pocosin", "NC Pocosin", "Loxahatchee", "Loxahatchee", "Boreal bogs", "Boreal bogs", "Mendaram", "Mendaram", "Boreal bogs", "Boreal fens", "Boreal bogs", "Boreal bogs")
    ),
    by = "core_label"
  )

d_veg_pred_best <-
  d_veg_pred_best %>% 
  dplyr::left_join(
    tibble::tibble(
      site_name = unique(d_veg_pred_best$site_name),
      barchart_label = c("Stordalen", "Boreal bogs", "Boreal fens", "NC Pocosin", "Loxahatchee", "Mendaram", "Mendaram")
    ),
    by = "site_name"
  )

d_barchart1 <-
  dplyr::bind_rows(
    d_peat_pred_best %>%
      dplyr::filter(sample_depth_lower <= 50) %>%
      dplyr::mutate(sample_type = "Peat"),
    d_veg_pred_best %>%
      dplyr::mutate(sample_type = "Vegetation"),
  ) %>%
  dplyr::mutate(barchart_label = factor(barchart_label, levels = c("Stordalen", "Boreal bogs", "Boreal fens", "NC Pocosin", "Loxahatchee", "Mendaram") %>% rev()),
                sample_type = factor(sample_type, levels = c("Vegetation", "Peat", "Difference")))

# compute difference between median vegetation and peat signature
d_barchart1 <-
  d_barchart1 %>%
  dplyr::group_by(variable_y, model, barchart_label, sample_type) %>%
  dplyr::summarise(fit_median = median(fit), lwr = fit_median - sd(fit), upr = fit_median + sd(fit))
d_barchart1 <-
  dplyr::bind_rows(
    d_barchart1,
    d_barchart1 %>%
      dplyr::group_by(variable_y, model, barchart_label) %>%
      dplyr::summarise(fit_median = fit_median[sample_type == "Peat"] - fit_median[sample_type == "Vegetation"],
                       sample_type = "Difference")
  )

# plots
p1 <- 
  d_barchart1 %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Best all peaks") %>%
  ggplot(aes(y = fit_median, ymin = lwr, ymax = upr,
             x = barchart_label,
             fill = sample_type)) +
  geom_errorbar(colour = "grey", width = 0.2, position = position_dodge(.9)) +
  geom_hline(yintercept = c(-0.2, 0, 0.2, 0.4), colour = "grey", linetype = 2) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"["*g/g[sample]*"]"),
       x = "Site category") +
  facet_wrap(~ model) +
  guides(fill = guide_legend(title = "")) +
  scale_fill_grey()

p2 <- 
  d_barchart1 %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Best all peaks") %>%
  ggplot(aes(y = fit_median, ymin = lwr, ymax = upr,
             x = barchart_label,
             fill = sample_type)) +
  geom_errorbar(colour = "grey", width = 0.2, position=position_dodge(.9)) +
  geom_hline(yintercept = c(0, 0.2, 0.4), colour = "grey", linetype = 2) +
  geom_bar(stat = "identity", position=position_dodge()) +
  coord_flip() +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~"["*g/g[sample]*"]"),
       x = "Site category") +
  facet_wrap(~ model) +
  guides(fill = guide_legend(title = "")) +
  scale_fill_grey()

# combine plots
p_veg_peat_difference <-
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

### 2022-02-03 analysis/paper/005-paper-m-minerals.Rmd
Initial commit.

```{r m-minerals-1}
## Holocellulose

m1.12 <- 
  brms::brm(hol ~ carb + arom15 + arom16 + acids + aliph28 + aliph29 + trough16 + trough28, 
            data = d, 
            family = Beta(link = "logit", link_phi = "log"),
            prior = brms_beta_priors1,
            chains = chains,
            iter = iter,
            warmup = warmup,
            seed = seed)

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[4]], ", df_global = 1, autoscale = TRUE)"),
                       class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m1.13 <- 
  brms::brm(
    hol ~ ., 
    data = 
      d_underfitting[[4]] %>% 
      dplyr::select(-kl),
    family = Beta(link = "logit", link_phi = "log"),
    prior = brms_beta_priors2,
    cores = chains,
    chains = chains,
    warmup = warmup,
    iter = iter,
    seed = seed,
    control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 22 divergent transitions; adapt_delta = 0.99: There were 2000 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10
  ) 
```

```{r m-minerals-2}
# PSIS-LOO
m1.12_loo <- loo(m1.12, reloo = TRUE, cores = chains)
m1.13_loo <- loo(m1.13, reloo = TRUE, cores = chains)

loo_comparison_mineral <- loo::loo_compare(m1.12_loo, m1.13_loo)
```

```{r m-minerals-3}
# plot of fitted values against measured values
d_minerals_pred <-
  tibble::tibble(
    fit = c(
      posterior_predict(m1.2, newdata = d) %>%
        apply(2, median),
      posterior_predict(m1.10, newdata = d_underfitting[[4]]) %>%
        apply(2, median),
      posterior_predict(m1.6, newdata = d) %>%
        apply(2, median),
      posterior_predict(m1.12) %>%
        apply(2, median),
      posterior_predict(m1.13) %>%
        apply(2, median)
    ),
    model = 
      rep(c("Original", "Binned (no minerals)", "Peaks (no minerals)", "Peaks", "Binned"), each = nrow(d)) %>%
      factor(levels = c("Original", "Peaks (no minerals)", "Binned (no minerals)", "Peaks", "Binned"))
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:5, function(x) {
      d %>% 
        dplyr::select(sample_type, hol) %>%
        dplyr::mutate(mineral = ifelse(sample_type == "old magazines", TRUE, FALSE))
    })
  )

# plot
p_y_yhat_minerals <-
  d_minerals_pred %>%
  ggplot(aes(x = fit * 100, y = hol * 100, colour = mineral)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey") +
  geom_point() +
  coord_fixed() +
  facet_wrap( ~ model, nrow = 1) +
  labs(x = "Fitted",
       y = "Measured") +
  guides(colour = guide_legend(title = "Mineral rich samples")) +
  theme(legend.position = "bottom")
```

```{r m-p-coefficients}
## plot to compare coefficients of the best models

# Holocellulose
m1.10_pars <- 
  m1.10 %>%
  as.data.frame() %>%
  dplyr::select(dplyr::contains("b_V"))

m1.10_pars <- 
  m1.10_pars %>%
  purrr::map_df(function(x) {
    res <- quantile(x, c(0.5, 0.05, 0.95))
    tibble::tibble(
      median = res[[1]],
      lwr = res[[2]],
      upr = res[[3]]
    )
  }) %>%
  dplyr::mutate(
    par_name = colnames(m1.10_pars),
    wn = d_underfitting_flat[[4]]$x,
    variable_y = "Holocellulose (no minerals)"
  )

m1.13_pars <- 
  m1.13 %>%
  as.data.frame() %>%
  dplyr::select(dplyr::contains("b_V"))

m1.13_pars <- 
  m1.13_pars %>%
  purrr::map_df(function(x) {
    res <- quantile(x, c(0.5, 0.05, 0.95))
    tibble::tibble(
      median = res[[1]],
      lwr = res[[2]],
      upr = res[[3]]
    )
  }) %>%
  dplyr::mutate(
    par_name = colnames(m1.13_pars),
    wn = d_underfitting_flat[[4]]$x,
    variable_y = "Holocellulose (with minerals)"
  )

# Klason lignin
m2.7_pars <- 
  m2.7 %>%
  as.data.frame() %>%
  dplyr::select(dplyr::contains("b_V"))

m2.7_pars <- 
  m2.7_pars %>%
  purrr::map_df(function(x) {
    res <- quantile(x, c(0.5, 0.05, 0.95))
    tibble::tibble(
      median = res[[1]],
      lwr = res[[2]],
      upr = res[[3]]
    )
  }) %>%
  dplyr::mutate(
    par_name = colnames(m2.7_pars),
    wn = d_underfitting_flat[[1]]$x,
    variable_y = "Klason lignin"
  )

# combine
m_pars <-
  dplyr::bind_rows(
    m1.10_pars,
    m1.13_pars,
    m2.7_pars
  )

# plots
p_coefficients <-
  m_pars %>%
  ggplot(aes(x = wn, y = median)) +
  geom_ribbon(data = d_peat$spectra[[1]], aes(x = x, y = y * 200 + 0.4, ymax = y * 200 + 0.4, ymin = rep(0.4, length(y))), colour = NA, fill = "lightgrey", alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  geom_errorbar(aes(ymin = lwr, ymax = upr), colour = "lightblue", width = 0) +
  geom_point(shape = 21, fill = "white", colour = "grey50") +
  facet_wrap(~ variable_y, scales = "free_x") +
  geom_text_repel(data = m_pars %>%
                    dplyr::mutate(label = ifelse(abs(median) >= 0.02, wn %>% round(0), "")),
                  aes(label = label),
                  segment.colour = "grey50") +
  labs(x = expression(Bin~wavenumber~"["*cm^{-1}*"]"),
       y = "Coefficient")
```

```{r m-model-validation}
d_models <-
  tibble::tibble(
    model = list(m1.2, m1.3, m1.5, m1.6, m1.7, m1.8, m1.9, m1.10, m1.11, m1.12, m1.13, 
                 m2.2, m2.3, m2.5, m2.6, m2.7, m2.8, m2.9, m2.10, m2.11),
    model_type = purrr::map_chr(model, function(x) class(x)[[1]]),
    variable_y = c(rep("Holocellulose", 11), rep("Klason lignin", 9)),
    n_divergent_transitions = 
      purrr::map2_dbl(model, model_type, function(x, y) {
        switch(y,
               "brmsfit" = sum(rstan::get_divergent_iterations(x$fit)),
               "stanreg" = sum(rstan::get_divergent_iterations(x$stanfit)),
               stop("Unknown model type.")
        )
      }
      )
  ) %>%
  dplyr::select(-model)
```

### 2022-02-03 analysis/paper/006-paper-m-prediction-domain.Rmd
Initial commit.

```{r m-prediction-domain-1}
### assess the prediction domain of the original models

## Holocellulose

# orginal model
d_prediction_domain_hol_or <-
  tibble::tibble(
    variable_y = "Holocellulose",
    model = "Original",
    data_type = 
      c(
        rep("Training", sum(d$sample_type != "old magazines")),
        rep("Peat", nrow(d_peat_res)),
        rep("Vegetation", nrow(d_veg_res))
      ),
    predictor_value = 
      c(
        d$carb[d$sample_type != "old magazines"],
        d_peat_res$carb,
        d_veg_res$carb
      )
  )

d_prediction_domain_hol_or <-
  dplyr::bind_cols(
    d_prediction_domain_hol_or,
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_prediction_domain_hol_or %>%
                                    dplyr::mutate(carb = predictor_value)) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      })
    )

# binned spectra, best model
d_prediction_domain_hol_bin <-
  dplyr::bind_rows(
    # training data
    d_underfitting[[4]] %>% 
      dplyr::select(-kl, -hol) %>%
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_underfitting_flat[[4]]$x %>%
          rep(nrow(d_underfitting[[4]])),
        sample_id = 
          seq_len(nrow(d_underfitting[[4]])) %>% 
          rep(each = nrow(d_underfitting_flat[[4]])) %>% 
          as.character(),
        data_type = "Training"
      ),
    # peat data
     d_peat_bin50 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_peat_bin50_flat$x %>%
          rep(nrow(d_peat_bin50)),
        sample_id = 
          seq_len(nrow(d_peat_bin50)) %>% 
          rep(each = nrow(d_peat_bin50_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[4]])) %>%
          as.character(),
        data_type = "Peat"
      ),
    # vegetation data
     d_veg_bin50 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_veg_bin50_flat$x %>%
          rep(nrow(d_veg_bin50)),
        sample_id = 
          seq_len(nrow(d_veg_bin50)) %>% 
          rep(each = nrow(d_veg_bin50_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[4]]) + nrow(d_peat_bin50)) %>%
          as.character(),
        data_type = "Vegetation"
      )
  ) %>%
  dplyr::group_by(data_type, wn) %>%
  dplyr::summarise(value_x_min = min(value_x),
                   value_x_max = max(value_x),
                   .groups = "keep") %>%
  dplyr::mutate(y_variable = "Holocellulose",
                model = "Binned")

## Klason lignin

# original model
d_prediction_domain_kl_or <-
  tibble::tibble(
    variable_y = "Klason lignin",
    model = "Original",
    data_type = 
      c(
        rep("Training", sum(d$sample_type != "office paper")),
        rep("Peat", nrow(d_peat_res)),
        rep("Vegetation", nrow(d_veg_res))
      ),
    predictor_value = 
      c(
        d$arom15arom16[d$sample_type != "office paper"],
        d_peat_res$arom15arom16,
        d_veg_res$arom15arom16
      )
  )

d_prediction_domain_kl_or <-
  dplyr::bind_cols(
    d_prediction_domain_kl_or,
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_prediction_domain_kl_or %>%
                                    dplyr::mutate(arom15arom16 = predictor_value)) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      })
    )

# binned spectra, best model
d_prediction_domain_kl_bin <-
  dplyr::bind_rows(
    # training data
    d_underfitting[[1]] %>% 
      dplyr::select(-kl, -hol) %>%
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_underfitting_flat[[1]]$x %>%
          rep(nrow(d_underfitting[[1]])),
        sample_id = 
          seq_len(nrow(d_underfitting[[1]])) %>% 
          rep(each = nrow(d_underfitting_flat[[1]])) %>% 
          as.character(),
        data_type = "Training"
      ),
    # peat data
     d_peat_bin10 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_peat_bin10_flat$x %>%
          rep(nrow(d_peat_bin10)),
        sample_id = 
          seq_len(nrow(d_peat_bin10)) %>% 
          rep(each = nrow(d_peat_bin10_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[1]])) %>%
          as.character(),
        data_type = "Peat"
      ),
    # vegetation data
     d_veg_bin10 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_veg_bin10_flat$x %>%
          rep(nrow(d_veg_bin10)),
        sample_id = 
          seq_len(nrow(d_veg_bin10)) %>% 
          rep(each = nrow(d_veg_bin10_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[1]]) + nrow(d_peat_bin10)) %>%
          as.character(),
        data_type = "Vegetation"
      )
  ) %>%
  dplyr::group_by(data_type, wn) %>%
  dplyr::summarise(value_x_min = min(value_x),
                   value_x_max = max(value_x),
                   .groups = "keep") %>%
  dplyr::mutate(y_variable = "Klason lignin",
                model = "Binned")

## plots

# colors_prediction_domain <- scales::hue_pal()(3)
# colors_prediction_domain <- scales::grey_pal(0.1, 0.9)(3)[c(3, 2, 1)]
# colors_prediction_domain <- scales::viridis_pal(option = "plasma")(3)
# colors_prediction_domain <- c("grey", "orange1", "lightcoral") %>% rev()
colors_prediction_domain <- scales::brewer_pal(palette = "Paired")(3)
colors_prediction_domain[[3]] <- "salmon"
data_type_levels <- c("Training", "Vegetation", "Peat") %>% rev()

# make colors darker: source: https://gist.github.com/Jfortin1/72ef064469d1703c6b30
colors_prediction_domain_darker <- 
  colors_prediction_domain %>%
  col2rgb() %>%
  magrittr::divide_by(2) %>%
  t() %>%
  grDevices::rgb(maxColorValue = 255)

p1 <-
  d_prediction_domain_hol_or %>%
  dplyr::mutate(
    data_type = factor(data_type, levels = data_type_levels)
  ) %>%
  ggplot(aes(x = predictor_value, fill = data_type)) +
  geom_histogram(colour = NA, bins = 30) +
  labs(x = "carb", y = "Count", title = "Holocellulose") +
  guides(fill = guide_legend(title = "Dataset")) +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p2 <-
  d_prediction_domain_kl_or %>%
  dplyr::mutate(
    data_type = factor(data_type, levels = data_type_levels)
  ) %>%
  ggplot(aes(x = predictor_value, fill = data_type)) +
  geom_histogram(colour = NA, bins = 30) +
  labs(x = "arom15arom16", y = "Count", title = "Klason lignin") +
  guides(fill = guide_legend(title = "Dataset")) +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p3 <-
  d_prediction_domain_hol_bin %>%
  dplyr::mutate(
    data_type = factor(data_type, levels = data_type_levels)
  ) %>%
  ggplot(aes(x = wn, ymin = value_x_min, ymax = value_x_max, fill = data_type)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_ribbon(colour = NA, alpha = 1) +
  geom_vline(xintercept = 
               m_pars %>%
               dplyr::filter(stringr::str_detect(variable_y, "Holocellulose") & abs(median) >= 0.02) %>%
               dplyr::pull(wn),
             colour = "grey", linetype = 2) + 
  labs(x = expression("Bin wavenumber ["*cm^{-1}*"]"),
       y = "Predictor value",
       title = "Bin width = 50") +
  guides(fill = guide_legend(title = "Dataset")) +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p4 <-
  d_prediction_domain_kl_bin %>%
  dplyr::mutate(
    data_type = factor(data_type, levels = data_type_levels)
  ) %>%
  ggplot(aes(x = wn, ymin = value_x_min, ymax = value_x_max, fill = data_type)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_ribbon(colour = NA, alpha = 1) +
  geom_vline(xintercept =  
               m_pars %>%
               dplyr::filter(stringr::str_detect(variable_y, "Klason lignin") & abs(median) >= 0.02) %>%
               dplyr::pull(wn),
             colour = "grey", linetype = 2) + # ---todo: remove hard coding of values
  labs(x = expression("Bin wavenumber ["*cm^{-1}*"]"),
       y = "Predictor value",
       title = "Bin width = 10") +
  guides(fill = guide_legend(title = "Dataset")) +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine plots
pp1 <-
  (p1 | p2) +
  plot_annotation(title = "Original models")

pp2 <-
  p3 / p4 +
  plot_annotation(title = "Models using binned spectra")

p_prediction_domain <- 
  pp1 / pp2 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(heights = c(0.4, 2), guides = "collect") &
  theme(legend.position = "bottom",
        legend.box = "vertical")
```

### 2022-02-03 analysis/paper/007-paper-m-prediction-differences.Rmd
Initial commit.

```{r m-prediction-differences-1}
# create a data frame for plotting
d_pred_diff <-
  dplyr::bind_rows(
    dplyr::bind_rows(
    ## training data
    # Original Gaussian
    d %>%
      dplyr::mutate(
        hol_hat =
          m1.2 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.2 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Gaussian"
      ) %>%
      dplyr::select(-spectra),
    # Original Beta
    d %>%
      dplyr::mutate(
        hol_hat =
          m1.3 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.3 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Beta"
      ) %>%
      dplyr::select(-spectra),
    # Best all peaks
    d %>%
      dplyr::mutate(
        hol_hat =
          m1.6 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.6 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best all peaks"
      ) %>%
      dplyr::select(-spectra),
    # Best binned spectra
    d %>%
      dplyr::mutate(
        hol_hat =
          m1.10 %>%
          posterior_predict(newdata = d_underfitting[[4]]) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.7 %>%
          posterior_predict(newdata = d_underfitting[[1]]) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best binned spectra"
      ) %>%
      dplyr::select(-spectra)
    ) %>%
      dplyr::mutate(data_type = "Training"),

    dplyr::bind_rows(
    ## peat data
    # Original Gaussian
    d_peat %>%
      dplyr::mutate(
        hol_hat =
          m1.2 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.2 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Gaussian"
      ) %>%
      dplyr::select(-spectra),
    # Original Beta
    d_peat %>%
      dplyr::mutate(
        hol_hat =
          m1.3 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.3 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Beta"
      ) %>%
      dplyr::select(-spectra),
    # Best all peaks
    d_peat %>%
      dplyr::mutate(
        hol_hat =
          m1.6 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.6 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best all peaks"
      ) %>%
      dplyr::select(-spectra),
    # Best binned spectra
    d_peat %>%
      dplyr::mutate(
        hol_hat =
          m1.10 %>%
          posterior_predict(newdata = d_peat_bin50) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.7 %>%
          posterior_predict(newdata = d_peat_bin10) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best binned spectra"
      ) %>%
      dplyr::select(-spectra)
    ) %>%
      dplyr::mutate(data_type = "Peat") %>%
      dplyr::select(-dplyr::contains(c("holocellulose_hodgkins", "klason_lignin_hodgkins"))) %>%
      dplyr::bind_cols(purrr::map_df(1:4, function(x) d_peat_res)),

    dplyr::bind_rows(
    ## vegetation data
    # Original Gaussian
    d_veg %>%
      dplyr::mutate(
        hol_hat =
          m1.2 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.2 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Gaussian"
      ) %>%
      dplyr::select(-spectra),
    # Original Beta
    d_veg %>%
      dplyr::mutate(
        hol_hat =
          m1.3 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.3 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Beta"
      ) %>%
      dplyr::select(-spectra),
    # Best all peaks
    d_veg %>%
      dplyr::mutate(
        hol_hat =
          m1.6 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.6 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best all peaks"
      ) %>%
      dplyr::select(-spectra),
    # Best binned spectra
    d_veg %>%
      dplyr::mutate(
        hol_hat =
          m1.10 %>%
          posterior_predict(newdata = d_veg_bin50) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.7 %>%
          posterior_predict(newdata = d_veg_bin10) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best binned spectra"
      ) %>%
      dplyr::select(-spectra)
    ) %>%
      dplyr::mutate(data_type = "Vegetation") %>%
      dplyr::select(-dplyr::contains(c("holocellulose_hodgkins", "klason_lignin_hodgkins"))) %>%
      dplyr::bind_cols(purrr::map_df(1:4, function(x) d_veg_res))
  )

# add additional variables
d_pred_diff <-
  d_pred_diff %>%
  dplyr::mutate(
    hol_residuals = hol - hol_hat,
    kl_residuals = kl - kl_hat,
    hol_hat_or = 
      purrr::map(unique(d_pred_diff$data_type), function(x) {
        d_pred_diff %>% 
          dplyr::filter(model == "Original Gaussian" & data_type == x) %>%
          dplyr::pull(hol_hat) %>%
          rep(4)
      }) %>% 
      unlist(),
    kl_hat_or = 
      purrr::map(unique(d_pred_diff$data_type), function(x) {
        d_pred_diff %>% 
          dplyr::filter(model == "Original Gaussian" & data_type == x) %>%
          dplyr::pull(kl_hat) %>%
          rep(4)
      }) %>% 
      unlist(),
    arom15 = as.numeric(arom15) * res_scale$scale[res_scale$name == "arom15"] + res_scale$center[res_scale$name == "arom15"],
    arom16 = as.numeric(arom16) * res_scale$scale[res_scale$name == "arom16"] + res_scale$center[res_scale$name == "arom16"],
    arom15arom16 = arom15 + arom16,
    arom15arom16_rel = arom15/arom15arom16
  ) %>% 
  dplyr::mutate(
    model = factor(model, levels = c("Original Gaussian", "Original Beta", "Best all peaks", "Best binned spectra")),
    data_type = factor(data_type, levels = c("Training", "Peat", "Vegetation")),
    sample_type = factor(sample_type, levels = c("needles", "hardwood", "leaves and grasses", "softwood", "old corrugated cardboard", "old newsprint", "office paper", "old magazines"))
  )

# spectra
d_pred_diff_spectra <-
  ir::ir_new_ir(
    spectra = 
      dplyr::bind_rows(
        purrr::map_df(1:4, function(x) d %>% dplyr::select(spectra)),
        purrr::map_df(1:4, function(x) d_peat %>% dplyr::select(spectra)),
        purrr::map_df(1:4, function(x) d_veg %>% dplyr::select(spectra))
      ) %>% dplyr::pull(spectra),
    sample_id =
      d_pred_diff %>%
      nrow() %>%
      seq_len() %>%
      as.character()
  )

d_pred_diff_intensities <-
  d_pred_diff_spectra %>%
  ir::ir_get_intensity(wavenumber = c(3400, 1200, 1550)) %>%
  ir::ir_drop_spectra() %>%
  tidyr::unnest(cols = "intensity") %>%
  tidyr::pivot_wider(names_from = x, values_from = y, names_prefix = "intensity_")
```

```{r m-prediction-differences-2}
## plot of measured vs fitted values for the training data

# Holocellulose
p_index <- with(d_pred_diff, data_type == "Training" & 
                  sample_type != "old magazines" & 
                  model != "Original Beta")

p_pred_diff_train1_p1 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = 
                  d_pred_diff_intensities %>%
                  dplyr::pull(intensity_1550) %>%
                  scale()
  ) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, y = hol * 100)) +
  geom_point(aes(colour = scale(carb),
                 size = carb_rel)) +
  #geom_point(data = 
  #             d_pred_diff %>%
  #             dplyr::mutate(carb_rel = 
  #                
  #                  d_pred_diff_intensities %>%
  #                    dplyr::pull(intensity_1550)) %>%
  #             dplyr::filter(p_index & sample_id %in% d$sample_id[d$sample_type != "old magazines"][hol_sel]), 
  #           aes(x = hol_hat * 100, y = hol * 100,
  #               size = carb_rel), 
  #           fill = NA, shape = 21, colour = "grey50") +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = expression(Absorbance~at~1550~cm^{-1})),
         colour = guide_colourbar(title = "carb")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Holocellulose") +
  coord_fixed() +
  theme(legend.position = "bottom")

## Klason lignin

p_index <- with(d_pred_diff, data_type == "Training" & 
                  sample_type != "office paper" & 
                  model != "Original Beta")

p_pred_diff_train1_p2 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = arom15arom16_rel) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl * 100)) +
  geom_point(aes(colour = scale(arom15arom16),
                 size = carb_rel)) +
  #geom_point(data = 
  #             d_pred_diff %>%
  #             dplyr::mutate(carb_rel = arom15arom16_rel) %>%
  #             dplyr::filter(p_index & sample_id %in% d$sample_id[d$sample_type != "old magazines"][kl_sel]), 
  #           aes(x = kl_hat * 100, y = kl * 100,
  #               size = carb_rel), 
  #           fill = NA, shape = 21, colour = "grey50") +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = expression("arom15/(arom15arom16)")),
         colour = guide_colourbar(title = "arom15arom16")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Klason lignin") +
  coord_fixed() +
  theme(legend.position = "bottom")

# combine
p_pred_diff_train1 <-
  p_pred_diff_train1_p1 / p_pred_diff_train1_p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")")

### Start: test

hol_sel <- c(40, 18, 49)
hol_sel <- c(30, 39, 15)

d[d$sample_type != "old magazines", ][hol_sel, ]  %>%
  irpeat:::irp_content_klh_hodgkins_prepare() %>%
  hodgkins_ggplot() %>%
  wrap_plots()

d_pred_diff %>%
  dplyr::mutate(carb_rel = 
                  d_pred_diff_intensities %>%
                  dplyr::pull(intensity_1550)) %>%
  dplyr::filter(model == "Best binned spectra") %>%
  ggplot(aes(x = carb, y = (hol_hat - hol_hat_or) * 100)) +
  geom_point() +
  facet_grid(data_type ~ cut_number(carb_rel, 5))

p1_index <- with(d_pred_diff, (sample_type != "old magazines" | is.na(sample_type)) & 
                   model != "Original Beta")

p1 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = 
                  d_pred_diff_intensities %>%
                  dplyr::pull(intensity_1550)) %>%
  dplyr::filter(p1_index) %>%
  ggplot(aes(x = hol_hat * 100, y = hol_hat_or * 100)) +
  geom_point(aes(colour = carb,
                 size = carb_rel,
                 shape = data_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = "carb/OH-peak")) +
  labs(x = "Fitted values", y = "Measured values") +
  coord_fixed()

hol_sel <- c(1, 11, 12, 6) # transect from underestimation to overestimation for the best binned model.

### End: test
```

```{r m-prediction-differences-3}
## plot of measured vs fitted values for the training data

# Holocellulose
p_index <- with(d_pred_diff, data_type == "Training" & 
                  sample_type != "old magazines" & 
                  model != "Original Beta")

p1 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, 
             y = hol * 100)) +
  geom_point(aes(colour = sample_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Sample type")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Holocellulose") +
  coord_fixed() +
  scale_colour_manual(values = scales::hue_pal()(length(levels(d_pred_diff$sample_type))),
                    labels = levels(d_pred_diff$sample_type), 
                    drop = FALSE)

## Klason lignin

p_index <- with(d_pred_diff, data_type == "Training" & 
                  sample_type != "office paper" & 
                  model != "Original Beta")

p2 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl * 100)) +
  geom_point(aes(colour = sample_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Sample type")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Klason lignin") +
  coord_fixed() +
  scale_colour_manual(values = scales::hue_pal()(length(levels(d_pred_diff$sample_type))),
                    labels = levels(d_pred_diff$sample_type), 
                    drop = FALSE)

# combine
p_pred_diff_train2 <-
  p1 / p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "right")

### Test start:

# in the plots above, one can see that the model makes different predictions for two OM classes: needles and leaves and grasses. How do these spectra differ? It seem that this is mainly th region around 1550
d %>% 
  dplyr::filter(sample_type %in% c("needles", "leaves and grasses")) %>%
  plot() + 
  geom_path(aes(x = x, y = y, colour = sample_type, group = measurement_id)) +
  geom_vline(xintercept = 1550)
```

```{r m-prediction-differences-4}
## plot of fitted (original model) vs fitted values for the training data

# Holocellulose
p_index <- with(d_pred_diff, 
                data_type == "Training" & 
                  sample_type != "old magazines" & 
                  model != "Original Beta")

p1 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = 
                  d_pred_diff_intensities %>%
                  dplyr::pull(intensity_1550) %>%
                  scale()
  ) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, y = hol_hat_or * 100)) +
  geom_point(aes(colour = scale(carb),
                 size = carb_rel)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = expression(Absorbance~at~1550~cm^{-1})),
         colour = guide_colourbar(title = "carb")) +
  labs(x = "Fitted values", y = "Fitted values (original model)",
       title = "Holocellulose") +
  coord_fixed() +
  theme(legend.position = "bottom")

## Klason lignin

p_index <- with(d_pred_diff, 
                data_type == "Training" & 
                  sample_type != "office paper" & 
                  model != "Original Beta")

p2 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = scale(arom15arom16_rel)) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl_hat_or * 100)) +
  geom_point(aes(colour = scale(arom15arom16),
                 size = carb_rel)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = expression("arom15/(arom15arom16)")),
         colour = guide_colourbar(title = "arom15arom16")) +
  labs(x = "Fitted values", y = "Fitted values (original model)",
       title = "Klason lignin") +
  coord_fixed() +
  theme(legend.position = "bottom")

# combine
p_pred_diff_train3 <-
  p1 / p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")")
```

```{r m-prediction-differences-5}
## plot of measured vs fitted values for the training data, peat data and vegetation data

# Holocellulose
p_index <- 
  with(
    d_pred_diff %>%
      dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
      dplyr::arrange(dplyr::desc(data_type)),
    (sample_type != "old magazines" | is.na(sample_type)) & 
      model != "Original Beta"
  )

p1 <-
  d_pred_diff %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  dplyr::arrange(dplyr::desc(data_type)) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, y = hol_hat_or * 100)) +
  geom_point(aes(colour = data_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Data set")) +
  labs(x = "Predicted values", y = "Predicted values (original model)",
       title = "Holocellulose") +
  coord_fixed() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

## Klason lignin

p_index <- 
  with(
    d_pred_diff %>%
      dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
      dplyr::arrange(dplyr::desc(data_type)),
    (sample_type != "office paper" | is.na(sample_type)) & 
      model != "Original Beta"
  )

p2 <-
  d_pred_diff %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  dplyr::arrange(dplyr::desc(data_type)) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl_hat_or * 100)) +
  geom_point(aes(colour = data_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Data set")) +
  labs(x = "Predicted values", y = "Predicted values (original model)",
       title = "Klason lignin") +
  coord_fixed() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine
p_pred_diff_all1 <-
  p1 / p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r m-prediction-differences-6}
## for Klason lignin: residuals versus arom15arom16 conditional on arom15arom16_rel

p_index <- with(d_pred_diff, 
                data_type == "Training" & 
                  sample_type != "office paper" & 
                  model != "Original Beta")

p_pred_diff_kl_residuals <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = arom15arom16 %>% scale(), y = kl_residuals)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, colour = "dimgrey") +
  geom_hline(yintercept = 0, colour = "grey50") +
  facet_grid(cut_number(arom15arom16_rel, n = 4) ~ model) +
  labs(x = "arom15arom16", y = "Residuals") + 
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Standardized normalized arom15arom16 peak height", breaks = NULL, labels = NULL)) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Model", breaks = NULL, labels = NULL)) +
  theme(
    axis.line.y.right = element_line(color = "white"),
    axis.line.x.top = element_line(color = "white")
    )
```

```{r m-prediction-differences-7}
## for holocellulose: spectra across the gradient in predictions differences between the original model and the improved models

p_index <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Best binned spectra")

p_pred_diff_hol_spectra_diff1 <-
  d_pred_diff_spectra %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  plot() +
  geom_path(aes(x = x, y = y, group = measurement_id), 
            colour = "white") +
  geom_path(aes(x = x, y = y, group = measurement_id, colour = hol_hat_or - hol_hat), 
            alpha = 0.3) +
  scale_colour_gradient2(low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))  +
  facet_wrap(~ data_type, ncol = 1) +
  guides(colour = guide_colorbar(title = "Prediction difference")) +
  theme(legend.position = "bottom") +
  labs(x = expression("Wavenumber ["*cm^{-1}*"]"), y = "Absorbance")

p_pred_diff_hol_spectra_diff2 <-
  d_pred_diff_spectra %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  plot() +
  geom_path(aes(x = x, y = y, group = measurement_id), 
            colour = "white") +
  geom_path(aes(x = x, y = y, group = measurement_id, colour = hol_hat_or - hol_hat), 
            alpha = 0.3) +
  scale_colour_gradient2(low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))  +
  facet_grid(cut_number(carb, 4) ~ data_type) +
  guides(colour = guide_colorbar(title = "Prediction difference")) +
  labs(x = expression("Wavenumber ["*cm^{-1}*"]"), y = "Absorbance") + 
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Standardized normalized carb peak height", breaks = NULL, labels = NULL)) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Sample type", breaks = NULL, labels = NULL)) +
  theme(
    legend.position = "bottom",
    axis.line.y.right = element_line(color = "white"),
    axis.line.x.top = element_line(color = "white")
  )

## same for Klason lignin
p_index <- with(d_pred_diff, 
                  (sample_type != "office paper" | is.na(sample_type)) & 
                  model == "Best binned spectra")

p_pred_diff_kl_spectra_diff1 <-
  d_pred_diff_spectra %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, kl_hat_or, kl_hat, arom15arom16)
  ) %>%
  dplyr::filter(p_index) %>%
  plot() +
  geom_path(aes(x = x, y = y, group = measurement_id), 
            colour = "white") +
  geom_path(aes(x = x, y = y, group = measurement_id, colour = kl_hat_or - kl_hat), 
            alpha = 0.3) +
  scale_colour_gradient2(low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))  +
  facet_wrap(~ data_type, ncol = 1) +
  guides(colour = guide_colorbar(title = "Prediction difference")) +
  theme(legend.position = "bottom") +
  labs(x = expression("Wavenumber ["*cm^{-1}*"]"), y = "Absorbance")

p_pred_diff_kl_spectra_diff2 <-
  d_pred_diff_spectra %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, kl_hat_or, kl_hat, arom15arom16)
  ) %>%
  dplyr::filter(p_index) %>%
  plot() +
  geom_path(aes(x = x, y = y, group = measurement_id), 
            colour = "white") +
  geom_path(aes(x = x, y = y, group = measurement_id, colour = kl_hat_or - kl_hat), 
            alpha = 0.3) +
  scale_colour_gradient2(low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))  +
  facet_grid(cut_number(arom15arom16, 4) ~ data_type) +
  guides(colour = guide_colorbar(title = "Prediction difference")) +
  labs(x = expression("Wavenumber ["*cm^{-1}*"]"), y = "Absorbance") + 
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Standardized normalized arom15arom16 peak height", breaks = NULL, labels = NULL)) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Sample type", breaks = NULL, labels = NULL)) +
  theme(
    legend.position = "bottom",
    axis.line.y.right = element_line(color = "white"),
    axis.line.x.top = element_line(color = "white")
  )
```

```{r m-prediction-differences-8}
## for holocellulose: plot prediction differences versus carb peak height conditional on different spectral variables

p_index <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Best binned spectra" &
                  (!core_label %in% c("MB-775", "MB-930") | is.na(core_label)))
facet_n_level <- 3

p_pred_diff_hol_intensity1 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_3400 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
       title = expression("Intensity at 3400 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p_pred_diff_hol_intensity2 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1200 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 1200 cm"^{-1}), 
       title = expression("Intensity at 1200 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p_pred_diff_hol_intensity3 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1550 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 1550 cm"^{-1}), 
       title = expression("Intensity at 1550 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine
p_pred_diff_hol_intensity1 <-
  p_pred_diff_hol_intensity1 /
  p_pred_diff_hol_intensity2 /
  p_pred_diff_hol_intensity3 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

## same plots with Mer Bleue cores

p_index <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Best binned spectra")

p_pred_diff_hol_intensity4 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1200 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
       title = expression("Intensity at 3400 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p_pred_diff_hol_intensity5 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1200 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 1200 cm"^{-1}), 
       title = expression("Intensity at 1200 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p_pred_diff_hol_intensity6 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1550 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 1550 cm"^{-1}), 
       title = expression("Intensity at 1550 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine
p_pred_diff_hol_intensity2 <-
  p_pred_diff_hol_intensity4 /
  p_pred_diff_hol_intensity5 /
  p_pred_diff_hol_intensity6 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# plots just showing boxplots of OH peak height between different data types
p_oh_height_difference <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(y = intensity_3400, x = data_type, fill = data_type)) +
  geom_boxplot(width = 0.5) +
  labs(y = expression(Intensity~at~3400~cm^{-1}), x = "Data type") +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  ) +
  guides(fill = guide_legend(title = "Dataset"))

```

```{r, eval=FALSE}
## Test start
p_index_test <- with(d_pred_diff, 
                  (sample_type != "old magazines"))

d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb, hol_residuals, model)
  ) %>%
  dplyr::filter(p_index_test) %>%
  ggplot(aes(x = intensity_3400 %>% scale(), y = hol_residuals, colour = model)) + 
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(model ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
       title = expression("Intensity at 3400 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb, hol_residuals, model)
  ) %>%
  dplyr::filter(p_index_test) %>%
  ggplot(aes(x = intensity_1200 %>% scale(), y = hol_residuals, colour = model)) + 
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(model ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 1200 cm"^{-1}), 
       title = expression("Intensity at 1200 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb, hol_residuals, model)
  ) %>%
  dplyr::filter(p_index_test) %>%
  ggplot(aes(x = intensity_1550 %>% scale(), y = hol_residuals, colour = model)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(model ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 1550 cm"^{-1}), 
       title = expression("Intensity at 1550 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

# same plots as originals again, but with restricted carb values

p_index_test2 <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Best binned spectra" &
                    carb >= -1.45)

p_pred_diff_hol_intensity7 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index_test2) %>%
  ggplot(aes(x = intensity_3400 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
       title = expression("Intensity at 3400 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

p_pred_diff_hol_intensity8 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index_test2) %>%
  ggplot(aes(x = intensity_1200 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 1200 cm"^{-1}), 
       title = expression("Intensity at 1200 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

p_pred_diff_hol_intensity9 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index_test2) %>%
  ggplot(aes(x = intensity_1550 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 1550 cm"^{-1}), 
       title = expression("Intensity at 1550 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")
```

```{r m-prediction-differences-9}
## Klason lignin: histogram of arom15arom16 conditional on arom15arom16_rel

p_index <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Original Gaussian") # model is in principle unimportant, this just avoids duplicates

# plot
p_pred_diff_kl_histogram1 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels)) %>%
  ggplot() +
  geom_histogram(aes(x = arom15arom16 %>% scale(), fill = data_type), bins = 30) +
  facet_wrap(~ cut_number(arom15arom16_rel, n = 4)) +
  labs(y = "Count", x = "arom15arom16") +
  guides(fill = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )
```

### 2022-02-03 analysis/paper/008-paper-supplementary.Rmd
Initial commit.

---
title: 'Supplementary Information to "Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra"'
author: "Henning Teickner, Klaus-Holger Knorr"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
header-includes:
   - \usepackage{float}
   - \renewcommand\thesection{S\arabic{section}}
---

```{r sup-setup, echo=FALSE}
knitr::opts_chunk$set(fig.align = "center", 
                      fig.pos = "H",
                      echo = FALSE,
                      cache = TRUE,
                      dev = "pdf")
```

# Comparison of original non-Bayesian and original Bayesian regression models

Throughout the main text, we analyzed weaknesses of the models as developed by @Hodgkins.2018 using a Bayesian implementation of these original models. Here, we show that working within a Bayesian framework does not create the bias or any other weaknesses we discussed. To this end, we compare predictions of the original non-Bayesian models (implemented with `lm` [@RCoreTeam.2020]) with those of the original Bayesian model we used as baseline model throughout the main text.  
Figure \@ref(fig:res-sup-original-models) compares median fitted values, and lower and upper 90% prediction intervals for both models.

```{r sup-original-models, echo=FALSE, include=FALSE}
# data frame with predicted values from both models
d_compare_original <- 
  dplyr::bind_cols(
    dplyr::bind_rows(
    m1.1 %>%
      predict(interval = "prediction",
              level = 0.9) %>%
      as.data.frame() %>%
      dplyr::mutate(
        variable_y = "Holocellulose"
      ),
    m2.1 %>%
      predict(interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        variable_y = "Klason lignin"
        )
    ) %>%
      tidyr::pivot_longer(cols = dplyr::one_of(c("fit", "lwr", "upr")),
                          names_to = "variable_quantile",
                          values_to = "value"),
    dplyr::bind_rows(
      m1.2 %>%
      rstanarm::posterior_predict() %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit_bayes = median(x),
          lwr_bayes = quantile(x, prob = 0.05),
          upr_bayes = quantile(x, prob = 0.95)
        )
      }),
    m2.2 %>%
      rstanarm::posterior_predict() %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit_bayes = median(x),
          lwr_bayes = quantile(x, prob = 0.05),
          upr_bayes = quantile(x, prob = 0.95)
        )
      })
    ) %>%
      tidyr::pivot_longer(cols = dplyr::one_of(c("fit_bayes", "lwr_bayes", "upr_bayes")),
                          names_to = "variable_quantile",
                          values_to = "value_bayes") %>%
      dplyr::select(-variable_quantile)
  ) %>%
  dplyr::mutate(
    variable_quantile = dplyr::case_when(
      variable_quantile == "fit" ~ "Median",
      variable_quantile == "lwr" ~ "Lower 90%",
      variable_quantile == "upr" ~ "Upper 90%",
    )
  )

# annotations
d_compare_original_annotation <-
  d_compare_original %>%
  dplyr::group_by(variable_y, variable_quantile) %>%
  dplyr::summarise(correlation = cor(value, value_bayes, method = "pearson"),
                   .groups = "drop")

# plot
p1 <-
  d_compare_original %>%
  dplyr::filter(variable_y == "Holocellulose") %>%
  ggplot(aes(x = value * 100, y = value_bayes * 100)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey") +
  geom_point() +
  labs(title = "Holocellulose",
       x = "Non-Bayesian",
       y = "Bayesian") +
  facet_wrap(~ variable_quantile) +
  coord_fixed()

p2 <-
  d_compare_original %>%
  dplyr::filter(variable_y == "Klason lignin") %>%
  ggplot(aes(x = value * 100, y = value_bayes * 100)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey") +
  geom_point() +
  labs(title = "Klason lignin",
       x = "Non-Bayesian",
       y = "Bayesian") +
  facet_wrap(~ variable_quantile) +
  coord_fixed() 
```

```{r res-sup-original-models, echo=FALSE, out.width="70%", fig.height=5, fig.width=6, fig.cap="Comparison of the original non-Bayesian model (the models used by \\citet{Hodgkins.2018}) and original Bayesian model used in the analyses of the main text. Shown are lower and upper 90\\% prediction intervals of both models plotted against each other, as well as median predicitons, for holocellulose (a) and Klason lignin (b), respectively. All values have a Pearson correlation of 1."}
p1 / p2 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

\clearpage

# Simulation experiment

Here, we present results of a simulation study that shows exemplarily how changes in heights of other peaks can affect the normalized peak height of the `carb` peak and absorbances at other wavenumbers. With this, we want to demonstrate that both the original and improved models are sensitive to the overall spectra and therefore differences between training data and sample data can cause differences in predictions.  
In the simulation study, we simulated spectra with the same `carb` peak height, but a differently large OH-peak height. We used the script from @Hodgkins.2018 to compute the normalized `carb` peak height in the spectra and analyzed also the absorbance at other wavenumbers. Figure \@ref(fig:res-sup-spectra-sim-p) shows the simulated spectra.

```{r sup-simulation, echo=FALSE, include=FALSE}
# simulate spectra
wn <- d$spectra[[1]]$x    # wavenumber range
carb_range <- 900:1190
carb_pos <- mean(carb_range) # approximate position of carb peak
carb_width <- 0.0003 # defines width of carb peak
carb_height <- seq(from = 20, to = 75, by = 5) # sequence of carb peak heights
oh_range <- 2500:3700
oh_pos <- mean(oh_range) # approximate position of OH peak
oh_width <- 0.00002 # defines width of OH peak
oh_height <- seq(from = 20, to = 75, by = 20) # sequence of OH peak heights
bl_height <- 10 # absorbance in other parts of spectrum

# simulate spectra
d_spectra_sim_raw <- rep(bl_height, length(wn))
d_spectra_sim <-
  expand.grid(carb_pos = carb_pos, carb_width = carb_width, carb_height = carb_height, oh_pos = oh_pos, oh_width = oh_width, oh_height = oh_height, bl_height = bl_height) %>%
  dplyr::mutate(measurement_id = seq_len(nrow(.)),
                sample_id = as.character(measurement_id))
d_spectra_sim$spectra <- 
  purrr::map(seq_len(nrow(d_spectra_sim)), function(i) {
    res <- 
      tibble::tibble(
        x = wn,
        y = d_spectra_sim_raw
      )
    res$y[wn %in% carb_range] <- (d_spectra_sim$carb_height[i] -d_spectra_sim$carb_width[i] * (carb_range - d_spectra_sim$carb_pos[i])^2)
    res$y[wn %in% oh_range] <- (d_spectra_sim$oh_height[i] -d_spectra_sim$oh_width[i] * (oh_range - d_spectra_sim$oh_pos[i])^2)
    res
  })

class(d_spectra_sim) <- c("ir", class(d_spectra_sim))
d_spectra_sim <-
  d_spectra_sim %>%
  ir::ir_normalise()

d_spectra_sim_res <-
  d_spectra_sim %>%
  irpeat:::irp_content_klh_hodgkins_prepare() %>%
  irpeat:::irp_content_klh_hodgkins_main(export = NULL,
                                         verbose = FALSE,
                                         make_plots = FALSE)

d_spectra_sim$carb_height_sim <- d_spectra_sim_res$norm.Acorr$carb

# plot extracted carb peak height versus simulated peak height in dependency of OH peak height
spectra_sim_p1 <-
  d_spectra_sim %>%
  ggplot(aes(x = carb_height, y = carb_height_sim, colour = oh_height)) +
  geom_point() +
  labs(y = "carb peak height (extracted and normalized)",
       x = "carb peak height (simulated)") +
  guides(colour = guide_colorbar(title = "OH-peak height (simulated)")) +
  theme(legend.position = "bottom")

# plot absorbance at 1250 in dependency of OH peak height
spectra_sim_p2 <-
  d_spectra_sim %>%
  ir::ir_get_intensity(wavenumber = 1250) %>%
  dplyr::mutate(intensity = 
                  intensity %>%
                  bind_rows() %>%
                  dplyr::pull(y)) %>%
  ggplot(aes(x = carb_height, y = intensity, colour = oh_height)) +
  geom_point() +
  labs(y = expression("Absorbance at 1250 cm"^{-1}),
       x = "carb peak height (simulated)") +
  guides(colour = guide_colorbar(title = "OH-peak height (simulated)")) +
  theme(legend.position = "bottom")

# plot of simulated spectra
spectra_sim_p3 <-
  d_spectra_sim %>%
  plot() +
  geom_path(aes(x = x, y = y, colour = carb_height, group = measurement_id)) +
  facet_wrap(~ oh_height) +
  theme(legend.position = "bottom") +
  guides(colour = guide_colorbar(title = "carb peak height (simulated)")) +
  labs(y = "Absorbance (normalized)", x = expression("Wavenumber ["*cm^{-1}*"]"))

spectra_sim_p <-
  (spectra_sim_p1 +
     spectra_sim_p2) +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

spectra_sim_p <-
  spectra_sim_p3 /
  spectra_sim_p +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")")
```

```{r res-sup-spectra-sim-p, echo=FALSE, out.width="80%", fig.height=7, fig.width=7, fig.cap="Results of the simulation experiment. (a) Plots of the simulated spectra with different OH-peak heights (columns) and different carb peak heights (colour gradients). Due to the normalization (division of each absorbance value by the sum of all absorbance value), the normalized simulated carb peak heights get smaller as the OH-peak gets larger. For this reason, the extracted normalized carb peak height depends not only on the simulated carb peak height, but is also sensitive to the OH-peak height (b). This is also true for absorbance values extracted at specific wavenumbers (c)."}
spectra_sim_p
```

\clearpage

# Supplementary figures

Figure \@ref(fig:res-p-gaussian-beta-depth-profile) shows depth profiles of holocellulose and Klason lignin contents for the peat cores from @Hodgkins.2018, predicted either with the original normal regression model or the beta regression model with the same model structure.

```{r res-p-gaussian-beta-depth-profile, echo=FALSE, out.width="70%", fig.height=9, fig.width=6, fig.cap="Depth profiles of predicted Holocellulose (a) and Klason lignin contents (b) for peat core data collected by \\cite{Hodgkins.2018}. Different panels represent different peat cores. Coloured curves are median point predictions, shaded regions are 90\\% prediction intervals, dashed vertical lines represent a content of 0  mass-\\%. Colours differentiate predictions of a Gaussian Bayesian model (blue) and a Bayesian beta regression model (red). All models had the same model structure as in \\cite{Hodgkins.2018}."}
p_gaussian_beta_depth_profile
```

\clearpage

Figure \@ref(fig:res-sup-kl-hol) summarizes the _measured_ holocellulose and Klason lignin contents in the OM samples of the training data set of @Hodgkins.2018.

```{r sup-kl-hol, echo=FALSE}
p1 <-
  d %>%
  ggplot(aes(y = (hol + kl) * 100, x = sample_type)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "Sample type", x = "Holocellulose + Klason lignin [mass-%]")

p2 <-
  d %>%
  ggplot(aes(y = kl * 100, x = hol * 100, colour = sample_type)) +
  geom_point() +
  labs(y = "Klason lignin [mass-%]", x = "Holocellulose [mass-%]") +
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  guides(colour = guide_legend(title = "Sample type", ncol = 3, title.position="top")) +
  theme(legend.position = "bottom")
```

```{r res-sup-kl-hol, echo=FALSE, out.width="100%", fig.height=4.5, fig.width=7, fig.cap='Measured holocellulose and Klason lignin contents in OM samples of the training data of \\citet{Hodgkins.2018}. A: Boxplots of the sum of the measured holocellulose and Klason lignin contents for different sample types in the training data. B: Klason lignin content versus holocellulose content for the training data samples. Colours represent different sample types and lines are linear regression lines fitted to the data.'}
p1 + p2 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

\clearpage

Figure \@ref(fig:res-p-gaussian-beta-depth-profile-best) shows holocellulose and Klason lignin contents for the peat core samples from @Hodgkins.2018 predicted by the original model, the improved model using binned spectra, and the improved model using extracted peaks.

```{r res-p-gaussian-beta-depth-profile-best, echo=FALSE, out.width="70%", fig.height=9, fig.width=6, fig.cap='Depth profiles of Holocellulose (a) and Klason lignin contents (b) for the same data as in figure \\ref{fig:res-p-gaussian-beta-depth-profile}, but now comparing predictions of the original model with those from improved models. Different panels represent different peat cores. Coloured curves are median point predictions, shaded regions are 90\\% prediction intervals, dashed vertical lines represent a content of 0 mass-\\%. Colours differentiate predictions of the original Gaussian model ("Original": blue), the best model using extracted peaks ("Best all peaks": red) and the best model using binned spectra ("Best binned spectra": green).'}
p_gaussian_beta_depth_profile_best
```

\clearpage

Figure \@ref(fig:res-p-y-yhat-best) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, and the improved model using extracted peaks.

```{r res-p-y-yhat-best, echo=FALSE, out.width="70%", fig.height=6.5, fig.width=6, fig.cap='Measured values of the training dataset from \\cite{Hodgkins.2018} versus fitted values of the original model (first column) in comparison to the improved models (second and third columns) for (a) holocellulose and (b) Klason lignin. Points are median predictions from the models, smooth lines and shaded regions are LOESS smoothers and corresponding 95\\% confidence intervals. The diagonal lines indicates ideal fit. Points above the line indicate overestimation by the model and points under the diagonal underestimation.'}
p_y_yhat_best
```

\clearpage

Figure \@ref(fig:res-p-pred-diff-hol-spectra-diff2) shows MIRS for the training data, vegetation samples, and peat samples from @Hodgkins.2018, classified according to different `carb` peak heights and colored according to the prediction differences between the original model and the improved model using binned spectra for holocellulose. This plot makes visible the differences between the spectra for peat (and vegetation samples) in comparison to the training data.

```{r res-p-pred-diff-hol-spectra-diff2, warning=FALSE, echo=FALSE, out.width="90%", fig.height=7, fig.width=7, fig.cap='Spectra for the training (left column), peat (middle column), and vegetation (right column) samples of \\cite{Hodgkins.2018} for different ranges of \\texttt{carb} peak heights (rows). The colour scale indicates prediction differences between the original and improved model using binned spectra for holocellulose. Larger values represent larger predicted holocellulose contents using the original model in comparison to the improved model using binned spectra.'}
p_pred_diff_hol_spectra_diff2
```

\clearpage

Figure \@ref(fig:res-p-pred-diff-hol-intensity2) shows prediction differences between the original model and the improved model using binned spectra for holocellulose in dependency of absorbance values at different wavenumbers and for different classes of `carb` peak heights.

```{r res-p-pred-diff-hol-intensity2, warning=FALSE, echo=FALSE, out.width="50%", fig.height=14, fig.width=7, fig.cap='Prediction differences between the original model and improved model using binned spectra for the training (first row), peat (second row), and vegetation (bottom row) samples of \\cite{Hodgkins.2018} for different ranges of \\texttt{carb} peak heights (columns) versus intensities at different wavenumber values on the x axis (a: 3400, b: 1200, c: 1550 cm$^{-1}$). Larger prediction differences mean that the original model made larger predictions than the improved model using binned spectra.'}
p_pred_diff_hol_intensity2
```

\clearpage

Figure \@ref(fig:res-p-pred-diff-kl-residuals) shows residuals of the original model, the improved model using binned spectra, and the improved model using extracted peaks versus standardized `arom15arom16` values and for different levels of the relative contribution of the `arom15` peak height to the `arom15arom16` value.

```{r res-p-pred-diff-kl-residuals, warning=FALSE, echo=FALSE, out.width="50%", fig.height=5, fig.width=6, fig.cap='Residuals of the original model (first column), improved model using extracted peaks (second column), and the improved model using binned spectra (third column) versus \\texttt{arom15arom16} values (is equivalent to values predicted by the original model) for different levels of the relative contribution of the \\texttt{arom15} peak to \\texttt{arom15arom16} (rows). Lines and shaded areas are regression lines fitted for each panel. For samples with large contribution of the \\texttt{arom15} peak to \\texttt{arom15arom16}, the original model is biased (bottom left panel), whereas the imprroved models are not biased.'}
p_pred_diff_kl_residuals
```

\clearpage

Figure \@ref(fig:res-p-pred-diff-kl-histogram1) shows `arom15arom16` values for the training data, peat, and vegetation samples from @Hodgkins.2018, for different levels of the relative contribution of the `arom15` peak height to the `arom15arom16` value. Conditional on this contribution, peat samples regularly have larger or smaller `arom15arom16` values.

```{r res-p-pred-diff-kl-histogram1, echo=FALSE, out.width="60%", fig.height=4, fig.width=5, fig.cap='Histograms of \\texttt{arom15arom16} values for the training, peat, and vegetation samples from \\cite{Hodgkins.2018} for different levels of the relative contribution of the \\texttt{arom15} peak to \\texttt{arom15arom16} (larger values mean the \\texttt{arom15} peak is larger relative to the \\texttt{arom16} peak).'}
p_pred_diff_kl_histogram1
```

\clearpage

Figure \@ref(fig:res-p-pred-diff-train2) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, and the improved model using extracted peaks, while differentiating the sample type classes defined in @Hodgkins.2018. This makes visible that for Klason lignin, the two classes of training samples in the main text comprise different sample types (class 1: wood samples, type 2: leaves and grasses).

```{r res-p-pred-diff-train2, echo=FALSE, out.width="90%", fig.height=6, fig.width=7, fig.cap='Measured values versus fitted values [mass-\\%] for the original model (first column), the improved model using extracted peaks (second column), and the improved model using binned spectra (third column) for holocellulose (a) and Klason lignin (b), respectively. Colours differentiate different sample types. The diagonal line represents values where measured and fitted values are identical. If points are below the line, the model overestimates contents.'}
p_pred_diff_train2
```

\clearpage

Figure \@ref(fig:res-p-y-yhat-minerals) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, the improved model using extracted peaks, the improved model using binned spectra trained with old magazine samples, and the improved model using extracted peaks trained with old magazine samples.

```{r res-p-y-yhat-minerals, echo=FALSE, out.width="100%", fig.height=3.5, fig.width=7.5, fig.cap='Predicted median holocellulose contents versus measured holocellulose contents for the Gaussian model with original model structure (first panel), the improved best models for both approaches (model including all peaks from the peak extraction procedure of \\cite{Hodgkins.2018}, model using binnes spectra with a bin width of 50 cm$^{-1}$), but fitted only to samples without mineral interferences (second and third panel), and the same models fitted also to samples with mineral interferences (last two panels). The diagonal line represent identical fitted and measured values. Colours differentiate samples with (blue) and without (red) mineral interferences.'}
p_y_yhat_minerals
```

<!---Figure \@ref(fig:res-p-veg-species) shows holocellulose and Klason lignin contents of vegetation samples from @Hosgkins.2018, predicted using the original model and the best models using binned spectra (for holocellulose: not calibrated with samples with mineral admixtures)

```{r res-p-veg-species, eval = FALSE, echo=FALSE, out.width="60%", fig.height=7, fig.width=4.5, fig.cap='Predicted vegetation median holocellulose and Klason lignin contents for the vegetation samples of \\cite{Hodgkins.2018}. Points are median predicted values, solid grey lines connect the same samples, shaded regions are 90\\% prediction intervals of individual samples.'}
p_veg_species
```
--->

\clearpage

Figure \@ref(fig:res-p-oh-height-difference) shows differences in the absorbance at 3400 cm$^{-1}$ in area normalized spectra between the training, peat, and vegetation data.

```{r, res-p-oh-height-difference, out.width="60%", fig.height=3.5, fig.width=5.5, fig.cap='Absorbance (intensity) at 3400 cm$^{-1}$ in area normalized spectra between the training, peat, and vegetation data.'}
p_oh_height_difference
```

### 2022-02-03 R/format_numbers.R

@@ -1,0 +1,18 @@
+#' Formats numbers for text output
+#'
+#' @param x A numeric value.
+#' @param capitalized A logical value indicating if the output should start with
+#' a capital letter (\code{TRUE}) or not (\code{FALSE}).
+#' @return A character value with the formatted value from \code{x}.
+#' @export
+format_numbers <- function(x, capitalized = FALSE) {
+  stopifnot(is.numeric(x) && length(x) == 1)
+  stopifnot(is.logical(capitalized) && length(capitalized) == 1)
+  if(x %% 1 == 0 && x <=10) {
+    x <- c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten")[[x]]
+    if(capitalized) {
+      x <- paste0(toupper(stringr::str_sub(x, 1, 1)), stringr::str_sub(x, 2, nchar(x)))
+    }
+  }
+  x
+}

### 2022-02-03 R/hkl_get_normalized_peak_heights.R

@@ -1,0 +1,36 @@
+#' Extracts normalized peak heights for a set of peaks from mid infrared spectra
+
+#'@param x An object of class \code{\link[ir:ir_new_ir]{ir}}.
+#'@param ... Further arguments passed to
+#'\code{pmird:::irp_content_klh_hodgkins_main}.
+#'@param scale A logical value indicating if peak heights should be
+#'z-transformed (\code{TRUE}) or not (\code{FALSE}), using
+#'\code{\link{scale}}.
+#'@return A data frame with a row for each spectrum in \code{x} and a column for
+#'each peak extracted by \code{pmird:::irp_content_klh_hodgkins_main}.
+#'@export
+hkl_get_normalized_peak_heights <- function(x, ..., do_scale = FALSE) {
+
+  stopifnot(inherits(x, "data.frame"))
+  stopifnot(length(do_scale) == 1 && is.logical(do_scale))
+
+  # extract peaks
+  res <-
+    irpeat:::irp_content_klh_hodgkins_main(data = x, ...)$norm.Acorr %>%
+    as.data.frame()
+
+  # define arom15arom16 peak
+  res <-
+    res %>%
+    dplyr::mutate(arom15arom16 = arom15 + arom16)
+
+  # scale
+  if(do_scale) {
+    res <-
+      res %>%
+      purrr::map_dfc(scale, center = TRUE, scale = TRUE)
+  }
+
+  res
+
+}

### 2022-02-03 R/hkl_hodgkins_ggplot.R

@@ -1,0 +1,520 @@
+#' Plots spectra with identified peaks as ggplot2 plots
+
+#' @details \code{hkl_hodgkins_ggplot} is a modified part of the code from
+#' \url{https://github.com/shodgkins/FTIRbaselines}. It extract a defined set of
+#' peaks from mid infrared spectra and creates a plot of these peaks as
+#' \code{\link[ggplot2:ggplot2]{ggplot2-object}}.
+
+#'@param x A data frame as returned by
+#'\code{irpeat:::irp_content_klh_hodgkins_prepare}.
+#'@return A list of \code{ggplot2-object} with a plot for each spectrum in
+#'\code{x}.
+#'
+#'@source The code is modified from
+#'\url{https://github.com/shodgkins/FTIRbaselines}.
+#'
+#'@export
+hkl_hodgkins_ggplot <- function (x) {
+
+    export <- NULL
+    verbose <- FALSE
+    make_plots <- TRUE
+    data <- x
+
+    irp_content_klh_hodgkins_minima <- function(region, default = NULL) {
+      success <- TRUE
+      offset.dy <- c(0, region$dy)[1:(length(region$dy))]
+      just.crossed <- which(sign(region$dy * offset.dy) ==
+                              -1 & region$d2y > 0)
+      crossing <- c(just.crossed, just.crossed - 1)
+      if (length(crossing) > 0) {
+        index.min <- which(region$y == min(region$y[crossing]) &
+                             region$x %in% region$x[crossing])
+        index.min <- which(region$y == min(region$y[(index.min -
+                                                       min(m, index.min - 1)):(index.min + min(m, nrow(region) -
+                                                                                                 index.min))]) & region$x %in% region$x[(index.min -
+                                                                                                                                           min(m, index.min - 1)):(index.min + min(m, nrow(region) -
+                                                                                                                                                                                     index.min))])
+        minimum <- region$x[index.min]
+      }
+      else {
+        if (max(region$d2y) > 0) {
+          minimum <- region$x[which(region$d2y == max(region$d2y))]
+        }
+        else {
+          minimum <- default
[…]

### 2022-02-03 R/hklmirs-package.R

@@ -1,0 +1,11 @@
+#' @keywords internal
+"_PACKAGE"
+
+#'@importFrom magrittr `%>%`
+NULL
+
+# The following block is used by usethis to automatically manage
+# roxygen namespace tags. Modify with care!
+## usethis namespace: start
+## usethis namespace: end
+NULL

### 2022-02-03 analysis/paper/001-paper-main.Rmd
Updated figure names.

@@
 To compute the models, the authors developed a procedure to extract peaks which are indicative for specific OM fractions, such as holocellulose and Klason lignin (Fig.
-\@ref(fig:hodgkins-peaks-sample)), from the MIRS.
+\@ref(fig:hodgkins-peaks-sample-fig01)), from the MIRS.
 The peaks are baseline corrected, their maximum height computed and divided by the sum of absorbance values in the spectra.
@@
 
-```{r hodgkins-peaks-sample, echo=FALSE, out.width="70%", fig.height=2.5, fig.width=6, fig.cap="Sample spectrum with the peaks and troughs, their heights, and their baselines (green) as detected by the script developed by \\cite{Hodgkins.2018}.
+```{r hodgkins-peaks-sample-fig01, echo=FALSE, out.width="70%", fig.height=2.5, fig.width=6, fig.cap="Sample spectrum with the peaks and troughs, their heights, and their baselines (green) as detected by the script developed by \\cite{Hodgkins.2018}.
 The plot was created with the data and code from \\cite{Hodgkins.2018} as implemented in ir and irpeat, respectively."}
@@
 We used all peak heights and trough heights returned by the peak identification algorithm of @Hodgkins.2018 (Fig.
-\@ref(fig:hodgkins-peaks-sample)).
+\@ref(fig:hodgkins-peaks-sample-fig01)).
 The advantage is that this makes the prediction model probably more robust against calibration transfer issues in comparison to the second approach (a hypothesis that remains to be tested).
@@
 Both models did not produce unrealistic predictions for the vegetation samples.
-Figure \@ref(fig:res-p-gaussian-beta-range) shows predicted medians and prediction intervals for both models across a potential range of the spectral predictor variables alongside the range covered by the peat and vegetation samples.
+Figure \@ref(fig:res-p-gaussian-beta-range-fig02) shows predicted medians and prediction intervals for both models across a potential range of the spectral predictor variables alongside the range covered by the peat and vegetation samples.
 The model with normal distribution produces negative lower prediction interval limits whenever the predicted median is lower than around `r d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose" & variable_limit == "fit_lwr") %>% dplyr::pull(fit) %>% magrittr::multiply_by(100) %>% round(0)` mass-% and upper prediction interval limits $>100$ mass-% whenever the predicted median is larger than around `r d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose" & variable_limit == "fit_upr") %>% dplyr::pull(fit) %>% magrittr::multiply_by(100) %>% round(0)` mass-%.
@@
 In addition, the beta regression model has smaller prediction uncertainties for extreme values (Fig.
-\@ref(fig:res-p-gaussian-beta-range)).
+\@ref(fig:res-p-gaussian-beta-range-fig02)).
 A consequence of this are larger predicted holocellulose contents with narrower prediction intervals for several (decomposed) peat samples under the beta regression model in comparison to the original normal model (Fig.
-\@ref(fig:res-p-gaussian-beta-range) and supplementary Fig.
+\@ref(fig:res-p-gaussian-beta-range-fig02) and supplementary Fig.
 3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->).
@@
 
-```{r res-p-gaussian-beta-range, warning=FALSE, echo=FALSE, out.width="70%", fig.height=3.5, fig.width=6.5, fig.cap="Predicted Holocellulose (a) and Klason lignin (b) contents using either the original Gaussian models or a beta regression model, versus predicitons of the original Gaussian model covering the entire interval [0, 100] mass-\\%.
+```{r res-p-gaussian-beta-range-fig02, warning=FALSE, echo=FALSE, out.width="70%", fig.height=3.5, fig.width=6.5, fig.cap="Predicted Holocellulose (a) and Klason lignin (b) contents using either the original Gaussian models or a beta regression model, versus predicitons of the original Gaussian model covering the entire interval [0, 100] mass-\\%.
 Coloured lines within shaded regions are median predictions.
@@
 To interpret the improved models using binned spectra, we plotted the median coefficients for the best models using binned spectra (Fig.
-\@ref(fig:res-p-coefficients)).
+\@ref(fig:res-p-coefficients-fig03)).
 From this plot, we identified the bins with the largest absolute median coefficients ($\ge0.2$, marked with points and labeled).
@@
 ~1250 cm$^{-1}$: aromatic in plane C-H bending) [@Stuart.2004] (Fig.
-\@ref(fig:res-p-coefficients), first panel).
+\@ref(fig:res-p-coefficients-fig03), first panel).
 This indicates that aromatics provide information on the holocellulose content in the training samples.
@@
 Instead bins with large absolute coefficients are located in the fingerprint region and probably related to aromatic in plane C-H bending (~1270 and ~1160 cm$^{-1}$) [@Stuart.2004] (Fig.
-\@ref(fig:res-p-coefficients), last panel).
+\@ref(fig:res-p-coefficients-fig03), last panel).
 We have no explanation why one coefficient is negative and the other positive.
@@
 
-```{r res-p-coefficients, echo=FALSE, out.width="90%", fig.height=3, fig.width=6.5, fig.cap='Coefficients of the best models using binned spectra for Holocellulose (first two panels, model trained without and with samples containing minerals, respectively) and Klason lignin (last panel), plotted against the average wavenumber of the bins.
+```{r res-p-coefficients-fig03, echo=FALSE, out.width="90%", fig.height=3, fig.width=6.5, fig.cap='Coefficients of the best models using binned spectra for Holocellulose (first two panels, model trained without and with samples containing minerals, respectively) and Klason lignin (last panel), plotted against the average wavenumber of the bins.
 Points are median coefficient estimates, error bars are 90\\% posterior intervals.
@@
 The training data set does not cover the range of the spectral variables relevant for predicting peat and peat forming vegetation holocellulose contents (Fig.
-\@ref(fig:res-p-prediction-domain) A and B).
+\@ref(fig:res-p-prediction-domain-fig04) (a) and (b)).
 For the original models, the training data `carb` and `arom15arom16` peak heights cover the range of the vegetation samples.
@@
 The same is true for the improved models using binned spectra (Fig.
-\@ref(fig:res-p-prediction-domain) C and D).
+\@ref(fig:res-p-prediction-domain-fig04) (c) and (d)).
 Even though it is not possible to establish a direct relation between extrapolation for individual spectral variables and predicted contents, it is visible that for several bins, peat samples --- and to a lesser extent --- vegetation samples have mostly larger standardized predictor values than covered by the training data set.
@@
 Samples with small `carb` peak, or for the binned spectra, large absorbance at ~`r m_pars %>% dplyr::filter(abs(median) >= 0.02 & variable_y == "Holocellulose (with minerals)") %>% dplyr::pull(wn) %>% sort() %>% knitr::combine_words()` cm$^{-1}$, are outside the prediction domain for holocellulose (Fig.
-\@ref(fig:res-p-prediction-domain)).
+\@ref(fig:res-p-prediction-domain-fig04)).
 Since the `carb` peak is small and absorbance at the selected wavenumbers is typically large in decomposed samples [@Cocozza.2003; @Tfaily.2014; @Hodgkins.2018] (and the respective coefficients are negative), this indicates that for holocellulose, extrapolation occurs mainly for more decomposed peat.
@@
 
-```{r res-p-prediction-domain, echo=FALSE, out.width="40%", fig.height=7, fig.width=4, fig.cap='Prediction domain for the original models (a, b), and the improved models using binned spectra (c, d).
+```{r res-p-prediction-domain-fig04, echo=FALSE, out.width="40%", fig.height=7, fig.width=4, fig.cap='Prediction domain for the original models (a, b), and the improved models using binned spectra (c, d).
 The prediction domain is the range of values covered by the predictor variables.
@@
 Dashed lines indicate the most important variables for the holocellulose (c) and Klason lignin model (d), respectively (compare with Fig.
-\\ref{fig:res-p-coefficients})'}
+\\ref{fig:res-p-coefficients-fig03})'}
 p_prediction_domain
@@
 There were considerable differences in predictions of the original and improved models using binned spectra for both Klason lignin and holocellulose for the peat samples (Fig.
-\@ref(fig:res-p-pred-diff-all1) and supplementary Fig.
+\@ref(fig:res-p-pred-diff-all1-fig05) and supplementary Fig.
 5 <!---\@ref(fig:res-p-gaussian-beta-depth-profile-best)--->).
 For holocellulose, the original model tends to predict larger contents for peat, especially for samples with larger holocellulose contents (Fig.
-\@ref(fig:res-p-pred-diff-all1)).
+\@ref(fig:res-p-pred-diff-all1-fig05)).
 Similar, albeit less pronounced patterns are visible for the vegetation and training data.
@@
 This happens particularly in a region where the original model predicts Klason lignin contents between 25 and 35 mass-% (Fig.
-\@ref(fig:res-p-pred-diff-all1)).
+\@ref(fig:res-p-pred-diff-all1-fig05)).
 This is surprising for two reasons: First, for the training data, both models make quite similar predictions (Fig.
-\@ref(fig:res-p-pred-diff-all1)).
+\@ref(fig:res-p-pred-diff-all1-fig05)).
 The peat samples have `aro15arom16` values similar to the training data (Fig.
-\@ref(fig:res-p-prediction-domain)), but nevertheless both models produce contrasting predictions.
+\@ref(fig:res-p-prediction-domain-fig04)), but nevertheless both models produce contrasting predictions.
 The same is true for holocellulose for a relatively large fraction of peat samples (Fig.
-\@ref(fig:res-p-prediction-domain)).
+\@ref(fig:res-p-prediction-domain-fig04)).
 Second, both models have comparatively small bias within the range of the training data relevant to peat samples (supplementary Fig.
@@
 
-```{r res-p-pred-diff-all1, echo=FALSE, out.width="80%", fig.height=6, fig.width=6, fig.cap='Predicted values of the original model versus predicted values of the original model (first column), and the best improved models using extracted peaks (second column) and binned spectra (last column), respectively, for holocellulose (a) and Klason lignin (b), respectively.
+```{r res-p-pred-diff-all1-fig05, echo=FALSE, out.width="80%", fig.height=6, fig.width=6, fig.cap='Predicted values of the original model versus predicted values of the original model (first column), and the best improved models using extracted peaks (second column) and binned spectra (last column), respectively, for holocellulose (a) and Klason lignin (b), respectively.
 Colours differentiate the training, peat, and vegetation data from \\cite{Hodgkins.2018}.'}
@@
 Samples of these classes differ in the relative contribution of the `arom15` and `arom16` peaks to `arom15arom16` (the sum of their heights) (Fig.
-\@ref(fig:res-p-pred-diff-train1-p2)).
+\@ref(fig:res-p-pred-diff-train1-p2-fig06)).
 The first class (class 1) has high contributions of the `arom15` peak ($>30$%) and typically smaller `arom15arom16` values; the second class (class 2) a smaller contribution of the `arom15` peak and larger `arom15arom16` values.
 Samples of both classes can have large or small actual Klason lignin contents (Fig.
-\@ref(fig:res-p-pred-diff-train1-p2)).
+\@ref(fig:res-p-pred-diff-train1-p2-fig06)).
 What we observed is that the model makes biased predictions for both classes: For small `arom15arom16` values, it underestimates contents, but for larger values, it increasingly overestimates Klason lignin contents (supplementary Fig.
@@
 For the improved model, this bias is much smaller (Fig.
-\@ref(fig:res-p-pred-diff-train1-p2) and supplementary Fig.
+\@ref(fig:res-p-pred-diff-train1-p2-fig06) and supplementary Fig.
 9<!---\@ref(fig:res-p-pred-diff-kl-residuals)--->).
@@
 If one extrapolates the biased predictions of the original model for the training data classes, this means larger under- and overestimation for the peat samples than for the training data using the original model (Fig.
-\@ref(fig:res-p-pred-diff-train1-p2)).
+\@ref(fig:res-p-pred-diff-train1-p2-fig06)).
 This is supported by measured peat Klason lignin contents: Peat from temperate peatlands was reported to have Klason lignin contents as high as ~70 mass-% and ~57 mass-% on average [@Hayes.2015].
@@
 This is why the original model predicts high Klason lignin contents for samples of class 2, even though some of these actually have low Klason lignin contents (Fig.
-\@ref(fig:res-p-pred-diff-train1-p2)).
+\@ref(fig:res-p-pred-diff-train1-p2-fig06)).
 The model balances this, but in doing so it introduces the observed bias.
@@
 The improved model using binned spectra gives more weight to variables related to aromatic skeletal structures and not in a region where proteins cause large absorbances [@Stuart.2004] (Fig.
-\@ref(fig:res-p-coefficients)).
+\@ref(fig:res-p-coefficients-fig03)).
 We suggest that these variables are better predictors for Klason lignin because they interfere less with other molecular structures, such as proteins [@Stuart.2004].
@@
 
-```{r res-p-pred-diff-train1-p2, echo=FALSE, out.width="90%", fig.height=4, fig.width=7, fig.cap='Measured training data Klason lignin contents versus fitted values [mass-\\%] of the original model (first column), the improved model using extracted peaks (second column), and the improved model using binned spectra (third column).
+```{r res-p-pred-diff-train1-p2-fig06, echo=FALSE, out.width="90%", fig.height=4, fig.width=7, fig.cap='Measured training data Klason lignin contents versus fitted values [mass-\\%] of the original model (first column), the improved model using extracted peaks (second column), and the improved model using binned spectra (third column).
 Points are scaled according to the relative contribution of the \\texttt{arom15} peak to \\texttt{arom15arom16}.
@@
 According to Fig.
-\@ref(fig:res-p-coefficients) (middle panel), the "mineral-rich model" down-weights bins near the `carb` peak, in contrast to the improved model not trained on old magazine samples, and instead has a larger coefficient for an additional peak related to aromatic C=C stretching and amide N-H bending and C-N stretching (~1550 cm$^{-1}$) [@Stuart.2004] which corresponds to the `arom15` peak.
+\@ref(fig:res-p-coefficients-fig03) (middle panel), the "mineral-rich model" down-weights bins near the `carb` peak, in contrast to the improved model not trained on old magazine samples, and instead has a larger coefficient for an additional peak related to aromatic C=C stretching and amide N-H bending and C-N stretching (~1550 cm$^{-1}$) [@Stuart.2004] which corresponds to the `arom15` peak.
 This provides further evidence that it is possible to infer the holocellulose content via aromatics.
@@
 Here, it was important that one modification of the original model often addressed multiple limitations at once due to interdependences between the limitations: Improving the models' predictive performance did not only address underfitting, but reduced also the bias in the models (Fig.
-\@ref(fig:res-p-pred-diff-train1-p2)).
+\@ref(fig:res-p-pred-diff-train1-p2-fig06)).
 Analyzing the prediction domain did not only show that the models extrapolate, but we could use this knowledge to analyze the practical impact of the bias.

### 2022-03-03 analysis/paper/001-paper-main.Rmd
Update manuscript 1, update formatting fo references.

@@
     To understand global soil organic matter (SOM) chemistry and its dynamics, we need tools to efficiently quantify SOM properties, for example prediction models using mid infrared spectra.
-All the advantages of such models rely on their critical validation.
-Recently, \cite{Hodgkins.2018} developed models to predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major compounds of organic matter in general.
+The advantages of such models rely on their validity and accuracy.
+Recently, \cite{Hodgkins.2018} developed models to predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major fractions of organic matter in general.
 The models can be used to understand large-scale SOM gradients and have been used in various studies.
@@
     The major limitation we identified is that the original training data are not representative for a range of peat samples which causes both biased estimates and extrapolation uncertainty under the original models.
-In addition, the original models can in parctice produce unrealistic predictions (negative values or values $>100$ mass-%).
+In addition, the original models can in practice produce unrealistic predictions (negative values or values $>100$ mass-%).
 Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid unrealistic predictions.
@@
   We like to thank @Hodgkins.2018 and @DelaCruz.2016 for provision of data and code to reproduce and modify their analyses, and @Allaire.2020 for provision of an Rmarkdown template.
-We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions.
+We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions which helped to improve a first version of this manuscript.
+We acknowledge support from the Open Access Publication Fund of the University of Münster.
 #appendix: |
@@
 library(loo)
-library(projpred)
 library(bayesplot)
@@
 library(scales)
-library(bib2df)
 library(kableExtra)
@@
 Moreover, both raw data and model code are publicly available [@Hodgkins.2018; @Teickner.2020a].
-For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020], most recently in @Verbeke.2022, and for future developments, e.g.
+For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020; @Cong.2022], most recently in @Verbeke.2022, and for future developments, e.g.
 by including additional data.
@@
 To compare models, one typically computes the posterior distribution of the difference in PSIS-LOO ELPD between the model with the largest PSIS-LOO ELPD and the other models, such that one can evaluate the probability of a certain predictive advantage of the best model in comparison to the other models [@Vehtari.2017; @Vehtari.2019].
-To further validate and interpret the models using binned spectra, we visually identified the most important bins (largest absolute median coefficients $\ge0.2$) and linked these to molecular structures and causal relations that most probably result in the correlation with holocellulose and Klason lignin, respectively.
+To further validate and interpret the models using binned spectra, we visually identified the most important bins in the models (largest absolute median coefficients $\ge0.2$) with bin a width of 20 cm$^{-1}$ (which are among the models with the best predictive accuracy; see below) and linked these to molecular structures and causal relations that most probably result in the correlation with holocellulose and Klason lignin, respectively.
+Based on the model validation, we defined the following set of models used in the subsequent analyses: "Best all peaks" denotes the models for holocellulose and Klason lignin with best average predictive accuracy of approach 1 described above.
+"Best binned spectra" denotes the models for holocellulose and Klason lignin with nearly the best average predictive accuracy of approach 2 described above.
+With "nearly the best predictive accuracy" we mean here that we used the models using binned spectra with a bin width of 20 cm$^{-1}$.
+This was the model with best average predictive accuracy for Klason lignin, but not holocellulose (Sect.
+3.2).
+However, the model for holocellulose had a predictive performance very similar to the on average better model using a bin width of 10 cm$^{-1}$ (Sect.
+3.2), and to keep the following analyses compact and easier to follow, we decided to use a bin width of 20 cm$^{-1}$ in all cases.
+These models are the models used during the subsequent analyses.
 
@@
 Finally, by identifying the most important variables in the improved models, we could qualitatively summarize how large the risk of this extrapolation is.
-We did not additionally investigate the prediction domain for the models using all extracted peaks and troughs since these models had no predictive advantage in comparison to the models using binned spectra.
+Since models with a bin width of 20 cm$^{-1}$ were among the best models, we performed this analysis only for these models ("Best binned spectra").
+We did not additionally investigate the prediction domain for the models using all extracted peaks and troughs ("Best all peaks") since these models had no predictive advantage in comparison to the models using binned spectra (Sect.
+3.2).
 
 ## Analyzing differences in predictions between the original models and the improved models 
-We analyzed how predictions of the improved models (best model for each of the two approaches) differ from the original models.
-We found considerable differences, even within the spectral range of the training data and therefore analyzed what factors probably cause these differences.
+We analyzed how predictions of the improved models ("Best all peaks" and "Best binned spectra" defined in Sect.
+2.2) differ from the original models.
+We found considerable differences, even within the spectral range of the training data, and therefore analyzed what factors probably cause these differences.
 In a first step, we were interested if the original or improved models are biased.
@@
 We therefore created such plots for the original model and the improved model and compared indication of bias between both.
-In a second step, we compared how predictions of both models differ in practice.
-To this end, we compared predictions of both for the peat and vegetation data from @Hodgkins.2018, similarly to how we compared the Gaussian and beta regression models.
+In a second step, we compared how predictions of the improved models differ in practice from the original models.
+To this end, we compared the models' predictions for the peat and vegetation data from @Hodgkins.2018, similarly to how we compared the Gaussian and beta regression models.
 Since we found considerable differences, we conducted a more targeted exploratory analysis to identify factors probably causing these differences.
@@
 By including more variables into the model than the `carb` peak (as described in the previous section), we tried to compute models that can also describe the holocellulose content in the training data samples with high clay content.
-For this, we fitted the best models for holocellulose from each of the two approaches described in the previous section again (binned spectra: `r loo_comparison_hol$bin_width[[1]]` cm$^{-1}$ bin width; extracted peaks: priors with standard deviation =`r loo_comparison_hol$prior_scale[loo_comparison_hol$variables == "peaks"][[1]]`), but this time including the four old magazine samples previously left out.
-We interpreted the coefficients of these binned models similarly to the versions not fitted with mineral rich samples.
+For this, we fitted the best models for holocellulose (the models in "Best all peaks" and "Best binned spectra" defined in Sect.
+2.2) from each of the two approaches described in the previous section again, but this time including the four old magazine samples previously left out.
+We interpreted the coefficients of these binned models similarly to the versions not fitted with mineral rich samples (see Sect.
+2.2).
 
@@
 
-Our analysis shows that both strategies to include more variables result on average in more accurate predictions.
-If uncertainties for the estimated predictive accuracy are considered, the advantage is clear only for the models for holocellulose using binned spectra (Table \@ref(tab:res-tab-loo-comparison-res)), but not for Klason lignin.
-Consequently, using the data available, we could improve the model for holocellulose, but our analysis remains inconclusive with respect to Klason lignin.
+Our analysis shows that both strategies to include more variables result in on average more accurate predictions (Table \@ref(tab:res-tab-loo-comparison-res)).
+This can be concluded based on the average estimated ELPD values (larger values indicate a better estimated predictive accuracy).
+Consequently, using the data available, we could improve both the model for holocellulose and Klason lignin.
 Is one of the strategies to include more variables advantageous?
-The on average overall best models, according to their PSIS-LOO ELPD, were those using binned spectra with bin widths of `r loo_comparison_hol$bin_width[[1]]` and `r loo_comparison_kl$bin_width[[1]]` for holocellulose and Klason lignin, respectively (Table \@ref(tab:res-tab-loo-comparison-res)).
-For Klason lignin, comparatively large prediction uncertainties result in the practical equivalence of both strategies.
-For holocellulose, using binned spectra (except bin width = 20) is more clearly better than using the extracted peaks.
-Summarized, this indicates that using binned spectra results in an improved predictive accuracy over the second approach to use all extracted peaks for holocellulose, whereas our validation remains inconclusive with respect to Klason lignin.
+The models with the largest predictive accuracy are models with small to moderate bin width (Table \@ref(tab:res-tab-loo-comparison-res)).
+These are better than the respective models using all extracted peaks: in addition to the ELPD, this can be derived from the standard errors of the ELPD differences relative to the best models ($2 \cdot \Delta \text{SE}$ is approximately the 95% confidence interval).
+In addition, the remaining binned models generally have a better average predictive accuracy than the models using all extracted peaks, except for the model with bin width = 100 cm$^{-1}$ for holocellulose.
+Summarized, this indicates that --- if a sufficiently high bin resolution (sufficiently small bin width) is chosen --- using binned spectra results in an improved predictive accuracy over the second approach to use all extracted peaks for both holocellulose and Klason lignin.
 To interpret the improved models using binned spectra, we plotted the median coefficients for the best models using binned spectra (Fig.
@@
 The reason for the (on average) improved performance is that the improved models can use information contained in the complete spectra which is not contained in the manually selected peaks for the original models:
-Similarly to the original model, a variable near the `carb` peak is important to predict the holocellulose content (bin at ~950 cm$^{-1}$) since it is related to cellulose [@Cocozza.2003; @Stuart.2004; @Artz.2008], but two peaks related to aromatic structures are also important (~1200 cm$^{-1}$: C-O stretching in phenols and aromatic ethers [@Stuart.2004; @Kubo.2005].
-~1250 cm$^{-1}$: aromatic in plane C-H bending) [@Stuart.2004] (Fig.
+Similarly to the original model, a variable near the `carb` peak is important to predict the holocellulose content (bin at ~970 cm$^{-1}$) since it is related to cellulose [@Cocozza.2003; @Stuart.2004; @Artz.2008], but one peak related to aromatic structures which is not extracted with the approach from @Hodgkins.2018 is also important (~1250 cm$^{-1}$: aromatic in plane C-H bending) [@Stuart.2004] (Fig.
 \@ref(fig:res-p-coefficients-fig03), first panel).
-This indicates that aromatics provide information on the holocellulose content in the training samples.
-A plausible explanation is that holocellulose and Klason lignin together are the major mass fractions in many OM types and both are controlled by the same processes, but often in different directions (decomposition, resource allocation trade-off during plant growth, selective removal during processing) [@Biester.2014; @Chen.2016] (supporting Fig.
+This indicates that aromatics provide information on the holocellulose content in the training samples and that extracting only selected peaks may omit useful information.
+A plausible explanation for the importance of this variable is that holocellulose and Klason lignin together are the major mass fractions in many OM types and both are controlled by the same processes, but often in different directions (decomposition, resource allocation trade-off during plant growth, selective removal during processing) [@Biester.2014; @Chen.2016] (supporting Fig.
 4<!---\@ref(fig:res-sup-kl-hol)--->).
 For Klason lignin, the improved model does not use bins near to the `arom15` and `arom16` peaks, in contrast to the original model.
-Instead bins with large absolute coefficients are located in the fingerprint region and probably related to aromatic in plane C-H bending (~1270 and ~1160 cm$^{-1}$) [@Stuart.2004] (Fig.
+Instead bins with large absolute coefficients are located in the fingerprint region and probably related to aromatic in plane C-H bending (~1150 and ~1270 cm$^{-1}$) [@Stuart.2004] (Fig.
 \@ref(fig:res-p-coefficients-fig03), last panel).
 We have no explanation why one coefficient is negative and the other positive.
-In section "What causes differences in predictions?", we provide a mechanistic interpretation for why the selected variables are important.
+In Sect.
+3.5<!--"What causes differences in predictions?"-->, we provide a mechanistic interpretation for why the selected variables are important.
+These selected bins are not represented in the extracted peaks.
+Consequently, models using all extracted peaks cannot use this information and tend to underfit.
+Similarly, binning with a broad bin width may result in a too coarse spectral resolution, as indicated by the relatively weak predictive accuracy of the model for holocellulose using binned spectra with a bin width of 100 cm$^{-1}$ (Table \@ref(tab:res-tab-loo-comparison-res)).
+
+```{r res-p-coefficients-fig03, echo=FALSE, out.width="90%", fig.height=3, fig.width=6.5, fig.cap='Coefficients of the best models using binned spectra for Holocellulose (first two panels, model trained without and with samples containing minerals, respectively) and Klason lignin (last panel), plotted against the average wavenumber of the bins.
+Points are median coefficient estimates, error bars are 90\\% posterior intervals.
+Points with an absolute median coefficient $\\ge 0.2$ are labeled with the respective average wavenumber value.
+The grey shaded region is a reference spectrum.'}
+p_coefficients
+```
 
@@
 
-\clearpage
-
-```{r res-p-coefficients-fig03, echo=FALSE, out.width="90%", fig.height=3, fig.width=6.5, fig.cap='Coefficients of the best models using binned spectra for Holocellulose (first two panels, model trained without and with samples containing minerals, respectively) and Klason lignin (last panel), plotted against the average wavenumber of the bins.
-Points are median coefficient estimates, error bars are 90\\% posterior intervals.
-Points with an absolute median coefficient $\\ge 0.2$ are labeled with the respective average wavenumber value.
-The grey shaded region is a reference spectrum.'}
-p_coefficients
-```
-
 ## Is the prediction domain of the training data representative for peat samples?
@@
 The same is true for the improved models using binned spectra (Fig.
-\@ref(fig:res-p-prediction-domain-fig04) (c) and (d)).
+\@ref(fig:res-p-prediction-domain-fig04) (c)).
 Even though it is not possible to establish a direct relation between extrapolation for individual spectral variables and predicted contents, it is visible that for several bins, peat samples --- and to a lesser extent --- vegetation samples have mostly larger standardized predictor values than covered by the training data set.
-This is also true for bins with the largest absolute median coefficients (including the model for holocellulose fitted with mineral-rich samples; see section "Can holocellulose contents also be predicted for samples with mineral interferences?" below).
+This is also true for bins with the largest absolute median coefficients (including the model for holocellulose fitted with mineral-rich samples; see Sect.
+3.6 <!--"Can holocellulose contents also be predicted for samples with mineral interferences?"--> below).
 Under what conditions does extrapolation occur?
@@
 This is probably also true for non-peat SOM.
-Therefore the models' prediction can represent extrapolations in practice.
+Therefore the models' predictions can represent extrapolations in practice; the training data is not in general representative for peat and peat-forming vegetation.
 
-```{r res-p-prediction-domain-fig04, echo=FALSE, out.width="40%", fig.height=7, fig.width=4, fig.cap='Prediction domain for the original models (a, b), and the improved models using binned spectra (c, d).
+```{r res-p-prediction-domain-fig04, echo=FALSE, out.width="100%", fig.height=3, fig.width=9, fig.cap='Prediction domain for the original models (a, b), and the improved models using binned spectra with a bind width of 20 cm$^{-1}$ (c).
 The prediction domain is the range of values covered by the predictor variables.
@@
 Counts are number of samples in the respective datasets.
-The improved models include bins across the complete spectra (c, d).
-Therefore, the range in standardized predictor variable values (y axis) covered by the training data for each bin is shown as orange shaded area for the dataset with a bin width of 50 (c) and 10 cm$^{-1}$ (d), respectively, as used by the best improved models for holocellulose and Klason lignin.
+The improved models include bins across the complete spectra (c).
+Therefore, the range in standardized predictor variable values (y axis) covered by the training data for each bin is shown as orange shaded area for the dataset with a bin width of 20 cm$^{-1}$, as used by the (nearly) best improved models for holocellulose and Klason lignin.
 Extrapolation occurs if other samples (blue shaded regions) exceed these areas.
-Dashed lines indicate the most important variables for the holocellulose (c) and Klason lignin model (d), respectively (compare with Fig.
+Dashed lines indicate the most important variables for the holocellulose and Klason lignin model (compare with Fig.
 \\ref{fig:res-p-coefficients-fig03})'}
@@
 
-In comparison to the training data, spectra for peat typically have larger absorption values at around 1200 cm$^{-1}$ and between (~1200 to 1500 cm$^{-1}$), but a smaller absorption in the region of the "OH-peak" (~3400 cm$^{-1}$) (supplementary Fig.
-7 <!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).
+In comparison to the training data, spectra for peat typically have larger absorption values at around 1250 cm$^{-1}$ and between (~1250 to 1500 cm$^{-1}$), but a smaller absorption in the region of the "OH-peak" (~3400 cm$^{-1}$) (supplementary Fig.
+7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).
 These spectral differences are directly related to the differences in predicted values of the original and improved model (supplementary Fig.
-8 <!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->): The smaller the OH-peak is, the larger are predicted values of the original model in comparison to the improved model using binned spectra, especially for larger `carb` peak heights.
-Similarly, the larger the trough at 1200 cm$^{-1}$ or the absorbance at 1550 cm$^{-1}$ is, the larger are predictions of the original model in comparison to the improved model using binned spectra.
+8<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->): The smaller the OH-peak is, the larger are predicted values of the original model in comparison to the improved model using binned spectra, especially for larger `carb` peak heights.
+Similarly, the larger the trough at 1250 cm$^{-1}$ or the absorbance at 1590 cm$^{-1}$ is, the larger are predictions of the original model in comparison to the improved model using binned spectra.
 This indicates that both the delineation of the `carb` peak and predictions using binned spectra are potentially sensitive to these differences and that this causes the observed differences in the predictions even in the range of the training data where both models yield similar predictions for the training samples.
@@
 9<!---\@ref(fig:res-p-pred-diff-kl-residuals)--->).
-We suggest that it is this bias which causes the difference in predictions of the original and improved models.
+We suggest that it is this bias which causes most of the difference in predictions of the original and improved models.
 Conditional on the relative contribution of the `arom15` peak to `arom15arom16`, peat samples have smaller and larger `arom15arom16` values than the training data (supplementary Fig.
@@
 Larger `arom16` peak heights may be indicative for Klason lignin, but can also be due to high protein contents.
-This is what distinguishes samples of class 1 and 2 (supplementary Fig.
+This is what distinguishes samples of class 1 and 2 in the training data (supplementary Fig.
 11<!---\@ref(fig:res-p-pred-diff-train2)--->): Samples of class 1 are wood samples.
@@
 According to Fig.
-\@ref(fig:res-p-coefficients-fig03) (middle panel), the "mineral-rich model" down-weights bins near the `carb` peak, in contrast to the improved model not trained on old magazine samples, and instead has a larger coefficient for an additional peak related to aromatic C=C stretching and amide N-H bending and C-N stretching (~1550 cm$^{-1}$) [@Stuart.2004] which corresponds to the `arom15` peak.
+\@ref(fig:res-p-coefficients-fig03) (middle panel), the "mineral-rich model" down-weights bins near the `carb` peak, in contrast to the improved model not trained on old magazine samples, and instead has a larger coefficient for an additional peak related to aromatic C=C stretching and amide N-H bending and C-N stretching (~1590 cm$^{-1}$) [@Stuart.2004] which corresponds to the `arom15` peak.
 This provides further evidence that it is possible to infer the holocellulose content via aromatics.
@@
 Therefore, to assess if training data (the prediction domain) is representative, whole spectra have to be compared.
-Third, our analysis also shows that re-evaluating existing models with their original training data can be an effective way to improve the models.
+Third, it is helpful to identify potential causal mechanisms which may affect differently the MIRS the model will be applied on than the MIRS it was trained on.
+As shown here (Sect.
+3.5), differences in the degree of decomposition or the relative contribution of proteins and aromatic skeletal structures cause differences in MIRS which can result in biased predictions.
+Consequently, providing a causal explanation what causes the correlation between  a target variable and specific MIRS variables is a useful tool to assess if a model may be applicable for new samples.
+Fourth, our analysis also shows that re-evaluating existing models with their original training data can be an effective way to improve the models.
 Here, it was important that one modification of the original model often addressed multiple limitations at once due to interdependences between the limitations: Improving the models' predictive performance did not only address underfitting, but reduced also the bias in the models (Fig.
@@
 ```
-
-```{r references-preparation, echo=FALSE, eval=FALSE}
-# get bibkeys
-bibkeys <- bib2df::bib2df("references.bib")
-
-# get texts to search for bibkey
-si_text <- 
-  readLines("008-paper-supplementary.Rmd") %>%
-  paste(collapse = " ")
-
-main_text <- 
-  readLines("001-paper-main.Rmd") %>%
-  paste(collapse = " ")
-
-bibkeys <-
-  bibkeys %>%
-  dplyr::mutate(
-    refs_only_si = 
-      purrr::map_lgl(BIBTEXKEY, function(x) {
-        stringr::str_detect(si_text, x) & !stringr::str_detect(main_text, x)
-      })
-  )
-
-\nocite{`r paste(bibkeys$BIBTEXKEY[!bibkeys$refs_only_si], collapse = ", ")`}
-\bibliography{references}
-
-\nocitesi{`r paste(bibkeys$BIBTEXKEY[bibkeys$refs_only_si], collapse = ", ")`}
-\bibliographystylesi{apacite}
-\bibliographysi{references}
-```

### 2022-03-03 analysis/paper/004-paper-m-reduce-underfitting.Rmd
Update manuscript 1, update formatting fo references.

@@
 # compute loos
-m1.7_loo <- loo(m1.7, reloo = TRUE, cores = chains)
+m1.7_loo <- loo(m1.7, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ```{r m-reduce-underfitting-5-2}
-m1.8_loo <- loo(m1.8, reloo = TRUE, cores = chains)
+m1.8_loo <- loo(m1.8, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ```{r m-reduce-underfitting-5-3}
-m1.9_loo <- loo(m1.9, reloo = TRUE, cores = chains)
+m1.9_loo <- loo(m1.9, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ```{r m-reduce-underfitting-5-4}
-m1.10_loo <- loo(m1.10, reloo = TRUE, cores = chains)
+m1.10_loo <- loo(m1.10, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ```{r m-reduce-underfitting-5-5}
-m1.11_loo <- loo(m1.11, reloo = TRUE, cores = chains)
+m1.11_loo <- loo(m1.11, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 # compute loos
-m2.7_loo <- loo(m2.7, reloo = TRUE, cores = chains)
+m2.7_loo <- loo(m2.7, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ```{r m-reduce-underfitting-6-2}
-m2.8_loo <- loo(m2.8, reloo = TRUE, cores = chains)
+m2.8_loo <- loo(m2.8, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ```{r m-reduce-underfitting-6-3}
-m2.9_loo <- loo(m2.9, reloo = TRUE, cores = chains)
+m2.9_loo <- loo(m2.9, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ```{r m-reduce-underfitting-6-4}
-m2.10_loo <- loo(m2.10, reloo = TRUE, cores = chains)
+m2.10_loo <- loo(m2.10, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ```{r m-reduce-underfitting-6-5}
-m2.11_loo <- loo(m2.11, reloo = TRUE, cores = chains)
+m2.11_loo <- loo(m2.11, reloo = TRUE, cores = chains, recompile = FALSE)
 ```
@@
 ## create table for printing
+loo_comparison_code <-
+  tibble::tibble(
+    model_code = paste0("m1.", c(2, 3, 5:11)),
+    bin_width = c(rep("-", 4), 10, 20, 30, 50, 100),
+    family = c("Gaussian", rep("Beta", 8)),
+    original_model = c(rep(TRUE, 2), rep(FALSE, 7)),
+    prior_scale = c(rep(2.5, 3), 0.5, rep("-", 5)),
+    variables = ifelse(bin_width == "-", "peaks", "bins")
+  )
 
@@
   dplyr::select(1:3) %>%
-  dplyr::mutate(dplyr::across(dplyr::everything(), round, 1)) %>%
-  dplyr::mutate(
-    y_variable = "Holocellulose",
-    model_code = rownames(.),
-    bin_width = c(50, 100, 30, 10, rep("-", 4), 20),
-    family = c(rep("Beta", 7), "Gaussian", "Beta"),
-    original_model = c(rep(FALSE, 6), TRUE, TRUE, FALSE),
-    prior_scale = c(rep("-", 4), 0.5, 2.5, 2.5, 2.5, "-"),
-    variables = ifelse(bin_width == "-", "peaks", "bins")
-  )
+  dplyr::mutate(model_code = rownames(.)) %>%
+  dplyr::mutate(dplyr::across(1:3, round, 1)) %>%
+  dplyr::left_join(loo_comparison_code, by = "model_code")
 
@@
   dplyr::select(1:3) %>%
-  dplyr::mutate(dplyr::across(dplyr::everything(), round, 1)) %>%
-  dplyr::mutate(
-    y_variable = "Klason lignin",
-    model_code = rownames(.),
-    bin_width = c(10, rep("-", 4), 30, 20, 100, 50),
-    family = c(rep("Beta", 3), "Gaussian", rep("Beta", 5)),
-    original_model = c(rep(FALSE, 3), TRUE, TRUE, rep(FALSE, 4)),
-    prior_scale = c("-", 0.5, 2.5, 2.5, 2.5, rep("-", 4)),
-    variables = ifelse(bin_width == "-", "peaks", "bins")
-  )
+  dplyr::mutate(model_code = rownames(.)) %>%
+  dplyr::mutate(dplyr::across(1:3, round, 1)) %>%
+  dplyr::left_join(
+    loo_comparison_code %>%
+      dplyr::mutate(model_code =
+                      model_code %>% stringr::str_replace(pattern = "^m1", replacement = "m2")), 
+    by = "model_code")
 
@@
 
-# bin width: 50
-d_peat_bin50_scale <-
-  d_underfitting_flat[[4]] %>%
+# bin width: 20
+d_peat_bin20_scale <-
+  d_underfitting_flat[[2]] %>%
   dplyr::select(-1) %>%
@@
 
-d_peat_bin50_scale <-
-  tibble::tibble(
-    center = attr(d_peat_bin50_scale, "scaled:center"),
-    scale = attr(d_peat_bin50_scale, "scaled:scale")
-  )
-
-d_peat_bin50_flat <-
+d_peat_bin20_scale <-
+  tibble::tibble(
+    center = attr(d_peat_bin20_scale, "scaled:center"),
+    scale = attr(d_peat_bin20_scale, "scaled:scale")
+  )
+
+d_peat_bin20_flat <-
   d_peat %>%
-  ir::ir_bin(width = 50) %>%
+  ir::ir_bin(width = 20) %>%
   ir::ir_flatten()
 
-d_peat_bin50 <- 
-  d_peat_bin50_flat %>%
+d_peat_bin20 <- 
+  d_peat_bin20_flat %>%
   dplyr::select(-1) %>%
@@
   scale(
-    center = d_peat_bin50_scale$center, 
-    scale = d_peat_bin50_scale$scale
+    center = d_peat_bin20_scale$center, 
+    scale = d_peat_bin20_scale$scale
   ) %>%
@@
   dplyr::bind_rows(
-    m1.10 %>%
-      rstanarm::posterior_predict(newdata = d_peat_bin50) %>%
+    m1.8 %>%
+      rstanarm::posterior_predict(newdata = d_peat_bin20) %>%
       as.data.frame() %>%
@@
       ),
-    m2.7 %>%
-      rstanarm::posterior_predict(newdata = d_peat_bin10) %>%
+    m2.8 %>%
+      rstanarm::posterior_predict(newdata = d_peat_bin20) %>%
       as.data.frame() %>%
@@
       yhat = 
-        list(m1.2, m1.5, m1.10) %>%
+        list(m1.2, m1.5, m1.8) %>%
         purrr::map(function(x) {
@@
       yhat = 
-        list(m2.2, m2.6, m2.7) %>%
+        list(m2.2, m2.6, m2.8) %>%
         purrr::map(function(x) {
@@
 # bin width: 50
-d_veg_bin50_flat <-
+d_veg_bin20_flat <-
   d_veg %>%
-  ir::ir_bin(width = 50) %>%
+  ir::ir_bin(width = 20) %>%
   ir::ir_flatten()
 
-d_veg_bin50 <- 
-  d_veg_bin50_flat %>%
+d_veg_bin20 <- 
+  d_veg_bin20_flat %>%
   dplyr::select(-1) %>%
@@
   as.data.frame() %>%
-  scale(center = d_peat_bin50_scale$center, 
-        scale = d_peat_bin50_scale$scale) %>%
+  scale(center = d_peat_bin20_scale$center, 
+        scale = d_peat_bin20_scale$scale) %>%
   as.data.frame()
@@
   dplyr::bind_rows(
-    m1.10 %>%
-      rstanarm::posterior_predict(newdata = d_veg_bin50) %>%
+    m1.8 %>%
+      rstanarm::posterior_predict(newdata = d_veg_bin20) %>%
       as.data.frame() %>%
@@
       ),
-    m2.7 %>%
-      rstanarm::posterior_predict(newdata = d_veg_bin10) %>%
+    m2.8 %>%
+      rstanarm::posterior_predict(newdata = d_veg_bin20) %>%
       as.data.frame() %>%

### 2022-03-03 analysis/paper/005-paper-m-minerals.Rmd
Update manuscript 1, update formatting fo references.

@@
   c(
-    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[4]], ", df_global = 1, autoscale = TRUE)"),
+    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[2]], ", df_global = 1, autoscale = TRUE)"),
                        class = "b"),
@@
     data = 
-      d_underfitting[[4]] %>% 
+      d_underfitting[[2]] %>% 
       dplyr::select(-kl),
@@
     seed = seed,
-    control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 22 divergent transitions; adapt_delta = 0.99: There were 2000 transitions after warmup that exceeded the maximum treedepth.
-Increase max_treedepth above 10
+    control = list(adapt_delta = 0.99, max_treedepth = 15)
   ) 
@@
         apply(2, median),
-      posterior_predict(m1.10, newdata = d_underfitting[[4]]) %>%
+      posterior_predict(m1.8, newdata = d_underfitting[[2]]) %>%
         apply(2, median),
@@
 # Holocellulose
-m1.10_pars <- 
-  m1.10 %>%
+m1.8_pars <- 
+  m1.8 %>%
   as.data.frame() %>%
@@
 
-m1.10_pars <- 
-  m1.10_pars %>%
+m1.8_pars <- 
+  m1.8_pars %>%
   purrr::map_df(function(x) {
@@
   dplyr::mutate(
-    par_name = colnames(m1.10_pars),
-    wn = d_underfitting_flat[[4]]$x,
+    par_name = colnames(m1.8_pars),
+    wn = d_underfitting_flat[[2]]$x,
     variable_y = "Holocellulose (no minerals)"
@@
     par_name = colnames(m1.13_pars),
-    wn = d_underfitting_flat[[4]]$x,
+    wn = d_underfitting_flat[[2]]$x,
     variable_y = "Holocellulose (with minerals)"
@@
 # Klason lignin
-m2.7_pars <- 
-  m2.7 %>%
+m2.8_pars <- 
+  m2.8 %>%
   as.data.frame() %>%
@@
 
-m2.7_pars <- 
-  m2.7_pars %>%
+m2.8_pars <- 
+  m2.8_pars %>%
   purrr::map_df(function(x) {
@@
   dplyr::mutate(
-    par_name = colnames(m2.7_pars),
-    wn = d_underfitting_flat[[1]]$x,
+    par_name = colnames(m2.8_pars),
+    wn = d_underfitting_flat[[2]]$x,
     variable_y = "Klason lignin"
@@
   dplyr::bind_rows(
-    m1.10_pars,
+    m1.8_pars,
     m1.13_pars,
-    m2.7_pars
+    m2.8_pars
   )

### 2022-03-03 analysis/paper/006-paper-m-prediction-domain.Rmd
Update manuscript 1, update formatting fo references.

@@
     # training data
-    d_underfitting[[4]] %>% 
+    d_underfitting[[2]] %>% 
       dplyr::select(-kl, -hol) %>%
@@
         wn = 
-          d_underfitting_flat[[4]]$x %>%
-          rep(nrow(d_underfitting[[4]])),
+          d_underfitting_flat[[2]]$x %>%
+          rep(nrow(d_underfitting[[2]])),
         sample_id = 
-          seq_len(nrow(d_underfitting[[4]])) %>% 
-          rep(each = nrow(d_underfitting_flat[[4]])) %>% 
+          seq_len(nrow(d_underfitting[[2]])) %>% 
+          rep(each = nrow(d_underfitting_flat[[2]])) %>% 
           as.character(),
@@
     # peat data
-     d_peat_bin50 %>% 
+     d_peat_bin20 %>% 
       tidyr::pivot_longer(cols = dplyr::everything(),
@@
         wn = 
-          d_peat_bin50_flat$x %>%
-          rep(nrow(d_peat_bin50)),
+          d_peat_bin20_flat$x %>%
+          rep(nrow(d_peat_bin20)),
         sample_id = 
-          seq_len(nrow(d_peat_bin50)) %>% 
-          rep(each = nrow(d_peat_bin50_flat)) %>% 
-          magrittr::add(nrow(d_underfitting[[4]])) %>%
+          seq_len(nrow(d_peat_bin20)) %>% 
+          rep(each = nrow(d_peat_bin20_flat)) %>% 
+          magrittr::add(nrow(d_underfitting[[2]])) %>%
           as.character(),
@@
     # vegetation data
-     d_veg_bin50 %>% 
+     d_veg_bin20 %>% 
       tidyr::pivot_longer(cols = dplyr::everything(),
@@
         wn = 
-          d_veg_bin50_flat$x %>%
-          rep(nrow(d_veg_bin50)),
+          d_veg_bin20_flat$x %>%
+          rep(nrow(d_veg_bin20)),
         sample_id = 
-          seq_len(nrow(d_veg_bin50)) %>% 
-          rep(each = nrow(d_veg_bin50_flat)) %>% 
-          magrittr::add(nrow(d_underfitting[[4]]) + nrow(d_peat_bin50)) %>%
+          seq_len(nrow(d_veg_bin20)) %>% 
+          rep(each = nrow(d_veg_bin20_flat)) %>% 
+          magrittr::add(nrow(d_underfitting[[2]]) + nrow(d_peat_bin20)) %>%
           as.character(),
@@
     # training data
-    d_underfitting[[1]] %>% 
+    d_underfitting[[2]] %>% 
       dplyr::select(-kl, -hol) %>%
@@
         wn = 
-          d_underfitting_flat[[1]]$x %>%
-          rep(nrow(d_underfitting[[1]])),
+          d_underfitting_flat[[2]]$x %>%
+          rep(nrow(d_underfitting[[2]])),
         sample_id = 
-          seq_len(nrow(d_underfitting[[1]])) %>% 
-          rep(each = nrow(d_underfitting_flat[[1]])) %>% 
+          seq_len(nrow(d_underfitting[[2]])) %>% 
+          rep(each = nrow(d_underfitting_flat[[2]])) %>% 
           as.character(),
@@
     # peat data
-     d_peat_bin10 %>% 
+     d_peat_bin20 %>% 
       tidyr::pivot_longer(cols = dplyr::everything(),
@@
         wn = 
-          d_peat_bin10_flat$x %>%
-          rep(nrow(d_peat_bin10)),
+          d_peat_bin20_flat$x %>%
+          rep(nrow(d_peat_bin20)),
         sample_id = 
-          seq_len(nrow(d_peat_bin10)) %>% 
-          rep(each = nrow(d_peat_bin10_flat)) %>% 
-          magrittr::add(nrow(d_underfitting[[1]])) %>%
+          seq_len(nrow(d_peat_bin20)) %>% 
+          rep(each = nrow(d_peat_bin20_flat)) %>% 
+          magrittr::add(nrow(d_underfitting[[2]])) %>%
           as.character(),
@@
     # vegetation data
-     d_veg_bin10 %>% 
+     d_veg_bin20 %>% 
       tidyr::pivot_longer(cols = dplyr::everything(),
@@
         wn = 
-          d_veg_bin10_flat$x %>%
-          rep(nrow(d_veg_bin10)),
+          d_veg_bin20_flat$x %>%
+          rep(nrow(d_veg_bin20)),
         sample_id = 
-          seq_len(nrow(d_veg_bin10)) %>% 
-          rep(each = nrow(d_veg_bin10_flat)) %>% 
-          magrittr::add(nrow(d_underfitting[[1]]) + nrow(d_peat_bin10)) %>%
+          seq_len(nrow(d_veg_bin20)) %>% 
+          rep(each = nrow(d_veg_bin20_flat)) %>% 
+          magrittr::add(nrow(d_underfitting[[2]]) + nrow(d_peat_bin20)) %>%
           as.character(),
@@
              colour = "grey", linetype = 2) + 
+  geom_vline(xintercept =  
+               m_pars %>%
+               dplyr::filter(stringr::str_detect(variable_y, "Klason lignin") & abs(median) >= 0.02) %>%
+               dplyr::pull(wn),
+             colour = "grey", linetype = 2) + 
   labs(x = expression("Bin wavenumber ["*cm^{-1}*"]"),
        y = "Predictor value",
-       title = "Bin width = 50") +
+       title = "Bin width = 20") +
   guides(fill = guide_legend(title = "Dataset")) +
@@
                dplyr::pull(wn),
-             colour = "grey", linetype = 2) + # ---todo: remove hard coding of values
+             colour = "grey", linetype = 2) + 
   labs(x = expression("Bin wavenumber ["*cm^{-1}*"]"),
@@
 # combine plots
-pp1 <-
-  (p1 | p2) +
-  plot_annotation(title = "Original models")
-
-pp2 <-
-  p3 / p4 +
-  plot_annotation(title = "Models using binned spectra")
-
 p_prediction_domain <-
-  pp1 / pp2 +
+  (p1 | p2 | p3) + 
   plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
-  plot_layout(heights = c(0.4, 2), guides = "collect") &
+  plot_layout(widths = c(1, 1, 3), guides = "collect") &
   theme(legend.position = "bottom",
         legend.box = "vertical")
+
+## old version wit two different plots for holocellulose and Klason lignin
+## combine plots
+#pp1 <-
+#  (p1 | p2) +
+#  plot_annotation(title = "Original models")
+#
+#pp2 <-
+#  p3 / p4 +
+#  plot_annotation(title = "Models using binned spectra")
+#
+#p_prediction_domain <- 
+#  pp1 / pp2 +
+#  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
+#  plot_layout(heights = c(0.4, 2), guides = "collect") &
+#  theme(legend.position = "bottom",
+#        legend.box = "vertical")
 ```

### 2022-03-03 analysis/paper/007-paper-m-prediction-differences.Rmd
Update manuscript 1, update formatting fo references.

@@
         hol_hat =
-          m1.10 %>%
-          posterior_predict(newdata = d_underfitting[[4]]) %>%
-          as.data.frame() %>%
-          purrr::map_dbl(median),
-         kl_hat = 
-          m2.7 %>%
-          posterior_predict(newdata = d_underfitting[[1]]) %>%
+          m1.8 %>%
+          posterior_predict(newdata = d_underfitting[[2]]) %>%
+          as.data.frame() %>%
+          purrr::map_dbl(median),
+         kl_hat = 
+          m2.8 %>%
+          posterior_predict(newdata = d_underfitting[[2]]) %>%
           as.data.frame() %>%
@@
         hol_hat =
-          m1.10 %>%
-          posterior_predict(newdata = d_peat_bin50) %>%
-          as.data.frame() %>%
-          purrr::map_dbl(median),
-         kl_hat = 
-          m2.7 %>%
-          posterior_predict(newdata = d_peat_bin10) %>%
+          m1.8 %>%
+          posterior_predict(newdata = d_peat_bin20) %>%
+          as.data.frame() %>%
+          purrr::map_dbl(median),
+         kl_hat = 
+          m2.8 %>%
+          posterior_predict(newdata = d_peat_bin20) %>%
           as.data.frame() %>%
@@
         hol_hat =
-          m1.10 %>%
-          posterior_predict(newdata = d_veg_bin50) %>%
-          as.data.frame() %>%
-          purrr::map_dbl(median),
-         kl_hat = 
-          m2.7 %>%
-          posterior_predict(newdata = d_veg_bin10) %>%
+          m1.8 %>%
+          posterior_predict(newdata = d_veg_bin20) %>%
+          as.data.frame() %>%
+          purrr::map_dbl(median),
+         kl_hat = 
+          m2.8 %>%
+          posterior_predict(newdata = d_veg_bin20) %>%
           as.data.frame() %>%
@@
   d_pred_diff_spectra %>%
-  ir::ir_get_intensity(wavenumber = c(3400, 1200, 1550)) %>%
+  ir::ir_get_intensity(wavenumber = c(3400, 1250, 1590)) %>%
   ir::ir_drop_spectra() %>%
@@
                   d_pred_diff_intensities %>%
-                  dplyr::pull(intensity_1550) %>%
+                  dplyr::pull(intensity_1590) %>%
                   scale()
@@
   #                  d_pred_diff_intensities %>%
-  #                    dplyr::pull(intensity_1550)) %>%
+  #                    dplyr::pull(intensity_1590)) %>%
   #             dplyr::filter(p_index & sample_id %in% d$sample_id[d$sample_type != "old magazines"][hol_sel]), 
@@
   facet_wrap( ~ model) +
-  guides(size = guide_legend(title = expression(Absorbance~at~1550~cm^{-1})),
+  guides(size = guide_legend(title = expression(Absorbance~at~1590~cm^{-1})),
          colour = guide_colourbar(title = "carb")) +
@@
   irpeat:::irp_content_klh_hodgkins_prepare() %>%
-  hodgkins_ggplot() %>%
+  hkl_hodgkins_ggplot() %>%
   wrap_plots()
@@
                   d_pred_diff_intensities %>%
-                  dplyr::pull(intensity_1550)) %>%
+                  dplyr::pull(intensity_1590)) %>%
   dplyr::filter(model == "Best binned spectra") %>%
@@
                   d_pred_diff_intensities %>%
-                  dplyr::pull(intensity_1550)) %>%
+                  dplyr::pull(intensity_1590)) %>%
   dplyr::filter(p1_index) %>%
@@
 How do these spectra differ?
-It seem that this is mainly th region around 1550
+It seem that this is mainly th region around 1590
 d %>% 
@@
   geom_path(aes(x = x, y = y, colour = sample_type, group = measurement_id)) +
-  geom_vline(xintercept = 1550)
+  geom_vline(xintercept = 1590)
 ```
@@
   dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
-  ggplot(aes(x = intensity_1200 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
+  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
   geom_hline(yintercept = 0, colour = "grey50") +
@@
   facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
-  labs(y = "Prediction difference", x = expression("Intensity at 1200 cm"^{-1}), 
-       title = expression("Intensity at 1200 cm"^{-1})) +
+  labs(y = "Prediction difference", x = expression("Intensity at 1250 cm"^{-1}), 
+       title = expression("Intensity at 1250 cm"^{-1})) +
   guides(colour = guide_legend(title = "Dataset")) +
@@
   dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
-  ggplot(aes(x = intensity_1550 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
+  ggplot(aes(x = intensity_1950 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
   geom_hline(yintercept = 0, colour = "grey50") +
@@
   facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
-  labs(y = "Prediction difference", x = expression("Intensity at 1550 cm"^{-1}), 
-       title = expression("Intensity at 1550 cm"^{-1})) +
+  labs(y = "Prediction difference", x = expression("Intensity at 1590 cm"^{-1}), 
+       title = expression("Intensity at 1590 cm"^{-1})) +
   guides(colour = guide_legend(title = "Dataset")) +
@@
   dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
-  ggplot(aes(x = intensity_1550 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
+  ggplot(aes(x = intensity_1590 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
   geom_hline(yintercept = 0, colour = "grey50") +
@@
   dplyr::filter(p_index_test) %>%
-  ggplot(aes(x = intensity_1200 %>% scale(), y = hol_residuals, colour = model)) + 
+  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_residuals, colour = model)) + 
   geom_hline(yintercept = 0, colour = "grey50") +
@@
   facet_grid(model ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
-  labs(y = "Prediction difference", x = expression("Intensity at 1200 cm"^{-1}), 
-       title = expression("Intensity at 1200 cm"^{-1})) +
+  labs(y = "Prediction difference", x = expression("Intensity at 1250 cm"^{-1}), 
+       title = expression("Intensity at 1250 cm"^{-1})) +
   guides(colour = guide_legend(title = "Data type")) +
@@
   dplyr::filter(p_index_test) %>%
-  ggplot(aes(x = intensity_1550 %>% scale(), y = hol_residuals, colour = model)) +
+  ggplot(aes(x = intensity_1590 %>% scale(), y = hol_residuals, colour = model)) +
   geom_hline(yintercept = 0, colour = "grey50") +
@@
   facet_grid(model ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
-  labs(y = "Prediction difference", x = expression("Intensity at 1550 cm"^{-1}), 
-       title = expression("Intensity at 1550 cm"^{-1})) +
+  labs(y = "Prediction difference", x = expression("Intensity at 1590 cm"^{-1}), 
+       title = expression("Intensity at 1590 cm"^{-1})) +
   guides(colour = guide_legend(title = "Data type")) +
@@
   dplyr::filter(p_index_test2) %>%
-  ggplot(aes(x = intensity_1200 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
+  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
   geom_hline(yintercept = 0, colour = "grey50") +
@@
   facet_grid(data_type ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
-  labs(y = "Prediction difference", x = expression("Intensity at 1200 cm"^{-1}), 
-       title = expression("Intensity at 1200 cm"^{-1})) +
+  labs(y = "Prediction difference", x = expression("Intensity at 1250 cm"^{-1}), 
+       title = expression("Intensity at 1250 cm"^{-1})) +
   guides(colour = guide_legend(title = "Data type")) +
@@
   dplyr::filter(p_index_test2) %>%
-  ggplot(aes(x = intensity_1550 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
+  ggplot(aes(x = intensity_1590 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
   geom_hline(yintercept = 0, colour = "grey50") +
@@
   facet_grid(data_type ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
-  labs(y = "Prediction difference", x = expression("Intensity at 1550 cm"^{-1}), 
-       title = expression("Intensity at 1550 cm"^{-1})) +
+  labs(y = "Prediction difference", x = expression("Intensity at 1590 cm"^{-1}), 
+       title = expression("Intensity at 1590 cm"^{-1})) +
   guides(colour = guide_legend(title = "Data type")) +

### 2022-03-03 analysis/paper/008-paper-supplementary.Rmd
Update manuscript 1, update formatting fo references.

@@
                       cache = TRUE,
-                      dev = "pdf")
+                      dev = "png")
 ```
@@
 
-```{r res-p-pred-diff-hol-intensity2, warning=FALSE, echo=FALSE, out.width="50%", fig.height=14, fig.width=7, fig.cap='Prediction differences between the original model and improved model using binned spectra for the training (first row), peat (second row), and vegetation (bottom row) samples of \\cite{Hodgkins.2018} for different ranges of \\texttt{carb} peak heights (columns) versus intensities at different wavenumber values on the x axis (a: 3400, b: 1200, c: 1550 cm$^{-1}$).
+```{r res-p-pred-diff-hol-intensity2, warning=FALSE, echo=FALSE, out.width="50%", fig.height=14, fig.width=7, fig.cap='Prediction differences between the original model and improved model using binned spectra for the training (first row), peat (second row), and vegetation (bottom row) samples of \\cite{Hodgkins.2018} for different ranges of \\texttt{carb} peak heights (columns) versus intensities at different wavenumber values on the x axis (a: 3400, b: 1250, c: 1590 cm$^{-1}$).
 Larger prediction differences mean that the original model made larger predictions than the improved model using binned spectra.'}

### 2022-03-03 analysis/paper/001-reply-main.Rmd
Initial commit for manuscript 2.

---
title: 'Supplementary Information to: "Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging"'
author: "Henning Teickner, Klaus-Holger Knorr"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
---

```{r setup-reply, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", 
                      fig.pos = "H",
                      echo = FALSE,
                      cache = TRUE)
```

# Methods

For our analysis, we used three models for each holocellulose and Klason lignin:

1. The models described in @Hodgkins.2018 to predict holocellulose and Klason lignin contents from mid infrared spectra, but recomputed in a Bayesian framework. This is the "original Bayesian model" in [---todo:ref]. These models were used to reproduce the computations in @Hodgkins.2018. Here, these models are called "original" models.

2. A modified model to predict each, holocellulose and Klason lignin contents from mid infrared spectra. These are the models called "best binned spectra" models in [---todo:ref]. A brief description of these models is given in section [Computation of modified models to predict organic matter holocellulose and Klason lignin contents from mid infrared spectra]. These models are used to reanalyze holocellulose and Klason lignin depth profiles for the two latitude classes defined in @Hodgkins.2018, as described in section [Depth profiles].

3. The same models as in number 2, but extended such that it does not only predict holocellulose or Klason lingin contents, but also describes the latitudinal gradient in average near-surface holocellulose or Klason lingin contents. These models are described in detail in section [Latitudinal gradients].

To differentiate the models, we call the models from number 1 "original models", the models from number 2 "modified models using only the mid infrared spectra" and "Modified, only MIRS-based" in figure \@ref(fig:reply-res-p-latitude-poc-slope), and the models from number 3 (whenever it is not clear from the context which model we refer to) "our models for the latitudinal gradient" and "Modified" in figure \@ref(fig:reply-res-p-latitude-poc-slope).

## Computation of modified models to predict organic matter holocellulose and Klason lignin contents from mid infrared spectra

Computation of the modified models using only the mid infrared spectra (MIRS) is described in detail in [---todo:ref]. In brief, we used the same training data as @Hodgkins.2018 and a Bayesian linear regression model which we computed with Markov Chain Monte Carlo (MCMC) sampling using brms [@Burkner.2017; @Burkner.2018] and rstan [@StanDevelopmentTeam.2020] which are both interfaces to Stan [@StanDevelopmentTeam.2021]. In contrast to the original models of @Hodgkins.2018, we (1) used a beta distribution as likelihood function instead of assuming a Gaussian distribution, because holocellulose and Klason lignin contents cannot be negative or larger than 100 mass-%, we (2) did not use manually selected peaks extracted from the spectra, but all the $z$-transformed absorbance values after baseline correction, normalization, and binning (bin width = 20 cm$^{-1}$), and we (3) did not use simple linear regression, but regularized linear regression by using a regularized horseshoe prior [@Piironen.2017a; @Piironen.2017b] for coefficients for spectral variables. For the intercept and for the scale parameter of the beta distribution, we used weakly informative default priors of rstanarm [@Goodrich.2020] and brms [@Burkner.2017; @Burkner.2018].

## Reanalysis of latitudinal and depth patterns, considering prediction uncertainty of the mid infrared-based models

### Depth profiles

Predictions of the modified models using MIRS only were used to create figure 2<!---\@ref(fig:reply-res-p-depth-profiles-latitude)---> in the main text. For this, we extracted posterior draws from the models and for each MCMC iteration fitted a LOESS smoother with the same parameters as @Hodgkins.2018. From the fitted values (average predictions), we computed predictions across depth and from these 95% uncertainty intervals. We contrast these predictions and intervals with the intervals as computed by @Hodgkins.2018 where the prediction uncertainty of the model is not considered.

We acknowledge that this is no full Bayesian analysis. In a fully Bayesian analysis and uncertainty propagation, parameters for the smoother would have assigned prior distributions and the distribution of the parameters for the LOESS smoother would inform also the parameters of the regression model predicting holocellulose and Klason lignin contents, respectively, from the spectral data.

We did not compute a full Bayesian model here for the following reasons:

1. Since representative training data are missing, both the original and modified models cannot be fully validated and prediction uncertainties are probably larger than computed with these models [---todo:ref]. Since representative test data are missing, these additional uncertainties cannot be estimated. We assume that these uncertainties are larger than those induced by parameter uncertainty from models for the depth profiles.

2. The 45$^\circ$ latitude threshold applied in separating the data is a heuristic threshold to analyze some differences, but there is no specific reason about choosing this or some other value, except that sample numbers would be less balanced for this particular dataset. The goal therefore was not to compute a model which describes the underlying data optimally, but to show how depth profiles of the two heuristic latitude classes differ if one uses the same method as @Hodgkins.2018 (to compute the depth profile), but considers the prediction uncertainty of the models.

The modified models using only the MIRS were also used to compute median values and confidence intervals mentioned in the main text in section "Klason lignin content and residual enrichment in (sub)tropical peatlands probably are larger". In these computations, we also considered the predictive uncertainty of the original models (the Bayesian formulation described in ---todo:ref).

### Latitudinal gradients

```{r m-reply-latitude-fit-1, include=FALSE}
# compile the stan model
s1 <- rstan::stan_model("../stan_models/reply_stan_latitude.stan")

## create data list for the stan model

# peat MIRS, bin width: 20
d_peat_bin20_scale <-
  d_underfitting_flat[[2]] %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(center = TRUE, scale = TRUE)
d_peat_bin20_scale <-
  tibble::tibble(
    center = attr(d_peat_bin20_scale, "scaled:center"),
    scale = attr(d_peat_bin20_scale, "scaled:scale")
  )

d_peat_bin20_flat <-
  d_peat %>%
  ir::ir_bin(width = 20) %>%
  ir::ir_flatten()

d_peat_bin20 <- 
  d_peat_bin20_flat %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(center = d_peat_bin20_scale$center, 
        scale = d_peat_bin20_scale$scale) %>%
  as.data.frame()

# index for near-surface peat samples
core_sample_count <- table(d_peat[d_peat$sample_depth_lower <= 50, ]$core_label)
cores_to_keep <- names(core_sample_count)[core_sample_count >= 0]
index <- d_peat$core_label %in% cores_to_keep & d_peat$sample_depth_lower <= 50

index <- d_peat$sample_depth_lower <= 50
index_core <-
  d_peat[index, ]$core_label %>%
  as.factor() %>%
  as.integer()

# holocellulose
s1_data_holocellulose <-
  list(
    N = 
      d_underfitting[[2]] %>% 
      dplyr::select(-kl) %>% 
      dplyr::filter(d$sample_type != "old magazines") %>%
      nrow(),
    Y = 
      d_underfitting[[2]] %>% 
      dplyr::select(-kl) %>% 
      dplyr::filter(d$sample_type != "old magazines") %>%
      dplyr::pull("hol"),
    K = 
      d_underfitting[[2]] %>% 
      dplyr::select(-kl, -hol) %>% 
      ncol(),
    X = 
      d_underfitting[[2]] %>% 
      dplyr::select(-kl, -hol) %>%
      dplyr::filter(d$sample_type != "old magazines") %>%
      as.matrix(),
    hs_df = 1,
    hs_df_global = 1,
    hs_df_slab = 4,
    hs_scale_global = prior_hs_tau0[[2]],
    hs_scale_slab = 2,
    prior_only = 0,
    N1 = nrow(d_peat %>% dplyr::filter(index)),
    X1 = 
      d_peat_bin20 %>% 
      dplyr::filter(index) %>%
      as.matrix(),
    index_core = index_core,
    N1_cores = length(unique(index_core)),
    latitudes = 
      d_peat %>%
      dplyr::filter(index) %>%
      dplyr::filter(!duplicated(index_core)) %>%
      dplyr::pull(sampling_latitude) %>%
      scale() %>%
      as.numeric(),
    a_latitude_mu = -0.8, # -1.1
    a_latitude_sigma = 0.08,
    b_latitude_mu = 0,
    b_latitude_sigma = 0.3,
    phi_pop_a = 120,
    phi_pop_b = 15,
    phi_pop_scale = 5,
    phi_core_a = 50,
    phi_core_b = 5,
    phi_core_scale = 5,
    phi_a = 90,
    phi_b = 20,
    phi_scale = 20,
    Intercept_sigma = 0.25
  )

# Klason lignin
s1_data_klason_lignin <- s1_data_holocellulose
s1_data_klason_lignin$N <- 
  d_underfitting[[2]] %>% 
  dplyr::select(-hol) %>% 
  dplyr::filter(d$sample_type != "office paper") %>%
  nrow()
s1_data_klason_lignin$Y <-
  d_underfitting[[2]] %>% 
  dplyr::select(-hol) %>% 
  dplyr::filter(d$sample_type != "office paper") %>%
  dplyr::pull("kl")
s1_data_klason_lignin$K <-
  d_underfitting[[2]] %>% 
  dplyr::select(-kl, -hol) %>% 
  ncol()
s1_data_klason_lignin$X <-
  d_underfitting[[2]] %>% 
  dplyr::select(-kl, -hol) %>%
  dplyr::filter(d$sample_type != "office paper") %>%
  as.matrix()
s1_data_klason_lignin$X1 <- 
  d_peat_bin20 %>% 
  dplyr::filter(index) %>%
  as.matrix()
s1_data_klason_lignin$hs_scale_global <- prior_hs_tau0[[2]]
s1_data_klason_lignin$a_latitude_mu <- -0.2
s1_data_klason_lignin$a_latitude_sigma <- 0.08 # with 0.2, ALL contents get shrunk toward smaller values
s1_data_klason_lignin$b_latitude_sigma <- 0.3
s1_data_klason_lignin$phi_pop_a <- 120
s1_data_klason_lignin$phi_core_a <- 10
```

To reproduce the latitudinal gradients (figure 1<!---\@ref(fig:reply-res-p-sites-latitude)---> in the main text), we computed a hierarchical Bayesian model which is an extension to the modified model to predict holocellulose and Klason lignin contents using only the spectra, respectively. The extension is that we generate predictions for the peat samples by treating their holocellulose content (or Klason lignin content) as missing data. These predicted values are linked to a second model which is a linear regression model for the latitudinal gradient in average surface peat contents of holocellulose (or Klason lignin). Holocellulose (or Klason lignin) contents of individual peat samples are modeled with a beta distribution for each peat core. The mean value of these beta models is informed by a beta distribution describing the average surface peat content of an assumed population of peat cores. Similarly, the scale parameter of the beta models for each cores is informed by a gamma distribution describing the average scale parameter of an assumed population of peat cores. The full model for holocellulose content is:

$$
\begin{aligned}
\text{Y}_1[i] &\sim \text{beta}(\mu_{1}[i] \phi, (1-\mu_{1}[i])\phi)\\
Y_{2}[j] &\sim \text{beta}(\mu_{2}[j] \phi, (1-\mu_{2}[j])\phi)\\
\mu_{1} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}\right)\\
\mu_{2} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_1\right)\\
\alpha &\sim \text{normal(0, 0.2)}\\
\beta &\sim \text{horseshoe}(zb, hs_{local}, hs_{global}, hs_{scale\_slab}^2 \cdot hs_{slab})\\
\phi &\sim \text{beta}(`r s1_data_holocellulose$phi_a`, `r s1_data_holocellulose$phi_b/s1_data_holocellulose$phi_scale`)\\
Y_{2}[j] &\sim \text{beta}(\mu_{\text{core}}[\text{index\_core}[j]] \phi_{\text{core}}[\text{index\_core}[j]], (1-\mu_{\text{core}}[\text{index\_core}[j]])\phi_{\text{core}}[\text{index\_core}[j]])\\
\mu_{\text{core}}[k] &\sim \text{beta}(\mu_{\text{pop}}[k] \phi_{\text{pop}}, (1-\mu_{\text{pop}}[k])\phi_{\text{pop}})\\
\mu_{\text{pop}}[k] &= \text{logit}^{-1}\left(\alpha_{\text{latitude}} + \beta_{\text{latitude}}\text{x}_{\text{latitude}}[k]\right)\\
\phi_{\text{core}} &\sim \text{gamma}(`r s1_data_holocellulose$phi_core_a`, `r s1_data_holocellulose$phi_core_b/s1_data_holocellulose$phi_core_scale`)\\
\phi_{\text{pop}} &\sim \text{gamma}(`r s1_data_holocellulose$phi_pop_a`, `r s1_data_holocellulose$phi_pop_b/s1_data_holocellulose$phi_pop_scale`)\\
\alpha_{\text{latitude}} &\sim \text{normal}(`r s1_data_holocellulose$a_latitude_mu`, `r s1_data_holocellulose$a_latitude_sigma`)\\
\beta_{\text{latitude}} &\sim \text{normal}(`r s1_data_holocellulose$b_latitude_mu`, `r s1_data_holocellulose$b_latitude_sigma`)\\
\end{aligned}
$$

Where $\text{Y}_1$ are the measured holocellulose contents of the training data; $\mu_{1}$ are the respective mean values modeled with the MIRS, only; $i$ is an index for each sample of the training data; $Y_{2}$ are the estimated peat holocellulose contents modeled simultaneously from the MIRS ($\mu_{2}$) and from the average average near-surface peat holocellulose contents per core ($\mu_{\text{core}}$); $j$ is an index for each sample in the peat data; $\text{index\_core}$ is a vector assigning each sample in the peat data to the respective core index; $k$ is an index for each unique core in the peat data. The average average near-surface peat holocellulose contents per core ($\mu_{\text{core}}$) are in turn modeled in dependency on the latitude.

The full model for Klason lignin is:

$$
\begin{aligned}
\text{Y}_1[i] &\sim \text{beta}(\mu_{1}[i] \phi, (1-\mu_{1}[i])\phi)\\
Y_{2}[j] &\sim \text{beta}(\mu_{2}[j] \phi, (1-\mu_{2}[j])\phi)\\
\mu_{1} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}\right)\\
\mu_{2} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_1\right)\\
\alpha &\sim \text{normal(0, 0.2)}\\
\beta &\sim \text{horseshoe}(zb, hs_{local}, hs_{global}, hs_{scale\_slab}^2 \cdot hs_{slab})\\
\phi &\sim \text{beta}(`r s1_data_klason_lignin$phi_a`, `r s1_data_klason_lignin$phi_b/s1_data_klason_lignin$phi_scale`)\\
Y_{2}[j] &\sim \text{beta}(\mu_{\text{core}}[\text{index\_core}[j]] \phi_{\text{core}}[\text{index\_core}[j]], (1-\mu_{\text{core}}[\text{index\_core}[j]])\phi_{\text{core}}[\text{index\_core}[j]])\\
\mu_{\text{core}}[k] &\sim \text{beta}(\mu_{\text{pop}}[k] \phi_{\text{pop}}, (1-\mu_{\text{pop}}[k])\phi_{\text{pop}})\\
\mu_{\text{pop}}[k] &= \text{logit}^{-1}\left(\alpha_{\text{latitude}} + \beta_{\text{latitude}}\text{x}_{\text{latitude}}[k]\right)\\
\phi_{\text{core}} &\sim \text{gamma}(`r s1_data_klason_lignin$phi_core_a`, `r s1_data_klason_lignin$phi_core_b/s1_data_klason_lignin$phi_core_scale`)\\
\phi_{\text{pop}} &\sim \text{gamma}(`r s1_data_klason_lignin$phi_pop_a`, `r s1_data_klason_lignin$phi_pop_b/s1_data_klason_lignin$phi_pop_scale`)\\
\alpha_{\text{latitude}} &\sim \text{normal}(`r s1_data_klason_lignin$a_latitude_mu`, `r s1_data_klason_lignin$a_latitude_sigma`)\\
\beta_{\text{latitude}} &\sim \text{normal}(`r s1_data_klason_lignin$b_latitude_mu`, `r s1_data_klason_lignin$b_latitude_sigma`)\\
\end{aligned}
$$

```{r m-reply-latitude-ppc}
## prior preditive check

# function to perform the prior predictive check
ppc_latitude <- 
  function(x_latitude, 
           a_latitude_mu, 
           a_latitude_sigma, 
           b_latitude_mu, 
           b_latitude_sigma, 
           phi_pop_a, 
           phi_pop_b, 
           phi_pop_scale, 
           phi_core_a, 
           phi_core_b, 
           phi_core_scale, 
           nsim = 1e4) {

    a_latitude <- rnorm(nsim, a_latitude_mu, a_latitude_sigma)
    b_latitude <- rnorm(nsim, b_latitude_mu, b_latitude_sigma)
    phi_pop <- rgamma(nsim, phi_pop_a, phi_pop_b) * phi_pop_scale
    phi_core <-  purrr::map(x_latitude, function(x) rgamma(nsim, phi_core_a, phi_core_b) * phi_core_scale)

    mu_pop <- purrr::map(x_latitude, function(x) brms::inv_logit_scaled(a_latitude + b_latitude * x))
    mu_core <- purrr::map(mu_pop, function(x) rbeta(nsim, x * phi_pop, (1 - x) * phi_pop))
    Y_1 <- purrr::map2(mu_core, phi_core, function(x, y) rbeta(nsim, x * y, (1 - x) * y))

    tibble::tibble(
      iter = rep(seq_len(nsim), length(!!x_latitude)),
      x_latitude = rep(x_latitude, each = nsim),
      a_latitude = rep(a_latitude, length(!!x_latitude)),
      b_latitude = rep(b_latitude, length(!!x_latitude)),
      phi_pop = rep(phi_pop, length(!!x_latitude)),
      phi_core = unlist(phi_core),
      mu_pop = unlist(mu_pop),
      mu_core = unlist(mu_core),
      Y_1 = unlist(Y_1)
    )
  }

x_latitude <- 
  seq(min(scale(unique(d_peat$sampling_latitude))),
      max(scale(unique(d_peat$sampling_latitude))), 
      length.out = 10)

d_ppc_latitude <- 
  dplyr::bind_rows(
    ppc_latitude(x_latitude = x_latitude,
               a_latitude_mu = s1_data_holocellulose$a_latitude_mu, 
               a_latitude_sigma = s1_data_holocellulose$a_latitude_sigma, 
               b_latitude_mu = s1_data_holocellulose$b_latitude_mu, 
               b_latitude_sigma = s1_data_holocellulose$b_latitude_sigma, 
               phi_pop_a = s1_data_holocellulose$phi_pop_a, 
               phi_pop_b = s1_data_holocellulose$phi_pop_b,
               phi_pop_scale = s1_data_holocellulose$phi_pop_scale,
               phi_core_a = s1_data_holocellulose$phi_core_a,
               phi_core_b = s1_data_holocellulose$phi_core_b,
               phi_core_scale = s1_data_holocellulose$phi_core_scale,
               nsim = 4 * 1e3) %>%
      dplyr::mutate(variable_y = "Holocellulose"),
    ppc_latitude(x_latitude = x_latitude,
               a_latitude_mu = s1_data_klason_lignin$a_latitude_mu, 
               a_latitude_sigma = s1_data_klason_lignin$a_latitude_sigma, 
               b_latitude_mu = s1_data_klason_lignin$b_latitude_mu, 
               b_latitude_sigma = s1_data_klason_lignin$b_latitude_sigma, 
               phi_pop_a = s1_data_klason_lignin$phi_pop_a, 
               phi_pop_b = s1_data_klason_lignin$phi_pop_b,
               phi_pop_scale = s1_data_klason_lignin$phi_pop_scale,
               phi_core_a = s1_data_klason_lignin$phi_core_a,
               phi_core_b = s1_data_klason_lignin$phi_core_b,
               phi_core_scale = s1_data_klason_lignin$phi_core_scale,
               nsim = 4 * 1e3) %>%
      dplyr::mutate(variable_y = "Klason lignin")
  )

# histogram of average surface peat content for a core with average latitude
p_ppc_latitude1 <-
  d_ppc_latitude %>%
  dplyr::filter(x_latitude == unique(x_latitude)[[1]]) %>%
  ggplot(aes(x = brms::inv_logit_scaled(a_latitude) * 100)) +
  geom_histogram(bins = 30) +
  labs(y = "Counts", x = expression(alpha[latitude]~"(inverse logit) [mass-%]")) +
  facet_wrap(~ variable_y)

# plot of 100 regression lines drawn from the joint prior distribution
x_latitude_center <- attr(scale(unique(d_peat$sampling_latitude)), "scaled:center")
x_latitude_scale <- attr(scale(unique(d_peat$sampling_latitude)), "scaled:scale")
p_ppc_latitude2 <-
  d_ppc_latitude %>%
  dplyr::filter(iter <= 100) %>%
  ggplot(aes(x = x_latitude * x_latitude_scale + x_latitude_center, y = mu_core * 100, group = iter)) +
  geom_line(stat = "smooth", formula = y ~ x, method = "lm", se = FALSE, colour = "black", alpha = 0.3) +
  labs(y = expression(mu[core]~"[mass-%]"), x = expression(Latitude~"[°]")) +
  facet_wrap(~ variable_y)

# plot range in which content for individual samples may lie
p_ppc_latitude3 <-
  d_ppc_latitude %>%
  dplyr::filter(iter <= 100) %>%
  ggplot(aes(x = x_latitude * x_latitude_scale + x_latitude_center, y = Y_1 * 100, group = iter)) +
  geom_point(alpha = 0.3) +
  labs(y = expression(Y[2]*"["*j*"]"~"[mass-%]"), x = expression(Latitude~"[°]")) +
  facet_wrap(~ variable_y)

# histogram for a marginal peat sample
p_ppc_latitude4 <-
  d_ppc_latitude %>%
  dplyr::filter(x_latitude == unique(x_latitude)[[1]]) %>%
  ggplot(aes(x = Y_1 * 100)) +
  geom_histogram(bins = 30) +
  labs(y = "Counts", x = expression(Y[2]*"["*j*"]"~"[mass-%]")) +
  facet_wrap(~ variable_y)
```

Parameter values for prior distributions were chosen based on a prior predictive simulation by plotting histograms and prior model predictions within the range of the predictor values (figure \@ref(fig:reply-res-p-latitude-ppc)). These plots allow to check implications of prior choices in light of the knowledge on holocellulose and Klason lignin contents available: We know e.g. from @Hayes.2015 that holocellulose (total sugar) and Klason lignin contents of peat in Ireland (from unknown depths) measured with a procedure very similar to that used by @Hodgkins.2018 [@DelaCruz.2016] range between 8 to 29 and 42 to 72 mass-%, respectively. Holocellulose and Klason lignin contents of peat forming vegetation (of boreal peatlands) have been reported to range between 25.8 to 70.9, and 8.1 to 54.3 mass-%, respectively [@Turetsky.2008; @Strakova.2010; @Bengtsson.2018]. Surface peat contents should be similar to these values and holocellulose contents generally smaller and Klason lignin contents larger due to decomposition losses [@Benner.1984; @Worrall.2017]. In addition, we adjusted prior choices (particularly by reducing the standard deviation of the global intercept of the model for the latitudinal gradient and other parameters (see below)) to avoid computational issues (see below). We assured via prior and posterior predictive checks that these adjustments neither result in too informative priors, nor in a bad fit to the data.

(ref:reply-res-p-latitude-ppc-caption) Prior predictive checks for the modified models describing the latitudinal gradient of peat core average near-surface holocellulose and Klason lignin contents. (a) Marginal prior distribution for the intercept of the latitudinal gradient model (average peat core average near-surface content at the average latitude of the sampled peat cores). (b) 100 random lines from the joint prior distribution for the intercept and slope for the latitudinal gradient (difference between average peat core average near-surface content at different latitudes). (c) Marginal prior distribution for the content of an individual peat sample at a latitude of ~`r round(unique(x_latitude)[[1]] * x_latitude_scale + x_latitude_center, 1)` $^\circ$N. (d) At a sequence of discrete latitude values 100 randomly sampled contents for individual near-surface peat samples.

```{r reply-res-p-latitude-ppc, echo=FALSE, out.width="70%", fig.height=5, fig.width=6.5, fig.cap='(ref:reply-res-p-latitude-ppc-caption)'}
p_ppc_latitude1 + p_ppc_latitude2 + p_ppc_latitude4 + p_ppc_latitude3 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")")
```

Our final prior choice assumes a population average near-surface content of holocellulose of ~`r brms::inv_logit_scaled(s1_data_holocellulose$a_latitude_mu) %>% magrittr::multiply_by(10) %>% round(0) %>% magrittr::multiply_by(10)` mass-% with a standard deviation of ~`r brms::inv_logit_scaled(rnorm(1e4, s1_data_holocellulose$a_latitude_mu, s1_data_holocellulose$a_latitude_sigma)) %>% sd() %>% magrittr::multiply_by(100) %>% round(0)` mass-% and of Klason lignin of ~`r brms::inv_logit_scaled(s1_data_klason_lignin$a_latitude_mu) %>% magrittr::multiply_by(10) %>% round(0) %>% magrittr::multiply_by(10)` mass-% with a standard deviation of ~`r brms::inv_logit_scaled(rnorm(1e4, s1_data_klason_lignin$a_latitude_mu, s1_data_klason_lignin$a_latitude_sigma)) %>% sd() %>% magrittr::multiply_by(100) %>% round(0)` mass-% (figure \@ref(fig:reply-res-p-latitude-ppc)). Assuming a normal(`r s1_data_holocellulose$b_latitude_mu`, `r s1_data_holocellulose$b_latitude_sigma`) prior for $\beta_{\text{latitude}}$ for both models resulted in a broad range of possible relations between contents and latitude (figure \@ref(fig:reply-res-p-latitude-ppc)). Similarly, the parameter choices for $\phi$, $\phi_{\text{core}}$, and $\phi_{\text{pop}}$ resulted in reasonable ranges of contents on the level of averages for individual cores and samples, respectively (figure \@ref(fig:reply-res-p-latitude-ppc)).

As mentioned above, some regularization was necessary to avoid computational issues: Less regularizing priors for $\alpha_{\text{latitude}}$, $\phi_{\text{core}}$, and $\phi_{\text{pop}}$ resulted in (more) divergent transitions and badly mixing MCMC chains. This behavior occurred because average population and core contents tended to 0 mass-%. This is a consequence of the inverse logit transformation which can cause high prior mass around 0 and 100 mass-% if the variability in its parameters is too large. We can exclude 0 mass-% as implausible and therefore can justify our regularization because it avoided a drift of the majority of the MCMC draws into implausible regions.

With this model, we could estimate the average difference in surface peat holocellulose (or Klason lignin) content for different latitudes while considering (1) the prediction uncertainty for each individual peat sample, (2) the uncertainty in the average surface peat contents for each peat core (which is heterogeneous across cores due to a finite number of samples and due to core-dependent depth gradients), and (3) parameter uncertainty of the regression model for the latitudinal gradient. In contrast, the original analysis did not consider uncertainty from predictions, and from computing average surface peat contents per core.

To create figure 1<!--\@ref(fig:reply-res-p-sites-latitude)--> in the main text, we used the models for the latitudinal gradient and samples from the posterior distribution to compute the predicted average and 95% confidence interval for a range of latitude values and we added predictions for the average contents for each of the sampled cores. Next, we recomputed the models from @Hodgkins.2018 by (1) predicting average peat sample contents with the original models (the Bayesian version described in ---todo:ref), (2) averaging the near surface peat sample average contents for each core, and (3) fitting an ordinary least squares regression line to these average core values. For this linear model, we plotted the confidence interval across latitude and we added the computed average core near-surface contents. Thus, figure 1<!---\@ref(fig:reply-res-p-sites-latitude)---> in the main text compares the graphed values and model prediction from figure 3 in @Hodgkins.2018 with predictions from our modified analysis considering prediction uncertainty from the MIRS-based model part, and the averaging per core.  
Additional sources for differences in predictions and prediction uncertainties in comparison to the analysis in @Hodgkins.2018 are the use of a beta distribution which constraints uncertainty intervals for extreme (near 0 and 100 mass-%) values, and the use of all spectral variables which resulted in reduced prediction uncertainties [---todo:ref].

```{r m-reply-latitude-fit-2, include=FALSE}
# fit model for holocellulose
s1_fit_holocellulose <-
  rstan::sampling(s1, 
                  data = s1_data_holocellulose, 
                  chains = 4, 
                  cores = 4, 
                  iter = iter, 
                  warmup = iter %/% 2, 
                  control = list(adapt_delta = 0.99, max_treedepth = 15),
                  seed = seed)

s1_res_holocellulose <- as.data.frame(rstan::extract(s1_fit_holocellulose))
```

```{r m-reply-latitude-fit-3, include=FALSE}
# fit model for klason_lignin
s1_fit_klason_lignin <-
  rstan::sampling(s1, 
                  data = s1_data_klason_lignin, 
                  chains = 4, 
                  cores = 4, 
                  iter = iter, 
                  warmup = iter %/% 2, 
                  control = list(adapt_delta = 0.99, max_treedepth = 15),
                  seed = seed)

s1_res_klason_lignin <- as.data.frame(rstan::extract(s1_fit_klason_lignin))
```

```{r m-reply-latitude-poc}
## posterior predictive checks
m1.8_pred <-
  predict(m1.8, newdata = d_peat_bin20)[d_peat$sample_depth_lower <= 50, ]
m2.8_pred <-
  predict(m2.8, newdata = d_peat_bin20)[d_peat$sample_depth_lower <= 50, ]

# function to perform the posterior predictive check
poc_latitude <- 
  function(x_latitude, 
           s1_fit, 
           s1_data, 
           core_i = 1) {

    s1_pars <- as.data.frame(rstan::extract(s1_fit))
    nsim <- nrow(s1_pars)

    a_latitude <- s1_pars$a_latitude * s1_data$a_latitude_sigma + s1_data$a_latitude_mu 
    b_latitude <- s1_pars$b_latitude * s1_data$b_latitude_sigma + s1_data$b_latitude_mu
    phi_pop <- s1_pars$phi_pop * s1_data$phi_pop_scale
    phi_core <-  s1_pars[stringr::str_detect(colnames(s1_pars), "phi_core\\.")] %>%
      purrr::map_df(function(x) x * s1_data$phi_core_scale)

    mu_pop <- purrr::map(x_latitude, function(x) brms::inv_logit_scaled(a_latitude + b_latitude * x))
    mu_core <- purrr::map(mu_pop, function(x) rbeta(nsim, x * phi_pop, (1 - x) * phi_pop))
    Y_1 <- purrr::map(mu_core, function(x) rbeta(nsim, x * phi_core[[core_i]], (1 - x) * phi_core[[core_i]]))

    tibble::tibble(
      iter = rep(seq_len(nsim), length(!!x_latitude)),
      x_latitude = rep(x_latitude, each = nsim),
      a_latitude = rep(a_latitude, length(!!x_latitude)),
      b_latitude = rep(b_latitude, length(!!x_latitude)),
      phi_pop = rep(phi_pop, length(!!x_latitude)),
      phi_core = unlist(rep(phi_core[core_i], length(!!x_latitude))),
      mu_pop = unlist(mu_pop),
      mu_core = unlist(mu_core),
      Y_1 = unlist(Y_1)
    )

  }

# compute posterior predictive check
d_poc_latitude <- 
  dplyr::bind_rows(
  dplyr::bind_rows(
    poc_latitude(x_latitude = x_latitude, 
                 s1_fit = s1_fit_holocellulose, 
                 s1_data = s1_data_holocellulose, 
                 core_i = 9) %>%
      dplyr::mutate(variable_y = "Holocellulose"),
    poc_latitude(x_latitude = x_latitude, 
                 s1_fit = s1_fit_klason_lignin, 
                 s1_data = s1_data_klason_lignin, 
                 core_i = 9) %>%
      dplyr::mutate(variable_y = "Klason lignin")
  ) %>%
    dplyr::mutate(sample_type = "Posterior"),
  d_ppc_latitude %>%
    dplyr::mutate(sample_type = "Prior")
  )

# histogram of average surface peat content for a core with average latitude
p_poc_latitude1 <-
  d_poc_latitude %>%
  dplyr::filter(x_latitude == unique(x_latitude)[[1]]) %>%
  ggplot(aes(x = brms::inv_logit_scaled(a_latitude) * 100)) +
  geom_histogram(bins = 30) +
  labs(y = "Counts", x = expression(alpha[latitude]~"(inverse logit) [mass-%]")) +
  facet_grid(sample_type ~ variable_y, scales = "free_x")

# plot of 100 regression lines drawn from the joint prior distribution
x_latitude_center <- attr(scale(unique(d_peat$sampling_latitude)), "scaled:center")
x_latitude_scale <- attr(scale(unique(d_peat$sampling_latitude)), "scaled:scale")
p_poc_latitude2 <-
  d_poc_latitude %>%
  dplyr::filter(iter <= 100) %>%
  ggplot(aes(x = x_latitude * x_latitude_scale + x_latitude_center, y = mu_core * 100, group = iter)) +
  geom_line(stat = "smooth", formula = y ~ x, method = "lm", se = FALSE, colour = "black", alpha = 0.3) +
  labs(y = expression(mu[core]~"[mass-%]"), x = expression(Latitude~"[°]"))  +
  facet_grid(sample_type ~ variable_y)

# plot range in which content for individual samples may lie
p_poc_latitude3 <-
  d_poc_latitude %>%
  dplyr::filter(iter <= 100) %>%
  ggplot(aes(x = x_latitude * x_latitude_scale + x_latitude_center, 
             y = Y_1 * 100, 
             group = iter)) +
  geom_point(alpha = 0.3) +
  labs(y = expression(Y[2]*"["*j*"]"~"[mass-%]"), x = expression(Latitude~"[°]")) +
  facet_grid(sample_type ~ variable_y)

# histogram for a marginal peat sample
p_poc_latitude4 <-
  d_poc_latitude %>%
  dplyr::filter(x_latitude == unique(x_latitude)[[1]]) %>%
  ggplot(aes(x = Y_1 * 100)) +
  geom_histogram(bins = 30) +
  labs(y = "Counts", x = expression(Y[2]*"["*j*"]"~"[mass-%]")) +
  facet_grid(sample_type ~ variable_y, scales = "free_x")

# histogram of marginal slope of latitudinal gradient
p_poc_latitude5 <-
  d_poc_latitude %>%
  dplyr::filter(x_latitude == unique(x_latitude)[[1]]) %>%
  ggplot(aes(x = b_latitude)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = 0, colour = "grey") +
  labs(y = "Counts", x = expression(beta[latitude])) +
  facet_grid(sample_type ~ variable_y, scales = "free_x")

# summarize the data to compare predictions to those for the unconstrained MIRS-based model
d_poc_latitude_summary <-
  d_poc_latitude %>%
  dplyr::group_by(sample_type, variable_y, x_latitude) %>%
  dplyr::summarise(mu_pop_lwr = quantile(mu_pop, probs = 0.025),
                   mu_pop_fit = quantile(mu_pop, probs = 0.5),
                   mu_pop_upr = quantile(mu_pop, probs = 0.975),
                   .groups = "drop")

d_y_yhat <-
  dplyr::bind_rows(
    s1_res_holocellulose %>%
      dplyr::select(dplyr::contains("Y1.")) %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = mean(x),
          lwr = quantile(x, probs = 0.025),
          upr = quantile(x, probs = 0.975)
        )
      }) %>%
      dplyr::bind_cols(as.data.frame(m1.8_pred)),
    s1_res_klason_lignin %>%
      dplyr::select(dplyr::contains("Y1.")) %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = mean(x),
          lwr = quantile(x, probs = 0.025),
          upr = quantile(x, probs = 0.975)
        )
      }) %>%
      dplyr::bind_cols(as.data.frame(m2.8_pred))
  ) %>%
  dplyr::mutate(
    x_latitude = rep((d_peat$sampling_latitude[d_peat$sample_depth_lower <= 50] - x_latitude_center)/x_latitude_scale, 2),
    variable_y = rep(c("Holocellulose", "Klason lignin"), each = sum(d_peat$sample_depth_lower <= 50))
  )

d_y_yhat_long <-
  dplyr::bind_rows(
    s1_res_holocellulose %>%
      dplyr::select(dplyr::contains("Y1.")) %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = mean(x),
          lwr = quantile(x, probs = 0.025),
          upr = quantile(x, probs = 0.975)
        )
      })  %>%
      dplyr::bind_rows(
        m1.8_pred %>% 
          as.data.frame() %>%
          dplyr::rename(fit = "Estimate",
                        lwr = "Q2.5",
                        upr = "Q97.5"),
        m1.2 %>%
          posterior_predict(newdata = d_peat_res[d_peat$sample_depth_lower <= 50, ] %>% dplyr::mutate(carb = as.numeric(carb))) %>%
          as.data.frame() %>%
          purrr::map_df(function(x) {
            tibble::tibble(
              fit = mean(x),
              lwr = quantile(x, probs = 0.025),
              upr = quantile(x, probs = 0.975)
            ) 
          })
        ),
    s1_res_klason_lignin %>%
      dplyr::select(dplyr::contains("Y1.")) %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = mean(x),
          lwr = quantile(x, probs = 0.025),
          upr = quantile(x, probs = 0.975)
        )
      }) %>%
      dplyr::bind_rows(
        m2.8_pred %>% 
          as.data.frame() %>%
          dplyr::rename(fit = "Estimate",
                        lwr = "Q2.5",
                        upr = "Q97.5"),
        m2.2 %>%
          posterior_predict(newdata = d_peat_res[d_peat$sample_depth_lower <= 50, ] %>% dplyr::mutate(arom15arom16 = as.numeric(arom15arom16))) %>%
          as.data.frame() %>%
          purrr::map_df(function(x) {
            tibble::tibble(
              fit = mean(x),
              lwr = quantile(x, probs = 0.025),
              upr = quantile(x, probs = 0.975)
            ) 
          })
      )
  ) %>%
  dplyr::mutate(
    core_label = rep(d_peat$core_label[d_peat$sample_depth_lower <= 50], 6),
    x_latitude = rep((d_peat$sampling_latitude[d_peat$sample_depth_lower <= 50] - x_latitude_center)/x_latitude_scale, 6),
    variable_y = rep(c("Holocellulose", "Klason lignin"), each = sum(d_peat$sample_depth_lower <= 50) * 3),
    model = rep(c("Modified", "Modified, only MIRS-based", "Original", "Modified", "Modified, only MIRS-based", "Original"), each = sum(d_peat$sample_depth_lower <= 50))
  )

# summary of predictions for sites
d_y_yhat_long_sites <-
  dplyr::bind_rows(
    s1_res_holocellulose %>%
      dplyr::select(dplyr::contains("mu_core.")) %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit_mean = mean(x),
          fit_lwr = quantile(x, probs = 0.025),
          fit_upr = quantile(x, probs = 0.975)
        )
      }) %>%
      dplyr::mutate(core_label = unique(d_peat$core_label[d_peat$sample_depth_lower <= 50])),
    m1.2 %>%
      posterior_predict(newdata = d_peat_res[d_peat$sample_depth_lower <= 50, ] %>% dplyr::mutate(carb = as.numeric(carb))) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit_mean = mean(x),
          fit_lwr = quantile(x, probs = 0.025),
          fit_upr = quantile(x, probs = 0.975)
        ) 
      }) %>%
        dplyr::mutate(core_label = d_peat$core_label[d_peat$sample_depth_lower <= 50]) %>%
        dplyr::group_by(core_label) %>%
        dplyr::summarise(
          fit_lwr = mean(fit_mean) - 2 * sd(fit_mean),
          fit_upr = mean(fit_mean) + 2 * sd(fit_mean),
          fit_mean = mean(fit_mean),
          .groups = "drop"
        ),
    s1_res_klason_lignin %>%
      dplyr::select(dplyr::contains("mu_core.")) %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit_mean = mean(x),
          fit_lwr = quantile(x, probs = 0.025),
          fit_upr = quantile(x, probs = 0.975)
        )
      })%>%
      dplyr::mutate(core_label = unique(d_peat$core_label[d_peat$sample_depth_lower <= 50])),
        m2.2 %>%
          posterior_predict(newdata = d_peat_res[d_peat$sample_depth_lower <= 50, ] %>% dplyr::mutate(arom15arom16 = as.numeric(arom15arom16))) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit_mean = mean(x),
          fit_lwr = quantile(x, probs = 0.025),
          fit_upr = quantile(x, probs = 0.975)
        ) 
      }) %>%
        dplyr::mutate(core_label = d_peat$core_label[d_peat$sample_depth_lower <= 50]) %>%
        dplyr::group_by(core_label) %>%
        dplyr::summarise(
          fit_lwr = mean(fit_mean) - 2 * sd(fit_mean),
          fit_upr = mean(fit_mean) + 2 * sd(fit_mean),
          fit_mean = mean(fit_mean),
          .groups = "drop"
        )
  ) %>%
  dplyr::mutate(
    core_label = rep(unique(d_peat$core_label[d_peat$sample_depth_lower <= 50]), 4),
    x_latitude = rep((d_peat$sampling_latitude[d_peat$sample_depth_lower <= 50 & !duplicated(d_peat$core_label)] - x_latitude_center)/x_latitude_scale, 4),
    variable_y = rep(c("Holocellulose", "Klason lignin"), each = sum(!duplicated(d_peat$core_label)) * 2),
    model = rep(c("Modified", "Original", "Modified", "Original"), each = sum(!duplicated(d_peat$core_label)))
  )

d_y_yhat_long_sites <-
  d_y_yhat_long %>%
  dplyr::group_by(variable_y, model, core_label) %>%
  dplyr::summarise(x_latitude = mean(x_latitude),
                   fit_mean = mean(fit),
                   fit_lwr = mean(fit) - sd(fit),
                   fit_upr = mean(fit) + sd(fit),
                   .groups = "drop")

# add predictions from the original model
d_poc_latitude_summary <-
  d_poc_latitude_summary %>%
  dplyr::bind_rows(
    {
      m <- lm(fit_mean ~ x_latitude, 
              data = 
                d_y_yhat_long_sites %>%
                dplyr::filter(model == "Original" & variable_y == "Holocellulose"))

      m_pred <- predict(m, interval = "confidence", level = 0.95, 
                        newdata = d_poc_latitude_summary %>%
                          dplyr::filter(sample_type == "Posterior" & variable_y == "Holocellulose")
      )

      tibble::tibble(
        variable_y = "Holocellulose",
        sample_type = "Original",
        x_latitude = x_latitude,
        mu_pop_fit = m_pred[, 1],
        mu_pop_lwr = m_pred[, 2],
        mu_pop_upr = m_pred[, 3]
      )
    },
    {
      m <- lm(fit_mean ~ x_latitude, 
              data = 
                d_y_yhat_long_sites %>%
                dplyr::filter(model == "Original" & variable_y == "Klason lignin"))

      m_pred <- predict(m, interval = "confidence", level = 0.95, 
                        newdata = d_poc_latitude_summary %>%
                          dplyr::filter(sample_type == "Posterior" & variable_y == "Klason lignin")
      )

      tibble::tibble(
        variable_y = "Klason lignin",
        sample_type = "Original",
        x_latitude = x_latitude,
        mu_pop_fit = m_pred[, 1],
        mu_pop_lwr = m_pred[, 2],
        mu_pop_upr = m_pred[, 3]
      )
    }
  )

p_poc_latitude_6_1 <-
  d_y_yhat_long %>%
  ggplot(aes(x = core_label, y = fit, ymin = lwr, ymax = upr, colour = model)) +
  geom_point(stat = "identity", position = position_dodge(0.8)) +
  geom_errorbar(width = 0, stat = "identity", position = position_dodge(0.8)) +
  facet_wrap(~ variable_y) +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(x = "Core", y = "Estimted content [mass-%]") +
  guides(colour = guide_legend(title = "Model", reverse = TRUE)) +
  theme(legend.position = "bottom")

## plot of predictions from the modified model in comparison to the original model across latitude

# Holocellulose
p1 <- 
  d_poc_latitude_summary %>%
  dplyr::filter(sample_type %in% c("Posterior", "Original") & variable_y == "Holocellulose") %>%
  dplyr::mutate(model = 
                  dplyr::case_when(
                    sample_type == "Posterior" ~ "Modified",
                    sample_type == "Prior" ~ "Prior",
                    sample_type == "Original" ~ "Original"
                  )) %>%
  ggplot() +
  geom_ribbon(aes(x = x_latitude, y = mu_pop_fit, 
                  ymin = mu_pop_lwr, ymax = mu_pop_upr,
                  fill = model), 
              alpha = 0.3) +
  geom_errorbar(data = d_y_yhat_long_sites %>%
                  dplyr::filter(model != "Modified, only MIRS-based" & variable_y == "Holocellulose"), 
                aes(y = fit_mean, x = x_latitude + ifelse(model == "Original", 0, 0.1), 
                    ymin = fit_lwr, ymax = fit_upr),
                colour = "grey", width = 0) +
  geom_path(aes(x = x_latitude, y = mu_pop_fit, colour = model), size = 0.7) +
  geom_point(data = d_y_yhat_long_sites %>%
               dplyr::filter(model != "Modified, only MIRS-based" & variable_y == "Holocellulose"), aes(y = fit_mean, x = x_latitude + ifelse(model == "Original", 0, 0.1), colour = model)) +
  scale_y_continuous(labels = function(x) x * 100) +
  scale_x_continuous(labels = function(x) round(x * x_latitude_scale + x_latitude_center, 0)) +
  labs(x = "Latitude [°]", y = "Content [mass-%]", title = "Holocellulose") +
  theme(legend.position = "bottom")

# Klason lignin
p2 <- 
  d_poc_latitude_summary %>%
  dplyr::filter(sample_type %in% c("Posterior", "Original") & variable_y == "Klason lignin") %>%
  dplyr::mutate(model = 
                  dplyr::case_when(
                    sample_type == "Posterior" ~ "Modified",
                    sample_type == "Prior" ~ "Prior",
                    sample_type == "Original" ~ "Original"
                  )) %>%
  ggplot() +
  geom_ribbon(aes(x = x_latitude, y = mu_pop_fit, 
                  ymin = mu_pop_lwr, ymax = mu_pop_upr,
                  fill = model), 
              alpha = 0.3) +
  geom_errorbar(data = d_y_yhat_long_sites %>%
                  dplyr::filter(model != "Modified, only MIRS-based" & variable_y == "Klason lignin"), 
                aes(y = fit_mean, x = x_latitude + ifelse(model == "Original", 0, 0.1), 
                    ymin = fit_lwr, ymax = fit_upr),
                colour = "grey", width = 0) +
  geom_path(aes(x = x_latitude, y = mu_pop_fit, colour = model), size = 0.7) +
  geom_point(data = d_y_yhat_long_sites %>%
               dplyr::filter(model != "Modified, only MIRS-based" & variable_y == "Klason lignin"), aes(y = fit_mean, x = x_latitude + ifelse(model == "Original", 0, 0.1), colour = model)) +
  scale_y_continuous(labels = function(x) x * 100) +
  scale_x_continuous(labels = function(x) round(x * x_latitude_scale + x_latitude_center, 0)) +
  labs(x = "Latitude [°]", y = "Content [mass-%]", title = "Klason lignin") +
  theme(legend.position = "bottom")

p_poc_latitude6 <- 
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# posterior predictive check with density plots for individual samples
p_poc_latitude7 <- 
  ppc_dens_overlay(y = m1.8_pred[, 1] %>%
                     magrittr::multiply_by(100), 
                   yrep = as.matrix(s1_res_holocellulose[, stringr::str_detect(colnames(s1_res_holocellulose), "Y1\\.")])[1:100, ] %>%
                     magrittr::multiply_by(100)) +
  labs(x = "Holocellulose content [mass-%]", y = "Density")

p_poc_latitude8 <- 
  ppc_dens_overlay(y = m2.8_pred[, 1] %>%
                     magrittr::multiply_by(100), 
                   yrep = as.matrix(s1_res_klason_lignin[, stringr::str_detect(colnames(s1_res_klason_lignin), "Y1\\.")])[1:100, ] %>%
                     magrittr::multiply_by(100)) +
  labs(x = "Klason lignin content [mass-%]", y = "Density")

# posterior predictive check with density plots for individual cores
p_poc_latitude10 <- 
  ppc_dens_overlay(y = 
                     d_peat[d_peat$sample_depth_lower <= 50, ] %>%
                     dplyr::mutate(y = m1.8_pred[, 1]) %>%
                     dplyr::group_by(core_label) %>%
                     dplyr::summarise(y = mean(y), .groups = "drop") %>%
                     dplyr::pull(y) %>%
                     magrittr::multiply_by(100),  
                   yrep = as.matrix(s1_res_holocellulose[, stringr::str_detect(colnames(s1_res_holocellulose), "mu_core\\.")])[1:100, ] %>%
                     magrittr::multiply_by(100)) +
  labs(x = "Hollocellulose content [mass-%]", y = "Density") +
  theme(legend.position = "bottom")

p_poc_latitude11 <- 
  ppc_dens_overlay(y = 
                     d_peat[d_peat$sample_depth_lower <= 50, ] %>%
                     dplyr::mutate(y = m2.8_pred[, 1]) %>%
                     dplyr::group_by(core_label) %>%
                     dplyr::summarise(y = mean(y), .groups = "drop") %>%
                     dplyr::pull(y) %>%
                     magrittr::multiply_by(100), 
                   yrep = as.matrix(s1_res_klason_lignin[, stringr::str_detect(colnames(s1_res_klason_lignin), "mu_core\\.")])[1:100, ] %>%
                     magrittr::multiply_by(100)) +
  labs(x = "Klason lignin content [mass-%]", y = "Density") +
  theme(legend.position = "bottom")

# plot of predicted versus fitted values
p_y_yhat <-                  
  d_y_yhat %>%                  
  ggplot(mapping = aes(y = Estimate, x = fit)) +
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), colour = "grey", width = 0) +
  geom_errorbarh(aes(xmin = lwr, xmax = upr), colour = "grey", height = 0) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  geom_point(colour = "dimgrey") +
  facet_wrap(~ variable_y) + 
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  coord_fixed() +
  labs(x = "Modified", y = "Modified, only MIRS-based")
```

# Criticizing our models for the latitudinal gradient

In spite of our suggested improvements, our model is far from perfect to describe the data. One issue is that there were `r hklmirs::format_numbers(sum(rstan::get_divergent_iterations(s1_fit_holocellulose)), capitalized = FALSE)` divergent `r ifelse(sum(rstan::get_divergent_iterations(s1_fit_holocellulose)) == 1, "transition", "transitions")` for the model for holocellulose which probably could be avoided by using tighter priors. Another issue is that the models for the latitudinal gradient provide only little information on the variability in average surface peat contents and on the variability of individual samples. For holocellulose, it is visible that the large uncertainty in predictions and average contents of cores (near-surface) results in strong shrinkage to the populations' average (figure \@ref(fig:reply-res-p-latitude-poc-shrinkage)). For Klason lignin, the distribution of individual samples' contents is modeled quite accurately relative to the only-MIRS-based reference model, but there is also strong shrinkage on the level of average core contents (figure \@ref(fig:reply-res-p-latitude-poc-shrinkage) (c)).

Shrinkage is generally an advantage of Bayesian hierarchical models because it improves the predictive performance of models [e.g. @Park.2004; @Lemoine.2019]. The amount of shrinkage here is not surprising because the predicted contents of individual samples come with large uncertainties and therefore relative large shrinkage does not contradict the distributions estimated for individual samples' contents. However, in this case, shrinkage is in our opinion also an indication for underfitting.  Underfitting occurs because the latitudinal gradient describes only a small proportion of the variability in peat surface holocellulose contents. The reason is that latitude is no causal factor and that known causal factors which control vegetation and decomposition can be quite diverse on regional to local scales (figure \@ref(fig:reply-res-p-latitude-poc-shrinkage)). Our analysis therefore shows that we need models which explain how these known causal factors control peat holocellulose and Klason lignin contents to accurately describe their patterns on larger spatial scales.

(ref:reply-res-p-latitude-poc-caption) Posterior predictive checks for the modified models describing the latitudinal gradient of peat core average near-surface holocellulose and Klason lignin contents. The individual plots show the same as figure \@ref(fig:reply-res-p-latitude-ppc), but now additionally with the draws from the posterior distribution for the models.

```{r reply-res-p-latitude-poc, echo=FALSE, out.width="70%", fig.height=5.5, fig.width=6.5, fig.cap='(ref:reply-res-p-latitude-poc-caption)'}
p_poc_latitude1 + p_poc_latitude2 + p_poc_latitude4 + p_poc_latitude3 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")")
```

(ref:reply-res-p-latitude-poc-slope-caption) Marginal prior and posterior distribution for the slopes for the latitudinal gradient.

```{r reply-res-p-latitude-poc-slope, echo=FALSE, out.width="70%", fig.height=3.5, fig.width=3.5, fig.cap='(ref:reply-res-p-latitude-poc-slope-caption)'}
p_poc_latitude5
```

(ref:reply-res-p-latitude-poc-shrinkage-caption) Shrinkage effects of the modified models describing the latitudinal gradient ("Modified") in comparison to the modified models using the MIRS only ("Modified, only MIRS-based"). (a) Predictions of the models for individual near-surface peat samples for all collected peat cores. As comparison, also the predictions of the original model (the Bayesian formulation of the models to predict holocellulose and Klason lignin contents of @Hodgkins.2018; "Original") are displayed. Points are average predictions and error bars 95 % prediction intervals. For holocellulose, shrinkage effects are visible both for the variability within cores and the variability across cores. (b, c) Posterior predictive densities of 100 random draws of average peat core near-surface contents ($y_{rep}$) in comparison to the average contents of the peat samples predicted with the modified model using only the MIRS ($y$). Due to shrinkage, all models produce a higher density of average peat core near-surface samples with average holocellulose and Klason lignin contents. (d) Plot of individual near-surface peat sample contents predicted by the modified models using the MIRS only versus the contents predicted by the modified models describing the latitudinal gradients. Points are average predictions and error bars 95 % prediction intervals. For holocellulose, shrinkage effects for samples with larger contents are visible.

```{r reply-res-p-latitude-poc-shrinkage, echo=FALSE, out.width="70%", fig.height=8.6, fig.width=6.5, fig.cap='(ref:reply-res-p-latitude-poc-shrinkage-caption)'}
p1 <- 
  p_poc_latitude_6_1 +
  plot_layout(guides = 'collect') &
  theme(legend.position = "bottom", legend.box="vertical")

p2 <- 
  (p_poc_latitude10 + p_poc_latitude11) +
  plot_layout(guides = 'collect') &
  theme(legend.position = "bottom", legend.box="vertical")

p3 <-
  p_y_yhat +
  plot_layout(guides = 'collect') &
  theme(legend.position = "bottom", legend.box="vertical")

p1 /p2 /p3 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") + 
  plot_layout(heights = c(2.5, 1, 1.7))
```

```{r m-reply-predictions, include=FALSE}
### aim: collect samples from the posterior distribution of the original and improved model for peat and vegetation samples such that key figures from Hodgkins.2018 can be reproduced

## peat data

# predictions
d_peat_pred_best <- 
  dplyr::bind_rows(
    # holocellulose
    m1.8 %>%
      rstanarm::posterior_predict(newdata = d_peat_bin20) %>%
      as.data.frame() %>%
      tidyr::pivot_longer(cols = dplyr::everything(), names_to = "sample_id", values_to = "y_hat") %>%
      dplyr::mutate(
        iter = seq_len(length(y_hat)/nrow(d_peat)) %>% rep(each = nrow(d_peat)),
        sample_id = d_peat$sample_id %>% rep(length(unique(iter))),
        model = "Best binned spectra",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      tidyr::pivot_longer(cols = dplyr::everything(), names_to = "sample_id", values_to = "y_hat") %>%
      dplyr::mutate(
        iter = seq_len(length(y_hat)/nrow(d_peat)) %>% rep(each = nrow(d_peat)),
        sample_id = d_peat$sample_id %>% rep(length(unique(iter))),
        model = "Original",
        variable_y = "Holocellulose"
      ),
    # klason lginin
    m2.8 %>%
      rstanarm::posterior_predict(newdata = d_peat_bin20) %>%
      as.data.frame() %>%
      tidyr::pivot_longer(cols = dplyr::everything(), names_to = "sample_id", values_to = "y_hat") %>%
      dplyr::mutate(
        iter = seq_len(length(y_hat)/nrow(d_peat)) %>% rep(each = nrow(d_peat)),
        sample_id = d_peat$sample_id %>% rep(length(unique(iter))),
        model = "Best binned spectra",
        variable_y = "Klason lignin"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      tidyr::pivot_longer(cols = dplyr::everything(), names_to = "sample_id", values_to = "y_hat") %>%
      dplyr::mutate(
        iter = seq_len(length(y_hat)/nrow(d_peat)) %>% rep(each = nrow(d_peat)),
        sample_id = d_peat$sample_id %>% rep(length(unique(iter))),
        model = "Original",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::left_join(d_peat %>% dplyr::select(-spectra), by = "sample_id") %>%
  dplyr::filter(iter <= 2000)

## summarize across depth values to reproduce key plots in Hodgkins et al. 2018
#d_peat_pred_best_summary <-
#  d_peat_pred_best %>%
#  dplyr::filter(sample_depth_lower <= 50) %>%
#  dplyr::group_by(variable_y, model, core_label) %>%
#  dplyr::summarise(fit = mean(y_hat),
#                   lwr = quantile(y_hat, probs = 0.05),
#                   upr = quantile(y_hat, probs = 0.95),
#                   sampling_latitude = mean(sampling_latitude),
#                   .groups = "drop") %>%
#  dplyr::mutate(
#    sampling_latitude = dplyr::case_when(model != "Original" ~ sampling_latitude + 3,
#                                         TRUE ~ sampling_latitude)
#    )

#d_peat_pred_best_summary1 <-
#  d_peat_pred_best %>%
#  dplyr::group_by(variable_y, model, core_label, sample_id) %>%
#  dplyr::summarise(fit = mean(y_hat),
#                   lwr = quantile(y_hat, probs = 0.05),
#                   upr = quantile(y_hat, probs = 0.95),
#                   sample_depth_lower = mean(sample_depth_lower),
#                   sampling_latitude = mean(sampling_latitude),
#                   .groups = "drop")

## compute average and prediction intervals for latitude surface peat contents linear model
#x_seq <- seq(min(d_peat_pred_best$sampling_latitude, na.rm = TRUE), max(d_peat_pred_best$sampling_latitude, na.rm = TRUE), by = 1)
#d_peat_pred_best_latitude <-
#  d_peat_pred_best %>%
#  dplyr::filter(iter <= 1000 & sample_depth_lower <= 50) %>%
#  dplyr::group_by(variable_y, model, iter, core_label) %>%
#  dplyr::summarise(y_hat = mean(y_hat),
#                   sampling_latitude = mean(sampling_latitude),
#                   .groups = "drop_last") %>%
#  dplyr::summarise(
#    lm_fit = {
#      m <- lm(y_hat ~ sampling_latitude, 
#                 data=tibble::tibble(y_hat = y_hat, sampling_latitude = sampling_latitude))
#      res <- predict(m, newdata = data.frame(sampling_latitude = x_seq), interval = "confidence", level = 0.95)
#      tibble::tibble(y_hat = res[, 1],
#                     lwr = res[, 2],
#                     upr = res[, 3],
#                     sampling_latitude = x_seq) %>% list()
#      }
#  ) %>%
#  tidyr::unnest(cols = dplyr::contains("lm_fit")) %>%
#  dplyr::group_by(variable_y, model, sampling_latitude) %>%
#  dplyr::summarise(fit = mean(y_hat),
#                   lwr = quantile(lwr, 0.025, na.rm = TRUE),
#                   upr = quantile(upr, 0.975, na.rm = TRUE),
#                   .groups = "drop")

#d_peat_pred_best_latitude_slopes <-
#  d_peat_pred_best %>%
#  dplyr::filter(iter <= 1000 & sample_depth_lower <= 50) %>%
#  dplyr::group_by(variable_y, model, iter, core_label) %>%
#  dplyr::summarise(y_hat = mean(y_hat),
#                   sampling_latitude = mean(sampling_latitude),
#                   .groups = "drop_last") %>%
#  dplyr::summarise(
#    lm_fit = {
#      m <- lm(y_hat ~ sampling_latitude, 
#                 data=tibble::tibble(y_hat = y_hat, sampling_latitude = sampling_latitude))
#      m <- summary(m)
#      tibble::tibble(intercept = coef(m)[1, 1],
#                     slope = coef(m)[2, 1],
#                     slope_p = coef(m)[2, 4]) %>% list()
#      }
#  ) %>%
#  tidyr::unnest(cols = dplyr::contains("lm_fit")) %>%
#  dplyr::group_by(variable_y, model) %>%
#  dplyr::summarise(intercept_mean = mean(intercept),
#                   slope_mean = mean(slope),
#                   slope_lwr = quantile(slope, 0.025, na.rm = TRUE),
#                   slope_upr = quantile(slope, 0.975, na.rm = TRUE),
#                   slope_p_mean = mean(slope_p),
#                   slope_p_lwr = quantile(slope_p, 0.025, na.rm = TRUE),
#                   slope_p_upr = quantile(slope_p, 0.975, na.rm = TRUE),
#                   .groups = "drop")

# compute average and prediction intervals for latitude depth profile LOESS smoother
x_seq <- seq(min(d_peat_pred_best$sample_depth_lower, na.rm = TRUE), max(d_peat_pred_best$sample_depth_lower, na.rm = TRUE), by = 5)
d_peat_pred_best_latitude_depth <-
  d_peat_pred_best %>%
  dplyr::filter(iter <= 1000) %>%
  dplyr::mutate(latitude_class = sampling_latitude > 45) %>%
  dplyr::group_by(variable_y, model, iter, latitude_class) %>%
  dplyr::summarise(
    loess_fit = {
      m <- loess(y_hat ~ sample_depth_lower, 
                 data=tibble::tibble(y_hat = y_hat, sample_depth_lower = sample_depth_lower),
                 span = 0.75, degree = 2, family = "gaussian", method = "loess", model = FALSE)
            res <- predict(m, newdata = data.frame(sample_depth_lower = x_seq), se = FALSE)
      tibble::tibble(y_hat = res,
                     sample_depth_lower = x_seq) %>% list()
      }
  ) %>%
  tidyr::unnest(cols = dplyr::contains("loess_fit")) %>%
  dplyr::group_by(variable_y, model, latitude_class, sample_depth_lower) %>%
  dplyr::summarise(fit = mean(y_hat),
                   lwr = quantile(y_hat, 0.025, na.rm = TRUE),
                   upr = quantile(y_hat, 0.975, na.rm = TRUE),
                   .groups = "drop")

# compute average differences in prediction between the models
d_peat_pred_best_diff <-
  d_peat_pred_best %>%
  dplyr::filter(model == "Best binned spectra" & iter <= 1000) %>%
  dplyr::select(variable_y, model, iter, sampling_latitude, sample_depth_lower, y_hat) %>%
  dplyr::mutate(y_hat_or = 
                  d_peat_pred_best %>% 
                  dplyr::filter(model == "Original" & iter <= 1000) %>%
                  dplyr::pull(y_hat),
                high_latitude = sampling_latitude > 45,
                surface_peat = sample_depth_lower <= 50,
                y_hat_diff = y_hat - y_hat_or)

d_peat_pred_best_diff_mean <-
  d_peat_pred_best_diff %>%
  dplyr::group_by(variable_y, model, iter, high_latitude, surface_peat) %>%
  dplyr::summarise(y_hat_diff = mean(y_hat_diff),
                   y_hat = mean(y_hat),
                   y_hat_or = mean(y_hat_or),
                   .groups = "drop")

d_peat_pred_best_diff_mean1 <-
  d_peat_pred_best_diff %>%
  dplyr::group_by(variable_y, model, iter, high_latitude, surface_peat) %>%
  dplyr::summarise(y_hat = mean(y_hat),
                   y_hat_or = mean(y_hat_or),
                   .groups = "drop") %>%
  dplyr::group_by(variable_y, model, iter, high_latitude) %>%
  dplyr::summarise(y_hat_diff = abs(y_hat[surface_peat] - y_hat[!surface_peat]),
                   y_hat_or_diff = abs(y_hat_or[surface_peat] - y_hat_or[!surface_peat]),
                   .groups = "drop")
```

```{r m-reply-predictions-veg, include=FALSE, eval=FALSE}
## vegetation data

# bin width: 20
d_veg_bin20_flat <-
  d_veg %>%
  ir::ir_bin(width = 50) %>%
  ir::ir_flatten()

d_veg_bin20 <- 
  d_veg_bin20_flat %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(center = d_peat_bin20_scale$center, 
        scale = d_peat_bin20_scale$scale) %>%
  as.data.frame()

# bin width: 10
d_veg_bin10_flat <-
  d_veg %>%
  ir::ir_bin(width = 10) %>%
  ir::ir_flatten()

d_veg_bin10 <- 
  d_veg_bin10_flat %>%
  dplyr::select(-1) %>%
  t() %>%
  as.data.frame() %>%
  scale(center = d_peat_bin10_scale$center, 
        scale = d_peat_bin10_scale$scale) %>%
  as.data.frame()

# predictions
d_veg_pred_best <- 
  dplyr::bind_rows(
    m1.8 %>%
      rstanarm::posterior_predict(newdata = d_veg_bin20) %>%
      as.data.frame() %>%
      tidyr::pivot_longer(cols = dplyr::everything(), 
                          names_to = "sample_id", 
                          values_to = "y_hat") %>%
      dplyr::mutate(
        iter = seq_len(length(y_hat)/nrow(d_veg)) %>% rep(each = nrow(d_veg)),
        sample_id = d_veg$sample_id %>% rep(length(unique(iter))),
        model = "Best binned spectra",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      tidyr::pivot_longer(cols = dplyr::everything(), 
                          names_to = "sample_id", 
                          values_to = "y_hat") %>%
      dplyr::mutate(
        iter = seq_len(length(y_hat)/nrow(d_veg)) %>% rep(each = nrow(d_veg)),
        sample_id = d_veg$sample_id %>% rep(length(unique(iter))),
        model = "Original",
        variable_y = "Holocellulose"
      ),
    m2.8 %>%
      rstanarm::posterior_predict(newdata = d_veg_bin10) %>%
      as.data.frame() %>%
      tidyr::pivot_longer(cols = dplyr::everything(), 
                          names_to = "sample_id", 
                          values_to = "y_hat") %>%
      dplyr::mutate(
        iter = seq_len(length(y_hat)/nrow(d_veg)) %>% rep(each = nrow(d_veg)),
        sample_id = d_veg$sample_id %>% rep(length(unique(iter))),
        model = "Best binned spectra",
        variable_y = "Klason lignin"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      tidyr::pivot_longer(cols = dplyr::everything(), 
                          names_to = "sample_id", 
                          values_to = "y_hat") %>%
      dplyr::mutate(
        iter = seq_len(length(y_hat)/nrow(d_veg)) %>% rep(each = nrow(d_veg)),
        sample_id = d_veg$sample_id %>% rep(length(unique(iter))),
        model = "Original",
        variable_y = "Klason lignin"
      ),
  ) %>%
  dplyr::left_join(d_veg %>% dplyr::select(-spectra), by = "sample_id") %>%
  dplyr::filter(iter <= 2000)

# classify sites so that they match the site-vegetation classification in Hodgkins et al. 2018
d_peat_pred_best <-
  d_peat_pred_best %>%
  dplyr::left_join(
    tibble::tibble(
      core_label = unique(d_peat_pred_best$core_label),
      barchart_label = c("Boreal fens", "Stordalen", "NC Pocosin", "NC Pocosin", "Loxahatchee", "Loxahatchee", "Boreal bogs", "Boreal bogs", "Mendaram", "Mendaram", "Boreal bogs", "Boreal fens", "Boreal bogs", "Boreal bogs")
    ),
    by = "core_label"
  )

d_veg_pred_best <-
  d_veg_pred_best %>% 
  dplyr::left_join(
    tibble::tibble(
      site_name = unique(d_veg_pred_best$site_name),
      barchart_label = c("Stordalen", "Boreal bogs", "Boreal fens", "NC Pocosin", "Loxahatchee", "Mendaram", "Mendaram")
    ),
    by = "site_name"
  )

# compute average values per site
d_peat_pred_best_turnover_summary <-
  d_peat_pred_best %>%
  dplyr::filter(sample_depth_lower <= 50) %>%
  dplyr::group_by(variable_y, model, barchart_label, iter) %>%
  dplyr::summarise(y_hat = mean(y_hat),
                   .groups = "drop") %>%
  dplyr::arrange(barchart_label, iter)

d_veg_pred_best_turnover_summary <-
  d_veg_pred_best %>%
  dplyr::group_by(variable_y, model, barchart_label, iter) %>%
  dplyr::summarise(y_hat = mean(y_hat),
                   .groups = "drop") %>%
  dplyr::arrange(barchart_label, iter)

d_turnover <-
  d_peat_pred_best_turnover_summary %>%
  dplyr::mutate(
    Peat = y_hat,
    Vegetation = d_veg_pred_best_turnover_summary$y_hat,
    Difference = Peat - Vegetation
  ) %>%
  dplyr::mutate(barchart_label = factor(barchart_label, levels = c("Stordalen", "Boreal bogs", "Boreal fens", "NC Pocosin", "Loxahatchee", "Mendaram") %>% rev())) %>%
  tidyr::pivot_longer(cols = dplyr::one_of(c("Peat", "Vegetation", "Difference")),
                      names_to = "sample_type",
                      values_to = "values") %>%
  dplyr::group_by(variable_y, model, barchart_label, sample_type) %>%
  dplyr::summarise(.groups = "drop",
                   fit = median(values) * 100,
                   lwr = quantile(values, prob = 0.025) * 100,
                   upr = quantile(values, prob = 0.975) * 100)
```

```{r m-reply-p-sites-latitude, include=FALSE, eval = FALSE}
### Reproduce plots in Hodgkins et al. 2018

## contents vs latitude across sites
p1 <-
  d_peat_pred_best_latitude %>%
  dplyr::filter(variable_y == "Holocellulose") %>%
  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
  ggplot(aes(y = fit * 100, ymin = lwr * 100, ymax = upr * 100,
             x = sampling_latitude, 
             colour = model, 
             fill = model)) +
  geom_path() +
  geom_ribbon(alpha = 0.2, colour = NA) +
  geom_errorbar(data = 
                  d_peat_pred_best_summary %>%
                  dplyr::filter(variable_y == "Holocellulose")  %>%
                  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
                  dplyr::mutate(sampling_latitude = ifelse(model == "Original", sampling_latitude, sampling_latitude + 2)),
                aes(y = fit * 100, ymin = lwr * 100, ymax = upr * 100, 
                    x = sampling_latitude, 
                    colour = model)) +
  geom_point(data = 
               d_peat_pred_best_summary %>%
               dplyr::filter(variable_y == "Holocellulose") %>%
               dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
               dplyr::mutate(sampling_latitude = ifelse(model == "Original", sampling_latitude, sampling_latitude + 2)),
             aes(y = fit * 100, x = sampling_latitude, colour = model)) +

  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = expression("Latitude ["*degree*N*"]")) +
  guides(colour = guide_legend(title = "Model"),
         fill = guide_legend(title = "Model"))

p2 <-
  d_peat_pred_best_latitude %>%
  dplyr::filter(variable_y == "Klason lignin") %>%
  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
  ggplot(aes(y = fit * 100, ymin = lwr * 100, ymax = upr * 100,
             x = sampling_latitude, 
             colour = model, 
             fill = model)) +
  geom_path() +
  geom_ribbon(alpha = 0.2, colour = NA) +
  geom_errorbar(data = 
                  d_peat_pred_best_summary %>%
                  dplyr::filter(variable_y == "Klason lignin") %>%
                  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
                  dplyr::mutate(sampling_latitude = ifelse(model == "Original", sampling_latitude, sampling_latitude + 2)),
                aes(y = fit * 100, ymin = lwr * 100, ymax = upr * 100, 
                    x = sampling_latitude, 
                    colour = model)) +
  geom_point(data = 
               d_peat_pred_best_summary %>%
               dplyr::filter(variable_y == "Klason lignin") %>%
               dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
               dplyr::mutate(sampling_latitude = ifelse(model == "Original", sampling_latitude, sampling_latitude + 2)),
             aes(y = fit * 100, x = sampling_latitude, colour = model)) +

  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = expression("Latitude ["*degree*N*"]")) +
  guides(colour = guide_legend(title = "Model"),
         fill = guide_legend(title = "Model"))

# combine
p_sites_latitude <- 
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r m-reply-p-depth-profiles-latitude, include=FALSE}
## plots of average depth profiles for two classes of latitudes

p1 <-
  d_peat_pred_best_latitude_depth %>%
  dplyr::filter(variable_y == "Holocellulose") %>%
  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
  dplyr::arrange(sample_depth_lower) %>%
  ggplot(aes(y = fit, 
             ymin = lwr,
             ymax = upr,
             x = sample_depth_lower, 
             colour = latitude_class, 
             fill = latitude_class)) +
  geom_hline(yintercept = c(0.11, 0.16, 0.21, 0.37), colour = "grey", linetype = 2) +
  geom_path() +
  geom_ribbon(alpha = 0.2, colour = NA) +
  coord_flip() +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ model, 
             scales = "free_y", 
             ncol = 1) +
  scale_x_reverse() +
  scale_y_continuous(labels = function(x) x * 100) +
  guides(colour = guide_legend(title = expression(Latitude>45~degree*N)),
         fill = guide_legend(title = expression(Latitude>45~degree*N)))

p2 <-
  d_peat_pred_best_latitude_depth %>%
  dplyr::filter(variable_y == "Klason lignin") %>%
  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
  dplyr::arrange(sample_depth_lower) %>%
  ggplot(aes(y = fit, 
             ymin = lwr,
             ymax = upr,
             x = sample_depth_lower, 
             colour = latitude_class, 
             fill = latitude_class)) +
  geom_hline(yintercept = c(0.22, 0.32, 0.35, 0.48), colour = "grey", linetype = 2) +
  geom_path() +
  geom_ribbon(alpha = 0.2, colour = NA) +
  coord_flip() +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ model, 
             scales = "free_y", 
             ncol = 1) +
  scale_x_reverse() + 
  scale_y_continuous(labels = function(x) x * 100) +
  guides(colour = guide_legend(title = expression(Latitude>45~degree*N)),
         fill = guide_legend(title = expression(Latitude>45~degree*N)))

# combine
p_depth_profiles_latitude <- 
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r m-reply-p-veg-peat-difference, eval=FALSE, include=FALSE}
## plots of difference between vegetation and surface peat contents
p1 <-
  d_turnover %>%
  dplyr::filter(variable_y == "Holocellulose") %>%
  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr,
             x = barchart_label,
             fill = sample_type)) +
  geom_errorbar(colour = "grey", width = 0.2, position = position_dodge(.9)) +
  geom_hline(yintercept = c(-0.2, 0, 0.2, 0.4) * 100, colour = "grey", linetype = 2) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Site category") +
  facet_wrap(~ model) +
  guides(fill = guide_legend(title = "")) +
  scale_fill_grey()

p2 <-
  d_turnover %>%
  dplyr::filter(variable_y == "Klason lignin") %>%
  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr,
             x = barchart_label,
             fill = sample_type)) +
  geom_hline(yintercept = c(-0.2, 0, 0.2, 0.4) * 100, colour = "grey", linetype = 2) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(colour = "grey", width = 0.2, position = position_dodge(.9)) +
  coord_flip() +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~"[mass-%]"),
       x = "Site category") +
  facet_wrap(~ model) +
  guides(fill = guide_legend(title = "")) +
  scale_fill_grey()

# combine plots
p_veg_peat_difference <-
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r reply-2, echo=FALSE, include=FALSE, eval = TRUE, cache = FALSE}
rmarkdown::render("002-reply-main.Rmd", clean = FALSE)
```

### 2022-03-03 analysis/paper/002-reply-main.Rmd
Initial commit for manuscript 2.

---
title: 'Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging'
author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
---

```{r setup-reply-2, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", 
                      fig.pos = "H",
                      echo = FALSE,
                      cache = TRUE)
```

$^1$ ILÖK, Ecohydrology and Biogeochemistry Group, University of Münster, Heisenbergstr. 2, 48149 Münster, Germany  
$^*$ Corresponding author: Henning Teickner (henning.teickner@uni-muenster.de)

The resilience of the large global peatland carbon stocks against climate change may depend on how peat chemistry controls decomposition rates [@Wang.2015; @Hodgkins.2018]. In their study, @Hodgkins.2018 argue that there is a latitudinal gradient in peat recalcitrance which may explain the stability of tropical peat in spite of warmer temperatures and which points toward a susceptibility of labile surface peat at high latitudes to degradation, but not deeper peat which has a chemistry more similar to tropical peat. This gradient can be described by differences in holocellulose contents --- representing relatively easily degradable carbohydrates --- and Klason lignin contents --- representing aromatics --- which @Hodgkins.2018 predicted from mid infrared spectra [@Hodgkins.2018].  
We acknowledge this conceptual model as an important step in understanding the stability of peat deposits and their susceptibility to degradation. However, we challenge the accuracy of the models used to predict peat holocellulose and Klason lignin contents. This inaccuracy has two major implications:

1. Peat Klason lignin contents and relative accumulation during decomposition probably are underestimated by @Hodgkins.2018 which may have important consequences for decomposition of high-latitude peat deposits due to climate change.

2. The inaccuracies imply that the predicted values are only very rough estimates for holocellulose and Klason lignin contents and at least partly biased. An interpretation of the values given in @Hodgkins.2018 in terms of quantitative concepts for organic matter chemistry (i.e. holocellulose and Klason lignin) therefore is not possible.

Instead, we assume that if the study is reinterpreted in terms of qualitative concepts for carbohydrates and aromatics as used in the qualitative interpretation of mid infrared spectra for peat, there still is evidence for general latitudinal patterns in peat recalcitrance. However, interpretation and generalization of these findings are less straightforward. Moreover, we suggest improvements for uncertainty propagation for the models for latitudinal and temperature gradients. Remaining open questions to this approach should be addressed in future studies.

## Patterns in peat chemistry should not be interpreted in terms of holocellulose and Klason lignin contents {.unnumbered}

In a recent evaluation of the models to predict peat holocellulose and Klason lignin contents presented in @Hodgkins.2018, we found that the training data used for the models (leaves, paper products) are not representative for peat [---todo: ref]. The peat samples are more decomposed than the training data samples. Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [---todo:ref]. Since these features are not present in the training data, predictions of the original model are extrapolations. Following our re-evaluation, this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >~30 mass-% [---todo:ref]. Because of these uncertainties and biases, the analyses of @Hodgkins.2018 cannot be interpreted in terms of absolute holocellulose and Klason lignin contents.  
Instead, we suggest to interpret differences in peat chemistry presented by @Hodgkins.2018 not in terms of specific organic matter fractions (i.e. holocellulose and Klason lignin), but rather in terms of more qualitative concepts typically used in studies interpreting mid infrared spectra of peat (e.g. @Artz.2008; @Broder.2012; @Tfaily.2014; @Biester.2014), i.e. as "carbohydrates" and "aromatics". We assume that under such an interpretation, the conclusions of @Hodgkins.2018 probably are valid.  
The proposed reinterpretation is possible if values given in @Hodgkins.2018 are only interpreted in terms of the normalized peak heights used in the original prediction models, and not as absolute holocellulose and Klason lignin contents. Each model is a linear model with one independent variable (the normalized height of one peak); therefore a direct reinterpretation of the predicted absolute contents in terms of normalized peak heights is possible if the figure axes are relabeled as "Relative content of carbohydrates/aromatics" instead of "Estimated % carbohydrates/aromatics" (implying an absolute content of holocellulose and Klason lignin, respectively) [@Hodgkins.2018].   
An important open question to this approach of interpreting large-scale gradients in peat chemistry based on mid infrared spectra should be addressed in future studies: Are differences in specific mid infrared spectra peaks reliable indicators for differences in a specific concept for peat chemistry? For example, are differences in the peak around 1040 cm$^{-1}$ indicative for differences in carbohydrate contents (also across larger scales)?  
A similar problem arose in the discussion on the use of the C/N ratio as a good indicator for peat mass loss. This may only hold true if there have been no or negligible shifts in the vegetation composition which causes differences in the C/N ratio already in initial plant litter (e.g. @Biester.2014). For mid infrared spectra, such knowledge still has to be acquired in future studies, especially if large environmental gradients are considered as in @Hodgkins.2018.

## Klason lignin content and residual enrichment in (sub)tropical peatlands probably are larger {.unnumbered}

We suggest that @Hodgkins.2018 have underestimated peat Klason lignin contents, especially in (sub)tropical peat. As described above, the original model for Klason lignin underestimates peat Klason lignin contents. A modified model has less bias [---todo:ref]. In spite of several remaining limitations (see below and the supplementary information), we argue that this modified model can serve as rough indicator for the underestimation of peat Klason lignin contents by the original model. We therefore present values to provide tentative estimates for the underestimation of Klason lignin contents, but we do not claim that these estimates are optimal.  
With this modified model, near-surface ($\le50$ cm) peat Klason lignin contents are on average ~10 to 15 mass-% larger across the latitudinal gradient (figure \@ref(fig:reply-res-p-sites-latitude)). Specifically, average (sub)tropical near-surface peat ($<45 ^\circ$N, $\le50$ cm) Klason lignin contents are `r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% (median, lower and upper 95% prediction interval limit) larger than with the original model. There is also tentative indication for on average larger differences between deep (sub)tropical ($<45 ^\circ$N) and high-latitude ($\ge45 ^\circ$N) peat (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).    
We also suggest that the residual enrichment of Klason lignin during decomposition probably has been underestimated. A rough estimate for the residual enrichment of Klason lignin during decomposition is the difference in Klason lignin content between near-surface peat ($\le50$ cm) and deeper peat. With the modified model, this difference is on average for (sub)tropical peatlands `r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% larger (figure \@ref(fig:reply-res-p-depth-profiles-latitude)). For high latitude peatlands the difference is smaller on average (`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-%).  
Consequently, in general --- and especially in (sub)tropical peatlands --- the average residual enrichment of Klason lignin due to decomposition probably has been underestimated by the original model. A consequence of this in light of climate change is that high latitude peat deposits may be more prone to shifts in decomposition rates than suggested by the original results.

## Larger uncertainties due to a lack of uncertainty propagation and unrepresentative data {.unnumbered}

Besides a reinterpretation in terms of qualitative concepts for organic matter recalcitrance, we also believe that a more comprehensive quantification and propagation of uncertainties is important to further scrutinize the results of @Hodgkins.2018. In the original analysis, the following uncertainties are omitted: (1) As described above, the training dataset is not representative for peat and therefore predictions are biased. This uncertainty is not considered in the analyses of @Hodgkins.2018. (2) The second source of uncertainty which was omitted is the prediction uncertainty of the models used to predict holocellulose and Klason lignin contents. (3) A third source of uncertainty not considered yet is that in computing the regression model to analyze the latitudinal gradient based on average core near-surface contents, the uncertainty for these average values was not propagated.  
The first two sources of uncertainty can be omitted if only normalized peak heights (as qualitative indicators) are used instead of predicted holocellulose and Klason lignin contents, as described above. But the third source of uncertainty should also be considered when reinterpreting the results in terms of area normalized peak heights.  
To support our concerns, we reanalyzed the latitudinal gradient for holocellulose and Klason lignin, respectively, using the modified models described in [---todo:ref] and fully considering the last two sources of uncertainty using a Bayesian approach (described in more detail in the supplementary information). With the modified models, the estimated slopes for the latitudinal gradient are more uncertain (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->). A similar reanalysis of the depth profiles between the two latitude categories shows that differences are also more uncertain (figure \@ref(fig:reply-res-p-depth-profiles-latitude)). Our analysis therefore indicates larger uncertainties both in the depth profiles and the latitudinal gradient. Considering these uncertainties is important when peat properties predicted with a model are analyzed.
Again, the values presented here are only tentative because the modified models probably are also biased.

## Future directions {.unnumbered}

Notwithstanding our critical remarks, the idea of @Hodgkins.2018 to relate mid infrared spectra to precisely quantifiable and defined organic matter fractions is important. Such fractions are relatively well understood and therefore allow us to better deduce what effects climate shifts can have on peat decomposition and recalcitrance. It is much more informative if we are able to precisely estimate a change in Klason lignin contents of e.g. 5 mass-% if average annual temperatures increase by 1$^{\circ}$C than if we are only able to make qualitative statements that the relative amount of aromatics (without describing in detail what this fraction means) increases, as is currently the case.  
When interpreting mid infrared spectra, distinguishing between precisely quantifiable concepts for organic matter chemistry, such as holocellulose and Klason lignin contents, and qualitative concepts which currently are used in interpreting peat MIRS, such as "carbohydrates and aromatics" (without specifying how this fraction could actually be measured), is crucial: Models relying on quantitative concepts are more testable and allow to actually quantify fractional changes in organic matter chemistry, whereas models built on qualitative concepts can only be refuted if there are blatant discrepancies relative to expectations (i.e. a decrease in "carbohydrates" with latitude). Ultimately, we need more precise models for peat variations in chemistry than can be developed based on qualitative concepts.  
We suggest the following steps to gain deeper insights into latitudinal gradient in peat chemistry: As outlined above, we suggest to interpret the results of @Hodgkins.2018 in terms of qualitative concepts for peat chemistry, i.e. "carbohydrates" and "aromatics". These suggestions also apply to a follow-up study using the same models [@Verbeke.2022]. Further studies are needed to check if specific peaks are good indicators for these qualitative concepts for peat chemistry, how the relation may be confounded e.g. by vegetation composition or other organic matter fractions, as could be the case especially for peaks related to aromatics [---todo:ref]. This could be done by computing models to reliably predict holocellulose and Klason lignin contents of peat from mid infrared spectra and analyzing which spectral variables are relevant for this purpose and if there are confounding variables. Such models will also help us to get more precise insights into the degree to which high latitude peat accumulation may be decreased by decomposition due to rising temperatures, because they allow replacing more costly measurements. Finally, future studies should carefully consider uncertainties and fully propagate these into model estimates. Predicting absolute peat holocellulose and Klason lignin contents from infrared spectra remains challenging.

(ref:reply-res-p-sites-latitude-caption) Predicted average surface ($\le 50$ cm) peat holocellulose (a) and Klason lignin (b) contents plotted against latitude (compare with Fig. 3 in @Hodgkins.2018). Lines and shaded areas represent average predictions from regression models and 95% confidence intervals. "Modified" is the modified Bayesian hierarchical regression model which simultaneously models individual samples' contents from mid infrared spectra and the latitudinal gradient of average core near-surface peat contents. This model considers prediction uncertainty from the mid infrared spectra and from computing per-core averages. "Original" is the original linear regression model [@Hodgkins.2018] computing only the latitudinal gradient of average core near-surface contents. This model does not consider prediction uncertainty from mid infrared spectra, nor uncertainty from computing per-core averages. Points are average core near-surface contents predicted from the model ("Modified") or computed from the average predictions for individual samples ("Original") with error bars representing 95% confidence intervals. Points for "Modified" are shifted by $+0.1^\circ$.

```{r reply-res-p-sites-latitude, echo=FALSE, out.width="70%", fig.height=3.5, fig.width=6.5, fig.cap='(ref:reply-res-p-sites-latitude-caption)'}
p_poc_latitude6
```

(ref:reply-res-p-depth-profiles-latitude-caption) Predicted median holocellulose (A) and Klason lignin (B) depth profiles of peat core data classified into two latitude categories following @Hodgkins.2018 (compare with Fig. 2 in @Hodgkins.2018). Lines are averages of LOESS smoothers fitted to the predicted values by the models. Shaded regions are corresponding 95% confidence intervals, comprising prediction uncertainty in the holocellulose and Klason lignin estimates, respectively. Vertical dashed lines in each columns represent approximate surface and average mean contents as predicted using the improved models ("Modified").

```{r reply-res-p-depth-profiles-latitude, echo=FALSE, warning=FALSE, out.width="70%", fig.height=5.5, fig.width=6.5, fig.cap='(ref:reply-res-p-depth-profiles-latitude-caption)'}
p_depth_profiles_latitude
```

```{r reply-res-p-veg-peat-differnece, echo=FALSE, eval=FALSE, out.width="90%", fig.height=3.5, fig.width=7.5, fig.cap='Predicted surface ($\\le 50$ cm) peat and vegetation median holocellulose (A) and Klason lignin (B) contents for the peat cores shown in figure \\ref{fig:res-p-gaussian-beta-depth-profile-best} grouped according to \\cite{Hodgkins.2018}. Predictions of the (biased) original models are equivalent to the data shown in \\cite{Hodgkins.2018} (figure 5). Averages are median values of predictions, error bars are 95\\% prediction intervals. "Difference" is the difference between predictions for vegetation and peat.'}
p_veg_peat_difference
```

# Acknowledgements {.unnumbered}

We like to thank Jeffrey Chanton and Suzanne Hodgkins for a constructive dicussion which greatly helped to improve a first version of this manuscript.

# Competing interests {.unnumbered}

The authors declare no competing interests.

# Author contributions {.unnumbered}

HT performed the calculations, prepared the figures, and wrote the original text. Both HT and KHK revised and edited the text. KHK provided funding for this study.

### 2022-03-03 README.Rmd
Updated formatting.
Update .gitignore, Dockerfile, renv.lock, and the compendium package description and readme.

@@
 
-This repository contains the data and code for our submitted manuscript:
+This repository contains the data and code for our two manuscripts (in preparation):
 
-> `r paste(manuscript1$authors, "(submitted)", manuscript1$title)`
+> `r paste(manuscript1$authors, "(in preparation):", manuscript1$title)`
+
+> `r paste(manuscript2$authors, "(in preparation):", manuscript2$title)`
 
@@
 > Henning Teickner and Klaus-Holger Knorr, (`r format(Sys.Date(), "%Y")`).
-_Compendium of R code and data for "`r manuscript1_yaml$title`"_.
+_Compendium of R code and data for "`r manuscript1_yaml$title`" and "`r manuscript2_yaml$title`"_.
 Accessed `r format(Sys.Date(), "%d %b %Y")`.
@@
 
-  - [:file\_folder: paper](/analysis/paper): R Markdown source documents
-    needed to reproduce the manuscript, including figures and tables. The main 
-    script is [001-paper-main.Rmd](analysis/paper/001-paper-main.Rmd). Additional scripts are:
+- [:file\_folder: paper](/analysis/paper): R Markdown source documents needed to reproduce the manuscript, including figures and tables.
+The main script is [001-paper-main.Rmd](analysis/paper/001-paper-main.Rmd).
+This script produces both manuscripts and the corresponding supplementary information.
+Additional scripts are:
     - [002-paper-m-original-models.Rmd](analysis/paper/002-paper-m-original-models.Rmd): Computes the original models used in @Hodgkins.2018 and models with the same model structure, but as Bayesian models.
@@
     - [007-paper-m-prediction-differences.Rmd](analysis/paper/007-paper-m-prediction-differences.Rmd): Compares predictions for the training data and the peat and vegetation data from @Hodgkins.2018 for the original models from @Hodgkins.2018 and the modified models from `004-paper-m-reduce-underfitting.Rmd`.
-        - [008-paper-supplementary.Rmd](analysis/paper/008-paper-supplementary.Rmd): Computes supplementary analyses and figures.
+    - [008-paper-supplementary.Rmd](analysis/paper/008-paper-supplementary.Rmd): Computes supplementary analyses and figures for the first manuscript.
+    - [001-reply-main.Rmd](analysis/paper/001-reply-main.Rmd): This is the main script for manuscript 2.
+It is run from within `001-paper-main.Rmd` and produces the supplementary information for manuscript 2.
+    - [002-reply-main.Rmd](analysis/paper/002-reply-main.Rmd): This script produces the document for manuscript 2. It is run from within `001-reply-main.Rmd`.
+
 - [:file\_folder: data](/analysis/data): Data used in the analysis.
@@
 The raw data are derived from @Hodgkins.2018.
-  - [:file\_folder: figures](/analysis/figures): Plots and other
-    illustrations (empty folder).
+
+- [:file\_folder: stan_models](/analysis/stan_models): The Stan model used in `001-reply-main.Rmd`.
 
@@
 - `R`, `man`, `inst`, `data-raw`, `data`, `src`: Default folders for making the R package run.
-
 - Folder `inst/extdata`: Folder with the raw data used for the analyses. All files in this folder are derived from @Hodgkins.2018.
@@
 
-Running the whole script takes about two hours and occupies additional disk space of ~2 Gb.
+Running the whole script takes about 12 hours and occupies additional disk space of ~2 Gb.
 
@@
 
-The [Dockerfile](Dockerfile) provides instructions how to build a Docker image from the Dockerfile and how to run the image in a Docker container. It occupies disk space of ~5 Gb.
+The [Dockerfile](Dockerfile) provides instructions how to build a Docker image from the Dockerfile and how to run the image in a Docker container. It occupies disk space of ~6 Gb.
 

### 2022-03-03 README.Rmd
Update readme, update Dockerfile.

@@
 manuscript1_yaml <- rmarkdown::yaml_front_matter("analysis/paper/001-paper-main.Rmd")
-manuscript2_yaml <- rmarkdown::yaml_front_matter("analysis/paper/001-reply-main.Rmd")
+manuscript2_yaml <- rmarkdown::yaml_front_matter("analysis/paper/002-reply-main.Rmd")
 
@@
   paste0(".")
+
+# manuscript2
+manuscript2 <- list()
+manuscript2$authors <- "Henning Teickner and Klaus-Holger Knorr"
+
+manuscript2$year <- 
+  paste0("(",  eval(parse(text = stringr::str_extract(manuscript2_yaml$date, pattern = "^\\d+"))), "):")
+
+manuscript2$title <-
+  manuscript2_yaml$title %>%
+  paste0(".")
+
+manuscript2$journal <-
+  eval(parse(text = manuscript2_yaml$journal %>% stringr::str_remove("^`r ") %>% stringr::str_remove("`$"))) %>%
+  names()
+
+manuscript2$doi <-
+  manuscript2_yaml$doi %>%
+  paste0(".")
 ```
@@
 Accessed `r format(Sys.Date(), "%d %b %Y")`.
-Online at <https://doi.org/xxx/xxx>
+Online at <https://github.com/henningte/hklmirs/>
 
@@
 
-You can download the compendium as a zip from from this URL: <https://doi.org/> ---todo
+You can download the compendium as a zip from from this URL: <https://github.com/henningte/hklmirs/>
 
@@
 
-The [Dockerfile](Dockerfile) provides instructions how to build a Docker image from the Dockerfile and how to run the image in a Docker container. It occupies disk space of ~6 Gb.
+The [Dockerfile](Dockerfile) provides instructions how to build a Docker image from the Dockerfile and how to run the image in a Docker container. It occupies disk space of ~7 Gb.
 

### 2022-03-03 analysis/paper/001-reply-main.Rmd
Cleanup.

@@
 
-## summarize across depth values to reproduce key plots in Hodgkins et al. 2018
-#d_peat_pred_best_summary <-
-#  d_peat_pred_best %>%
-#  dplyr::filter(sample_depth_lower <= 50) %>%
-#  dplyr::group_by(variable_y, model, core_label) %>%
-#  dplyr::summarise(fit = mean(y_hat),
-#                   lwr = quantile(y_hat, probs = 0.05),
-#                   upr = quantile(y_hat, probs = 0.95),
-#                   sampling_latitude = mean(sampling_latitude),
-#                   .groups = "drop") %>%
-#  dplyr::mutate(
-#    sampling_latitude = dplyr::case_when(model != "Original" ~ sampling_latitude + 3,
-#                                         TRUE ~ sampling_latitude)
-#    )
-
-#d_peat_pred_best_summary1 <-
-#  d_peat_pred_best %>%
-#  dplyr::group_by(variable_y, model, core_label, sample_id) %>%
-#  dplyr::summarise(fit = mean(y_hat),
-#                   lwr = quantile(y_hat, probs = 0.05),
-#                   upr = quantile(y_hat, probs = 0.95),
-#                   sample_depth_lower = mean(sample_depth_lower),
-#                   sampling_latitude = mean(sampling_latitude),
-#                   .groups = "drop")
-
-## compute average and prediction intervals for latitude surface peat contents linear model
-#x_seq <- seq(min(d_peat_pred_best$sampling_latitude, na.rm = TRUE), max(d_peat_pred_best$sampling_latitude, na.rm = TRUE), by = 1)
-#d_peat_pred_best_latitude <-
-#  d_peat_pred_best %>%
-#  dplyr::filter(iter <= 1000 & sample_depth_lower <= 50) %>%
-#  dplyr::group_by(variable_y, model, iter, core_label) %>%
-#  dplyr::summarise(y_hat = mean(y_hat),
-#                   sampling_latitude = mean(sampling_latitude),
-#                   .groups = "drop_last") %>%
-#  dplyr::summarise(
-#    lm_fit = {
-#      m <- lm(y_hat ~ sampling_latitude, 
-#                 data=tibble::tibble(y_hat = y_hat, sampling_latitude = sampling_latitude))
-#      res <- predict(m, newdata = data.frame(sampling_latitude = x_seq), interval = "confidence", level = 0.95)
-#      tibble::tibble(y_hat = res[, 1],
-#                     lwr = res[, 2],
-#                     upr = res[, 3],
-#                     sampling_latitude = x_seq) %>% list()
-#      }
-#  ) %>%
-#  tidyr::unnest(cols = dplyr::contains("lm_fit")) %>%
-#  dplyr::group_by(variable_y, model, sampling_latitude) %>%
-#  dplyr::summarise(fit = mean(y_hat),
-#                   lwr = quantile(lwr, 0.025, na.rm = TRUE),
-#                   upr = quantile(upr, 0.975, na.rm = TRUE),
-#                   .groups = "drop")
-
-#d_peat_pred_best_latitude_slopes <-
-#  d_peat_pred_best %>%
-#  dplyr::filter(iter <= 1000 & sample_depth_lower <= 50) %>%
-#  dplyr::group_by(variable_y, model, iter, core_label) %>%
-#  dplyr::summarise(y_hat = mean(y_hat),
-#                   sampling_latitude = mean(sampling_latitude),
-#                   .groups = "drop_last") %>%
-#  dplyr::summarise(
-#    lm_fit = {
-#      m <- lm(y_hat ~ sampling_latitude, 
-#                 data=tibble::tibble(y_hat = y_hat, sampling_latitude = sampling_latitude))
-#      m <- summary(m)
-#      tibble::tibble(intercept = coef(m)[1, 1],
-#                     slope = coef(m)[2, 1],
-#                     slope_p = coef(m)[2, 4]) %>% list()
-#      }
-#  ) %>%
-#  tidyr::unnest(cols = dplyr::contains("lm_fit")) %>%
-#  dplyr::group_by(variable_y, model) %>%
-#  dplyr::summarise(intercept_mean = mean(intercept),
-#                   slope_mean = mean(slope),
-#                   slope_lwr = quantile(slope, 0.025, na.rm = TRUE),
-#                   slope_upr = quantile(slope, 0.975, na.rm = TRUE),
-#                   slope_p_mean = mean(slope_p),
-#                   slope_p_lwr = quantile(slope_p, 0.025, na.rm = TRUE),
-#                   slope_p_upr = quantile(slope_p, 0.975, na.rm = TRUE),
-#                   .groups = "drop")
-
 # compute average and prediction intervals for latitude depth profile LOESS smoother
@@
                    .groups = "drop")
-```
-
-```{r m-reply-predictions-veg, include=FALSE, eval=FALSE}
-## vegetation data
-
-# bin width: 20
-d_veg_bin20_flat <-
-  d_veg %>%
-  ir::ir_bin(width = 50) %>%
-  ir::ir_flatten()
-
-d_veg_bin20 <- 
-  d_veg_bin20_flat %>%
-  dplyr::select(-1) %>%
-  t() %>%
-  as.data.frame() %>%
-  scale(center = d_peat_bin20_scale$center, 
-        scale = d_peat_bin20_scale$scale) %>%
-  as.data.frame()
-
-# bin width: 10
-d_veg_bin10_flat <-
-  d_veg %>%
-  ir::ir_bin(width = 10) %>%
-  ir::ir_flatten()
-
-d_veg_bin10 <- 
-  d_veg_bin10_flat %>%
-  dplyr::select(-1) %>%
-  t() %>%
-  as.data.frame() %>%
-  scale(center = d_peat_bin10_scale$center, 
-        scale = d_peat_bin10_scale$scale) %>%
-  as.data.frame()
-
-# predictions
-d_veg_pred_best <- 
-  dplyr::bind_rows(
-    m1.8 %>%
-      rstanarm::posterior_predict(newdata = d_veg_bin20) %>%
-      as.data.frame() %>%
-      tidyr::pivot_longer(cols = dplyr::everything(), 
-                          names_to = "sample_id", 
-                          values_to = "y_hat") %>%
-      dplyr::mutate(
-        iter = seq_len(length(y_hat)/nrow(d_veg)) %>% rep(each = nrow(d_veg)),
-        sample_id = d_veg$sample_id %>% rep(length(unique(iter))),
-        model = "Best binned spectra",
-        variable_y = "Holocellulose"
-      ),
-    m1.2 %>%
-      rstanarm::posterior_predict(newdata = d_veg_res) %>%
-      as.data.frame() %>%
-      tidyr::pivot_longer(cols = dplyr::everything(), 
-                          names_to = "sample_id", 
-                          values_to = "y_hat") %>%
-      dplyr::mutate(
-        iter = seq_len(length(y_hat)/nrow(d_veg)) %>% rep(each = nrow(d_veg)),
-        sample_id = d_veg$sample_id %>% rep(length(unique(iter))),
-        model = "Original",
-        variable_y = "Holocellulose"
-      ),
-    m2.8 %>%
-      rstanarm::posterior_predict(newdata = d_veg_bin10) %>%
-      as.data.frame() %>%
-      tidyr::pivot_longer(cols = dplyr::everything(), 
-                          names_to = "sample_id", 
-                          values_to = "y_hat") %>%
-      dplyr::mutate(
-        iter = seq_len(length(y_hat)/nrow(d_veg)) %>% rep(each = nrow(d_veg)),
-        sample_id = d_veg$sample_id %>% rep(length(unique(iter))),
-        model = "Best binned spectra",
-        variable_y = "Klason lignin"
-      ),
-    m2.2 %>%
-      rstanarm::posterior_predict(newdata = d_veg_res) %>%
-      as.data.frame() %>%
-      tidyr::pivot_longer(cols = dplyr::everything(), 
-                          names_to = "sample_id", 
-                          values_to = "y_hat") %>%
-      dplyr::mutate(
-        iter = seq_len(length(y_hat)/nrow(d_veg)) %>% rep(each = nrow(d_veg)),
-        sample_id = d_veg$sample_id %>% rep(length(unique(iter))),
-        model = "Original",
-        variable_y = "Klason lignin"
-      ),
-  ) %>%
-  dplyr::left_join(d_veg %>% dplyr::select(-spectra), by = "sample_id") %>%
-  dplyr::filter(iter <= 2000)
-
-# classify sites so that they match the site-vegetation classification in Hodgkins et al. 2018
-d_peat_pred_best <-
-  d_peat_pred_best %>%
-  dplyr::left_join(
-    tibble::tibble(
-      core_label = unique(d_peat_pred_best$core_label),
-      barchart_label = c("Boreal fens", "Stordalen", "NC Pocosin", "NC Pocosin", "Loxahatchee", "Loxahatchee", "Boreal bogs", "Boreal bogs", "Mendaram", "Mendaram", "Boreal bogs", "Boreal fens", "Boreal bogs", "Boreal bogs")
-    ),
-    by = "core_label"
-  )
-
-d_veg_pred_best <-
-  d_veg_pred_best %>% 
-  dplyr::left_join(
-    tibble::tibble(
-      site_name = unique(d_veg_pred_best$site_name),
-      barchart_label = c("Stordalen", "Boreal bogs", "Boreal fens", "NC Pocosin", "Loxahatchee", "Mendaram", "Mendaram")
-    ),
-    by = "site_name"
-  )
-
-# compute average values per site
-d_peat_pred_best_turnover_summary <-
-  d_peat_pred_best %>%
-  dplyr::filter(sample_depth_lower <= 50) %>%
-  dplyr::group_by(variable_y, model, barchart_label, iter) %>%
-  dplyr::summarise(y_hat = mean(y_hat),
-                   .groups = "drop") %>%
-  dplyr::arrange(barchart_label, iter)
-
-d_veg_pred_best_turnover_summary <-
-  d_veg_pred_best %>%
-  dplyr::group_by(variable_y, model, barchart_label, iter) %>%
-  dplyr::summarise(y_hat = mean(y_hat),
-                   .groups = "drop") %>%
-  dplyr::arrange(barchart_label, iter)
-
-d_turnover <-
-  d_peat_pred_best_turnover_summary %>%
-  dplyr::mutate(
-    Peat = y_hat,
-    Vegetation = d_veg_pred_best_turnover_summary$y_hat,
-    Difference = Peat - Vegetation
-  ) %>%
-  dplyr::mutate(barchart_label = factor(barchart_label, levels = c("Stordalen", "Boreal bogs", "Boreal fens", "NC Pocosin", "Loxahatchee", "Mendaram") %>% rev())) %>%
-  tidyr::pivot_longer(cols = dplyr::one_of(c("Peat", "Vegetation", "Difference")),
-                      names_to = "sample_type",
-                      values_to = "values") %>%
-  dplyr::group_by(variable_y, model, barchart_label, sample_type) %>%
-  dplyr::summarise(.groups = "drop",
-                   fit = median(values) * 100,
-                   lwr = quantile(values, prob = 0.025) * 100,
-                   upr = quantile(values, prob = 0.975) * 100)
-```
-
-```{r m-reply-p-sites-latitude, include=FALSE, eval = FALSE}
-### Reproduce plots in Hodgkins et al. 2018
-
-## contents vs latitude across sites
-p1 <-
-  d_peat_pred_best_latitude %>%
-  dplyr::filter(variable_y == "Holocellulose") %>%
-  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
-  ggplot(aes(y = fit * 100, ymin = lwr * 100, ymax = upr * 100,
-             x = sampling_latitude, 
-             colour = model, 
-             fill = model)) +
-  geom_path() +
-  geom_ribbon(alpha = 0.2, colour = NA) +
-  geom_errorbar(data = 
-                  d_peat_pred_best_summary %>%
-                  dplyr::filter(variable_y == "Holocellulose")  %>%
-                  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
-                  dplyr::mutate(sampling_latitude = ifelse(model == "Original", sampling_latitude, sampling_latitude + 2)),
-                aes(y = fit * 100, ymin = lwr * 100, ymax = upr * 100, 
-                    x = sampling_latitude, 
-                    colour = model)) +
-  geom_point(data = 
-               d_peat_pred_best_summary %>%
-               dplyr::filter(variable_y == "Holocellulose") %>%
-               dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
-               dplyr::mutate(sampling_latitude = ifelse(model == "Original", sampling_latitude, sampling_latitude + 2)),
-             aes(y = fit * 100, x = sampling_latitude, colour = model)) +
-
-  labs(title = "Holocellulose", 
-       y = expression(Holocellulose~content~"[mass-%]"),
-       x = expression("Latitude ["*degree*N*"]")) +
-  guides(colour = guide_legend(title = "Model"),
-         fill = guide_legend(title = "Model"))
-
-p2 <-
-  d_peat_pred_best_latitude %>%
-  dplyr::filter(variable_y == "Klason lignin") %>%
-  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
-  ggplot(aes(y = fit * 100, ymin = lwr * 100, ymax = upr * 100,
-             x = sampling_latitude, 
-             colour = model, 
-             fill = model)) +
-  geom_path() +
-  geom_ribbon(alpha = 0.2, colour = NA) +
-  geom_errorbar(data = 
-                  d_peat_pred_best_summary %>%
-                  dplyr::filter(variable_y == "Klason lignin") %>%
-                  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
-                  dplyr::mutate(sampling_latitude = ifelse(model == "Original", sampling_latitude, sampling_latitude + 2)),
-                aes(y = fit * 100, ymin = lwr * 100, ymax = upr * 100, 
-                    x = sampling_latitude, 
-                    colour = model)) +
-  geom_point(data = 
-               d_peat_pred_best_summary %>%
-               dplyr::filter(variable_y == "Klason lignin") %>%
-               dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
-               dplyr::mutate(sampling_latitude = ifelse(model == "Original", sampling_latitude, sampling_latitude + 2)),
-             aes(y = fit * 100, x = sampling_latitude, colour = model)) +
-
-  labs(title = "Klason lignin", 
-       y = expression(Klason~lignin~content~"[mass-%]"),
-       x = expression("Latitude ["*degree*N*"]")) +
-  guides(colour = guide_legend(title = "Model"),
-         fill = guide_legend(title = "Model"))
-
-# combine
-p_sites_latitude <- 
-  p1 + p2 + 
-  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
-  plot_layout(guides = "collect") &
-  theme(legend.position = "bottom")
 ```
@@
 
-```{r m-reply-p-veg-peat-difference, eval=FALSE, include=FALSE}
-## plots of difference between vegetation and surface peat contents
-p1 <-
-  d_turnover %>%
-  dplyr::filter(variable_y == "Holocellulose") %>%
-  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
-  ggplot(aes(y = fit, ymin = lwr, ymax = upr,
-             x = barchart_label,
-             fill = sample_type)) +
-  geom_errorbar(colour = "grey", width = 0.2, position = position_dodge(.9)) +
-  geom_hline(yintercept = c(-0.2, 0, 0.2, 0.4) * 100, colour = "grey", linetype = 2) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  coord_flip() +
-  labs(title = "Holocellulose", 
-       y = expression(Holocellulose~content~"[mass-%]"),
-       x = "Site category") +
-  facet_wrap(~ model) +
-  guides(fill = guide_legend(title = "")) +
-  scale_fill_grey()
-
-p2 <-
-  d_turnover %>%
-  dplyr::filter(variable_y == "Klason lignin") %>%
-  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
-  ggplot(aes(y = fit, ymin = lwr, ymax = upr,
-             x = barchart_label,
-             fill = sample_type)) +
-  geom_hline(yintercept = c(-0.2, 0, 0.2, 0.4) * 100, colour = "grey", linetype = 2) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(colour = "grey", width = 0.2, position = position_dodge(.9)) +
-  coord_flip() +
-  labs(title = "Klason lignin", 
-       y = expression(Klason~lignin~"[mass-%]"),
-       x = "Site category") +
-  facet_wrap(~ model) +
-  guides(fill = guide_legend(title = "")) +
-  scale_fill_grey()
-
-# combine plots
-p_veg_peat_difference <-
-  p1 + p2 + 
-  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
-  plot_layout(guides = "collect") &
-  theme(legend.position = "bottom")
-```
-
 ```{r reply-2, echo=FALSE, include=FALSE, eval = TRUE, cache = FALSE}

### 2022-03-03 analysis/paper/003-paper-m-gaussian-beta.Rmd
Cleanup.

@@
 m2.3_loo <- loo(m2.3, k_threshold = k_threshold)
-
-# ---todo: brms and rstanarm yield very different uncertainty estimates.
-brms has smaller prediction intervals.
-The reason seems to be the different priors for phi (normal(0, 2.5) for rstanarm; gamma(0.01, 0.01) for brms).
 ```

### 2022-03-03 README.Rmd
Add DOI for zenodo upload.

@@
 
+[![DOI](https://zenodo.org/badge/465653387.svg)](https://zenodo.org/badge/latestdoi/465653387)
+
 # hklmirs
@@
 Accessed `r format(Sys.Date(), "%d %b %Y")`.
-Online at <https://github.com/henningte/hklmirs/>
+Online at <https://doi.org/10.5281/zenodo.6325760>
 
@@
 
-You can download the compendium as a zip from from this URL: <https://github.com/henningte/hklmirs/>
+You can download the compendium as a zip from from these URLs: <https://github.com/henningte/hklmirs/> or <https://doi.org/10.5281/zenodo.6325760>
 

### 2022-03-07 README.Rmd
Add funding information.

@@
 
-[![DOI](https://zenodo.org/badge/465653387.svg)](https://zenodo.org/badge/latestdoi/465653387)
+[![DOI](https://zenodo.org/badge/465653387.svg)](https://zenodo.org/badge/latestdoi/465653387) ![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)
 
@@
 
+### Funding
+
+This study was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no.
+KN 929/23-1 to Klaus-Holger Knorr and grant no.
+PE 1632/18-1 to Edzer Pebesma.
+We acknowledge support from the Open Access Publication Fund of the University of Münster.
+
 ### References

### 2022-03-07 analysis/paper/001-paper-main.Rmd
Add DOI for zenodo upload.
Add funding information.

@@
   codedata: |
-    Data and code to reproduce our analyzes are available via [---todo: add link].
+    Data and code to reproduce our analyses are available via <https://doi.org/10.5281/zenodo.6325760>.
 #videosupplement: |
@@
 We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions which helped to improve a first version of this manuscript.
+This study was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no.
+KN 929/23-1 to Klaus-Holger Knorr and grant no.
+PE 1632/18-1 to Edzer Pebesma.
 We acknowledge support from the Open Access Publication Fund of the University of Münster.

### 2022-03-07 analysis/paper/001-paper-main.Rmd
Update `renv.lock`, update figure format in supporting info for manuscript 1. Update the text. Update the readme.

@@
     To understand global soil organic matter (SOM) chemistry and its dynamics, we need tools to efficiently quantify SOM properties, for example prediction models using mid infrared spectra.
-The advantages of such models rely on their validity and accuracy.
-Recently, \cite{Hodgkins.2018} developed models to predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major fractions of organic matter in general.
-The models can be used to understand large-scale SOM gradients and have been used in various studies.
+However, the advantages of such models rely on their validity and accuracy.
+Recently, \cite{Hodgkins.2018} developed models to quantitatively predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major fractions of organic matter.
+The models may have the potential to understand large-scale SOM gradients and have been used in various studies.
     A research gap to fill is that these models have not been validated in detail yet.
 What are their limitations and how can we improve them?
-This study provides a validation with the aim to identify concrete steps to improve the models.
+This study provides a validation with the aim to identify concrete steps to improve these models.
 As a first step, we provide several improvements using the original training data.
-    The major limitation we identified is that the original training data are not representative for a range of peat samples which causes both biased estimates and extrapolation uncertainty under the original models.
+    The major limitation we identified is that the original training data are not representative for a range of diverse peat samples.
+This causes both biased estimates and extrapolation uncertainty under the original models.
 In addition, the original models can in practice produce unrealistic predictions (negative values or values $>100$ mass-%).
-Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid unrealistic predictions.
+Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid such unrealistic predictions.
 Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples.
@@
 combustion and gas chromatographic analysis of resulting gases.
-All the advantages of MIRS-based models rely on the accuracy of the computed models.
+However, all the advantages of MIRS-based models rely on the accuracy of the computed models.
 This makes model validation a crucial step during model development.
@@
 Most of these models consider only wood, material from few species, or specific vegetation organs [@Elle.2019].
-Another problem is that most of these models cannot be easily reproduced and used since the raw data has not been published.
+Another problem is that most of these models cannot be easily reproduced and used since the raw data and code have not been published.
 Recently, @Hodgkins.2018 developed models to predict Klason lignin and holocellulose contents from attenuated total reflectance MIRS.
@@
 Moreover, both raw data and model code are publicly available [@Hodgkins.2018; @Teickner.2020a].
-For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020; @Cong.2022], most recently in @Verbeke.2022, and for future developments, e.g.
+For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020; @Cong.2022], most recently in @Verbeke.2022 and @Baysinger.2022, and for future developments, e.g.
 by including additional data.
-A problem is that neither @Hodgkins.2018, nor a later study using the models provided a thorough model validation.
+A major problem is that neither @Hodgkins.2018, nor a later study using the models provided a thorough model validation.
 To compute the models, the authors developed a procedure to extract peaks which are indicative for specific OM fractions, such as holocellulose and Klason lignin (Fig.
@@
 As model validation, only the linearity between the target variables and predictors was assessed (supplementary material to @Hodgkins.2018), but not for example how predicted values match measured values.
-Moreover, no information has been provided if the training data is representative for SOM, particularly the peat samples analyzed in the study.
-Similarly, it is unclear if pre-selecting peaks may reduce the predictive accuracy of the models.
+Moreover, no information has been provided whether the training data is representative for SOM, particularly the peat samples analyzed in the study.
+Furthermore, it is unclear if pre-selecting peaks may reduce the predictive accuracy of the models.
 Many of these concerns have also been raised by an anonymous reviewer of the paper (supplementary information to @Hodgkins.2018).
@@
 Our goals are to identify limitations of the original models and to give concrete recommendations for improvements.
-Moreover, we use the original data from @Hodgkins.2018 to provide improvements where possible.
+Moreover, we use the original data from @Hodgkins.2018 to provide improvements on the original models where possible.
 To this end, we conducted an exploratory analysis using the peat and peat-forming vegetation data provided by @Hodgkins.2018.
@@
 3.
-_Is the prediction domain of the training data representative for peat samples?_ If a model extrapolates outside the range of predictor variable values in the training data set, it is unclear if the predictions are valid, for instance because it cannot be evaluated if there is bias.
+_Is the prediction domain of the training data representative for peat samples?_ If a model extrapolates outside the range of predictor variable values in the training data set, it is unclear if the predictions are valid.
 An advantage of the independent reference sample set used by @Hodgkins.2018 is that the model is applicable on various OM types and e.g.
@@
 
-Our aim is to provide a validation of the original models of @Hodgkins.2018 and concrete suggestions for improvements.
-To this end, we conducted a series of statistical analyzes to query each of our research questions listed in the introduction.
+We conducted a series of statistical analyzes to query each of our research questions listed in the introduction.
 During this, we also computed improved models using the same training data that was used to compute the original models.
@@
 Whenever we refer to "original model", we mean these Bayesian translations, and otherwise refer to "original non-Bayesian models".
-To show that we did not introduce any of the identified weaknesses using the Bayesian approach, we compared the predictions of the Bayesian models to the original non-Bayesian models (supplementary Fig.
+To check that we did not introduce any of the identified weaknesses using the Bayesian approach, we compared the predictions of the Bayesian models to the original non-Bayesian models (supplementary Fig.
 1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).
@@
 We were interested if and under which conditions unrealistic predictions can occur.
-Here, we differentiate two ways in which a prediction can be unrealistic: It can be an unrealistic point estimate (median predictions), and the prediction interval can cover values outside [0, 100] mass-%, even though the median prediction can be in [0, 100] mass-%.
+Here, we differentiate two ways in which a prediction can be unrealistic: It can be either an unrealistic point estimate (median predictions), or the prediction interval can cover values outside [0, 100] mass-%, even though the median prediction can be in [0, 100] mass-%.
 For our analysis, we first computed a sequence of normalized `carb` and `arom15arom16` peak heights such that the original model predict values completely covering the interval [0, 100] mass-% for holocellulose and Klason lignin, respectively.
@@
 
-The original model for holocellulose produced unrealistic predictions: It had a negative point prediction for one strongly decomposed sample from a tropical peat core.
+The original model for holocellulose indeed produced unrealistic predictions: It had a negative point prediction for one strongly decomposed sample from a tropical peat core.
 Moreover, for `r d_peat_pred %>% dplyr::filter(variable_y == "Holocellulose" & model == "Gaussian, Bayesian") %>% dplyr::mutate(negative = lwr < 0) %>% dplyr::summarise(negative = sum(negative)/length(negative)) %>% dplyr::pull(negative) %>% magrittr::multiply_by(100) %>% round(0)`% of the peat samples in the data from @Hodgkins.2018, the lower 90% prediction interval covered negative values (supplementary Fig.
@@
 This can be concluded based on the average estimated ELPD values (larger values indicate a better estimated predictive accuracy).
-Consequently, using the data available, we could improve both the model for holocellulose and Klason lignin.
+Consequently, using the data available, we could improve both the model for holocellulose and the model for Klason lignin.
 Is one of the strategies to include more variables advantageous?
@@
 \@ref(fig:res-p-coefficients-fig03), last panel).
-We have no explanation why one coefficient is negative and the other positive.
+We have no explanation for why one coefficient is negative and the other positive.
 In Sect.
 3.5<!--"What causes differences in predictions?"-->, we provide a mechanistic interpretation for why the selected variables are important.
-These selected bins are not represented in the extracted peaks.
+The bins with especially large absolute coefficients in the models using binned spectra are not represented in the extracted peaks because these cover different wavenumber ranges (compare Fig.
+\@ref(fig:res-p-coefficients-fig03) to Fig.
+\@ref(fig:hodgkins-peaks-sample-fig01)).
 Consequently, models using all extracted peaks cannot use this information and tend to underfit.
@@
 
-The training data set does not cover the range of the spectral variables relevant for predicting peat and peat forming vegetation holocellulose contents (Fig.
+The training data do not cover the range of the spectral variables relevant for predicting peat and peat forming vegetation holocellulose contents (Fig.
 \@ref(fig:res-p-prediction-domain-fig04) (a) and (b)).
@@
 However, few peat samples have larger `arom15arom16` and many peat samples have smaller `carb` peak heights than covered by the training data.
-This suggests that the original models extrapolate to predict holocellulose contents in peat samples with small holocellulose contents and Klason lignin contents in peat samples with large Klason lignin contents.
+This suggests that the original models extrapolate holocellulose contents in peat samples with small holocellulose contents and also extrapolate Klason lignin contents in peat samples with large Klason lignin contents.
 The same is true for the improved models using binned spectra (Fig.
@@
 This is also true for Klason lignin since extrapolation occurs for samples with large `arom15arom16` peak, or for the binned spectra, large absorbance at ~`r m_pars %>% dplyr::filter(abs(median) >= 0.02 & variable_y == "Klason lignin") %>% dplyr::pull(wn) %>% sort() %>% knitr::combine_words()` cm$^{-1}$, all of which are larger for more decomposed peat [@Cocozza.2003; @Tfaily.2014; @Hodgkins.2018].
-Overall, this indicates that the prediction domain formed by the training data does not cover peat --- particularly decomposed peat --- and partly peat forming vegetation samples.
-This is probably also true for non-peat SOM.
+Overall, this indicates that the prediction domain formed by the training data does not cover the range needed for peat --- particularly decomposed peat --- and partly also does not cover the range needed for peat forming vegetation samples.
+We assume that this is probably also true for non-peat SOM.
 Therefore the models' predictions can represent extrapolations in practice; the training data is not in general representative for peat and peat-forming vegetation.
@@
 6 <!---\@ref(fig:res-p-y-yhat-best)--->).
-Thus, predictions are different even within the range covered by the training data where the models actually make similar predictions and these differences are not due to misfit of either model to the training data.
+Thus, predictions are different for the peat samples even within the holocellulose and Klason lignin range covered by the training data, where the models actually make similar predictions for the training data.
+And these differences are not due to misfit of either model to the training data.
 This indicates that the spectra of the training data are not entirely representative for the peat samples, even if they are within the prediction domain (compare also with the previous section).
@@
 For holocellulose, we could not find indications which model is better.
-For Klason lignin, we provide mechanistic evidence for why the prediction of the improved models probably are more correct.
-Therefore, a key result of our analysis is that additional training and validation data is required to compute models to predict SOM holocellulose and Klason lignin contents from MIRS.
+For Klason lignin, we provide mechanistic evidence for why the predictions of the improved models probably are more correct.
+Therefore, a key result of our analysis is that additional training and validation data is required to compute models to accurately predict SOM holocellulose and Klason lignin contents from MIRS.
 
@@
 7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).
-For decomposed OM, the same `carb` peak height or absorbance at specific wavenumbers therefore represents different holocellulose contents than for the training data.
-A consequence therefore is that it is questionable if the models (both original and improved) can be applied to peat and other SOM samples in general.
+For decomposed OM, the same normalized `carb` peak height or absorbance at specific wavenumbers therefore represents different holocellulose contents than for the training data, as the comparatively broad OH-peak strongly influences the normalization of the spectra.
+A consequence therefore is that it is questionable if the models (both original and improved) can be applied to peat and other SOM samples in general, without adapting the training data by including more representative samples.
 
@@
 \@ref(fig:res-p-pred-diff-train1-p2-fig06)).
-The model balances this, but in doing so it introduces the observed bias.
+The original model balances this, but in doing so it introduces the observed bias.
 As a consequence, `arom15arom16` alone is a poor predictor for Klason lignin content of OM.
@@
 
-## What can we learn in general from this study?
+## General implications of our results
 
@@
 Second, a linear relation between the target variable to predict and a predictor variable is no sufficient validation of a spectral prediction model if the training data is not representative for the samples on which the model is applied.
-Most importantly, due to normalization, predictions can even be sensitive to variables not included in the model.
+Most importantly, due to spectral normalization, predictions can even be sensitive to variables not included in the model.
 Therefore, to assess if training data (the prediction domain) is representative, whole spectra have to be compared.
@@
 We therefore suggest that continuous re-evaluation of existing models can be an effective way to develop better models and should be part of a general model development workflow [@Gelman.2020].
-
-## What are weaknesses of our analyses and opportunities for future improvements?
-
 There are several problems we could not solve.
@@
 This results in biased predictions for holocellulose and Klason lignin for SOM, such as peat.
+Results from currently published studies using the original models should be interpreted with caution.
 Manual variable selection favored this bias because it excluded information critical for better model fits.

### 2022-06-01 analysis/paper/001-reply-main.Rmd
Add missing references in manuscript 2 to manuscript 1 (available as preprint).

@@
 The models described in @Hodgkins.2018 to predict holocellulose and Klason lignin contents from mid infrared spectra, but recomputed in a Bayesian framework.
-This is the "original Bayesian model" in [---todo:ref].
+This is the "original Bayesian model" in @Teickner.2022a.
 These models were used to reproduce the computations in @Hodgkins.2018.
@@
 A modified model to predict each, holocellulose and Klason lignin contents from mid infrared spectra.
-These are the models called "best binned spectra" models in [---todo:ref].
+These are the models called "best binned spectra" models in @Teickner.2022a.
 A brief description of these models is given in section [Computation of modified models to predict organic matter holocellulose and Klason lignin contents from mid infrared spectra].
@@
 
-Computation of the modified models using only the mid infrared spectra (MIRS) is described in detail in [---todo:ref].
+Computation of the modified models using only the mid infrared spectra (MIRS) is described in detail in @Teickner.2022a.
 In brief, we used the same training data as @Hodgkins.2018 and a Bayesian linear regression model which we computed with Markov Chain Monte Carlo (MCMC) sampling using brms [@Burkner.2017; @Burkner.2018] and rstan [@StanDevelopmentTeam.2020] which are both interfaces to Stan [@StanDevelopmentTeam.2021].
@@
 1.
-Since representative training data are missing, both the original and modified models cannot be fully validated and prediction uncertainties are probably larger than computed with these models [---todo:ref].
+Since representative training data are missing, both the original and modified models cannot be fully validated and prediction uncertainties are probably larger than computed with these models [@Teickner.2022a].
 Since representative test data are missing, these additional uncertainties cannot be estimated.
@@
 The modified models using only the MIRS were also used to compute median values and confidence intervals mentioned in the main text in section "Klason lignin content and residual enrichment in (sub)tropical peatlands probably are larger".
-In these computations, we also considered the predictive uncertainty of the original models (the Bayesian formulation described in ---todo:ref).
+In these computations, we also considered the predictive uncertainty of the original models (the Bayesian formulation described in @Teickner.2022a).
 
@@
 To create figure 1<!--\@ref(fig:reply-res-p-sites-latitude)--> in the main text, we used the models for the latitudinal gradient and samples from the posterior distribution to compute the predicted average and 95% confidence interval for a range of latitude values and we added predictions for the average contents for each of the sampled cores.
-Next, we recomputed the models from @Hodgkins.2018 by (1) predicting average peat sample contents with the original models (the Bayesian version described in ---todo:ref), (2) averaging the near surface peat sample average contents for each core, and (3) fitting an ordinary least squares regression line to these average core values.
+Next, we recomputed the models from @Hodgkins.2018 by (1) predicting average peat sample contents with the original models (the Bayesian version described in @Teickner.2022a), (2) averaging the near surface peat sample average contents for each core, and (3) fitting an ordinary least squares regression line to these average core values.
 For this linear model, we plotted the confidence interval across latitude and we added the computed average core near-surface contents.
 Thus, figure 1<!---\@ref(fig:reply-res-p-sites-latitude)---> in the main text compares the graphed values and model prediction from figure 3 in @Hodgkins.2018 with predictions from our modified analysis considering prediction uncertainty from the MIRS-based model part, and the averaging per core.
-Additional sources for differences in predictions and prediction uncertainties in comparison to the analysis in @Hodgkins.2018 are the use of a beta distribution which constraints uncertainty intervals for extreme (near 0 and 100 mass-%) values, and the use of all spectral variables which resulted in reduced prediction uncertainties [---todo:ref].
+Additional sources for differences in predictions and prediction uncertainties in comparison to the analysis in @Hodgkins.2018 are the use of a beta distribution which constraints uncertainty intervals for extreme (near 0 and 100 mass-%) values, and the use of all spectral variables which resulted in reduced prediction uncertainties [@Teickner.2022a].
 

### 2022-06-01 analysis/paper/002-reply-main.Rmd
Add DOI for zenodo upload.
Add funding information.
Add missing references in manuscript 2 to manuscript 1 (available as preprint).

@@
 
-In a recent evaluation of the models to predict peat holocellulose and Klason lignin contents presented in @Hodgkins.2018, we found that the training data used for the models (leaves, paper products) are not representative for peat [---todo: ref].
+In a recent evaluation of the models to predict peat holocellulose and Klason lignin contents presented in @Hodgkins.2018, we found that the training data used for the models (leaves, paper products) are not representative for peat [@Teickner.2022a].
 The peat samples are more decomposed than the training data samples.
-Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [---todo:ref].
+Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [@Teickner.2022a].
 Since these features are not present in the training data, predictions of the original model are extrapolations.
-Following our re-evaluation, this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >~30 mass-% [---todo:ref].
+Following our re-evaluation, this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >~30 mass-% [@Teickner.2022a].
 Because of these uncertainties and biases, the analyses of @Hodgkins.2018 cannot be interpreted in terms of absolute holocellulose and Klason lignin contents.
@@
 As described above, the original model for Klason lignin underestimates peat Klason lignin contents.
-A modified model has less bias [---todo:ref].
+A modified model has less bias [@Teickner.2022a].
 In spite of several remaining limitations (see below and the supplementary information), we argue that this modified model can serve as rough indicator for the underestimation of peat Klason lignin contents by the original model.
@@
 But the third source of uncertainty should also be considered when reinterpreting the results in terms of area normalized peak heights.
-To support our concerns, we reanalyzed the latitudinal gradient for holocellulose and Klason lignin, respectively, using the modified models described in [---todo:ref] and fully considering the last two sources of uncertainty using a Bayesian approach (described in more detail in the supplementary information).
+To support our concerns, we reanalyzed the latitudinal gradient for holocellulose and Klason lignin, respectively, using the modified models described in @Teickner.2022a and fully considering the last two sources of uncertainty using a Bayesian approach (described in more detail in the supplementary information).
 With the modified models, the estimated slopes for the latitudinal gradient are more uncertain (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->).
@@
 Further studies are needed to check if specific peaks are good indicators for these qualitative concepts for peat chemistry, how the relation may be confounded e.g.
-by vegetation composition or other organic matter fractions, as could be the case especially for peaks related to aromatics [---todo:ref].
+by vegetation composition or other organic matter fractions, as could be the case especially for peaks related to aromatics [@Teickner.2022a].
 This could be done by computing models to reliably predict holocellulose and Klason lignin contents of peat from mid infrared spectra and analyzing which spectral variables are relevant for this purpose and if there are confounding variables.
@@
 We like to thank Jeffrey Chanton and Suzanne Hodgkins for a constructive dicussion which greatly helped to improve a first version of this manuscript.
+This study was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no.
+KN 929/23-1 to Klaus-Holger Knorr and grant no.
+PE 1632/18-1 to Edzer Pebesma.
 
@@
 HT performed the calculations, prepared the figures, and wrote the original text. Both HT and KHK revised and edited the text. KHK provided funding for this study.
+
+# Data and Code availability {.unnumbered}
+
+Data and code to reproduce our analyses are available via <https://doi.org/10.5281/zenodo.6325760.>

### 2022-07-24 analysis/paper/001-reply-main.Rmd
Update manuscript 2.

@@
 Y_{2}[j] &\sim \text{beta}(\mu_{2}[j] \phi, (1-\mu_{2}[j])\phi)\\
-\mu_{1} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}\right)\\
-\mu_{2} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_1\right)\\
+\mu_{1} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_1\right)\\
+\mu_{2} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_2\right)\\
 \alpha &\sim \text{normal(0, 0.2)}\\
 \beta &\sim \text{horseshoe}(zb, hs_{local}, hs_{global}, hs_{scale\_slab}^2 \cdot hs_{slab})\\
-\phi &\sim \text{beta}(`r s1_data_holocellulose$phi_a`, `r s1_data_holocellulose$phi_b/s1_data_holocellulose$phi_scale`)\\
+\phi &\sim \text{gamma}(`r s1_data_holocellulose$phi_a`, `r s1_data_holocellulose$phi_b/s1_data_holocellulose$phi_scale`)\\
 Y_{2}[j] &\sim \text{beta}(\mu_{\text{core}}[\text{index\_core}[j]] \phi_{\text{core}}[\text{index\_core}[j]], (1-\mu_{\text{core}}[\text{index\_core}[j]])\phi_{\text{core}}[\text{index\_core}[j]])\\
@@
 
-Where $\text{Y}_1$ are the measured holocellulose contents of the training data; $\mu_{1}$ are the respective mean values modeled with the MIRS, only; $i$ is an index for each sample of the training data; $Y_{2}$ are the estimated peat holocellulose contents modeled simultaneously from the MIRS ($\mu_{2}$) and from the average average near-surface peat holocellulose contents per core ($\mu_{\text{core}}$); $j$ is an index for each sample in the peat data; $\text{index\_core}$ is a vector assigning each sample in the peat data to the respective core index; $k$ is an index for each unique core in the peat data.
-The average average near-surface peat holocellulose contents per core ($\mu_{\text{core}}$) are in turn modeled in dependency on the latitude.
-
-The full model for Klason lignin is:
+All variables and parameters with subscript "1" refer to the training data for the MIRS prediction model (diverse organic matter samples, including leave, wood, and paper products):
+
+- $\text{Y}_1$ are the measured holocellulose contents of the training data.
+- $\mu_{1}$ are the respective average values modeled with MIRS.
+- $\text{X}_1$ is a matrix containing the MIRS variables (variables in columns and samples in rows) for the training data. 
+- $i$ is an index for each sample of the training data.
+- $\phi$ is the precision parameter for the Beta distribution with which values of $\text{Y}_1$ are modeled.
+- $\alpha$ and $\beta$ are intercept and regression coefficients for the MIRS prediction model.
+The $\beta$s are modeled via a regularized horseshoe prior parameterized following @Piironen.2017b as described in @Teickner.2022a.
+
+All variables and parameters with subscript "2" refer to the peat samples on which the MIRS prediciton model is applied to describe the latitudinal gradient:
+
+- $Y_{2}$ are the estimated peat holocellulose contents modeled simultaneously from the MIRS (via $\mu_{2}$ --- this is the same model as for $\mu_{1}$) and from the average average near-surface peat holocellulose contents per core ($\mu_{\text{core}}$).
+- $\text{X}_2$ is a matrix containing the MIRS variables (variables in columns and samples in rows) for the peat samples. 
+- $\mu_{\text{core}}$ are average average near-surface peat holocellulose contents per core and is assumed to follow a Beta distribution which average value is modeled in dependency of latitude.
+- $\mu_{\text{pop}}$ are average near-surface peat holocellulose contents computed for each latitude.
+- $j$ is an index for each sample in the peat data. 
+- $\text{index\_core}$ is a vector assigning each sample in the peat data to the respective core index. 
+- $k$ is an index for each unique core in the peat data. 
+- $\alpha_{\text{latitude}}$ and $\beta_{\text{latitude}}$ are intercept and regression coefficient of the linear model with which the latitudinal gradient in average near-surface peat holocellulose contents are described.
+- $x_{\text{latitude}}[k]$ is the latitude for core $k$.
+- $\phi_{\text{pop}}$ and $\phi_{\text{core}}$ are precision parameters for the gamma distributions for the average core near surface average holocellulose content and the core near surface average holocellulose contents, respectively.
+
+The full model for Klason lignin is analogous to that for holocellulose:
 
@@
 \beta &\sim \text{horseshoe}(zb, hs_{local}, hs_{global}, hs_{scale\_slab}^2 \cdot hs_{slab})\\
-\phi &\sim \text{beta}(`r s1_data_klason_lignin$phi_a`, `r s1_data_klason_lignin$phi_b/s1_data_klason_lignin$phi_scale`)\\
+\phi &\sim \text{gamma}(`r s1_data_klason_lignin$phi_a`, `r s1_data_klason_lignin$phi_b/s1_data_klason_lignin$phi_scale`)\\
 Y_{2}[j] &\sim \text{beta}(\mu_{\text{core}}[\text{index\_core}[j]] \phi_{\text{core}}[\text{index\_core}[j]], (1-\mu_{\text{core}}[\text{index\_core}[j]])\phi_{\text{core}}[\text{index\_core}[j]])\\

### 2022-07-24 analysis/paper/002-reply-main.Rmd
Update manuscript 2.

@@
 1.
-Peat Klason lignin contents and relative accumulation during decomposition probably are underestimated by @Hodgkins.2018 which may have important consequences for decomposition of high-latitude peat deposits due to climate change.
+Peat Klason lignin contents and relative accumulation during decomposition probably are underestimated by @Hodgkins.2018.
+This may have important consequences for decomposition of high-latitude peat deposits due to climate change.
 
@@
 holocellulose and Klason lignin) therefore is not possible.
-
 Instead, we assume that if the study is reinterpreted in terms of qualitative concepts for carbohydrates and aromatics as used in the qualitative interpretation of mid infrared spectra for peat, there still is evidence for general latitudinal patterns in peat recalcitrance.
 However, interpretation and generalization of these findings are less straightforward.
-Moreover, we suggest improvements for uncertainty propagation for the models for latitudinal and temperature gradients.
-Remaining open questions to this approach should be addressed in future studies.
+
+Moreover, we suggest improvements for uncertainty propagation for the models.
+Remaining open questions to this approach of analyzing global peat chemistry patterns should be addressed in future studies.
 
@@
 Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [@Teickner.2022a].
-Since these features are not present in the training data, predictions of the original model are extrapolations.
-Following our re-evaluation, this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >~30 mass-% [@Teickner.2022a].
+Since these features are not present in the training data, predictions of the original models are extrapolations.
+Our re-evaluation showed that this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >30 mass-% [@Teickner.2022a].
 Because of these uncertainties and biases, the analyses of @Hodgkins.2018 cannot be interpreted in terms of absolute holocellulose and Klason lignin contents.
-Instead, we suggest to interpret differences in peat chemistry presented by @Hodgkins.2018 not in terms of specific organic matter fractions (i.e.
-holocellulose and Klason lignin), but rather in terms of more qualitative concepts typically used in studies interpreting mid infrared spectra of peat (e.g.
-@Artz.2008; @Broder.2012; @Tfaily.2014; @Biester.2014), i.e.
-as "carbohydrates" and "aromatics".
+Instead, we suggest to interpret differences in peat chemistry presented by @Hodgkins.2018 not in terms of the quantitative concepts (i.e.
+holocellulose and Klason lignin), but rather in terms of more qualitative concepts typically used in studies interpreting mid infrared spectra of peat: as "carbohydrates" and "aromatics" (e.g.
+@Artz.2008; @Broder.2012; @Tfaily.2014; @Biester.2014).
 We assume that under such an interpretation, the conclusions of @Hodgkins.2018 probably are valid.
@@
 
-Besides a reinterpretation in terms of qualitative concepts for organic matter recalcitrance, we also believe that a more comprehensive quantification and propagation of uncertainties is important to further scrutinize the results of @Hodgkins.2018.
+Besides a reinterpretation in terms of qualitative concepts for organic matter recalcitrance, we also suggest that a more comprehensive quantification and propagation of uncertainties is important to further scrutinize the results of @Hodgkins.2018 because these uncertainties are not negligibly small.
 In the original analysis, the following uncertainties are omitted: (1) As described above, the training dataset is not representative for peat and therefore predictions are biased.
-This uncertainty is not considered in the analyses of @Hodgkins.2018.
-(2) The second source of uncertainty which was omitted is the prediction uncertainty of the models used to predict holocellulose and Klason lignin contents.
-(3) A third source of uncertainty not considered yet is that in computing the regression model to analyze the latitudinal gradient based on average core near-surface contents, the uncertainty for these average values was not propagated.
-The first two sources of uncertainty can be omitted if only normalized peak heights (as qualitative indicators) are used instead of predicted holocellulose and Klason lignin contents, as described above.
+(2) The prediction uncertainty of the models used to predict holocellulose and Klason lignin contents.
+(3) In computing the regression model to analyze the latitudinal gradient based on average core near-surface contents, the uncertainty for these average values was not propagated.
+The first two sources of uncertainty could be omitted if only normalized peak heights (as qualitative indicators) are used instead of predicted holocellulose and Klason lignin contents, as described above.
 But the third source of uncertainty should also be considered when reinterpreting the results in terms of area normalized peak heights.
@@
 With the modified models, the estimated slopes for the latitudinal gradient are more uncertain (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->).
-A similar reanalysis of the depth profiles between the two latitude categories shows that differences are also more uncertain (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
+A similar reanalysis of the depth profiles between the two latitude categories shows that differences in depth patterns are also more uncertain (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
 Our analysis therefore indicates larger uncertainties both in the depth profiles and the latitudinal gradient.
@@
 Notwithstanding our critical remarks, the idea of @Hodgkins.2018 to relate mid infrared spectra to precisely quantifiable and defined organic matter fractions is important.
-Such fractions are relatively well understood and therefore allow us to better deduce what effects climate shifts can have on peat decomposition and recalcitrance.
-It is much more informative if we are able to precisely estimate a change in Klason lignin contents of e.g.
+Such fractions are relatively well understood and therefore allow us to better deduce what effects climate shifts can have on peat decomposition and recalcitrance: It is much more informative if we are able to precisely estimate a change in Klason lignin contents of e.g.
 5 mass-% if average annual temperatures increase by 1$^{\circ}$C than if we are only able to make qualitative statements that the relative amount of aromatics (without describing in detail what this fraction means) increases, as is currently the case.
@@
 a decrease in "carbohydrates" with latitude).
-Ultimately, we need more precise models for peat variations in chemistry than can be developed based on qualitative concepts.
+Ultimately, we need more precise models for variations in peat chemistry which can be developed based on qualitative concepts.
 We suggest the following steps to gain deeper insights into latitudinal gradient in peat chemistry: As outlined above, we suggest to interpret the results of @Hodgkins.2018 in terms of qualitative concepts for peat chemistry, i.e.
@@
 Such models will also help us to get more precise insights into the degree to which high latitude peat accumulation may be decreased by decomposition due to rising temperatures, because they allow replacing more costly measurements.
-Finally, future studies should carefully consider uncertainties and fully propagate these into model estimates.
+Finally, future studies should carefully consider uncertainties and fully propagate these into model estimates if it is unknown whether such uncertainties are negligible.
 Predicting absolute peat holocellulose and Klason lignin contents from infrared spectra remains challenging.

### 2022-07-25 analysis/peer_review_manuscript_1/1/reply-reviewer_1-1.Rmd
Address reviewer comments to manuscript 1.

---
title: "Reply to Comments of Reviewer 1"
author: "Henning Teickner, Klaus-Holger Knorr"
date: "2022-07-24"
output: 
  html_document: default
  pdf_document:
    latex_engine: xelatex
    fig_caption: true
    keep_tex: true
bibliography: ../../paper/references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reviewer comments have prefix "Q" and our answers prefix "A". "old" in combination with line numbers refers to lines in the previous version of the manuscript. "new" in combination with line numbers refers to lines in the updated version of the manuscript.

## General comments

> **Q1:** "This is an excellent manuscript based on an excellent rationale with sound research
questions. It is very well structured and very well written. Building on previous data and
models, Teickner & Knorr provide detailed explanations of the steps followed to evaluate
the models, describing the limitations (representative of the training data sets, validation
of the models, availability of key data of the models’ output, biases and uncertainties,
etc.) and potential improvements. This research also shows the key importance of a
detailed analysis of the models’ residuals as a way to identify model’s deficiencies. It is
also a key finding that OM composition of mineral soils can also be accurately modeled.
This research also shows the potential of MIR for SOM characterization and modelling and
the importance of making data and code available for further improvements.

> Regarding the general issue of the preselection of peaks before modelling, one possible
alternative is to perform PCA on all detected peaks to reduce the dimension and extract
the most relevant spectral signals. PCA is usually efficient in allocating into different
components the overlapping effects of more than one compound on a given spectral
band/region."

**A1:** Thank you for this encouraging comment. It is true that there are many more modeling approaches which probably would have a similar predictive accuracy as the approach we used here (especially since the sample size is small). For example, reviewer 2 suggested partial least squares regression (PLSR) as yet another alternative approach and there are many more approaches one could have tried (e.g. supervised PCR, iterative supervised PCR, interval PLSR, moving window interval PLS, etc; see e.g. @Xiaobo.2010).  
We have the impression that PLSR is more commonly used than PCR since PLSR is more efficient in identifying latent variables related to the target variable being predicted (since it maximizes also the covariance of the computed latent variables to the target variable, whereas PCR only maximizes the variance of the computed latent variables, which comprise only the spectral data). That said, we think that it is good that different tools exist and, depending on the problem, one or the other approach may result in a better predictive accuracy. PCR (or a different dimension reduction approach) may be particularly useful if more data becomes available so that the computational expenses of our approach get too heavy [@Piironen.2020].  
One advantage of the approach we used in comparison to dimension reduction approaches (e.g. PCR, PLSR), which summarize the original predictors into latent variables, is that coefficients for individual predictors are estimated more independently, since no latent variables are computed as summary of all predictors. Therefore, multiple regression with regularizing priors has the advantage of facilitating model interpretation (see e.g. Fig. 3).  
We think that it is good to at least briefly mention these aspects to provide readers a better orientation. To this end, we added the following paragraph (old: l. 175; new: l. 185): "An alternative, popular, approach to approach 2 would be dimension reduction, for example via partial least squares regression, principal component regression, or variants of these [@Xiaobo.2010]. In general, there are many alternative approaches which could be tested to use more information contained within the spectra than the original models, and many of these probably would result in similar predictive performances as approach 2 (regularization), especially when sample sizes are small [@Xiaobo.2010; @Teickner.2022]. An advantage of regularization is that model coefficients are estimated more independently than in dimension reduction approaches, which makes it more straightforward to interpret model coefficients. The key is that the approaches we chose are suitable to analyze our research questions."

--------------------------------------------

## Specific comments

> **Q2:** "L277-279. Positive and negative coefficients for peaks at 1150 and 1270 cm-1 (aromatic CH bending). This may be due to they reflecting different lignin structures: 1270 cm-1 surely corresponds to guaiacol (G) moieties and 1150 cm-1 could correspond to syringol (S) moieties (although S most representative band is at 1310 cm-1). Both differ in content in the source peat vegetation and also show differences in resistance to degradation in peat depending on oxygen availability (see for example, Schellekens et al. 2012 Soil Biology and Biochemistry 53, 32-42). G lignin is less prone to degradation than S lignin and the G/S ratio of fen peat (more decomposed) is larger than that of bog peat (less decomposed) (see for example Martínez Cortizas et al., 2021 Boreas 50, 1161-1178)."

**A2:** Thank you for this information, this is indeed interesting. Building on this: The G/S ratio is also different in the training dataset sample types. Softwood samples and needle samples have a higher G content than hardwood samples and leaves and grasses [@DelaCruz.2016]. They also have a (slightly) lower Klason lignin content than hardwood samples and leaves and grasses (supplementary Fig. S4).  
This indicates that since the bin at 1150 cm$^{-1}$ is caused to some extent by S, and S has a larger abundance in samples with smaller (than average) Klason lignin contents, this may explain why the coefficient for this bin is negative. Likewise, the bin at 1270 cm$^{-1}$ is caused to some extent by G, G has a larger abundance in samples with larger (than average) Klason lignin content, and this may explain why the coefficient for this bin is positive.  
To consider this hypothesis in the main text, we changed the text to (old: l. 277 to 279; new: l. 288): "A plausible explanation for the negative sign of the coefficient for the 1150 cm$^{-1}$ bin is that absorbance in this range is partly caused by syringyl units (S) [@Kubo.2005] and that training samples with higher S content have smaller Klason lignin contents (supplementary Fig. 4, @DelaCruz.2016), making this bin indicative of smaller Klason lignin contents. Likewise, a plausible explanation for the positive sign of the coefficient for the 1270 cm$^{-1}$ bin is that absorbance in this range is caused predominantly by guaiacyl units (G) [@Kubo.2005] and that training samples with higher G content have smaller Klason lignin contents (supplementary Fig. 4, @DelaCruz.2016), making this bin indicative of larger Klason lignin contents."

--------------------------------------------

> **Q3:** "L.424 I find this “implication” to be quite important and other ways of spectral normalization should be tested in the future."

**A3** We completely agree with this statement. We currently are not aware of a solution to this problem, other than to specify precisely under which conditions a prediction model produces erroneous predictions.

--------------------------------------------

```{r child="reply-additional.Rmd"}
```

## References

### 2022-07-25 analysis/peer_review_manuscript_1/1/reply-reviewer_2-1.Rmd
Address reviewer comments to manuscript 1.

---
title: "Reply to Comments of Reviewer 2"
author: "Henning Teickner, Klaus-Holger Knorr"
date: '2022-07-24'
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
    fig_caption: true
    keep_tex: true
bibliography: ../../paper/references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center')

# packages
library(ir)
library(magrittr)
library(dplyr)
library(brms)
library(ggplot2)
```

Reviewer comments have prefix "Q" and our answers prefix "A". "old" in combination with line numbers refers to lines in the previous version of the manuscript. "new" in combination with line numbers refers to lines in the updated version of the manuscript.

## General Comments

>
**Q1:** "Several previous studies have used mid infra-red spectra to assess peat organic matter
chemistry, particularly holocellulose and lignin content. One in particular had calibrated
the spectral data with a set of chemical analyses using various (non-peat) organic
materials. This paper aimed to re-examine the calibration using more refined models and
gauge the suitability for applying them to peat and peat vegetation. The authors do a
thorough job in applying some sophisticated statistical techniques to the data used by
Hodgkins et al. (2018) and show how the calibration models can be improved and what
the limitations of the current dataset are. From this point of view, the paper is useful and
adds to the precision and accuracy of using infra-red spectra to rapidly assess peat
chemistry, with possible application to other organic matter types. A weakness is that the
authors have restricted themselves to the training chemical data from the Hodgkins et al.
paper which are a curious set of paper, wood and plant leaves. In fact, a prior referee of
the Hodgkins et al. paper described the set as being “bizarre” and actually it is surprizing
that such reasonable results are obtained from such an unlikely dataset. One would
normally expect the training set for peat analysis to be based upon samples of actual
peat, from as wide a set of sources as possible. Clearly, the authors have limited
themselves to what was available from the previous study and have not engaged in any
further chemical analysis. To be fair, this weakness is acknowledged (L462) and the need
for further representative training data expressed as the next step. Additionally, given the
quite detailed statistics, which use a Bayesian approach, some of the methodology might
be made clearer to the more general reader as several concepts are not explained and
taken for granted. Also, in places, the English is a little unclear or awkwardly expressed; I
have tried to indicate some of the more difficult passages below. Finally, there are quite a
few errors in the supplementary information."

**A1:** We thank you for your critical appraisal of the scope of our manuscript. It is completely correct that it would have been nice if we would provide already in this study improved models which use peat material as training samples. As you point out, we do not hide this fact throughout the manuscript and fully acknowledge that this issue was discussed by a reviewer of the study of @Hodgkins.2018 (old: l. 53 to 54).

There are three pragmatic reasons and one scientific reason why the scope of the manuscript is limited to what we have analyzed and presented and we take the chance to describe these in more detail here, but also to justify the current scope of the study:

1. This manuscript is a by-product of a PhD project which was conducted with the aim to assess if the prediction models can be reliably used during this project. For this reason, there are limited resources to conduct additional measurements.

2. The required holocellulose and Klason lignin measurements would have been outside our area of expertise.

3. Considering the growing number of studies which use the original models of @Hodgkins.2018, we considered it important to publish our analysis as soon as possible to avoid future misinterpretation of the models' predictions. An alternative option would have been to find collaborating partners who conduct (or conduct ourselves) the required measurements which would have cost additional time, but we would of course be open to that.

4. We think --- as you do --- that the manuscript in its present form makes genuine contributions to the scientific discourse, especially of researchers computing spectral prediction models for peat.

We build on the critical comment of a reviewer of the original study of @Hodgkins.2018 by giving these critical comments a quantitative, causal, and testable foundation.  
The quantitative foundation is that our critical appraisal of the original models builds on an actual analysis of the models with data and thus provides more than what critical comments in a peer review process can and should reasonably do.  
The causal foundation is that we provide --- based on our quantitative analysis --- a detailed causal explanation how and why the original training data are not representative for peat. It is easy to state a reasonable assumption that training data acquired from samples of a different type of material _may_ not be representative for peat. We argue that it is a genuine contribution to show that this is actually the case and to explain why. From this analysis, we learn something about the chemistry of the original training data, the peat samples, and about the relation between this chemistry and the performance of prediction models (and spectral data). We think that this knowledge will be important also when peat samples are available as training samples, because there is no standard stating what is a representative dataset or stating how to define a set of sources as wide as necessary to compute models with a certain predictive accuracy. Peat itself can be quite diverse. We show here --- as a hypothesis to be tested --- what aspects probably are especially important when computing and validating improved models.  
The testable foundation is that we provide hypotheses which can be tested. Our first hypothesis is that the normalized "carbohydrate" peak height (`carb`) is a good indicator for holocellulose content only when there are no (large) differences in OH bond types and abundances between the samples. We assume that relevant differences in OH bond types and abundances can also exist for peat samples because the chemistry and structure of cell walls of the peat forming vegetation can vary (both the amount and type of OH bonds and the holocellulose content) (e.g. comparing tropical peat with large fractions of wood, peat from (temperate) forest mires, or peat formed by *Sphagnum*). This hypothesis can be tested e.g. by comparing predictive performances of models fitted to either type of data. Testing this hypothesis would go beyond computing improved models: we can for example also learn something about how to interpret MIRS in terms of the decomposability of peat.  
Our second hypothesis is that the normalized "aromatic" peak height (`arom15arom16`) is a good indicator for Klason lignin only when samples do not differ in protein content. In direct analogy to the "carbohydrate" peak height, this hypothesis is relevant (because also peat samples can differ in their protein content), it can be tested in the same way as our hypothesis for the normalized "carbohydrate" peak height, and testing it can make us learn something about peat chemistry and MIRS interpretation: we can for example learn whether `arom15arom16` (or intensities of similar parts of MIRS as are widely used in the interpretation or computation of humification indices [e.g. @Broder.2012]) are good indicators for some concept of decomposability. This may ultimately lead to a similar discussion as whether the C/N ratio is a good indicator for decomposability (or degree of decompostion), a discussion which to our knoweldge has not been considered relevant yet.  
In addition, we apply tools to validate models (particularly the analysis of the prediction domain, the analysis of model coefficients, the analysis of model residuals) which have to our knowledge only seldomly been applied on spectral prediction models for peat. In particular where sample sizes are small, such analyses could in analogy to ours here show how small peat datasets are also not representative. We hope that our study helps to circulate such model validation steps by showing in detail how they can be useful and how to apply them.

To summarize: We agree that it would be ideal if we had already today models which use peat training data to predict peat holocellulose and Klason lignin contents. However, we argue that our study makes genuine contributions which go beyond this aim and which merit publication.

We would also like to thank you for the detailed recommendations to improve language and method descriptions. We addressed these points in the specific comments below.

## Specific Comments

> **Q2:** "Abstract (L1) I feel there is a need to distinguish peat from soil organic matter. True, peat is a type of soil (a Histosol) and the organic matter of peat is a type of ‘soil organic
matter’. However, the general expression of ‘soil organic matter’ is usually associated with
mineral soils. Thus, in L3, Hodgkins et al. set out to predict peat holocellulose and Klason
lignin contents, not organic matter holocellulose and Klason lignin contents; indeed, they
specifically removed the training samples containing silicates."

**A2:** We assume that you suggest to include the word "peat" (old: l. 3 to 4; new: l. 3 to 4): "Recently, Hodgkins et al. (2018) developed models to quantitatively predict _peat_ holocellulose and Klason lignin contents". We think that this is a good suggestion to avoid confusion and we changed the line as suggested.  
There are several reasons why we framed the manuscript in terms of SOM and not exclusively peat and we list them here to explain our rationale:

1. We suggest that our study builds a bridge to studies computing prediction models for mineral soils because we show as a proof-of-concept that it is possible to predict holocellulose content also for the mineral-rich training samples. We therefore think that our study is also of interest for colleagues working with mineral soils.

2. There are studies computing prediction models which predominantly include mineral soil data, but nevertheless also include peat samples or even focus on peat samples [e.g. @Helfenstein.2021]. These studies refer e.g. to soil organic carbon (SOC) which we also see as a term mostly used in the mineral soil community. However, as this study shows, these distinctions in terminology are less and less useful when soil C content is considered as continuum. We suggest that our study is of relevance also for mineral soils (because of our proof-of-concept mentioned above, because of the exemplary application of model validation approaches, and because of factors which may complicate prediciton of holocellulose and Klason lignin contents which may also be relevant in mineral soils).

--------------------------------------------

> **Q3:** "L13 “mineral-rich samples” – it is unclear whether this refers to peat or to mineral soil. Of course, it could be both. As mentioned later (L96), some peats can become contaminated with mineral material to varying levels, though these are relatively rare occurrences."

**A3:** We assume that you suggest to clarify already in l. 13 (old) what "mineral-rich samples" should refer to --- mineral soil or peat. In fact, we wanted to communicate that this is a proof-of-concept applying to both cases. One problem is that we have no information on the mineral content for the particular training samples and another problem is that the training data probably are not representative for either case. However, the mechanism by which minerals create peaks in the spectra should be the same for either soil type.  
We hope that the following change makes more clear that we refer to both soil types (old: l. 12 to 13; new: l. 12 to 14): "Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples (e.g. peat with mineral admixtures or potentially mineral soils)."

--------------------------------------------

> **Q4:** "L29 “OM derived from plants and SOM” – better to reword as “SOM and OM derived from
plants” (OM derived from SOM doesn’t make sense)."

**A4:** Thank you for pointing out this inconsistency. We actually wanted to write: "Plant and soil OM" (i.e. that the fractionation procedure is applied to both types of OM, not only SOM and not only plant OM). We changed the text accordingly (old: l. 29; new: l. 30).

--------------------------------------------

> **Q5:** "L37 “few species” – unclear if this is of wood or some other plants."

**A5:** Thank you for pointing out that this is not clearly written. We changed the text to make this clearer (old: l. 37; new: l. 38): "Most of these models consider only wood, material from few woody and non-woody species, or only specific vegetation organs [@Elle.2019]."

--------------------------------------------

> **Q6:** "L46 “a later study” – need to specify which."

**A6:** We here refer to all studies which were cited in the previous sentence. These are the studies which have cited @Hodgkins.2018 and which have used the models developed in @Hodgkins.2018 (and thus are the studies we are aware of which could have potentially validated the original models).  
We changed the sentence to state this explicitly and hope that this was what you had in mind (old: l. 46; new: l. 48): "A major problem is that neither @Hodgkins.2018, nor any later study which was cited above and which used the models provided a thorough validation of the models."

--------------------------------------------

> **Q7:** "L48 Replace “are” with “were”."

**A7:** We are not entirely sure whether you refer to l. 48 or actually wanted to refer to l. 49. In l. 48 (old), the sentence "The peaks are baseline corrected, their maximum height computed and divided by the sum of absorbance values in the spectra." refers to the procedure in general, and thus present tense should be correct. However, it is true that the next sentence (old: l. 49) "These values are used to calibrate the prediction models." refers to what was done in @Hodgkins.2018 and thus past tense should be used (as mentioned in Q8).  
We therefore changed the two sentences to (old: l.48 to 49; new: l. 50 to 52): "In the procedure, the peaks are baseline corrected, their maximum height computed and divided by the sum of absorbance values in the spectra. In @Hodgkins.2018, these values were used to calibrate the prediction models."

--------------------------------------------

> **Q8:** "L49 Replace “are” with “were”."

**A8:** Please see our reply to Q7.

--------------------------------------------

> **Q9:** "L51 Replace “for example” with “, for example,”."

**A9:** Thank you. We changed the text as suggested (new: l. 54).

--------------------------------------------

> **Q10:** "L56 Replace “of” with “in”."

**A10:** Thank you. We changed the text as suggested (new: l. 59). We made analogous changes in: old: l. 110, new: l. 117.

--------------------------------------------

> **Q11:** "L66 Replace “which” with “in which”."

**A11:** Thank you for mentioning this issue. We think that "for which" is more appropriate here since we refer to assumptions of a distribution and not _in_ a distribution (please correct us if you see this differently).  
We changed the sentence to (old: l. 66; new: l. 68): "In contrast, if unrealistic predictions occur, it is more reasonable to use a distribution for which assumptions are consistent with knowledge on the data generating process."

--------------------------------------------

> **Q12:** "L67 I suspect most readers of ‘SOIL’ will not be so familiar with the term “beta
distribution”; some explanation here is warranted."

**A12:** Thank you for pointing this out. We provide a brief explanation in the subsequent sentence (old: l. 67): "For example, the beta distribution assumes a lower and upper bound which can be mapped to the interval [0, 100] mass-%." This sentence mentions the values the Beta distribution is valid for and this is in our opinion the most relevant characteristic here.  
To make clearer what the role of the beta distribution in a model is, we elaborated this to (correcting the interval the beta distribution is valid for) (old: l. 67; new: l. 70): "For example, the beta distribution assumes a lower and upper bound which can be mapped to the interval (0, 100) mass-%. It can be used as replacement for the normal sampling distribution used in ordinary linear regression models, such as the original models."  
We hope that this is an explanation you had in mind, but we are open for detailed suggestions on how to improve the text here.

--------------------------------------------

> **Q13:** "L77 The meaning of “prediction domain” should be given."

**A13:** The term "prediction domain" is described in more detail in section 2.3. However, we agree that it is a good idea to explain the term already here: In principle, the prediction domain is the range of predictor variable values in the training data set.  
To make this clear, we changed the sentences in l. 77 to 78 (old) to (new: l. 81 to 83): "_Is the prediction domain of the training data representative for peat samples?_ If a model extrapolates outside the prediction domain --- the range of predictor variable values in the training data set [@Wadoux.2021] --- it is unclear if the predictions are valid."

--------------------------------------------

> **Q14:** "L79 Replace “on” with “to”, replace “e.g.” with “, for example,”."

**A14:** Thank you for pointing this out. We changed the text as suggested (new: l. 84). We made analogous changes in: (1) old: l. 197, new: l. 214; (2) old: l. 423, new: l. 446; (3) old: l. 426, new: 450.

--------------------------------------------

> **Q15:** "L80 The meaning of “no problem-specific correlations” is unclear."

**A15:** Thank you. We agree that this term is not clear enough. With "problem-specific correlations", we wanted to describe the dependence of a prediction model on spectral properties obviously specific for one type of material, only.  
We changed the sentence to make clearer what we mean (new: l. 84): "An advantage of the independent reference sample set used by @Hodgkins.2018 is that the model is applicable to various OM types and, for example, does not depend on obviously material-specific spectral characteristics [@Hodgkins.2018]."

--------------------------------------------

> **Q16:** "L82 Generally, it is not good to start a sentence with “But”; “However,” might be better here."

**A16:** Thank you. We changed the sentence as suggested (new: l. 87).

--------------------------------------------

> **Q17:** "L83 The “generality” of the training data set is not so obvious – it is quite wood based, since paper itself is derived from wood. Secondly, it could be argued to be a disadvantage since it overlooks any potential diagnostic peculiarities found only in peat."

**A17:** We partly agree here. We agree that we should stress more that the generality is a _potential_ advantage since it has not been tested so far. We tried to stress this aspect more as suggested (new: l. 88): "A potential advantage is therefore the potential generality of the training data set @Hodgkins.2018 used."  
However, we argue that the training data set is (at least one of the) most general training data sets available (based on the references cited in the introduction), and that the non-applicability to peat samples is also not so obvious without a detailed analysis.  
It is true that the data set contains many wood samples, but it also contains not a small amount of grass, leave, and needle samples (`r ir::ir_sample_data %>% dplyr::filter(sample_type == "leaves and grasses") %>% nrow() %>% magrittr::divide_by(nrow(ir::ir_sample_data)) %>% magrittr::multiply_by(100) %>% round(0)` % of the samples) which could potentially have similar spectral properties as peat samples. For this reason, we argue that without knowing our analysis, there are not many reliable arguments neither for, nor against the applicability of a model fitted to the data to peat samples.  
Of course we agree with you that the original study should have validated the applicability of the training data, but we do not think that without any analysis it is possible to say beforehand that the models could not work. This is the only point we want to make here.

--------------------------------------------

> **Q18:** "L97 Replace “predictions” with “prediction”?"

**A18:** Thank you for pointing this error out. We assume that this refers to l. 87 and corrected this as suggested (new: l. 91).

--------------------------------------------

> **Q19:** "L98 Replace “also for” with “that includes”."

**A19:** Thank you. We changed the text as suggested (new: l. 103).

--------------------------------------------

> **Q20:** "L99 Of course, as mentioned above, it would have been better to have a set of genuine
mineral soil samples."

**A20:** We agree that this important limitation should be stressed again here. We therefore added the following sentence (new: l. 105): "This is no replacement for a model using more training data, but it is a test whether it is at least in principle possible for MIRS to contain sufficient information to predict holocellulose contents of relatively diverse samples."  
However, we stress that it is not clear what a "set of genuine mineral soil samples" is unless one defines it via its prediction domain (or the molecular structures which eventually cause the prediction domain). This is probably not that easy since different minerals also differ in their spectral properties [e.g. @Stuart.2004], meaning that this requires detailed research. Our study exemplarily shows approaches to analyze such difficulties.

--------------------------------------------

> **Q21:** "L114 “The data are available…” – delete this; it is obvious."

**A21:** We agree that the formulation is not clear. We wanted to write that the data are accessible directly from the published paper (as supplementary information). It certainly is obvious that the data are used in @Hodgkins.2018, but we think that it is important to point to the exact data source used. Unfortunately, the supporting information has no own DOI.  
To stress that we refer here to the access to the supplementary information, we changed the sentence to (new: l. 121): "The data are accessible from the supplementary information of @Hodgkins.2018."

--------------------------------------------

> **Q22:** "L117 Replace “SOM” with “peat OM”."

**A22:** We changed the text as suggested (new: l. 125).

--------------------------------------------

> **Q23:** "L122 “informative priors” – again this term may not be readily understood and should be
made clearer. Secondly, what exactly these priors are should be made more specific."

**A23:** On the one hand, we agree that it would be good to communicate this to readers unfamiliar with Bayesian analysis. On the other hand, we have to assume a certain familiarity with Bayesian statistics to avoid repeating too much content from textbooks on Bayesian statistics.  
The term "informative" is no fixed term, but is a qualitative and problem-specific descriptor for the information content of a prior [@Gelman.2017a]. Here, we mean with a "weakly informative prior" a non-flat proper (meaning that the prior probability distribution integrates to 1) prior which has no strong effect on the predictions in comparison to the respective frequentist model. That this was indeed the case was tested in supplementary information 1.  
In particular, we assumed for both the intercept and regression coefficient a normal(0, 2.5) prior for $z$-transformed independent and dependent variable. This means that a one $\sigma$ (denoting the standard deviation of $z$-transformed variables) change in the independent variable (e.g. the $z$-transformed normalized `carb` peak height) can result in the same change in the dependent variable (e.g. the $z$-transformed holocellulose content) when the absolute coefficient is $\ge1$, which has a prior probability of approximately `r round(pnorm(-1, 0, 2.5) * 2, 2)`. The relevant information is that in comparison to the frequentist model, the Bayesian models produce the same prediction intervals.  
To describe this better in the text, we changed the text to (old: l. 121 to 123, new: l. 128): "For this reason, to facilitate model comparison, we also recomputed the original models as Bayesian models with weakly informative priors. With weakly informative priors, we mean here priors which result in approximately the same prediction intervals as the original frequentist models (supplementary Fig. 1)."  
We think that giving detailed parameter values would not support understanding unless you have more detailed knowledge on Bayesian statistics (and perhaps the software implementation) and we stress that these are available in full detail already via the reproducible research compendium (file `analysis/paper/002-paper-m-original-models.Rmd`, l. 40 to 41).

--------------------------------------------

> **Q24:** "L131 Replace “MCMC” with “Markov Chain Monte Carlo (MCMC)”."

**A24:** Thank you. We changed the text as suggested.

--------------------------------------------

> **Q25:** "L139 Is there not a third way? Surely (holocellulose + Klason lignin) ≤ 100%? In fact, it may be < 100% if there are other extracted components not included (I am not sure if
this might be the case, since I do not have ready access to the De la Cruz et al. (2016)
paper, describing the actual chemical analysis procedure). It is not clear if this constraint
has been included within the statistical analysis."

**A25:** Yes, you are right that (holocellulose + Klason lignin) $\le$ 100 mass-% should hold, or more specifically, (holocellulose + Klason lignin + all other compounds) = 1 should hold. One could consider this with a Dirichlet regression model which is the multivariate extension to the beta regression model [e.g. @Douma.2019]. We did not do this here for several reasons:

1. The analysis would become conceptually more complicated than it is now (which would make the analysis harder to understand) and we could not compare models for holocellulose and Klason lignin independently. This would also apply to the analysis of model coefficients.

2. The analysis would become computationally more difficult.

3. We are not aware of non-Bayesian approaches to Dirichlet regression for high dimensional data (e.g. PLSR) which would make such computations more feasible.

4. Not for all samples in the training data does (holocellulose + Klason lignin) $\le$ 100 mass-% hold (see supplementary Fig. 4 (a), this is the case for sample type "office paper"). This implies that the data generating process is not fully compatible with the Dirichlet distribution (at least for contents as measured for the training data). We assume that this implies that the measurement process needs to be improved since, apparently, the sum of the measured values cannot be unambiguously translated to mass-% in all cases. Note that this might also be a reason why Klason lignin contents for sample type "office paper" could not be appropriately modeled [@Hodgkins.2018].

```{r reply-reviewer-2-1-1}
knitr::opts_knit$set(output.dir = ".")

knitr::load_cache(label = "m-reduce-underfitting-11", path = "../../paper/001-paper-main_cache/latex/", notfound = NULL, lazy = TRUE)

d_pred_best_sum <-
  dplyr::bind_rows(
    d_peat_pred_best_sum %>%
      dplyr::mutate(sample_type = "peat"),
    d_veg_pred_best_sum %>%
      dplyr::mutate(sample_type = "vegetation")
  )
```

However, we fully agree that this would be a useful improvement of our approach. To at least check whether the condition is violated for the peat and vegetation samples for the best binned beta regression models, we added individual MCMC samples from these models for holocellulose and Klason lignin. The histograms of the upper 95% prediction interval limits are shown in Fig. 1 below. The maximum value across all samples is `r round(max(d_pred_best_sum$upr) * 100, 0)` mass-%.  
Both this maximum value and the histogram suggest that the new condition is not directly violated, but also that narrower prediction intervals can be expected from a Dirichlet regression model, since it is unlikely that holocellulose and Klason lignin contents are that high (i.e. 100 mass-%). Unfortunately, we are not aware of data for peat which would enable us to assess whether the median of the upper 95% prediction interval limit across all samples (`r round(median(d_pred_best_sum$upr) * 100, 0)` mass-%) (or the median of the median, `r round(median(d_pred_best_sum$fit) * 100, 0)` mass-%) is sensible or not.

```{r reply-reviewer-2-1-2, out.width="80%", fig.height=2, fig.width=5.5, fig.cap="Histogram of the 95% prediction intervals for the sum of holocellulose and Klason lignin mass contents predicted with the best binned models."}
d_pred_best_sum %>%
  ggplot(aes(x = upr, fill = sample_type)) +
  geom_histogram(bins = 60) +
  guides(fill = guide_legend(title = "Sample type")) +
  labs(x = "Holocellulose + Klason lignin (upper 95% prediction interval limit)", y = "Count") +
  theme_classic()
```

To make clear that this constraint has not been included within the statistical analysis, that this is a limitation, and why the constraint was not considered, we added the following sentences to the text (old: l. 449, new: l. 474):  
"A further limitation is that as the original models, the modified beta regression models do not consider the constraint that the contents of holocellulose, Klason lignin, and any remaining compounds should sum to 100 mass-%. This also represents a further test how realistic model predictions are (compare with Sect. 2.1). In principle, this constraint could be considered by using a Dirichlet regression model [e.g. @Douma.2019]. We have not used this approach here due to the higher computational costs, potential computational difficulties, to keep the present model validation straightforward, and since the training data do not fulfill this condition for all samples (indicating that the measurement procedure needs to be improved, too)."  
We did not add these sentences in l. 140 since in our opinion, this would have made the description of the methods unnecessarily complex.

--------------------------------------------

> **Q26:** "L140 Replace “in” with “within”."

**A26:** Thank you. We changed the text as suggested (new: l. 148).

--------------------------------------------

> **Q27:** "L142 Replace “covering” with “covered”."

**A27:** Thank you. We changed the text as suggested (new: l. 151).

--------------------------------------------

> **Q28:** "LL154-155 “in the form of depth profiles” – this is unclear; are you referring to actual peat depths here? Perhaps needs rewording."

**A28:** Yes, we refer to plots where predicted values for peat samples are plotted against the depths of the peat samples in the peat cores. We were not sure what rewording you would suggest in this case.  
We changed the text to hopefully make clearer what we mean (new: l. 163): "To facilitate practical comparison between the modeling approaches, we compared the predictions of all models for the peat samples by plotting predicted values versus depths of the peat samples."

--------------------------------------------

> **Q29:** "L165 Replace “amont” with “amounts”. The term “regularization” may need some defining.
Replace “for” with “of”?"

**A29:** Thank you for pointing out these errors. We changed the text as suggested (new: l. 174).  
Again, we had to face a trade-off between writing a short and concise manuscript which focuses on what we did and explaining in more detail statistical terms which may not be known to all readers. We hope that the following additions make the text better understandable (and we would like to thank you for mentioning such aspects!) (new: l. 174):  
"To avoid overfitting, we computed models using priors implying different amounts of regularization --- shrinking coefficients to 0 with the aim to avoid overfitting --- of model coefficients (standard deviation of 1 and 0.5, respectively) ..."

--------------------------------------------

> **Q30:** "L168 It might have been interesting to compare the approach given in 2. to a PLS (Partial Least Squares) analysis, which has been applied to similar data (see Artz et al.) and is a standard (non-Bayesian) approach for such multivariate data."

**A30:** It is true that there are many more modeling approaches which probably would have a similar predictive accuracy as the approach we used here (especially since the sample size is small). For example, reviewer 1 suggested principal component regression (PCR) as yet another alternative approach and there are many more approaches one could have tried (e.g. supervised PCR, iterative supervised PCR, interval PLSR, moving window interval PLS, etc; see e.g. @Xiaobo.2010).  
Based on a previous study [@Teickner.2022], we assume that for datasets with a similarly small sample size (around 50), the regression approach we used here and partial least squares regression (PLSR) often have similar predictive accuracies. Of course, this depends on the particular dataset used. Based on this previous study, we assume also here that PLSR would have improved the predictive accuracy in a similar way as our "best binned" models and would potentially have reduced the (potential) biases we found for the original models.  
One advantage of the approach we used in comparison to dimension reduction approaches (e.g. PCR, PLSR), which summarize the original predictors into latent variables, is that coefficients for individual predictors are estimated more independently, since no latent variables are computed as summary of all predictors. Therefore, multiple regression with regularizing priors has the advantage of facilitating model interpretation (see e.g. Fig. 3).  
A practical disadvantage we would have faced when using PLSR is that we are not aware of a fully working R implementation of a beta PLSR approach. This would have made, in our opinion, the implementation of some of the improvements we suggest here using PLSR difficult and/or conceptually similarly difficult as the Bayesian approach we used here.  
We do not claim that our approach is perfect yet, nor that we have filled all research gaps (however, we certainly have addressed the research questions set out in the introduction). We welcome any studies which want to further improve our models or compare them to other modeling approaches (all models can be reproduced via the reproducible research compendium).  
We think that it is good to at least briefly mention these aspects to provide readers a better orientation. To this end, we added the following paragraph (old: l. 175; new: l. 185): "An alternative, popular, approach to approach 2 would be dimension reduction, for example via partial least squares regression, principal component regression, or variants of these [@Xiaobo.2010]. In general, there are many alternative approaches which could be tested to use more information contained within the spectra than the original models, and many of these probably would result in similar predictive performances as approach 2 (regularization), especially when sample sizes are small [@Xiaobo.2010; @Teickner.2022]. An advantage of regularization is that model coefficients are estimated more independently than in dimension reduction approaches, which makes it more straightforward to interpret model coefficients. The key is that the approaches we chose are suitable to analyze our research questions."

--------------------------------------------

> **Q31:** "Figure 2. The key would be better expressed as “Bayesian models â¡ Beta â¡ Gaussian”, since both are Bayesian. The sentence “ Dashed grey…” could be omitted; it is pretty obvious."

**A31:** Thank you. We changed the figure as suggested (new: Fig. 2).

--------------------------------------------

> **Q32:** "L276 Define the “fingerprint” region (it could be shown in Figure 1)."

**A32:** We changed the sentence to (new: l. 293): "Instead, bins with large absolute coefficients are located in the fingerprint region (600 to 1500 cm$^{-1}$) and probably related to aromatic in plane C-H bending (~1150 and ~1270 cm$^{-1}$) [@Stuart.2004] ...".  
We think that it is not necessary to visualize this range in a figure as the term was just used to indicate the rough location of the selected bins. The fingerprint region is defined based on @Stuart.2004.

--------------------------------------------

> **Q33:** "Figure 3. Under “Holocellulose (no minerals)”, the blue line at ca. 1590 (negative
coefficient > 0.2) is unmarked – should it be?"

**A33:** Yes, it should be unmarked because the median absolute coefficient is not $>0.2$ (it is slightly smaller).

--------------------------------------------

> **Q34:** "Table 1. The column “Original model?” is not helpful, more confusing – omit (given in
legend, whether .2 or .3)."

**A34:** Thank you for this suggestion. We changed the table accordingly (new: Tab. 3).

--------------------------------------------

> **Q35:** "LL304-305 This is really unsurprising."

**A35:** You are right that one _can_ have a reasonably strong assumption/belief that it is _a priori_ unsurprising that the prediction domain of the training dataset does not cover the peat samples. We also explicitly described this assumption in the introduction (old: l. 84 to 86; new: l. 89 to 91) showing that we were not too surprised, too.   
However, in our opinion, a mere assumption needs to be backed up by a quantitative analysis in order to be suitable as an argument to suggest concrete improvements. We cannot criticize existing models by reasonable assumptions, we need tests of these assumptions.

--------------------------------------------

> **Q36:** "LL322-323 Again, this points to the training dataset not being suitable for the task in
hand."

**A36:** Yes, this is correct (as we mention in the manuscript (old: l. 327 to 329, 356 to 357; new: l. 349 to 351, 378 to 379)).  
However, we stress that our analysis does more than this because it explains why the training data are not suitable, i.e. which characteristics of the training samples cause which spectral features which cause the bias. This knowledge is important to choose suitable training data and to check under which conditions models computed with such data do make accurate predictions.  
We do not think that collecting a "genuine" training dataset is a trivial problem, exactly for the key arguments we provide in our study. These key arguments are that we need to pay attention to differences in protein content and to differences in other major spectral features, even if we do not include them as predictors into the model.

--------------------------------------------

> **Q37:** "L360 “classes” – this is a new term which has not been defined. We are left to guess what these classes are. They are given in Supplementary Figure 11, but at this point we are yet
to read this."

**A37:** After answering comment Q58 below, we noticed that the supplementary text and also the main text are written ambiguously, since we do not list all sample types belonging to class 1.  
To address this valuable point, we changed the main text as follows (old: l. 383 to 384; new: l. 405 to 406): "This is what distinguishes samples of class 1 and 2 in the training data (supplementary Fig. 11): Samples of class 1 are wood samples and paper product samples derived from wood."  
We think that this resulted in a misunderstanding. The classes are defined in the sentences following l. 360 (old): "Samples of these classes differ in the relative contribution of the `arom15` and `arom16` peaks to `arom15arom16` (the sum of their heights) (Fig. 6). The first class (class 1) has high contributions of the `arom15` peak ($>30$%) and typically smaller `arom15arom16` values; the second class (class 2) a smaller contribution of the `arom15` peak and larger `arom15arom16` values."  
In the main text, we refer to Fig. S11 in the context of the two classes only in l. 383 (old) (new: l. 405) and there, we describe that the two classes comprise different sample types. This order is due to how we defined the classes and how we build up our argument: first we describe what spectral features are related to the bias and then we explain how the bias arises with the help of knowledge of additional sample properties. Thus, this order is by intention.  
We hope that this makes more clear how the classes are defined, that we provide these definitions in the main text, and sample types are not the same as the two sample classes. We apologize that this has been written confusingly. If this is still confusing, we would be glad to receive suggestions for how to make this clearer.

--------------------------------------------

> **Q38:** "L395 “because they interfere less with” – no, the reverse applies: “because they are
interfered less by”."

**A38:** We see interference as a symmetrical interaction. But we see that in this context the wording could be improved as suggested. We changed the text as suggested (new: l. 418). Thank you.

--------------------------------------------

> **Q39:** "L396 Replace “that” with “why”."

**A39:** Thank you. We changed the text as suggested (new: l. 419).

--------------------------------------------

> **Q40:** "Figure 6 “If the points are…” – I think you can omit this; it is a rather trivial remark and obvious to all."

**A40:** We changed the figure caption as suggested (new: Fig. 6).

--------------------------------------------

> **Q41:** "L403 The ELPD values quoted seem to be the same – is this correct?"

**A41:** Yes, this is correct, but it is true that we could have written this less confusingly. The two statements say the same thing (the first says that the model using binned spectra had an average ELPD 17.24 larger than the model using all peaks; the second says that $\Delta$ELPD --- the difference in average ELPD of the model using all peaks relative to the model using binned spectra is -17.24).  
We changed the text by deleting the first statement (new: l. 425). Thank you for pointing this out.

--------------------------------------------

> **Q42:** "L422 Replace “to predict” with “being predicted”."

**A42:** Thank you. We changed the text as suggested (new: l. 445).

--------------------------------------------

> **Q43:** "L423 Replace “on” with “to”."

**A43** Thank you. We changed the text as suggested (new: l. 446).

--------------------------------------------

> **Q44:** "L427 Replace “on than” with “to than”."

**A44:** Thank you. We changed the text as suggested (new: l. 450).

--------------------------------------------

> **Q45:** "L429 Replace “what” with “as to what”."

**A45:** Thank you. We changed the text as suggested (new: l. 452).

--------------------------------------------

> **Q46:** "L439 Replace “problem is that also” with “problems are that”."

**A46:** Thank you for pointing out this error. We changed the text (new: l. 462) to: "The most important problem is ..." (as the sentence refers to one problem).

--------------------------------------------

> **Q47:** "L442 Replace “probably ca” with “can probably”. However, I think “probably” is inaccurate here; I think you say “certainly”."

**A47:** Thank you for pointing out this error. You are also right that we mean "certainly". We changed the text as suggested (new: l. 465).

--------------------------------------------

> **Q48:** "L444 “calibration transfer” needs definition or more explanation."

**A48**: Thank you. We changed the sentence to make this clearer (new: l. 465): "(2) It is unclear how robust the original models and our improved models --- especially models using binned spectra --- are in terms of calibration transfer (calibration transfer is the application of a model to spectra measured differently than the training data, e.g. with a different procedure, on a different device, in a different laboratory [@Workman.2018])."

--------------------------------------------

> **Q49:** "L450 I suggest replacing “SOM” with “actual peat samples” – the idea of application to
SOM (in general and including mineral soils) comes two sentences later. Hence in L451 I
suggest omitting “SOM and”."

**A49**: Thank you. We changed the sentence to (new: l. 481):"In summary, our analysis opens concrete and promising directions to further improve the models: We need training and validation data that include peat, particularly highly decomposed peat."

--------------------------------------------

> **Q50:** "L464 “interpreted with caution” – it would be useful to develop this thought more. The estimates given in Hodgkins et al. have been applied to give far-reaching conclusions regarding peat carbon storage, latitudinal trends and possible responses to climate change. Using the same basis, ideas have been amplified and extended in the Verbeke et al. (2022) paper. Can we still rely on the overall findings from these papers or are the conclusions now in doubt? I suspect that the revised models given here, while improving the accuracy of determinations to some minor degree, will not change the overall pictures presented in these (and perhaps other) derivative papers but some discussion on this would be very helpful and would give a wider context and application."

**A50:** We appreciate this suggestion as we also think that further discussions on this proposed gradient are useful.  
Indeed, we have elaborated this point further (and added additional remarks and suggestions regarding uncertainty handling) in a companion manuscript intended as a reply to @Hodgkins.2018 (this manuscript was written before @Verbeke.2022 was published). This reply manuscript is not yet published, since we first wanted to publish the manuscript discussed here which represents the technical basis for our reply manuscript. However, the script for the reply manuscript (containing also the text) is available via the reproducible research compendium to the manuscript discussed here.  
In brief, the results differ if the modified models are used due to the strong underestimation of Klason lignin contents (i.e. Klason lignin contents are strongly underestimated in @Hodgkins.2018). As a consequence, also the depth-related relative accumulation of Klason lignin during decomposition is probably underestimated. We think that this is an important finding since a larger amount of Klason lignin is required according to the modified models for the hypothesized stability of tropical peat and deeper peat than has been suggested.  
Another point is that because of the bias in the models, the original results can only be interpreted in terms of qualitative concepts for carbohydrates and aromatics which makes it much more difficult to connect the results to other studies, to generalize them, and to test them.  
To foreshadow more explicitly that we are preparing such an analysis, we modified the sentence (new: l. 496): "Results from currently published studies using the original models should be interpreted with caution (we are currently preparing a manuscript, see @Teickner.2022b, which explores the implications of our results here for the results of @Hodgkins.2018)."

--------------------------------------------

> **Q51:** "L466 Delete “, also in practice”."

**A51:** We changed the text as suggested (new: l. 500).

--------------------------------------------

## Supplementary Information

> **Q52:** "A general comment on the information here: apart from the introductory paragraphs to S1
and S2, there is a description of the figure followed by the legend to the figure and often
the material is a repeat. Really, it would be better just to have a descriptive legend."

**A52:** Thank you for making this suggestion. We changd the text as suggested.

--------------------------------------------

> **Q53:** "S1 Replace “as baseline” with “as a baseline”."

**A53:** Thank you. We changed the text as suggested.

--------------------------------------------

> **Q54:** "Figure 3. See same comment as for Figure 2 above."

**A54:** Thank you. We changed the figure as suggested (new: Fig. S3).

--------------------------------------------

> **Q55:** "Figure 4. The axis labels for (a) are the wrong way round! For “Sample type” is this the same as “class” in the main text? If so change to “class” (but see on Figure 11 below)."

**A55:** Thank you for pointing out this error. The axes labels are indeed the wrong way round. "Sample type" is not the same as "class" (the two classes defined in the manuscript text) in the main text. We changed the figure as suggested (new: Fig. S4).

--------------------------------------------

> **Q56:** "Figure 6. Replace “overestimation” with “underestimation” and “underestimation” with
“overestimation”."

**A56:** Thank you for pointing out this error. We changed the caption as suggested (new: Fig. S6).

--------------------------------------------

> **Q57:** "Figure 8. I would suggest having Training, Peat and Vegetation across the top and peak
heights down the side (as laid out in Figure 7). Also (a) and (b) seem to show the same
data; can this be right? In the legend it says peat (second row) but it is actually peat in
the third row."

**A57:** Thank you for spotting this error! Indeed, (a) did not show the data for the intensity at 3400 cm$^{-1}$, but for 1250 cm$^{-1}$. We corrected this and also changed the figure as suggested (new: Fig. S8).

--------------------------------------------

> **Q58:**  "Figure 10. Replace “6.31e-16” with “0”."

**A58:** Thank you! We changed the figure as suggested (new: Fig. S10).

--------------------------------------------

> **Q58:**  "Figure 11. Replace “type 2” with “class 2”. It is unclear what these classes are, and where does paper and cardboard fit in? Perhaps indicate in the list of sample types which are class 1, 2, 3, etc."

**A58:** Thank you. Since this text is deleted (see A52), we did not change the text. However, it is true that this was not described clearly enough and we think that this might also be relevant for the misunderstanding (we assume) for Q37. We hope that this comment was addressed in A37.

--------------------------------------------

> **Q59:** "Figure 12. The order of regressions mentioned in the description is not the same order as actually depicted. In the legend, replace “binnes” with “binned”, “50” with “20” (I presume), and “line” with “lines”."

**A59:** Thank you for pointing out these errors. We changed the caption as suggested (new: Fig. S12).

--------------------------------------------

> **Q60:** "Figure 13. The “Dataset” key could be omitted – it is superfluous."

**A60:** Thank you for the suggestion. It is true that the key is redundant. However, we would like to keep the redundancy here since the color scheme is applied across different figures and we think that this helps recognizing the different dataset types. If you think this is not only superfluous, but even confusing, we will remove the key here.

--------------------------------------------

```{r child="reply-additional.Rmd"}
```

## References

### 2022-07-25 analysis/paper/001-paper-main.Rmd
Address reviewer comments to manuscript 1.

@@
 However, the advantages of such models rely on their validity and accuracy.
-Recently, \cite{Hodgkins.2018} developed models to quantitatively predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major fractions of organic matter.
-The models may have the potential to understand large-scale SOM gradients and have been used in various studies.
+Recently, \cite{Hodgkins.2018} developed models to quantitatively predict peat holocellulose and Klason lignin contents, two indicators of SOM stability and major fractions of organic matter.
+The models may help to understand large-scale SOM gradients and have been used in various studies.
     A research gap to fill is that these models have not been validated in detail yet.
@@
 Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid such unrealistic predictions.
-Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples.
+Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples (e.g.
+peat with mineral admixtures or potentially mineral soils).
     A key step to improve the models will be to collect training data that is representative for SOM formed under various conditions.
@@
 This makes model validation a crucial step during model development.
-OM derived from plants and SOM is often characterized by step-wise chemical fractionation into holocellulose (acid soluble) --- a proxy for polysaccharides --- and lignin (acid insoluble) --- a proxy for aromatics --- [@DelaCruz.2016; @Elle.2019] and these variables are often important indicators for the chemical quality of OM in studies and models analyzing decomposition of SOM [@Leifeld.2012; @Biester.2014; @Worrall.2017; @Hodgkins.2018; @Bengtsson.2018; @Agren.1996a; @Bauer.2004; @Shipley.2021].
+Soil and plant OM is often characterized by step-wise chemical fractionation into holocellulose (acid soluble) --- a proxy for polysaccharides --- and lignin (acid insoluble) --- a proxy for aromatics --- [@DelaCruz.2016; @Elle.2019] and these variables are often important indicators for the chemical quality of OM in studies and models analyzing decomposition of SOM [@Leifeld.2012; @Biester.2014; @Worrall.2017; @Hodgkins.2018; @Bengtsson.2018; @Agren.1996a; @Bauer.2004; @Shipley.2021].
 If these fractions can be predicted from MIRS, it may be possible to understand decomposition across larger scales and at higher spatial resolution.
@@
 @Peltre.2011; @Sun.2011].
-Most of these models consider only wood, material from few species, or specific vegetation organs [@Elle.2019].
+Most of these models consider only wood, material from few woody and non-woody species, or only specific vegetation organs [@Elle.2019].
 Another problem is that most of these models cannot be easily reproduced and used since the raw data and code have not been published.
@@
 The models have several advantages over existing approaches.
-The training data comprises several different OM types, such as paper products (cardboard, office paper, magazines, newspaper), leaves from diverse species (trees and graminoids), and wood samples [@DelaCruz.2016; @Hodgkins.2018].
+The training data comprise several different OM types, such as paper products (cardboard, office paper, magazines, newspaper), leaves from diverse species (trees and graminoids), and wood samples [@DelaCruz.2016; @Hodgkins.2018].
 Moreover, both raw data and model code are publicly available [@Hodgkins.2018; @Teickner.2020a].
@@
 by including additional data.
-A major problem is that neither @Hodgkins.2018, nor a later study using the models provided a thorough model validation.
-To compute the models, the authors developed a procedure to extract peaks which are indicative for specific OM fractions, such as holocellulose and Klason lignin (Fig.
+A major problem is that neither @Hodgkins.2018, nor any later study which was cited above and which used the models provided a thorough validation of the models.
+To compute the models, @Hodgkins.2018 developed a procedure to extract peaks which are indicative for specific OM fractions, such as holocellulose and Klason lignin (Fig.
 \@ref(fig:hodgkins-peaks-sample-fig01)), from the MIRS.
-The peaks are baseline corrected, their maximum height computed and divided by the sum of absorbance values in the spectra.
-These values are used to calibrate the prediction models.
-As model validation, only the linearity between the target variables and predictors was assessed (supplementary material to @Hodgkins.2018), but not for example how predicted values match measured values.
-Moreover, no information has been provided whether the training data is representative for SOM, particularly the peat samples analyzed in the study.
+In the procedure, the peaks are baseline corrected, their maximum height computed, and normalized --- divided by the sum of absorbance values in the spectra.
+In @Hodgkins.2018, these values were used to calibrate the prediction models.
+As model validation, only the linearity between the target variables and predictors was assessed (supplementary material to @Hodgkins.2018), but not, for example, how predicted values match measured values.
+Moreover, no information has been provided whether the training data are representative for SOM, particularly the peat samples analyzed in the study.
 Furthermore, it is unclear if pre-selecting peaks may reduce the predictive accuracy of the models.
@@
 
-Our goals are to identify limitations of the original models and to give concrete recommendations for improvements.
+Our goals are to identify limitations in the original models and to give concrete recommendations for improvements.
 Moreover, we use the original data from @Hodgkins.2018 to provide improvements on the original models where possible.
@@
 This can potentially result in unrealistic predictions.
-However, if no unrealistic predictions occur for "representative" peat samples, using a model with normal distribution can be justifiable in practice due the central limit theorem.
-In contrast, if unrealistic predictions occur, it is more reasonable to use a distribution which assumptions are in accordance with knowledge on the data generating process.
-For example, the beta distribution assumes a lower and upper bound which can be mapped to the interval [0, 100] mass-%.
+However, if no unrealistic predictions occur for "representative" peat samples, using a model with normal distribution can be justifiable in practice.
+In contrast, if unrealistic predictions occur, it is more reasonable to use a distribution for which assumptions are consistent with knowledge on the data generating process.
+For example, the beta distribution assumes a lower and upper bound which can be mapped to the interval (0, 100) mass-%.
+It can be used as replacement for the normal sampling distribution used in ordinary linear regression models, such as the original models.
 For this reason, we were interested if and under which conditions the original models may produce unrealistic predictions for the peat and vegetation samples.
@@
 3.
-_Is the prediction domain of the training data representative for peat samples?_ If a model extrapolates outside the range of predictor variable values in the training data set, it is unclear if the predictions are valid.
-An advantage of the independent reference sample set used by @Hodgkins.2018 is that the model is applicable on various OM types and e.g.
-no problem-specific correlations are used by the models to make predictions [@Hodgkins.2018].
+_Is the prediction domain of the training data representative for peat samples?_ If a model extrapolates outside the prediction domain --- the range of predictor variable values in the training data set [@Wadoux.2021] --- it is unclear if the predictions are valid.
+An advantage of the independent reference sample set used by @Hodgkins.2018 is that the model is applicable to various OM types and, for example, does not depend on obviously material-specific spectral characteristics [@Hodgkins.2018].
 For example, in peat, holocellulose content is controlled by decomposition, and other structures also controlled by decomposition might therefore provide information on holocellulose contents [@Leifeld.2012; @Biester.2014].
-But this does not have to be the case for other samples, e.g.
+However, this does not have to be the case for other samples, e.g.
 peat forming vegetation.
-An advantage is therefore the generality of the training data set @Hodgkins.2018 used.
+A potential advantage is therefore the potential generality of the training data set @Hodgkins.2018 used.
 However, since it does not include peat samples, it is questionable if the prediction domain formed by the training samples covers peat samples and peat forming vegetation in general, and this can result in overlooking prediction failures due to extrapolation.
-For this reason, we were interested if the training data (its predictions domain, @Wadoux.2021) is representative for peat and peat forming vegetation.
+For this reason, we were interested if the training data (its prediction domain, @Wadoux.2021) is representative for peat and peat forming vegetation.
 
@@
 Model comparison often reveals detailed insights into the mechanics of a model and may provide hints to which model makes more correct predictions.
-We therefore were interested in analyzing differences between the models computed, what factors may cause differences in predictions, and if there are indications which models make more correct predictions.
+We therefore were interested in analyzing differences between the computed models, what factors may cause differences in predictions, and if there are indications which models make more correct predictions.
 
@@
 This decision is sensible, but since there are many organic soils with mineral admixtures --- for example due to volcanic ash, in base layers, or due to cryoturbation) [e.g.
-@Broder.2012; @Bockheim.2007; @Koven.2009; @Loisel.2014] --- we think that it is useful to have a model to predict holocellulose contents also for mineral rich samples.
+@Broder.2012; @Bockheim.2007; @Koven.2009; @Loisel.2014] --- we think that it is useful to have a model to predict holocellulose contents that includes mineral-rich samples.
 We therefore wanted to investigate if a suitable model can be computed by including the previously omitted old magazine samples during calibration.
+This is no replacement for a model using more training data, but it is a test whether it is at least in principle possible for MIRS to contain sufficient information to predict holocellulose contents of relatively diverse samples.
 
@@
 During this, we also computed improved models using the same training data that was used to compute the original models.
-Since we used these improved models to uncover and analyze the limitations of the original models, we first describe steps how we improved the original models and then how we analyzed the limitations.
+Since we used these improved models to uncover and analyze the limitations in the original models, we first describe steps how we improved the original models and then how we analyzed the limitations.
 We use the data and models collected and developed by @Hodgkins.2018.
 The training data used to calibrate the original models consist of various organic materials and comprise measured values for holocellulose and Klason lignin [@DelaCruz.2016].
-The data are available via @Hodgkins.2018.
+The data are accessible from the supplementary information of @Hodgkins.2018.
 In addition, we used a data set consisting of MIRS for `r length(unique(d_peat$core_label))` peat cores (`r nrow(d_peat)` samples in total) and `r nrow(d_veg)` vegetation samples collected at the same sites which are also provided by @Hodgkins.2018.
 We used these peat and vegetation data to evaluate the performance of the original and improved models for peat and peat forming vegetation in general.
-We assume that the peat samples are representative for a wide range of SOM because they cover a large latitudinal range and a wide range of degrees of decomposition, and were formed by different vegetation [@Hodgkins.2018].
+We assume that the peat samples are representative for a wide range of peat OM because they cover a large latitudinal range and a wide range of degrees of decomposition, and were formed by different vegetation [@Hodgkins.2018].
 We computed all models as Bayesian models.
 This allowed us to consider parameter uncertainty in predictions and to compute models with good predictive performance [@Raftery.2003; @Piironen.2017].
-For this reason, we also recomputed the original models as Bayesian models with weakly informative priors to facilitate model comparison using the same metrics and tools.
+For this reason, to facilitate model comparison, we also recomputed the original models as Bayesian models with weakly informative priors.
+With weakly informative priors, we mean here priors which result in approximately the same prediction intervals as the original frequentist models (supplementary Fig.
+1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).
 Whenever we refer to "original model", we mean these Bayesian translations, and otherwise refer to "original non-Bayesian models".
@@
 All predictor variables were $z$-transformed.
-MCMC sampling was validated in addition using bayesplot (`r packageVersion("bayesplot")`) [@Gabry.2020].
+Markov Chain Monte Carlo (MCMC) sampling was validated in addition using bayesplot (`r packageVersion("bayesplot")`) [@Gabry.2020].
 The out-of-sample predictive performance of the models was estimated using the expected log predictive density (ELPD) estimated using Pareto smoothed importance sampling-leave-one-out cross-validation (PSIS-LOO) [@Vehtari.2017] using the loo package (`r packageVersion("loo")`) [@Vehtari.2019].
@@
 We were interested if and under which conditions unrealistic predictions can occur.
-Here, we differentiate two ways in which a prediction can be unrealistic: It can be either an unrealistic point estimate (median predictions), or the prediction interval can cover values outside [0, 100] mass-%, even though the median prediction can be in [0, 100] mass-%.
-For our analysis, we first computed a sequence of normalized `carb` and `arom15arom16` peak heights such that the original model predict values completely covering the interval [0, 100] mass-% for holocellulose and Klason lignin, respectively.
+Here, we differentiate two ways in which a prediction can be unrealistic: It can be either an unrealistic point estimate (median predictions), or the prediction interval can cover values outside [0, 100] mass-%, even though the median prediction can be within [0, 100] mass-%.
+For our analysis, we first computed a sequence of normalized `carb` and `arom15arom16` peak heights such that the original model predict values which completely covered the interval [0, 100] mass-% for holocellulose and Klason lignin, respectively.
 For these simulated peak heights, we computed median model predictions and 90% prediction intervals using the original models ("original Gaussian models").
@@
 After identifying conditions under which unrealistic predictions occur, we compared this to the behavior of a model with assumptions in line with the data generating process.
-Beta regression models assume all values of the dependent variable to be in (0,100) mass-% which is a reasonable assumption given how the actual data is generated.
+Beta regression models assume all values of the dependent variable to be in (0,100) mass-% which is a reasonable assumption given how the actual data are generated.
 Therefore, we repeated the previous analyses, but now using beta regression models.
 We used the same model structure and priors for intercept and slopes as the original models ("original beta models").
-To facilitate practical comparison between the modeling approaches, we compared the predictions of all models in the form of depth profiles of the holocellulose and Klason lignin contents.
+To facilitate practical comparison between the modeling approaches, we compared the predictions of all models for the peat samples by plotting predicted values versus depths of the peat samples.
 Altogether, this allowed us to identify unrealistic predictions of the original models, and to analyze how the improved models (beta regression models) perform in comparison.
@@
 The disadvantage is that additional information contained within the spectra cannot be fully exploited if these are not contained within the extracted peak heights.
-To avoid overfitting, we computed models using priors implying different amount of regularization for model coefficients (standard deviation of 1 and 0.5, respectively) in addition to the default weakly informative prior which we derived from rstanarm (`r packageVersion("rstanarm")`) (standard deviation of 2.5) [@Goodrich.2020].
+To avoid overfitting, we computed models using priors implying different amounts of regularization --- shrinking coefficients to 0 with the aim to avoid overfitting --- of model coefficients (standard deviation of 1 and 0.5, respectively) in addition to the default weakly informative prior which we derived from rstanarm (`r packageVersion("rstanarm")`) (standard deviation of 2.5) [@Goodrich.2020].
 
 2.
-We used all variables in the spectra and Bayesian regularization [@Piironen.2020], similar to [@Teickner.2022].
+We used all variables in the spectra and Bayesian regularization [@Piironen.2020], similar to @Teickner.2022.
 The advantage is that this approach can more fully exploit the information contained within the spectra because it is not restricted to specific peaks, resulting potentially in a better predictive performance.
@@
 
+An alternative, popular, approach to approach 2 would be dimension reduction, for example via partial least squares regression, principal component regression, or variants of these [@Xiaobo.2010].
+In general, there are many alternative approaches which could be tested to use more information contained within the spectra than the original models, and many of these probably would result in similar predictive performances as approach 2 (regularization), especially when sample sizes are small [@Xiaobo.2010; @Teickner.2022].
+An advantage of regularization is that model coefficients are estimated more independently than in dimension reduction approaches, which makes it more straightforward to interpret model coefficients.
+The key is that the approaches we chose are suitable to analyze our research questions.
 Out-of-sample model performance was estimated with PSIS-LOO ELPD as described above [@Vehtari.2017; @Vehtari.2019].
@@
 Is the training data used to compute the models representative for the spectral properties of peat and peat forming vegetation?
-To answer this question, we needed to compare the spectral properties of the training samples to those of the peat and vegetation samples [@Wadoux.2021]: If for a sample the value of a spectral variable included in a model exceeds the range of values for the same variable in the training data, the model extrapolates if applied on the sample [@Wadoux.2021].
+To answer this question, we needed to compare the spectral properties of the training samples to those of the peat and vegetation samples [@Wadoux.2021]: If for a sample the value of a spectral variable included in a model exceeds the range of values for the same variable in the training data, the model extrapolates if applied to the sample [@Wadoux.2021].
 If extrapolation occurs, it is unclear if the predictions are valid estimates for holocellulose and Klason lignin contents.
@@
 Therefore, we could directly compare the area normalized heights of the `carb` and `arom15arom16` peak of the peat and vegetation samples to the range of area normalized heights of the same peaks for the training data.
-For the improved models using binned spectra, the predictor variables represent a multivariate prediction domain.
+For the improved models using binned spectra, the predictor variables form a multivariate prediction domain.
 We therefore created plots with which we could compare the value ranges for each bin for the training data with the respective values in the peat and vegetation spectra.
@@
 @Pineiro.2008].
-We therefore created such plots for the original model and the improved model and compared indication of bias between both.
+We therefore plotted measured values against predicted values for the original model and the improved model and compared indication of bias between both.
 In a second step, we compared how predictions of the improved models differ in practice from the original models.
@@
 Colors differentiate predictions of the original Gaussian model (blue) and the model with beta distribution (red).
-White filled points indicate the fitted values for which the 90\\% prediction interval of the Gaussian model contains unrealistic values.
-Dashed grey lines represent the limit for realistic point predictions."}
+White filled points indicate the fitted values for which the 90\\% prediction interval of the Gaussian model contains unrealistic values."}
 p_gaussian_beta_range
@@
 For Klason lignin, the improved model does not use bins near to the `arom15` and `arom16` peaks, in contrast to the original model.
-Instead bins with large absolute coefficients are located in the fingerprint region and probably related to aromatic in plane C-H bending (~1150 and ~1270 cm$^{-1}$) [@Stuart.2004] (Fig.
+Instead, bins with large absolute coefficients are located in the fingerprint region (600 to 1500 cm$^{-1}$) and probably related to aromatic in plane C-H bending (~1150 and ~1270 cm$^{-1}$) [@Stuart.2004] (Fig.
 \@ref(fig:res-p-coefficients-fig03), last panel).
-We have no explanation for why one coefficient is negative and the other positive.
+A plausible explanation for the negative sign of the coefficient for the 1150 cm$^{-1}$ bin is that absorbance in this range is partly caused by syringyl units (S) [@Kubo.2005] and that training samples with higher S content have smaller Klason lignin contents (supplementary Fig.
+4, @DelaCruz.2016), making this bin indicative of smaller Klason lignin contents.
+Likewise, a plausible explanation for the positive sign of the coefficient for the 1270 cm$^{-1}$ bin is that absorbance in this range is caused predominantly by guaiacyl units (G) [@Kubo.2005] and that training samples with higher G content have smaller Klason lignin contents (supplementary Fig.
+4, @DelaCruz.2016), making this bin indicative of larger Klason lignin contents.
 In Sect.
@@
 \@ref(fig:res-p-prediction-domain-fig04) (c)).
-Even though it is not possible to establish a direct relation between extrapolation for individual spectral variables and predicted contents, it is visible that for several bins, peat samples --- and to a lesser extent --- vegetation samples have mostly larger standardized predictor values than covered by the training data set.
+Even though it is not possible to establish a direct relation between extrapolation for individual spectral variables and predicted contents, it is visible that for several bins, peat samples and --- to a lesser extent --- vegetation samples have mostly larger standardized predictor values than covered by the training data set.
 This is also true for bins with the largest absolute median coefficients (including the model for holocellulose fitted with mineral-rich samples; see Sect.
@@
 We assume that this is probably also true for non-peat SOM.
-Therefore the models' predictions can represent extrapolations in practice; the training data is not in general representative for peat and peat-forming vegetation.
+Therefore the models' predictions can represent extrapolations in practice; the training data are not in general representative for peat and peat-forming vegetation.
 
@@
 For Klason lignin, we provide mechanistic evidence for why the predictions of the improved models probably are more correct.
-Therefore, a key result of our analysis is that additional training and validation data is required to compute models to accurately predict SOM holocellulose and Klason lignin contents from MIRS.
+Therefore, a key result of our analysis is that additional training and validation data are required to compute models to accurately predict SOM holocellulose and Klason lignin contents from MIRS.
 
@@
 For example, decomposition of phenols and disruption of tissue structures with hydrogen bonds result in a smaller absorbance around the OH-peak [@Kubo.2005; @Schellekens.2015].
-The training data contain only undecomposed or industrially processed OM [@DelaCruz.2016; @Hodgkins.2018] which do not reflect such decomposition changes and therefore generally have a larger OH peak (middle row in supplementary Fig.
+The training data contain only undecomposed or industrially processed OM [@DelaCruz.2016; @Hodgkins.2018] which do not reflect such decomposition changes and therefore generally have a larger OH peak (supplementary Fig.
 7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).
@@
 This is what distinguishes samples of class 1 and 2 in the training data (supplementary Fig.
-11<!---\@ref(fig:res-p-pred-diff-train2)--->): Samples of class 1 are wood samples.
+11<!---\@ref(fig:res-p-pred-diff-train2)--->): Samples of class 1 are wood samples and paper product samples derived from wood.
 Wood typically has smaller nitrogen contents, but larger Klason lignin contents [@Cowling.1966; @Aerts.1999; @DelaCruz.2016] which explains the large contribution of the `arom15` peak, indicative for aromatic skeletal vibrations (but not for proteins) [@Stuart.2004], to `arom15arom16`.
@@
 \@ref(fig:res-p-coefficients-fig03)).
-We suggest that these variables are better predictors for Klason lignin because they interfere less with other molecular structures, such as proteins [@Stuart.2004].
-Overall, this mechanistic explanation provides additional evidence that the improved model using binned spectra probably makes more correct predictions.
+We suggest that these variables are better predictors for Klason lignin because they are interfered less by other molecular structures, such as proteins [@Stuart.2004].
+Overall, this mechanistic explanation provides additional evidence why the improved model using binned spectra probably makes more correct predictions.
 
@@
 The color gradient represents \\texttt{arom15arom16} values.
-The diagonal line represents values where measured and fitted values are identical.
-If points are below the line, the model overestimates contents.'}
+The diagonal line represents values where measured and fitted values are identical.'}
 p_pred_diff_train1_p2
@@
 We see this as proof-of-concept that holocellulose contents can also be predicted from MIRS for SOM samples with mineral admixtures.
-If the clay rich old magazine samples are included, the model using binned spectra had the best average predictive performance (ELPD on average `r round(abs(loo_comparison_mineral[2, 1]), 2)` larger).
+If the clay rich old magazine samples are included, the model using binned spectra had the best average predictive performance.
 The model using extracted peaks had a worse, but similar performance ($\Delta\text{ELPD}= `r round(loo_comparison_mineral[2, 1], 2)`$, $\Delta\text{SE}= `r round(loo_comparison_mineral[2, 2], 2)`$).
@@
 \@ref(fig:res-p-coefficients-fig03) (middle panel), the "mineral-rich model" down-weights bins near the `carb` peak, in contrast to the improved model not trained on old magazine samples, and instead has a larger coefficient for an additional peak related to aromatic C=C stretching and amide N-H bending and C-N stretching (~1590 cm$^{-1}$) [@Stuart.2004] which corresponds to the `arom15` peak.
-This provides further evidence that it is possible to infer the holocellulose content via aromatics.
+This provides further evidence that it is possible to infer holocellulose contents via aromatics.
 
@@
 Instead, it is more effective to include a larger set of variables in a model and to use regularization to avoid overfitting and find relations between predictor variables and the target variable [@Piironen.2017].
-Second, a linear relation between the target variable to predict and a predictor variable is no sufficient validation of a spectral prediction model if the training data is not representative for the samples on which the model is applied.
-Most importantly, due to spectral normalization, predictions can even be sensitive to variables not included in the model.
-Therefore, to assess if training data (the prediction domain) is representative, whole spectra have to be compared.
-Third, it is helpful to identify potential causal mechanisms which may affect differently the MIRS the model will be applied on than the MIRS it was trained on.
+Second, a linear relation between the target variable being predicted and a predictor variable is no sufficient validation of a spectral prediction model if the training data are not representative for the samples to which the model is applied.
+Most importantly, due to spectral normalization, predictions can be sensitive even to variables not included in the model.
+Therefore, to assess if training data (the prediction domain) are representative, whole spectra have to be compared.
+Third, it is helpful to identify potential causal mechanisms which may affect differently the MIRS the model will be applied to than the MIRS it was trained on.
 As shown here (Sect.
 3.5), differences in the degree of decomposition or the relative contribution of proteins and aromatic skeletal structures cause differences in MIRS which can result in biased predictions.
-Consequently, providing a causal explanation what causes the correlation between  a target variable and specific MIRS variables is a useful tool to assess if a model may be applicable for new samples.
+Consequently, providing a causal explanation as to what causes the correlation between  a target variable and specific MIRS variables is a useful tool to assess if a model may be applicable to new samples.
 Fourth, our analysis also shows that re-evaluating existing models with their original training data can be an effective way to improve the models.
@@
 There are several problems we could not solve.
-The two most important problem is that also for the improved models, it is unclear if predictions are correct outside the prediction domain of the training data and therefore if the models are applicable to SOM in general.
-Further issues are: (1) The overall accuracy of the models probably can be improved with more training samples.
-(2) It is unclear how robust the original models and our improved models --- especially models using binned spectra --- are in terms of calibration transfer.
-We assume the binning and estimation procedure is relatively robust since bin widths are comparatively broad and all variables were standardized, meaning that only between sample variability in intensities are relevant.
+The most important problem is that also for the improved models, it is unclear if predictions are correct outside the prediction domain of the training data and therefore if the models are applicable to SOM in general.
+Further issues are: (1) The overall accuracy of the models can certainly be improved with more training samples.
+(2) It is unclear how robust the original models and our improved models --- especially models using binned spectra --- are in terms of calibration transfer (calibration transfer is the application of a model to spectra measured differently than the training data, e.g.
+with a different procedure, on a different device, in a different laboratory [@Workman.2018]).
+We assume that the binning and estimation procedure are relatively robust since bin widths are comparatively broad and all variables were standardized, meaning that only between sample variability in intensities are relevant.
 However, this remains to be tested.
-(3) Even though we computed a model that could fit and predict mineral rich samples, it is unclear if this holds in general.
+(3) Even though we computed a model that could fit and predict mineral rich samples, it is unclear whether this is possible in general.
 (4) Some of our models had computational issues (holocellulose: `r hklmirs::format_numbers(sum(d_models$n_divergent_transitions[d_models$variable_y == "Holocellulose"] > 0))` model with maximally `r hklmirs::format_numbers(max(d_models$n_divergent_transitions[d_models$variable_y == "Holocellulose"]))` divergent transitions, Klason lignin: `r hklmirs::format_numbers(sum(d_models$n_divergent_transitions[d_models$variable_y == "Klason lignin"] > 0))` models with maximally `r hklmirs::format_numbers(max(d_models$n_divergent_transitions[d_models$variable_y == "Klason lignin"]))` divergent transitions) that we could not resolve.
-Our analysis thus opens concrete and promising directions to further improve the models: We need training and validation data that includes SOM, particularly highly decomposed peat.
+A further limitation is that, as the original models, the modified beta regression models do not consider the constraint that the contents of holocellulose, Klason lignin, and any remaining compounds should sum to 100 mass-%.
+This also represents a further test how realistic model predictions are (compare with Sect.
+2.1).
+In principle, this constraint could be considered by using a Dirichlet regression model [e.g.
+@Douma.2019].
+We have not used this approach here due to the higher computational costs, potential computational difficulties, to keep the present model validation straightforward, and since the training data do not fulfill this condition for all samples (supplementary Fig.
+4; indicating that the measurement procedure needs to be improved, too).
+In summary, our analysis opens concrete and promising directions to further improve the models: We need training and validation data that include peat, particularly highly decomposed peat.
 Such data make it possible to analyze the impact of the bias and to compute models with less bias, higher representativity for SOM and peat, and potentially larger predictive accuracy.
@@
 Ideally, such improvements would be performed across multiple labs with an archive of sample materials such that calibration transfer of the models between different mid infrared spectrometers can be further explored.
-To support such developments, we implemented the best models using binned spectra and extracted peaks and also the model using binned spectra that could predict mineral-rich samples into the R package irpeat [@Teickner.2020a].
+To support such developments, we implemented the best models using binned spectra (for holocellulose the model which was also trained on mineral-rich samples) into the R package irpeat [@Teickner.2020a].
+The other models can be reproduced from the reproducible research compendium [@Teickner.2022b].
 
@@
 This results in biased predictions for holocellulose and Klason lignin for SOM, such as peat.
-Results from currently published studies using the original models should be interpreted with caution.
+Results from currently published studies using the original models should be interpreted with caution (we are currently preparing a manuscript, see @Teickner.2022b, which explores the implications of our results here for the results of @Hodgkins.2018).
 Manual variable selection favored this bias because it excluded information critical for better model fits.
-Finally, the original models can produce unrealistic (smaller than 0 or larger than 100 mass-%) predictions and prediction intervals, also in practice.
+Finally, the original models can produce unrealistic (smaller than 0 or larger than 100 mass-%) predictions and prediction intervals.
 Even though it was impossible to address the key problem of unrepresentative training data using the original data, we could address some of these issues, provide improved models, and develop a concrete strategy for future improvements.

### 2022-07-25 analysis/paper/002-reply-main.Rmd
Address reviewer comments to manuscript 1.

@@
 ---
-title: 'Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging'
+title: 'Comment on Hodgkins et al. (2018): Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging'
 author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"

### 2022-07-25 analysis/paper/003-paper-m-gaussian-beta.Rmd
Address reviewer comments to manuscript 1.

@@
   dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
+  dplyr::mutate(
+    model =
+      dplyr::case_when(
+        model == "Gaussian, Bayesian" ~ "Gaussian",
+        TRUE ~ model
+      )
+  ) %>%
   ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
@@
              ncol = 2) +
-  scale_x_reverse()
+  scale_x_reverse() +
+  guides(color = guide_legend(title = "Bayesian model"), fill = guide_legend(title = "Bayesian model"))
 
@@
   dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
+  dplyr::mutate(
+    model =
+      dplyr::case_when(
+        model == "Gaussian, Bayesian" ~ "Gaussian",
+        TRUE ~ model
+      )
+  ) %>%
   ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
@@
   dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
+  dplyr::mutate(
+    model =
+      dplyr::case_when(
+        model == "Gaussian, Bayesian" ~ "Gaussian",
+        TRUE ~ model
+      )
+  ) %>%
   ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
@@
   geom_path() +
-  geom_point(data = d_peat_pred %>% dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian"), 
-             mapping = aes(x = fit_gaussian, y = fit)) +
-  geom_point(data = d_veg_pred %>% dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian"), 
-             mapping = aes(x = fit_gaussian, y = fit), shape = 4) +
+  geom_point(
+    data = 
+      d_peat_pred %>% 
+      dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
+      dplyr::mutate(
+        model =
+          dplyr::case_when(
+            model == "Gaussian, Bayesian" ~ "Gaussian",
+            TRUE ~ model
+          )
+      ), 
+    mapping = aes(x = fit_gaussian, y = fit)
+  ) +
+  geom_point(
+    data = 
+      d_veg_pred %>% 
+      dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
+      dplyr::mutate(
+        model =
+          dplyr::case_when(
+            model == "Gaussian, Bayesian" ~ "Gaussian",
+            TRUE ~ model
+          )
+      ), 
+    mapping = aes(x = fit_gaussian, y = fit), shape = 4
+  ) +
   geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose"), 
@@
   scale_x_continuous(labels = function(x) x * 100) +
-  scale_y_continuous(labels = function(x) x * 100)
+  scale_y_continuous(labels = function(x) x * 100) +
+  guides(color = guide_legend(title = "Bayesian model"), fill = guide_legend(title = "Bayesian model"))
 
@@
   dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
+  dplyr::mutate(
+    model =
+      dplyr::case_when(
+        model == "Gaussian, Bayesian" ~ "Gaussian",
+        TRUE ~ model
+      )
+  ) %>%
   ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
@@
   geom_path() +
-  geom_point(data = d_peat_pred %>% dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian"), 
-             mapping = aes(x = fit_gaussian, y = fit)) +
-  geom_point(data = d_veg_pred %>% dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian"), 
-             mapping = aes(x = fit_gaussian, y = fit), shape = 4) +
+  geom_point(
+    data = 
+      d_peat_pred %>% 
+      dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
+      dplyr::mutate(
+        model =
+          dplyr::case_when(
+            model == "Gaussian, Bayesian" ~ "Gaussian",
+            TRUE ~ model
+          )
+      ), 
+    mapping = aes(x = fit_gaussian, y = fit)
+  ) +
+  geom_point(
+    data = 
+      d_veg_pred %>% 
+      dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
+      dplyr::mutate(
+        model =
+          dplyr::case_when(
+            model == "Gaussian, Bayesian" ~ "Gaussian",
+            TRUE ~ model
+          )
+      ), 
+    mapping = aes(x = fit_gaussian, y = fit), shape = 4
+  ) +
   geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Klason lignin"), 

### 2022-07-25 analysis/paper/004-paper-m-reduce-underfitting.Rmd
Address reviewer comments to manuscript 1.

@@
   ) %>%
+  dplyr::select(-original_model) %>%
   tibble::remove_rownames() %>%
   kableExtra::kable(booktabs = TRUE, 
-                    col.names = c("Model", "Original model?", "Distribution", "Predictors", "Prior scale", "Bin width", "ELPD", "$\\Delta$ELPD", "$\\Delta$SE"), 
+                    col.names = c("Model", "Distribution", "Predictors", "Prior scale", "Bin width", "ELPD", "$\\Delta$ELPD", "$\\Delta$SE"), 
                     escape = FALSE,
@@
 For each variable, the model with the best average predictive performance (largest ELPD) is at the top and the other models follow in descending order.
-Models ending in ".2" and ".3" are the models with the original model structure as developed by \\cite{Hodgkins.2018} (column "Original model?").
+Models ending in ".2" and ".3" are the models with the original model structure as developed by \\cite{Hodgkins.2018}.
 "Distribution" is the distribution assumed for the target variable.
@@
 "Bin width" is the width of bins (in wavenumber units).
-"ELPD" is the PSIS-LOO expected log-predictive density, $\\Delta$ELPD the difference in ELPD relative to the best average model, and $\\Delta$SE the standard error in the average $\\Delta$ELPD.
+"ELPD" is the PSIS-LOO expected log-predictive density, $\\Delta$ELPD the difference in ELPD relative to the average ELPD of the on average best model, and $\\Delta$SE the standard error in the average $\\Delta$ELPD.
 \\label{tab:res-tab-loo-comparison-res}') %>%
@@
 
-# bin width: 50
+# bin width: 20
 d_veg_bin20_flat <-
@@
 ```
+
+```{r m-reduce-underfitting-11}
+# predictions
+d_peat_pred_best_sum <- 
+  dplyr::bind_rows(
+    magrittr::add(
+      m1.8 %>%
+        rstanarm::posterior_predict(newdata = d_peat_bin20),
+      m2.8 %>%
+        rstanarm::posterior_predict(newdata = d_peat_bin20)
+    ) %>%
+      as.data.frame() %>%
+      purrr::map_df(function(x) {
+        tibble::tibble(
+          fit = median(x),
+          lwr = quantile(x, prob = 0.05),
+          upr = quantile(x, prob = 0.95)
+        )
+      }) %>%
+      dplyr::mutate(
+        model = "Best binned spectra",
+        variable_y = "Holocellulose + Klason lignin"
+      )
+  ) %>%
+  dplyr::bind_cols(
+    purrr::map_df(1, function(x) {
+      tibble::tibble(
+        sample_id = d_peat$sample_id, 
+        core_label = d_peat$core_label,
+        sampling_longitude = d_peat$sampling_longitude, 
+        sampling_latitude = d_peat$sampling_latitude, 
+        sampling_altitude = d_peat$sampling_altitude, 
+        sample_depth_lower = d_peat$sample_depth_lower, 
+        sample_depth_upper = d_peat$sample_depth_upper
+      )
+    })
+  )
+
+# predictions
+d_veg_pred_best_sum <- 
+  dplyr::bind_rows(
+    magrittr::add(
+      m1.8 %>%
+        rstanarm::posterior_predict(newdata = d_veg_bin20),
+      m2.8 %>%
+        rstanarm::posterior_predict(newdata = d_veg_bin20)
+    ) %>%
+      as.data.frame() %>%
+      purrr::map_df(function(x) {
+        tibble::tibble(
+          fit = median(x),
+          lwr = quantile(x, prob = 0.05),
+          upr = quantile(x, prob = 0.95)
+        )
+      }) %>%
+      dplyr::mutate(
+        model = "Best binned spectra",
+        variable_y = "Holocellulose + Klason lignin"
+      )
+  ) %>%
+  dplyr::bind_cols(
+    purrr::map_df(1, function(x) {
+      tibble::tibble(
+        sample_id = d_veg$sample_id, 
+        site_name = d_veg$site_name,
+        species_name = d_veg$species_name,
+        plant_part = d_veg$plant_part
+      )
+    })
+  )
+```

### 2022-07-25 analysis/paper/007-paper-m-prediction-differences.Rmd
Address reviewer comments to manuscript 1.

@@
   dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
-  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
+  ggplot(aes(x = intensity_3400 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
   geom_hline(yintercept = 0, colour = "grey50") +
@@
   geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
-  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
+  facet_grid(cut_number(carb, n = facet_n_level) ~ data_type) +
   labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
@@
   geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
-  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
+  facet_grid(cut_number(carb, n = facet_n_level) ~ data_type) +
   labs(y = "Prediction difference", x = expression("Intensity at 1250 cm"^{-1}), 
@@
   geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
-  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
+  facet_grid(cut_number(carb, n = facet_n_level) ~ data_type) +
   labs(y = "Prediction difference", x = expression("Intensity at 1590 cm"^{-1}), 
@@
   geom_histogram(aes(x = arom15arom16 %>% scale(), fill = data_type), bins = 30) +
-  facet_wrap(~ cut_number(arom15arom16_rel, n = 4)) +
+  facet_wrap(~ cut_number(ifelse(arom15arom16_rel < 6.32e-10, 0, arom15arom16_rel), n = 4)) +
   labs(y = "Count", x = "arom15arom16") +

### 2022-07-25 analysis/paper/008-paper-supplementary.Rmd
Update `renv.lock`, update figure format in supporting info for manuscript 1. Update the text. Update the readme.
Address reviewer comments to manuscript 1.

@@
                       cache = TRUE,
-                      dev = "png")
+                      dev = "pdf")
 ```
@@
 Here, we show that working within a Bayesian framework does not create the bias or any other weaknesses we discussed.
-To this end, we compare predictions of the original non-Bayesian models (implemented with `lm` [@RCoreTeam.2020]) with those of the original Bayesian model we used as baseline model throughout the main text.
+To this end, we compare predictions of the original non-Bayesian models (implemented with `lm` [@RCoreTeam.2020]) with those of the original Bayesian model we used as a baseline model throughout the main text.
 Figure \@ref(fig:res-sup-original-models) compares median fitted values, and lower and upper 90% prediction intervals for both models.
@@
 Figure \@ref(fig:res-sup-spectra-sim-p) shows the simulated spectra.
+The implication of the simulation is that, when the spectra are normalized by dividion by the sum of all absorbance values in each spectrum, all spectral variables which we use for interpretation or prediction (e.g.
+by selecting the `carb` peak, or a bin at a specific wavenumber) are dependent on all other spectral variables in the spectra, even if we do not include them in the model.
 
@@
 ```{r res-sup-spectra-sim-p, echo=FALSE, out.width="80%", fig.height=7, fig.width=7, fig.cap="Results of the simulation experiment.
-(a) Plots of the simulated spectra with different OH-peak heights (columns) and different carb peak heights (colour gradients).
-Due to the normalization (division of each absorbance value by the sum of all absorbance value), the normalized simulated carb peak heights get smaller as the OH-peak gets larger.
-For this reason, the extracted normalized carb peak height depends not only on the simulated carb peak height, but is also sensitive to the OH-peak height (b).
-This is also true for absorbance values extracted at specific wavenumbers (c)."}
+(a) Plots of the simulated spectra with different OH-peak heights (columns) and different \\texttt{carb} peak heights.
+In this panel, the colour gradient represents different \\texttt{carb} peak heights.
+(b) Due to the normalization (division of each absorbance value by the sum of all absorbance values), the normalized simulated \\texttt{carb} peak heights (y axis values) get smaller as the OH-peak gets larger (colour gradient).
+For this reason, the extracted normalized \\texttt{carb} peak height depends not only on the simulated \\texttt{carb} peak height, but is also sensitive to the OH-peak height.
+(c) This can also be seen when, for these simulated spectra, absorbance values are extracted at specific wavenumbers, e.g.
+1250 cm$^{-1}$."}
 spectra_sim_p
@@
 # Supplementary figures
-
-Figure \@ref(fig:res-p-gaussian-beta-depth-profile) shows depth profiles of holocellulose and Klason lignin contents for the peat cores from @Hodgkins.2018, predicted either with the original normal regression model or the beta regression model with the same model structure.
 
@@
 
-Figure \@ref(fig:res-sup-kl-hol) summarizes the _measured_ holocellulose and Klason lignin contents in the OM samples of the training data set of @Hodgkins.2018.
-
 ```{r sup-kl-hol, echo=FALSE}
@@
   coord_flip() +
-  labs(y = "Sample type", x = "Holocellulose + Klason lignin [mass-%]")
+  labs(x = "Sample type", y = "Holocellulose + Klason lignin [mass-%]")
 
@@
   plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
-  plot_layout(guides = "collect") &
+  plot_layout(guides = "collect", widths = c(1, 0.6)) &
   theme(legend.position = "bottom")
@@
 \clearpage
-
-Figure \@ref(fig:res-p-gaussian-beta-depth-profile-best) shows holocellulose and Klason lignin contents for the peat core samples from @Hodgkins.2018 predicted by the original model, the improved model using binned spectra, and the improved model using extracted peaks.
 
@@
 
-Figure \@ref(fig:res-p-y-yhat-best) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, and the improved model using extracted peaks.
-
 ```{r res-p-y-yhat-best, echo=FALSE, out.width="70%", fig.height=6.5, fig.width=6, fig.cap='Measured values of the training dataset from \\cite{Hodgkins.2018} versus fitted values of the original model (first column) in comparison to the improved models (second and third columns) for (a) holocellulose and (b) Klason lignin.
@@
 The diagonal lines indicates ideal fit.
-Points above the line indicate overestimation by the model and points under the diagonal underestimation.'}
+Points above the line indicate underestimation by the model and points under the diagonal overestimation.'}
 p_y_yhat_best
@@
 \clearpage
-
-Figure \@ref(fig:res-p-pred-diff-hol-spectra-diff2) shows MIRS for the training data, vegetation samples, and peat samples from @Hodgkins.2018, classified according to different `carb` peak heights and colored according to the prediction differences between the original model and the improved model using binned spectra for holocellulose.
-This plot makes visible the differences between the spectra for peat (and vegetation samples) in comparison to the training data.
 
@@
 
-Figure \@ref(fig:res-p-pred-diff-hol-intensity2) shows prediction differences between the original model and the improved model using binned spectra for holocellulose in dependency of absorbance values at different wavenumbers and for different classes of `carb` peak heights.
-
 ```{r res-p-pred-diff-hol-intensity2, warning=FALSE, echo=FALSE, out.width="50%", fig.height=14, fig.width=7, fig.cap='Prediction differences between the original model and improved model using binned spectra for the training (first row), peat (second row), and vegetation (bottom row) samples of \\cite{Hodgkins.2018} for different ranges of \\texttt{carb} peak heights (columns) versus intensities at different wavenumber values on the x axis (a: 3400, b: 1250, c: 1590 cm$^{-1}$).
@@
 \clearpage
-
-Figure \@ref(fig:res-p-pred-diff-kl-residuals) shows residuals of the original model, the improved model using binned spectra, and the improved model using extracted peaks versus standardized `arom15arom16` values and for different levels of the relative contribution of the `arom15` peak height to the `arom15arom16` value.
 
@@
 
-Figure \@ref(fig:res-p-pred-diff-kl-histogram1) shows `arom15arom16` values for the training data, peat, and vegetation samples from @Hodgkins.2018, for different levels of the relative contribution of the `arom15` peak height to the `arom15arom16` value.
-Conditional on this contribution, peat samples regularly have larger or smaller `arom15arom16` values.
-
 ```{r res-p-pred-diff-kl-histogram1, echo=FALSE, out.width="60%", fig.height=4, fig.width=5, fig.cap='Histograms of \\texttt{arom15arom16} values for the training, peat, and vegetation samples from \\cite{Hodgkins.2018} for different levels of the relative contribution of the \\texttt{arom15} peak to \\texttt{arom15arom16} (larger values mean the \\texttt{arom15} peak is larger relative to the \\texttt{arom16} peak).'}
@@
 \clearpage
-
-Figure \@ref(fig:res-p-pred-diff-train2) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, and the improved model using extracted peaks, while differentiating the sample type classes defined in @Hodgkins.2018.
-This makes visible that for Klason lignin, the two classes of training samples in the main text comprise different sample types (class 1: wood samples, type 2: leaves and grasses).
 
@@
 
-Figure \@ref(fig:res-p-y-yhat-minerals) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, the improved model using extracted peaks, the improved model using binned spectra trained with old magazine samples, and the improved model using extracted peaks trained with old magazine samples.
-
-```{r res-p-y-yhat-minerals, echo=FALSE, out.width="100%", fig.height=3.5, fig.width=7.5, fig.cap='Predicted median holocellulose contents versus measured holocellulose contents for the Gaussian model with original model structure (first panel), the improved best models for both approaches (model including all peaks from the peak extraction procedure of \\cite{Hodgkins.2018}, model using binnes spectra with a bin width of 50 cm$^{-1}$), but fitted only to samples without mineral interferences (second and third panel), and the same models fitted also to samples with mineral interferences (last two panels).
-The diagonal line represent identical fitted and measured values.
+```{r res-p-y-yhat-minerals, echo=FALSE, out.width="100%", fig.height=3.5, fig.width=7.5, fig.cap='Measured holocellulose contents versus predicted median holocellulose contents for the Gaussian model with original model structure (first panel), the improved best models for both approaches (model including all peaks from the peak extraction procedure of \\cite{Hodgkins.2018}, model using binned spectra with a bin width of 20 cm$^{-1}$), but fitted only to samples without mineral interferences (second and third panel), and the same models fitted also to samples with mineral interferences (last two panels).
+The diagonal lines represent identical measured and fitted values.
 Colours differentiate samples with (blue) and without (red) mineral interferences.'}
@@
 
-Figure \@ref(fig:res-p-oh-height-difference) shows differences in the absorbance at 3400 cm$^{-1}$ in area normalized spectra between the training, peat, and vegetation data.
-
 ```{r, res-p-oh-height-difference, out.width="60%", fig.height=3.5, fig.width=5.5, fig.cap='Absorbance (intensity) at 3400 cm$^{-1}$ in area normalized spectra between the training, peat, and vegetation data.'}

### 2022-07-25 analysis/peer_review_manuscript_1/1/reply-additional.Rmd
Address reviewer comments to manuscript 1.

## Additional changes

We made the following additional changes:

1. old: l. 4; new: l. 4: We changed the sentence "The models may have the potential to understand large-scale SOM gradients and have been used in various studies." to "The models may help to understand large-scale SOM gradients and have been used in various studies."

2. old: l. 40; new: l. 42: We changed "comprises" to "comprise".

4. old: l. 47; new: l. 49: We replaced "the authors" by "@Hodgkins.2018".

5. old: l. 49; new: l. 51: We replaced "... and divided by ..." by "..., and normalized --- divided by ..." to introduce the term "normalization" which is used later in the text and also in @Hodgkins.2018.

6. old: l. 52; new: l. 55: We replaced "data is" by "data are".

7. old: l. 65 to 66; new: l. 68: We removed "due the central limit theorem" because this applies only to mean values. However, we would see the application of a normal distribution only as practically justified when predictions for individual samples would not generate unrealistic values and prediction intervals.

8. old: l. 91; new: l. 96: We changed "models computed" to "computed models".

9. old: l. 151; new: l. 160: We replaced "data is" by "data are".

10. old: l. 168; new: l. 178: We changed "[@Teickner.2022]" to a textual citation.

11. old: l. 203; new: l. 220: We changed "represent" to "form".

12. old: l. 218; new: l. 235: We changed "We therefore created such plots" to "We therefore plotted measured values against predicted values".

14. old: l. 294; new: l. 316: We changed "samples --- and" to "samples and ---".

15. old: l. 307; new: l. 329: We replaced "data is" by "data are".

16. old: l. 328; new: l. 350: We replaced "data is" by "data are".

17. old: l. 352; new: l. 374: We removed "middle row in" as the described pattern is visible in rows 2 to 4 of the figure.

18. old: l. 414; new: l. 437: We changed "the holocellulose content" to "holocellulose contents".

19. old: l. 423 to 425; new: l. 444 to 448: We replaced "Most importantly, due to spectral normalization, predictions can even be sensitive to variables not included in the model. Therefore, to assess if training data (the prediction domain) is representative, whole spectra have to be compared." by "Most importantly, due to spectral normalization, predictions can be sensitive _even_ to variables not included in the model. Therefore, to assess if training data (the prediction domain) _are_ representative, whole spectra have to be compared." (changes emphasized). Additionally, we replaced "data is" by "data are" in the previous sentence.

20. old: l. 430; new: l. 453: We replaced "for" by "to".

21. old: l. 444; new: l. 468: We changed "We assume the binning and estimation procedure is relatively robust ..." to "We assume that the binning and estimation procedure are relatively robust ..."

22. old: l. 446; new: l. 471: We changed "if this holds in general" to "whether this is possible in general".

24. old: l. 456 to 457; new: l. 488 to 490: We corrected this statement to "To support such developments, we implemented the best models using binned spectra into the R package irpeat [@Teickner.2020a]. The other models can be reproduced from the reproducible research compendium [@Teickner.2022b]."

25. old: Tab. 1; new: Tab. 1: In the caption, we changed "$\Delta$ELPD the difference in ELPD relative to the best average model" to "$\Delta$ELPD the difference in ELPD relative to the average ELPD of the on average best model".

26. old : Fig. S2; new: Fig. S2: In the caption, we changed "all absorbance value" to "all absorbance values".

27. old : Fig. S2; new: Fig. S2: We improved the caption to make clearer what the different panels show.

28. old : Fig. S2; new: Fig. S2: We improved the text above this figure make clearer what the implication of the simulation is.

### 2022-07-29 README.Rmd
Add missing references in README to manuscript 1 (available as preprint).
Update README, increase version number.

@@
   manuscript2_yaml$title %>%
-  paste0(".")
+  paste0("_", ., "._")
 
@@
 
-> `r paste(manuscript1$authors, "(in preparation):", manuscript1$title)`
+> Henning Teickner, and Klaus-Holger Knorr (2022; preprint): _Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra_.
+Preprint.
+Soil and methods, June 1, 2022.
+DOI: [10.5194/soil-2022-27](https://doi.org/10.5194/soil-2022-27).
 
@@
     - [005-paper-m-minerals.Rmd](analysis/paper/005-paper-m-minerals.Rmd): Uses the models from `003-paper-m-gaussian-beta.Rmd` to test how accurate a model for holocellulose content is which is also calibrated on training samples with higher mineral contents.
-    - [006-paper-m-prediction-domain.Rmd](analysis/paper/006-paper-m-prediction-domain.Rmd): Analyzes the prediction domain [@Wadoux.2021] of the original models and the modified models and identifie under which conditions models extrapolate for peat and vegetation smaples from @Hodgkins.2018.
+    - [006-paper-m-prediction-domain.Rmd](analysis/paper/006-paper-m-prediction-domain.Rmd): Analyzes the prediction domain [@Wadoux.2021] of the original models and the modified models and identifies under which conditions models extrapolate for peat and vegetation samples from @Hodgkins.2018.
     - [007-paper-m-prediction-differences.Rmd](analysis/paper/007-paper-m-prediction-differences.Rmd): Compares predictions for the training data and the peat and vegetation data from @Hodgkins.2018 for the original models from @Hodgkins.2018 and the modified models from `004-paper-m-reduce-underfitting.Rmd`.
@@
     - [001-reply-main.Rmd](analysis/paper/001-reply-main.Rmd): This is the main script for manuscript 2.
-It is run from within `001-paper-main.Rmd` and produces the supplementary information for manuscript 2.
-    - [002-reply-main.Rmd](analysis/paper/002-reply-main.Rmd): This script produces the document for manuscript 2. It is run from within `001-reply-main.Rmd`.
+It is compiled from within `001-paper-main.Rmd` and produces the supplementary information for manuscript 2.
+    - [002-reply-main.Rmd](analysis/paper/002-reply-main.Rmd): This script produces the document for manuscript 2. It is compiled from within `001-reply-main.Rmd`.
 
@@
 
-To reproduce the analyses for the paper, open the Rstudio project included in this research compendium and run the Rmarkdown script in `analysis/paper/001-paper-main.rmd`.
+To reproduce the analyses for the paper, open the RStudio project included in this research compendium and run the Rmarkdown script in `analysis/paper/001-paper-main.rmd`.
 

### 2022-10-26 analysis/paper/001-paper-main.Rmd
Manuscript 1: Update figure numbering for supporting figures.

@@
 With weakly informative priors, we mean here priors which result in approximately the same prediction intervals as the original frequentist models (supplementary Fig.
-1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).
+S1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).
 Whenever we refer to "original model", we mean these Bayesian translations, and otherwise refer to "original non-Bayesian models".
 To check that we did not introduce any of the identified weaknesses using the Bayesian approach, we compared the predictions of the Bayesian models to the original non-Bayesian models (supplementary Fig.
-1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).
+S1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).
 All analyzes were performed in R (`r R.Version()$version.string %>% stringr::str_remove("^R ") %>% stringr::str_remove("\\)") %>% stringr::str_replace(" \\(", ", ")`) [@RCoreTeam.2020].
@@
 Moreover, for `r d_peat_pred %>% dplyr::filter(variable_y == "Holocellulose" & model == "Gaussian, Bayesian") %>% dplyr::mutate(negative = lwr < 0) %>% dplyr::summarise(negative = sum(negative)/length(negative)) %>% dplyr::pull(negative) %>% magrittr::multiply_by(100) %>% round(0)`% of the peat samples in the data from @Hodgkins.2018, the lower 90% prediction interval covered negative values (supplementary Fig.
-3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->).
+S3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->).
 No point prediction or 90% prediction interval was $>100$ mass-%.
 The original model for Klason lignin did not produce unrealistic predictions for the peat samples (supplementary Fig.
-3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->).
+S3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->).
 Both models did not produce unrealistic predictions for the vegetation samples.
@@
 \@ref(fig:res-p-gaussian-beta-range-fig02) and supplementary Fig.
-3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->).
+S3<!---\@ref(fig:res-p-gaussian-beta-depth-profile)--->).
 Thus, whereas our results indicate little differences in choosing a model distribution for Klason lignin, for holocellulose contents of peat samples it is crucial to use a beta regression model to avoid unrealistic predictions.
@@
 A plausible explanation for the negative sign of the coefficient for the 1150 cm$^{-1}$ bin is that absorbance in this range is partly caused by syringyl units (S) [@Kubo.2005] and that training samples with higher S content have smaller Klason lignin contents (supplementary Fig.
-4, @DelaCruz.2016), making this bin indicative of smaller Klason lignin contents.
+S4, @DelaCruz.2016), making this bin indicative of smaller Klason lignin contents.
 Likewise, a plausible explanation for the positive sign of the coefficient for the 1270 cm$^{-1}$ bin is that absorbance in this range is caused predominantly by guaiacyl units (G) [@Kubo.2005] and that training samples with higher G content have smaller Klason lignin contents (supplementary Fig.
-4, @DelaCruz.2016), making this bin indicative of larger Klason lignin contents.
+S4, @DelaCruz.2016), making this bin indicative of larger Klason lignin contents.
 In Sect.
@@
 \@ref(fig:res-p-pred-diff-all1-fig05) and supplementary Fig.
-5 <!---\@ref(fig:res-p-gaussian-beta-depth-profile-best)--->).
+S5 <!---\@ref(fig:res-p-gaussian-beta-depth-profile-best)--->).
 For holocellulose, the original model tends to predict larger contents for peat, especially for samples with larger holocellulose contents (Fig.
@@
 Second, both models have comparatively small bias within the range of the training data relevant to peat samples (supplementary Fig.
-6 <!---\@ref(fig:res-p-y-yhat-best)--->).
+S6 <!---\@ref(fig:res-p-y-yhat-best)--->).
 Thus, predictions are different for the peat samples even within the holocellulose and Klason lignin range covered by the training data, where the models actually make similar predictions for the training data.
@@
 In comparison to the training data, spectra for peat typically have larger absorption values at around 1250 cm$^{-1}$ and between (~1250 to 1500 cm$^{-1}$), but a smaller absorption in the region of the "OH-peak" (~3400 cm$^{-1}$) (supplementary Fig.
-7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).
+S7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).
 These spectral differences are directly related to the differences in predicted values of the original and improved model (supplementary Fig.
-8<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->): The smaller the OH-peak is, the larger are predicted values of the original model in comparison to the improved model using binned spectra, especially for larger `carb` peak heights.
+S8<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->): The smaller the OH-peak is, the larger are predicted values of the original model in comparison to the improved model using binned spectra, especially for larger `carb` peak heights.
 Similarly, the larger the trough at 1250 cm$^{-1}$ or the absorbance at 1590 cm$^{-1}$ is, the larger are predictions of the original model in comparison to the improved model using binned spectra.
@@
 Likewise, for the improved model using binned spectra a smaller OH-peak results in larger absorbances around ~1200 to 1600 cm$^{-1}$ where the most influential variables are located (supplementary Fig.
-7 <!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)---> and 2<!---\@ref(fig:res-sup-spectra-sim-p))--->).
+S7 <!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)---> and 2<!---\@ref(fig:res-sup-spectra-sim-p))--->).
 Why do peat samples have such different spectral properties?
@@
 The training data contain only undecomposed or industrially processed OM [@DelaCruz.2016; @Hodgkins.2018] which do not reflect such decomposition changes and therefore generally have a larger OH peak (supplementary Fig.
-7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).
+S7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).
 For decomposed OM, the same normalized `carb` peak height or absorbance at specific wavenumbers therefore represents different holocellulose contents than for the training data, as the comparatively broad OH-peak strongly influences the normalization of the spectra.
@@
 For Klason lignin, whereas the original model is unbiased across all samples (supplementary Fig.
-6<!---\@ref(fig:res-p-y-yhat-best)--->), it is not within two classes of training samples.
+S6<!---\@ref(fig:res-p-y-yhat-best)--->), it is not within two classes of training samples.
 Samples of these classes differ in the relative contribution of the `arom15` and `arom16` peaks to `arom15arom16` (the sum of their heights) (Fig.
@@
 What we observed is that the model makes biased predictions for both classes: For small `arom15arom16` values, it underestimates contents, but for larger values, it increasingly overestimates Klason lignin contents (supplementary Fig.
-9<!---\@ref(fig:res-p-pred-diff-kl-residuals)--->).
+S9<!---\@ref(fig:res-p-pred-diff-kl-residuals)--->).
 For the improved model, this bias is much smaller (Fig.
 \@ref(fig:res-p-pred-diff-train1-p2-fig06) and supplementary Fig.
-9<!---\@ref(fig:res-p-pred-diff-kl-residuals)--->).
+S9<!---\@ref(fig:res-p-pred-diff-kl-residuals)--->).
 We suggest that it is this bias which causes most of the difference in predictions of the original and improved models.
 Conditional on the relative contribution of the `arom15` peak to `arom15arom16`, peat samples have smaller and larger `arom15arom16` values than the training data (supplementary Fig.
-10<!---\@ref(fig:res-p-pred-diff-kl-histogram1)--->).
+S10<!---\@ref(fig:res-p-pred-diff-kl-histogram1)--->).
 If one extrapolates the biased predictions of the original model for the training data classes, this means larger under- and overestimation for the peat samples than for the training data using the original model (Fig.
@@
 This is what distinguishes samples of class 1 and 2 in the training data (supplementary Fig.
-11<!---\@ref(fig:res-p-pred-diff-train2)--->): Samples of class 1 are wood samples and paper product samples derived from wood.
+S11<!---\@ref(fig:res-p-pred-diff-train2)--->): Samples of class 1 are wood samples and paper product samples derived from wood.
 Wood typically has smaller nitrogen contents, but larger Klason lignin contents [@Cowling.1966; @Aerts.1999; @DelaCruz.2016] which explains the large contribution of the `arom15` peak, indicative for aromatic skeletal vibrations (but not for proteins) [@Stuart.2004], to `arom15arom16`.
@@
 Moreover, the models trained on mineral rich samples had a similar fit to the remaining samples as our improved models not trained with the mineral rich samples (supplementary Fig.
-12<!---\@ref(fig:res-p-y-yhat-minerals)--->).
+S12<!---\@ref(fig:res-p-y-yhat-minerals)--->).
 In comparison to this, the original model considerably overestimated the holocellulose content for these samples, as observed by @Hodgkins.2018.
 This was also the case for the improved models not trained with the mineral rich samples (supplementary Fig.
-12<!---\@ref(fig:res-p-y-yhat-minerals)--->).
+S12<!---\@ref(fig:res-p-y-yhat-minerals)--->).
 We see this as proof-of-concept that it is possible to predict holocellulose contents from MIRS even for samples with high mineral contents.
@@
 We have not used this approach here due to the higher computational costs, potential computational difficulties, to keep the present model validation straightforward, and since the training data do not fulfill this condition for all samples (supplementary Fig.
-4; indicating that the measurement procedure needs to be improved, too).
+S4; indicating that the measurement procedure needs to be improved, too).
 In summary, our analysis opens concrete and promising directions to further improve the models: We need training and validation data that include peat, particularly highly decomposed peat.

### 2022-10-26 analysis/paper/001-paper-main.Rmd
Adapt color schemes for color blind persons.

@@
   We like to thank @Hodgkins.2018 and @DelaCruz.2016 for provision of data and code to reproduce and modify their analyses, and @Allaire.2020 for provision of an Rmarkdown template.
-We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions which helped to improve a first version of this manuscript.
+We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions and Stephen Chapman and an anonymous reviewer for comments which helped to improve previous versions of this manuscript.
 This study was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no.
@@
 options(knitr.table.format = "latex")
+
+# color theme (source: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette)
+palette_cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 ```

### 2022-10-26 analysis/paper/003-paper-m-gaussian-beta.Rmd
Adapt color schemes for color blind persons.

--- 2022-03-03 analysis/paper/003-paper-m-gaussian-beta.Rmd
@@
              ncol = 2) +
-  scale_x_reverse()
+  scale_x_reverse() +
+  guides(color = guide_legend(title = "Bayesian model"), fill = guide_legend(title = "Bayesian model")) +
+  scale_color_manual(values = palette_cb[-1])
 
@@
   scale_x_continuous(labels = function(x) x * 100) +
-  scale_y_continuous(labels = function(x) x * 100)
+  scale_y_continuous(labels = function(x) x * 100) +
+  guides(color = guide_legend(title = "Bayesian model"), fill = guide_legend(title = "Bayesian model")) + 
+    scale_color_manual(values = palette_cb[-1]) +
+    scale_fill_manual(values = palette_cb[-1])
 

### 2022-10-26 analysis/paper/004-paper-m-reduce-underfitting.Rmd
Adapt color schemes for color blind persons.

@@
              ncol = 2) +
-  scale_x_reverse()
+  scale_x_reverse() +
+  scale_color_manual(values = palette_cb[-1]) +
+  scale_fill_manual(values = palette_cb[-1])
 

### 2022-10-26 analysis/paper/005-paper-m-minerals.Rmd
Adapt color schemes for color blind persons.

@@
   guides(colour = guide_legend(title = "Mineral rich samples")) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") +
+  scale_color_manual(values = palette_cb[-1])
 ```

### 2022-10-26 analysis/paper/007-paper-m-prediction-differences.Rmd
Adapt color schemes for color blind persons.

@@
   coord_fixed() +
-  scale_colour_manual(values = scales::hue_pal()(length(levels(d_pred_diff$sample_type))),
+  scale_color_manual(
+    values = palette_cb,
     labels = levels(d_pred_diff$sample_type), 
-                    drop = FALSE)
+    drop = FALSE
+  ) +
+  scale_fill_manual(
+    values = palette_cb,
+    labels = levels(d_pred_diff$sample_type), 
+    drop = FALSE
+  )
 

### 2022-10-26 analysis/paper/008-paper-supplementary.Rmd
Adapt color schemes for color blind persons.

@@
    - \renewcommand\thesection{S\arabic{section}}
+   - \renewcommand{\thefigure}{S\arabic{figure}}
 ---
@@
   guides(colour = guide_legend(title = "Sample type", ncol = 3, title.position="top")) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") +
+  scale_color_manual(values = palette_cb)
 ```

### 2022-10-26 analysis/paper/001-reply-main.Rmd
Minor textual changes to the second manuscript.

@@
 ---
-title: 'Supplementary Information to: "Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging"'
+title: 'Supplementary Information to: "Comment on Hodgkins et al. (2018): Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging"'
 author: "Henning Teickner, Klaus-Holger Knorr"
@@
    - \doublespacing
+fontsize: 12pt
 ---

### 2022-10-26 analysis/paper/002-reply-main.Rmd
Minor textual changes to the second manuscript.

@@
 An interpretation of the values given in @Hodgkins.2018 in terms of quantitative concepts for organic matter chemistry (i.e.
-holocellulose and Klason lignin) therefore is not possible.
+holocellulose and Klason lignin) is not possible.
 Instead, we assume that if the study is reinterpreted in terms of qualitative concepts for carbohydrates and aromatics as used in the qualitative interpretation of mid infrared spectra for peat, there still is evidence for general latitudinal patterns in peat recalcitrance.
@@
 @Biester.2014).
-For mid infrared spectra, such knowledge still has to be acquired in future studies, especially if large environmental gradients are considered as in @Hodgkins.2018.
+For mid infrared spectra, such knowledge still has to be acquired in future studies, especially if as large environmental gradients are considered as in @Hodgkins.2018.
 
-## Klason lignin content and residual enrichment in (sub)tropical peatlands probably are larger {.unnumbered}
+## Klason lignin contents and residual enrichment in (sub)tropical peatlands probably are larger {.unnumbered}
 

### 2022-11-23 README.Rmd
Rework manuscript 2, update the README.

@@
 
-> Henning Teickner, and Klaus-Holger Knorr (2022; preprint): _Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra_.
-Preprint.
-Soil and methods, June 1, 2022.
-DOI: [10.5194/soil-2022-27](https://doi.org/10.5194/soil-2022-27).
+> Henning Teickner, and Klaus-Holger Knorr (2022): _Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid-Infrared Spectra_.
+SOIL 8 (2): 699–715.
+DOI: [10.5194/soil-8-699-2022](https://doi.org/10.5194/soil-8-699-2022).
 

### 2022-11-23 analysis/paper/001-reply-main.Rmd
Rework manuscript 2, update the README.

--- 2022-07-24 analysis/paper/001-reply-main.Rmd
@@
 ---
-title: 'Supplementary Information to: "Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging"'
+title: 'Supplementary Information to: "The need to reinterpret Hodgkins et al. (2018)"'
 author: "Henning Teickner, Klaus-Holger Knorr"
@@
 The models described in @Hodgkins.2018 to predict holocellulose and Klason lignin contents from mid infrared spectra, but recomputed in a Bayesian framework.
-This is the "original Bayesian model" in @Teickner.2022a.
+This is the "original Bayesian model" in @Teickner.2022f.
 These models were used to reproduce the computations in @Hodgkins.2018.
@@
 A modified model to predict each, holocellulose and Klason lignin contents from mid infrared spectra.
-These are the models called "best binned spectra" models in @Teickner.2022a.
+These are the models called "best binned spectra" models in @Teickner.2022f.
 A brief description of these models is given in section [Computation of modified models to predict organic matter holocellulose and Klason lignin contents from mid infrared spectra].
@@
 
-Computation of the modified models using only the mid infrared spectra (MIRS) is described in detail in @Teickner.2022a.
+Computation of the modified models using only the mid infrared spectra (MIRS) is described in detail in @Teickner.2022f.
 In brief, we used the same training data as @Hodgkins.2018 and a Bayesian linear regression model which we computed with Markov Chain Monte Carlo (MCMC) sampling using brms [@Burkner.2017; @Burkner.2018] and rstan [@StanDevelopmentTeam.2020] which are both interfaces to Stan [@StanDevelopmentTeam.2021].
@@
 1.
-Since representative training data are missing, both the original and modified models cannot be fully validated and prediction uncertainties are probably larger than computed with these models [@Teickner.2022a].
+Since representative training data are missing, both the original and modified models cannot be fully validated and prediction uncertainties are probably larger than computed with these models [@Teickner.2022f].
 Since representative test data are missing, these additional uncertainties cannot be estimated.
@@
 The modified models using only the MIRS were also used to compute median values and confidence intervals mentioned in the main text in section "Klason lignin content and residual enrichment in (sub)tropical peatlands probably are larger".
-In these computations, we also considered the predictive uncertainty of the original models (the Bayesian formulation described in @Teickner.2022a).
+In these computations, we also considered the predictive uncertainty of the original models (the Bayesian formulation described in @Teickner.2022f).
 
@@
 - $\alpha$ and $\beta$ are intercept and regression coefficients for the MIRS prediction model.
-The $\beta$s are modeled via a regularized horseshoe prior parameterized following @Piironen.2017b as described in @Teickner.2022a.
+The $\beta$s are modeled via a regularized horseshoe prior parameterized following @Piironen.2017b as described in @Teickner.2022f.
 
@@
 To create figure 1<!--\@ref(fig:reply-res-p-sites-latitude)--> in the main text, we used the models for the latitudinal gradient and samples from the posterior distribution to compute the predicted average and 95% confidence interval for a range of latitude values and we added predictions for the average contents for each of the sampled cores.
-Next, we recomputed the models from @Hodgkins.2018 by (1) predicting average peat sample contents with the original models (the Bayesian version described in @Teickner.2022a), (2) averaging the near surface peat sample average contents for each core, and (3) fitting an ordinary least squares regression line to these average core values.
+Next, we recomputed the models from @Hodgkins.2018 by (1) predicting average peat sample contents with the original models (the Bayesian version described in @Teickner.2022f), (2) averaging the near surface peat sample average contents for each core, and (3) fitting an ordinary least squares regression line to these average core values.
 For this linear model, we plotted the confidence interval across latitude and we added the computed average core near-surface contents.
 Thus, figure 1<!---\@ref(fig:reply-res-p-sites-latitude)---> in the main text compares the graphed values and model prediction from figure 3 in @Hodgkins.2018 with predictions from our modified analysis considering prediction uncertainty from the MIRS-based model part, and the averaging per core.
-Additional sources for differences in predictions and prediction uncertainties in comparison to the analysis in @Hodgkins.2018 are the use of a beta distribution which constraints uncertainty intervals for extreme (near 0 and 100 mass-%) values, and the use of all spectral variables which resulted in reduced prediction uncertainties [@Teickner.2022a].
+Additional sources for differences in predictions and prediction uncertainties in comparison to the analysis in @Hodgkins.2018 are the use of a beta distribution which constraints uncertainty intervals for extreme (near 0 and 100 mass-%) values, and the use of all spectral variables which resulted in reduced prediction uncertainties [@Teickner.2022f].
 

### 2022-11-23 analysis/paper/002-reply-main.Rmd
Rework manuscript 2, update the README.

--- 2022-07-24 analysis/paper/002-reply-main.Rmd
@@
 ---
-title: 'Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging'
+title: 'The need to reinterpret Hodgkins et al. (2018)'
 author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
@@
 
-The resilience of the large global peatland carbon stocks against climate change may depend on how peat chemistry controls decomposition rates [@Wang.2015; @Hodgkins.2018].
-In their study, @Hodgkins.2018 argue that there is a latitudinal gradient in peat recalcitrance which may explain the stability of tropical peat in spite of warmer temperatures and which points toward a susceptibility of labile surface peat at high latitudes to degradation, but not deeper peat which has a chemistry more similar to tropical peat.
-This gradient can be described by differences in holocellulose contents --- representing relatively easily degradable carbohydrates --- and Klason lignin contents --- representing aromatics --- which @Hodgkins.2018 predicted from mid infrared spectra [@Hodgkins.2018].
-We acknowledge this conceptual model as an important step in understanding the stability of peat deposits and their susceptibility to degradation.
-However, we challenge the accuracy of the models used to predict peat holocellulose and Klason lignin contents.
-This inaccuracy has two major implications:
+@Hodgkins.2018 used spectral prediction models to estimate two major peat organic matter fractions --- holocellulose and Klason lignin --- in peat cores collected along a latitudinal gradient.
+Based on this, @Hodgkins.2018 argue that (sub)tropical peat deposits accumulate despite high temperatures due to higher Klason lignin over holocellulose contents, and that climate warming is a risk to northern surface peat deposits which would decompose faster due to higher holocellulose contents, but not to deeper peat which is more similar to tropical peat.
+In a recent study, we have shown that the spectral prediction models behind this conclusion are not valid for peat samples [@Teickner.2022f].
+In addition to that, we show here that @Hodgkins.2018 did not consider important uncertainty sources in their analysis.
+This raises the question whether the results of @Hodgkins.2018, and of a follow-up study with more data [@Verbeke.2022], remain valid or how they have to be reinterpreted.
 
-1.
-Peat Klason lignin contents and relative accumulation during decomposition probably are underestimated by @Hodgkins.2018.
-This may have important consequences for decomposition of high-latitude peat deposits due to climate change.
+We show that the results of @Hodgkins.2018 are to some extent still valid for holocellulose, but only when reinterpreted in terms of differences in peak heights, which is a major obstacle to a quantitative understanding of the control of peat chemistry on decomposition.
+For Klason lignin, remaining uncertainties are high even after reinterpretation in terms of peak heights.
+Even more, there is evidence that near-surface peat Klason lignin contents and residual enrichment of Klason lignin during decomposition have been underestimated.
+It thus remains uncertain whether and to what extent the reported latitudinal gradient in peat chemistry will control future peatland carbon losses.
 
-2.
-The inaccuracies imply that the predicted values are only very rough estimates for holocellulose and Klason lignin contents and at least partly biased.
-An interpretation of the values given in @Hodgkins.2018 in terms of quantitative concepts for organic matter chemistry (i.e.
-holocellulose and Klason lignin) therefore is not possible.
-Instead, we assume that if the study is reinterpreted in terms of qualitative concepts for carbohydrates and aromatics as used in the qualitative interpretation of mid infrared spectra for peat, there still is evidence for general latitudinal patterns in peat recalcitrance.
-However, interpretation and generalization of these findings are less straightforward.
+## Why are the results highly uncertain? {.unnumbered}
 
-Moreover, we suggest improvements for uncertainty propagation for the models.
-Remaining open questions to this approach of analyzing global peat chemistry patterns should be addressed in future studies.
+The original models to estimate peat holocellulose and Klason lignin contents are not valid and, highly uncertain, and thus likely biased --- especially for decomposed peat --- and therefore should not be used to estimate holocellulose and Klason lignin contents [@Teickner.2022f].
+Besides, there are that two important sources of uncertainty have been neglected in @Hodgkins.2018: first, the prediction uncertainty of the spectral prediction models, and second, the uncertainty introduced when computing near-surface average holocellulose and Klason lignin contents from estimates for individual peat layers.
+We illustrate the impact of all these factors by recomputing the latitudinal gradient and depth profiles with the modified, probably more accurate, but still biased, models from @Teickner.2022f, while considering these uncertainty sources (supporting information S1).
+This analysis shows that both the estimated slopes for the latitudinal gradient (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->) and differences in the depth profiles are highly uncertain due to these uncertainty sources (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
 
-## Patterns in peat chemistry should not be interpreted in terms of holocellulose and Klason lignin contents {.unnumbered}
+## Klason lignin contents and residual enrichment with decomposition probably are larger {.unnumbered}
 
-In a recent evaluation of the models to predict peat holocellulose and Klason lignin contents presented in @Hodgkins.2018, we found that the training data used for the models (leaves, paper products) are not representative for peat [@Teickner.2022a].
-The peat samples are more decomposed than the training data samples.
-Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [@Teickner.2022a].
-Since these features are not present in the training data, predictions of the original models are extrapolations.
-Our re-evaluation showed that this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >30 mass-% [@Teickner.2022a].
-Because of these uncertainties and biases, the analyses of @Hodgkins.2018 cannot be interpreted in terms of absolute holocellulose and Klason lignin contents.
-Instead, we suggest to interpret differences in peat chemistry presented by @Hodgkins.2018 not in terms of the quantitative concepts (i.e.
-holocellulose and Klason lignin), but rather in terms of more qualitative concepts typically used in studies interpreting mid infrared spectra of peat: as "carbohydrates" and "aromatics" (e.g.
-@Artz.2008; @Broder.2012; @Tfaily.2014; @Biester.2014).
-We assume that under such an interpretation, the conclusions of @Hodgkins.2018 probably are valid.
-The proposed reinterpretation is possible if values given in @Hodgkins.2018 are only interpreted in terms of the normalized peak heights used in the original prediction models, and not as absolute holocellulose and Klason lignin contents.
-Each model is a linear model with one independent variable (the normalized height of one peak); therefore a direct reinterpretation of the predicted absolute contents in terms of normalized peak heights is possible if the figure axes are relabeled as "Relative content of carbohydrates/aromatics" instead of "Estimated % carbohydrates/aromatics" (implying an absolute content of holocellulose and Klason lignin, respectively) [@Hodgkins.2018].
-An important open question to this approach of interpreting large-scale gradients in peat chemistry based on mid infrared spectra should be addressed in future studies: Are differences in specific mid infrared spectra peaks reliable indicators for differences in a specific concept for peat chemistry?
-For example, are differences in the peak around 1040 cm$^{-1}$ indicative for differences in carbohydrate contents (also across larger scales)?
-A similar problem arose in the discussion on the use of the C/N ratio as a good indicator for peat mass loss.
-This may only hold true if there have been no or negligible shifts in the vegetation composition which causes differences in the C/N ratio already in initial plant litter (e.g.
-@Biester.2014).
-For mid infrared spectra, such knowledge still has to be acquired in future studies, especially if large environmental gradients are considered as in @Hodgkins.2018.
+In a previous publication, we have shown that the spectral prediction model for Klason lignin underestimates Klason lignin contents especially when these are $>30$ mass-% [@Teickner.2022f].
+Our modified model is much less biased [@Teickner.2022f] and therefore enables a rough analysis of the order of magnitude of the bias.
+This allowed us to analyze what consequences the bias can have for the estimated near-surface ($\le50$ cm) peat Klason lignin contents and residual enrichment of Klason lignin during decomposition (supporting information S1).
+With our modified model, near-surface peat Klason lignin contents are on average ~10 to 15 mass-% larger across the latitudinal gradient (figure \@ref(fig:reply-res-p-sites-latitude)).
+Specifically, average (sub)tropical ($<45 ^\circ$N) near-surface peat Klason lignin contents are `r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% (median, lower and upper 95% prediction interval limit) larger than with the original model.
+There is also tentative indication for on average larger differences between deep (sub)tropical ($<45 ^\circ$N) and high-latitude ($\ge45 ^\circ$N) peat (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
+Residual enrichment of Klason lignin during decomposition probably has been underestimated.
+A rough estimate for the residual enrichment of Klason lignin during decomposition is the difference in Klason lignin content between near-surface peat and deeper peat.
+With the modified model, this difference is on average for (sub)tropical peatlands `r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% larger (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
+For high latitude peatlands the difference is smaller (`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-%).
+Consequently, in general --- and especially in (sub)tropical peatlands --- the average residual enrichment of Klason lignin due to decomposition probably has been underestimated by the original model.
+A consequence of this in light of climate change is that high latitude peat deposits may be even more prone to shifts in decomposition rates than suggested by the original results.
 
-## Klason lignin content and residual enrichment in (sub)tropical peatlands probably are larger {.unnumbered}
+## How to reinterpret @Hodgkins.2018 {.unnumbered}
 
-We suggest that @Hodgkins.2018 have underestimated peat Klason lignin contents, especially in (sub)tropical peat.
-As described above, the original model for Klason lignin underestimates peat Klason lignin contents.
-A modified model has less bias [@Teickner.2022a].
-In spite of several remaining limitations (see below and the supplementary information), we argue that this modified model can serve as rough indicator for the underestimation of peat Klason lignin contents by the original model.
-We therefore present values to provide tentative estimates for the underestimation of Klason lignin contents, but we do not claim that these estimates are optimal.
-With this modified model, near-surface ($\le50$ cm) peat Klason lignin contents are on average ~10 to 15 mass-% larger across the latitudinal gradient (figure \@ref(fig:reply-res-p-sites-latitude)).
-Specifically, average (sub)tropical near-surface peat ($<45 ^\circ$N, $\le50$ cm) Klason lignin contents are `r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% (median, lower and upper 95% prediction interval limit) larger than with the original model.
-There is also tentative indication for on average larger differences between deep (sub)tropical ($<45 ^\circ$N) and high-latitude ($\ge45 ^\circ$N) peat (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
-We also suggest that the residual enrichment of Klason lignin during decomposition probably has been underestimated.
-A rough estimate for the residual enrichment of Klason lignin during decomposition is the difference in Klason lignin content between near-surface peat ($\le50$ cm) and deeper peat.
-With the modified model, this difference is on average for (sub)tropical peatlands `r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% larger (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
-For high latitude peatlands the difference is smaller on average (`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-%).
-Consequently, in general --- and especially in (sub)tropical peatlands --- the average residual enrichment of Klason lignin due to decomposition probably has been underestimated by the original model.
-A consequence of this in light of climate change is that high latitude peat deposits may be more prone to shifts in decomposition rates than suggested by the original results.
+The mid-infrared spectra (MIRS), which have been used to compute the spectral prediction models, have so far mostly been interpreted in terms qualitative concepts for organic matter fractions (e.g.
+@Cocozza.2003): the `carb` peak [@Hodgkins.2018] --- the only predictor for holocellulose content in @Hodgkins.2018 --- is an indicator for "carbohydrates" in mineral-poor peat [@Hodgkins.2018]; the `arom15arom16` peak [@Hodgkins.2018] --- the only predictor for Klason lignin content in @Hodgkins.2018 --- is an indicator for "aromatics" in peat with low protein and carbonate contents [@Stuart.2004; @Tatzber.2007; @Teickner.2022f] (note that we write "carbohydrates" and "aromatics" because, when interpreted only in terms of peak heights, it is remains uncertain what these organic matter (OM) fractions represent chemically or how one could extract them).
 
-## Larger uncertainties due to a lack of uncertainty propagation and unrepresentative data {.unnumbered}
+Such a qualitative interpretation of MIRS enables us to reinterpret @Hodgkins.2018, circumventing some of the uncertainties we have uncovered: All holocellulose and Klason lignin values given in @Hodgkins.2018 are directly proportional to the heights of the `carb` and `arom15arom16` peak, respectively.
+This removes uncertainties due to invalid spectral prediction models from the analyses.
+Then, we can still say that @Hodgkins.2018 show systematic latitudinal and depth-related differences in peat "carbohydrate" and "aromatics" contents.
 
-Besides a reinterpretation in terms of qualitative concepts for organic matter recalcitrance, we also suggest that a more comprehensive quantification and propagation of uncertainties is important to further scrutinize the results of @Hodgkins.2018 because these uncertainties are not negligibly small.
-In the original analysis, the following uncertainties are omitted: (1) As described above, the training dataset is not representative for peat and therefore predictions are biased.
-(2) The prediction uncertainty of the models used to predict holocellulose and Klason lignin contents.
-(3) In computing the regression model to analyze the latitudinal gradient based on average core near-surface contents, the uncertainty for these average values was not propagated.
-The first two sources of uncertainty could be omitted if only normalized peak heights (as qualitative indicators) are used instead of predicted holocellulose and Klason lignin contents, as described above.
-But the third source of uncertainty should also be considered when reinterpreting the results in terms of area normalized peak heights.
-To support our concerns, we reanalyzed the latitudinal gradient for holocellulose and Klason lignin, respectively, using the modified models described in @Teickner.2022a and fully considering the last two sources of uncertainty using a Bayesian approach (described in more detail in the supplementary information).
-With the modified models, the estimated slopes for the latitudinal gradient are more uncertain (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->).
-A similar reanalysis of the depth profiles between the two latitude categories shows that differences in depth patterns are also more uncertain (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
-Our analysis therefore indicates larger uncertainties both in the depth profiles and the latitudinal gradient.
-Considering these uncertainties is important when peat properties predicted with a model are analyzed.
-Again, the values presented here are only tentative because the modified models probably are also biased.
+The trade-offs are first, that we have to make the assumption that there are no latitudinal or depth-related differences in mineral, carbonate, and protein contents, and second, that we have to make the assumption that differences in heights of the two peaks are not dominated by differences in other peaks in the spectra.
+The reason is that differences in peak heights can be attributed to "carbohydrates" or "aromatics" only when contents of minerals or proteins and carbonates, respectively, are the same, and when changes in the area normalized height of these peaks are not caused by changes in other peaks, which can be an effect of spectral normalization [@Stuart.2004; @Tatzber.2007; @Hodgkins.2018; @Teickner.2022f].
 
-## Future directions {.unnumbered}
+We think that only some of these assumptions can are supported at the moment.
+The `carb` peak is comparatively large and has low interferences with peaks caused by other molecular structures.
+It should therefore be a good indicator for "carbohydrates" because mineral contents in (near-surface) peat are typically low (as validated by @Hodgkins.2018).
+The `arom15arom16` peak has large contributions by proteins [@Stuart.2004; @Teickner.2022f], based on peat N contents, it is not unlikely that peat protein contents differ with depth and latitude [@Sowden.1978; @Loisel.2014], and it may be more affected by other peaks since it is smaller than the `carb` peak.
+For this reason, we suggest that more evaluation is necessary before accepting, based on @Hodgkins.2018, that there are systematic differences in peat "aromatic" contents.
 
-Notwithstanding our critical remarks, the idea of @Hodgkins.2018 to relate mid infrared spectra to precisely quantifiable and defined organic matter fractions is important.
-Such fractions are relatively well understood and therefore allow us to better deduce what effects climate shifts can have on peat decomposition and recalcitrance: It is much more informative if we are able to precisely estimate a change in Klason lignin contents of e.g.
-5 mass-% if average annual temperatures increase by 1$^{\circ}$C than if we are only able to make qualitative statements that the relative amount of aromatics (without describing in detail what this fraction means) increases, as is currently the case.
-When interpreting mid infrared spectra, distinguishing between precisely quantifiable concepts for organic matter chemistry, such as holocellulose and Klason lignin contents, and qualitative concepts which currently are used in interpreting peat MIRS, such as "carbohydrates and aromatics" (without specifying how this fraction could actually be measured), is crucial: Models relying on quantitative concepts are more testable and allow to actually quantify fractional changes in organic matter chemistry, whereas models built on qualitative concepts can only be refuted if there are blatant discrepancies relative to expectations (i.e.
-a decrease in "carbohydrates" with latitude).
-Ultimately, we need more precise models for variations in peat chemistry which can be developed based on qualitative concepts.
-We suggest the following steps to gain deeper insights into latitudinal gradient in peat chemistry: As outlined above, we suggest to interpret the results of @Hodgkins.2018 in terms of qualitative concepts for peat chemistry, i.e.
-"carbohydrates" and "aromatics".
-These suggestions also apply to a follow-up study using the same models [@Verbeke.2022].
-Further studies are needed to check if specific peaks are good indicators for these qualitative concepts for peat chemistry, how the relation may be confounded e.g.
-by vegetation composition or other organic matter fractions, as could be the case especially for peaks related to aromatics [@Teickner.2022a].
-This could be done by computing models to reliably predict holocellulose and Klason lignin contents of peat from mid infrared spectra and analyzing which spectral variables are relevant for this purpose and if there are confounding variables.
-Such models will also help us to get more precise insights into the degree to which high latitude peat accumulation may be decreased by decomposition due to rising temperatures, because they allow replacing more costly measurements.
-Finally, future studies should carefully consider uncertainties and fully propagate these into model estimates if it is unknown whether such uncertainties are negligible.
-Predicting absolute peat holocellulose and Klason lignin contents from infrared spectra remains challenging.
+The reinterpretation in terms of "carbohydrate" and "aromatics" instead of holocellulose and Klason lignin has the clear disadvantage that we sacrifice the quantitative precision @Hodgkins.2018 aimed to attribute to the results: it is much more informative to precisely estimate a change in Klason lignin content of e.g.
+5 mass-% when the average annual temperature increase by 1$^{\circ}$C than to only make qualitative statements that the relative amount of "aromatics" (without describing in detail what this OM fraction means) increases.
+
+We think @Hodgkins.2018 have advanced a very important idea to make the interpretation of MIRS more quantitative.
+Only with such a quantitative approach, which links different types of measurements and the knowledge we have about them, can we develop a quantitatively precise understanding.
+However, we have shown that there is some more way to go.
+We hope that our critical analysis will help to improve this approach.
 
@@
 
-We like to thank Jeffrey Chanton and Suzanne Hodgkins for a constructive dicussion which greatly helped to improve a first version of this manuscript.
+We like to thank Jeffrey Chanton and Suzanne Hodgkins for a constructive discussion which greatly helped to improve a first version of this manuscript.
 This study was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no.

### 2023-11-20 analysis/paper/001-reply-main.Rmd
Update figure layout and title.

@@
 ---
-title: 'Supplementary Information to: "The need to reinterpret Hodgkins et al. (2018)"'
+title: 'Supplementary Information to: "Comment on Hodgkins et al. (2018)"'
 author: "Henning Teickner, Klaus-Holger Knorr"
@@
   scale_x_continuous(labels = function(x) round(x * x_latitude_scale + x_latitude_center, 0)) +
+  scale_color_manual(values = colors_prediction_domain[c(3, 2)]) +
+  scale_fill_manual(values = colors_prediction_domain[c(3, 2)]) +
   labs(x = "Latitude [°]", y = "Content [mass-%]", title = "Holocellulose") +
@@
   scale_x_continuous(labels = function(x) round(x * x_latitude_scale + x_latitude_center, 0)) +
+  scale_color_manual(values = colors_prediction_domain[c(3, 2)]) +
+  scale_fill_manual(values = colors_prediction_domain[c(3, 2)]) +
   labs(x = "Latitude [°]", y = "Content [mass-%]", title = "Klason lignin") +
@@
   dplyr::filter(variable_y == "Holocellulose") %>%
-  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
+  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified model", "Original model")) %>%
   dplyr::arrange(sample_depth_lower) %>%
-  ggplot(aes(y = fit, 
+  ggplot(
+    aes(
+      y = fit, 
       ymin = lwr,
@@
       colour = latitude_class, 
-             fill = latitude_class)) +
-  geom_hline(yintercept = c(0.11, 0.16, 0.21, 0.37), colour = "grey", linetype = 2) +
+      fill = latitude_class
+    )
+  ) +
+  geom_hline(
+    yintercept = c(0.11, 0.16, 0.21, 0.37), 
+    colour = "grey", 
+    linetype = 2
+  ) +
   geom_path() +
@@
   coord_flip() +
-  labs(title = "Holocellulose", 
+  labs(
+    title = "Holocellulose", 
     y = expression(Holocellulose~content~"[mass-%]"),
-       x = "Lower layer boundary depth [cm]") +
-  facet_wrap(~ model, 
-             scales = "free_y", 
-             ncol = 1) +
+    x = "Lower layer boundary depth [cm]"
+  ) +
+  facet_wrap(~ model, scales = "free_y", ncol = 1) +
   scale_x_reverse() +
   scale_y_continuous(labels = function(x) x * 100) +
-  guides(colour = guide_legend(title = expression(Latitude>45~degree*N)),
-         fill = guide_legend(title = expression(Latitude>45~degree*N)))
+  scale_color_manual(values = c("indianred", "dodgerblue")) +
+  scale_fill_manual(values = c("indianred", "dodgerblue")) +
+  guides(
+    colour = guide_legend(title = expression(Latitude>45~degree*N)),
+    fill = guide_legend(title = expression(Latitude>45~degree*N))
+  ) +
+  theme(
+    strip.background = element_blank(),
+    strip.text.x = element_text(size = 12)
+  )
 
@@
   dplyr::filter(variable_y == "Klason lignin") %>%
-  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified", model)) %>%
+  dplyr::mutate(model = ifelse(model == "Best binned spectra", "Modified model", "Original model")) %>%
   dplyr::arrange(sample_depth_lower) %>%
-  ggplot(aes(y = fit, 
+  ggplot(
+    aes(
+      y = fit, 
       ymin = lwr,
@@
       colour = latitude_class, 
-             fill = latitude_class)) +
-  geom_hline(yintercept = c(0.22, 0.32, 0.35, 0.48), colour = "grey", linetype = 2) +
+      fill = latitude_class
+    )
+  ) +
+  geom_hline(
+    yintercept = c(0.22, 0.32, 0.35, 0.48), 
+    colour = "grey", 
+    linetype = 2
+  ) +
   geom_path() +
@@
   coord_flip() +
-  labs(title = "Klason lignin", 
+  labs(
+    title = "Klason lignin", 
     y = expression(Klason~lignin~content~"[mass-%]"),
-       x = "Lower layer boundary depth [cm]") +
-  facet_wrap(~ model, 
-             scales = "free_y", 
-             ncol = 1) +
+    x = "Lower layer boundary depth [cm]"
+  ) +
+  facet_wrap(~ model, scales = "free_y", ncol = 1) +
   scale_x_reverse() + 
   scale_y_continuous(labels = function(x) x * 100) +
-  guides(colour = guide_legend(title = expression(Latitude>45~degree*N)),
-         fill = guide_legend(title = expression(Latitude>45~degree*N)))
+  scale_color_manual(values = c("indianred", "dodgerblue")) +
+  scale_fill_manual(values = c("indianred", "dodgerblue")) +
+  guides(
+    colour = guide_legend(title = expression(Latitude>45~degree*N)),
+    fill = guide_legend(title = expression(Latitude>45~degree*N))
+  ) +
+  theme(
+    strip.background = element_blank(),
+    strip.text.x = element_text(size = 12)
+  )
 

### 2023-11-20 analysis/paper/002-reply-main.Rmd
Update the comment on Hodgkins et al. (2018).

@@
 ---
-title: 'The need to reinterpret Hodgkins et al. (2018)'
+title: 'Comment on Hodgkins et al. (2018)'
 author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
@@
 
-@Hodgkins.2018 used spectral prediction models to estimate two major peat organic matter fractions --- holocellulose and Klason lignin --- in peat cores collected along a latitudinal gradient.
-Based on this, @Hodgkins.2018 argue that (sub)tropical peat deposits accumulate despite high temperatures due to higher Klason lignin over holocellulose contents, and that climate warming is a risk to northern surface peat deposits which would decompose faster due to higher holocellulose contents, but not to deeper peat which is more similar to tropical peat.
-In a recent study, we have shown that the spectral prediction models behind this conclusion are not valid for peat samples [@Teickner.2022f].
-In addition to that, we show here that @Hodgkins.2018 did not consider important uncertainty sources in their analysis.
-This raises the question whether the results of @Hodgkins.2018, and of a follow-up study with more data [@Verbeke.2022], remain valid or how they have to be reinterpreted.
+@Hodgkins.2018 used mid infrared spectra (MIRS) to make inferences about the stability of peat against decomposition along a latitudinal gradient from the tropics to northern latitudes.
+Recently, we have shown that these spectral prediction models are biased and their predictions more uncertain than considered in @Hodgkins.2018 [@Teickner.2022f].
+Here, we show what consequences this bias and uncertainty and additional neglected uncertainty sources have for the main conclusions in @Hodgkins.2018.
 
-We show that the results of @Hodgkins.2018 are to some extent still valid for holocellulose, but only when reinterpreted in terms of differences in peak heights, which is a major obstacle to a quantitative understanding of the control of peat chemistry on decomposition.
-For Klason lignin, remaining uncertainties are high even after reinterpretation in terms of peak heights.
-Even more, there is evidence that near-surface peat Klason lignin contents and residual enrichment of Klason lignin during decomposition have been underestimated.
-It thus remains uncertain whether and to what extent the reported latitudinal gradient in peat chemistry will control future peatland carbon losses.
+In particular, we find that:
 
-## Why are the results highly uncertain? {.unnumbered}
+1.
+Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition and aromatics may accumulate at a higher rate with depth in tropical peatlands than estimated by @Hodgkis.2018.
 
-The original models to estimate peat holocellulose and Klason lignin contents are not valid and, highly uncertain, and thus likely biased --- especially for decomposed peat --- and therefore should not be used to estimate holocellulose and Klason lignin contents [@Teickner.2022f].
-Besides, there are that two important sources of uncertainty have been neglected in @Hodgkins.2018: first, the prediction uncertainty of the spectral prediction models, and second, the uncertainty introduced when computing near-surface average holocellulose and Klason lignin contents from estimates for individual peat layers.
-We illustrate the impact of all these factors by recomputing the latitudinal gradient and depth profiles with the modified, probably more accurate, but still biased, models from @Teickner.2022f, while considering these uncertainty sources (supporting information S1).
-This analysis shows that both the estimated slopes for the latitudinal gradient (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->) and differences in the depth profiles are highly uncertain due to these uncertainty sources (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
+2. Larger uncertainties indicate that also larger differences in aromatic contents between (sub)tropical and high latitude peat than estimated by @Hodgkins.2018 are possible, also between deeper peat.
 
-## Klason lignin contents and residual enrichment with decomposition probably are larger {.unnumbered}
+3.
+More research should focus on how estimates of carbohydrate and aromatic contents from MIRS may be confounded by other organic matter fractions, in particular proteins.
+As a first step, this requires accurate concepts to name organic matter fractions and variables used in the interpretation of MIRS.
 
-In a previous publication, we have shown that the spectral prediction model for Klason lignin underestimates Klason lignin contents especially when these are $>30$ mass-% [@Teickner.2022f].
-Our modified model is much less biased [@Teickner.2022f] and therefore enables a rough analysis of the order of magnitude of the bias.
-This allowed us to analyze what consequences the bias can have for the estimated near-surface ($\le50$ cm) peat Klason lignin contents and residual enrichment of Klason lignin during decomposition (supporting information S1).
-With our modified model, near-surface peat Klason lignin contents are on average ~10 to 15 mass-% larger across the latitudinal gradient (figure \@ref(fig:reply-res-p-sites-latitude)).
-Specifically, average (sub)tropical ($<45 ^\circ$N) near-surface peat Klason lignin contents are `r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% (median, lower and upper 95% prediction interval limit) larger than with the original model.
-There is also tentative indication for on average larger differences between deep (sub)tropical ($<45 ^\circ$N) and high-latitude ($\ge45 ^\circ$N) peat (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
-Residual enrichment of Klason lignin during decomposition probably has been underestimated.
-A rough estimate for the residual enrichment of Klason lignin during decomposition is the difference in Klason lignin content between near-surface peat and deeper peat.
-With the modified model, this difference is on average for (sub)tropical peatlands `r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% larger (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
-For high latitude peatlands the difference is smaller (`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-%).
-Consequently, in general --- and especially in (sub)tropical peatlands --- the average residual enrichment of Klason lignin due to decomposition probably has been underestimated by the original model.
-A consequence of this in light of climate change is that high latitude peat deposits may be even more prone to shifts in decomposition rates than suggested by the original results.
+These uncertainties have the potential to change predictions of models on global peat stability as well as what stabilizes peat against decomposition if temperatures rise and thus add in particular to the debate to what extent global warming may increase decomposition of northern deep peat deposits (e.g.
+@Dorrepaal.2009, @Wilson.2016, @Baysinger.2022).
 
-## How to reinterpret @Hodgkins.2018 {.unnumbered}
+That said, theoretical considerations of decomposition processes alone support the suggested gradient in peat chemistry and @Hodgkins.2018 made an important contribution in providing the first open access models and estimates for peat carbohydrate and aromatics content on a large scale.
+We hope that our comment highlights that accuracy matters and that it helps to focus research on the problems which hamper improving this accuracy.
 
-The mid-infrared spectra (MIRS), which have been used to compute the spectral prediction models, have so far mostly been interpreted in terms qualitative concepts for organic matter fractions (e.g.
-@Cocozza.2003): the `carb` peak [@Hodgkins.2018] --- the only predictor for holocellulose content in @Hodgkins.2018 --- is an indicator for "carbohydrates" in mineral-poor peat [@Hodgkins.2018]; the `arom15arom16` peak [@Hodgkins.2018] --- the only predictor for Klason lignin content in @Hodgkins.2018 --- is an indicator for "aromatics" in peat with low protein and carbonate contents [@Stuart.2004; @Tatzber.2007; @Teickner.2022f] (note that we write "carbohydrates" and "aromatics" because, when interpreted only in terms of peak heights, it is remains uncertain what these organic matter (OM) fractions represent chemically or how one could extract them).
+In the following, we will use the terms holocellulose instead of the more general term carbohydrates and Klason lignin instead of the more general term aromatics.
+These are the accurate terms for the organic matter fractions quantified in @Hodgkins.2018.
 
-Such a qualitative interpretation of MIRS enables us to reinterpret @Hodgkins.2018, circumventing some of the uncertainties we have uncovered: All holocellulose and Klason lignin values given in @Hodgkins.2018 are directly proportional to the heights of the `carb` and `arom15arom16` peak, respectively.
-This removes uncertainties due to invalid spectral prediction models from the analyses.
-Then, we can still say that @Hodgkins.2018 show systematic latitudinal and depth-related differences in peat "carbohydrate" and "aromatics" contents.
+# Uncertainty sources not considered in @Hodgkins.2018 {.unnumbered}
 
-The trade-offs are first, that we have to make the assumption that there are no latitudinal or depth-related differences in mineral, carbonate, and protein contents, and second, that we have to make the assumption that differences in heights of the two peaks are not dominated by differences in other peaks in the spectra.
-The reason is that differences in peak heights can be attributed to "carbohydrates" or "aromatics" only when contents of minerals or proteins and carbonates, respectively, are the same, and when changes in the area normalized height of these peaks are not caused by changes in other peaks, which can be an effect of spectral normalization [@Stuart.2004; @Tatzber.2007; @Hodgkins.2018; @Teickner.2022f].
+In a recent study, we have shown that the models used to estimate holocellulose and Klason lignin contents in @Hodgkins.2018 (original models) are not valid for peat and this will cause larger uncertainties and biased predictions (uncertainty source 1) --- especially for decomposed peat --- of the estimated holocellulose and Klason lignin contents [@Teickner.2022f].
 
-We think that only some of these assumptions can are supported at the moment.
-The `carb` peak is comparatively large and has low interferences with peaks caused by other molecular structures.
-It should therefore be a good indicator for "carbohydrates" because mineral contents in (near-surface) peat are typically low (as validated by @Hodgkins.2018).
-The `arom15arom16` peak has large contributions by proteins [@Stuart.2004; @Teickner.2022f], based on peat N contents, it is not unlikely that peat protein contents differ with depth and latitude [@Sowden.1978; @Loisel.2014], and it may be more affected by other peaks since it is smaller than the `carb` peak.
-For this reason, we suggest that more evaluation is necessary before accepting, based on @Hodgkins.2018, that there are systematic differences in peat "aromatic" contents.
+Besides this, the following two uncertainty sources have not been considered in @Hodgkins.2018: first, the prediction uncertainty of the spectral prediction models (uncertainty source 2), and second, the uncertainty introduced when computing near-surface average holocellulose and Klason lignin contents from estimates for individual peat layers (uncertainty source 3).
 
-The reinterpretation in terms of "carbohydrate" and "aromatics" instead of holocellulose and Klason lignin has the clear disadvantage that we sacrifice the quantitative precision @Hodgkins.2018 aimed to attribute to the results: it is much more informative to precisely estimate a change in Klason lignin content of e.g.
-5 mass-% when the average annual temperature increase by 1$^{\circ}$C than to only make qualitative statements that the relative amount of "aromatics" (without describing in detail what this OM fraction means) increases.
+We recomputed the models for the latitudinal gradients of near-surface peat holocellulose and Klason lignin contents while considering all three uncertainty sources (supporting information S1).
 
-We think @Hodgkins.2018 have advanced a very important idea to make the interpretation of MIRS more quantitative.
-Only with such a quantitative approach, which links different types of measurements and the knowledge we have about them, can we develop a quantitatively precise understanding.
-However, we have shown that there is some more way to go.
-We hope that our critical analysis will help to improve this approach.
+Since we do not have yet models to accurately predict peat holocellulose and Klason lignin contents from MIRS, it is of course difficult to quantify the uncertainty and bias introduced by uncertainty source 1.
+However, the modified models provided in @Teickner.2022f are very likely more accurate and therefore the difference between predictions of the original model and our modified models are a plausible approximation of this additional uncertainty and bias.
+
+The reanalysis shows that both the estimated slopes for the latitudinal gradient (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->) and differences in the depth profiles are highly uncertain due to these uncertainty sources (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
+
+Thus, our reanalysis shows that estimates of the latitudinal gradient and depth gradients are much more uncertain than previously assumed.
 
@@
 
-```{r reply-res-p-veg-peat-differnece, echo=FALSE, eval=FALSE, out.width="90%", fig.height=3.5, fig.width=7.5, fig.cap='Predicted surface ($\\le 50$ cm) peat and vegetation median holocellulose (A) and Klason lignin (B) contents for the peat cores shown in figure \\ref{fig:res-p-gaussian-beta-depth-profile-best} grouped according to \\cite{Hodgkins.2018}.
-Predictions of the (biased) original models are equivalent to the data shown in \\cite{Hodgkins.2018} (figure 5).
-Averages are median values of predictions, error bars are 95\\% prediction intervals.
-"Difference" is the difference between predictions for vegetation and peat.'}
-p_veg_peat_difference
-```
+# Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition {.unnumbered}
+
+Our previous study has shown that in particular the spectral prediction model for Klason lignin is biased, especially for more decomposed peat [@Teickner.2022f].
+Our reanalysis using the modified model for Klason lignin from @Teickner.2022f thus indicates that larger Klason lignin contents may be necessary to stabilize tropical near-surface peat against decomposition as well as that changes in Klason lignin with depth may be more pronounced in the tropics than estimated in @Hodgkins.2018 (supporting information S1).
+
+With our modified model, near-surface peat Klason lignin contents are on average ~10 to 15 mass-% larger across the latitudinal gradient (figure \@ref(fig:reply-res-p-sites-latitude)).
+Specifically, average (sub)tropical ($<45 ^\circ$N) near-surface peat Klason lignin contents are `r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% (median, lower and upper 95% prediction interval limit) larger than with the original model.
+The large uncertainties now made explicit would also allow on average larger differences between deep (sub)tropical ($<45 ^\circ$N) and high-latitude ($\ge45 ^\circ$N) peat (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
+
+Similarly, residual enrichment of Klason lignin during decomposition may have been underestimated.
+A rough estimate for the residual enrichment of Klason lignin during decomposition is the difference in Klason lignin content between near-surface peat and deeper peat.
+With the modified model, this difference is on average for (sub)tropical peatlands `r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% larger (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).
+For high latitude peatlands the difference is smaller and much more uncertain than previously stated (`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-%).
+
+Consequently, in general --- and especially in (sub)tropical peatlands --- the average residual enrichment of Klason lignin due to decomposition probably has been underestimated by the original model.
+A consequence of this in light of climate change is that high latitude peat deposits may be even more prone to shifts in decomposition rates than suggested in @Hodgkins.2018.
+
+# The need to use precise concepts for organic matter fractions {.unnumbered}
+
+We argue that we should differentiate between vague concepts such as carbohydrates and aromatics and precise concepts such as holocellulose and Klason lignin and we should always use the most precise concept possible, unless more general statements are explicitly warranted.
+
+For example, limitations of the extraction procedure by which holocellulose and Klason lignin are defined are known (e.g., @Hatfield.2005, @Bunzel.2011, @Abu-Omar.2021) and these limitations as well as differences to other procedures get obscured by using vague words such as carbohydrates and aromatics.
+
+Using precise concepts for organic matter fractions becomes especially important when interpreting peat chemistry based on mid infrared spectra (MIRS) or spectral prediction models because all variables derived from MIRS --- for example the widely used humification indices, as proxy for relative contents of recalcitrant organic matter fractions (e.g.
+@Broder.2012), or Klason lignin contents predicted from models --- are really only indirect estimates of organic matter fractions and there are abundant possibilities how they can be misleading if they are not sufficiently validated.
+
+Intensities recorded in MIRS really only represent the fraction of the incident infrared radiation which is absorbed by specific molecular structures which happen to absorb at that specific energy level [@Stuart.2004].
+Such molecular structures may for example be aromatic C=C bonds, C-N bonds in amides, or C-O bonds in alcohols (e.g.
+carbohydrates) [@Stuart.2004].
+In addition, absorption of infrared radiation (and hence peak heights) can be decoupled from abundances of molecular structures because also the electronegativity of the bonds controls absorption of mid infrared radiation [@Stuart.2004].
+
+One major reason why the model for Klason lignin in @Hodgkins.2018 is biased and invalid for peat is that the only predictor variable used in this model, `arom15arom16`, is not a good indicator for Klason lignin because also proteins can absorb in the same energy range [@Stuart.2004] and therefore, estimates are biased relative to the amount of proteins in the samples [@Teickner.2022f].
+Therefore, we should interpret `arom15arom16` and similar variables, such as specific humification indices (HI$_{1630/1090}$, HI$_{1510/1090}$) only then as good proxies for aromatics if we have shown that protein contents do not differ much between the peat samples (thresholds which still needs to be established).
+
+In all other cases, we should be precise in our wording by calling the variable `arom15arom16` to signal that this variable may be no good indicator for aromatics if it is confounded by proteins.
+Obviously, this recommendation equally applies to all other variables derived from MIRS.
+
+@Hodgkins.2018 have made an important first step to actually quantify holocellulose and Klason lignin contents, i.e.
+specific concepts of carbohydrates and aromatics, from MIRS and this is an important improvement over the qualitative interpretation of peak heights or peak ratios (such as humification indices) used in the past (e.g.
+@Cocozza.2003, @Broder.2012, @Tfaily.2014).
+Using precise words for what we have actually measured and what confounds we have considered is important to assure that this first step will actually be an improvement over the qualitative interpretation of MIRS.
+
+# Conclusions {.unnumbered}
+
+Our reanalysis of @Hodgkins.2018, taking into account previously unconsidered sources of uncertainties, shows that:
+
+1.
+Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition and aromatics may accumulate at a higher rate with depth in tropical peatlands than estimated by @Hodgkis.2018.
+
+2. Larger uncertainties indicate that also larger differences in aromatic contents between (sub)tropical and high latitude peat than estimated by @Hodgkins.2018 are possible, also between deeper peat.
+
+3.
+More research should focus on how estimates of carbohydrate and aromatic contents from MIRS may be confounded by other organic matter fractions, in particular proteins.
+As a first step, this requires accurate concepts to name organic matter fractions and variables used in the interpretation of MIRS.
+
+The results of this reanalysis also apply to a more recent extension of @Hodgkins.2018 with a larger dataset [@Verbeke.2022].
 
@@
 Data and code to reproduce our analyses are available via <https://doi.org/10.5281/zenodo.6325760.>
+
+```{r reply-res-p-veg-peat-differnece, echo=FALSE, eval=FALSE, out.width="90%", fig.height=3.5, fig.width=7.5, fig.cap='Predicted surface ($\\le 50$ cm) peat and vegetation median holocellulose (A) and Klason lignin (B) contents for the peat cores shown in figure \\ref{fig:res-p-gaussian-beta-depth-profile-best} grouped according to \\cite{Hodgkins.2018}.
+Predictions of the (biased) original models are equivalent to the data shown in \\cite{Hodgkins.2018} (figure 5).
+Averages are median values of predictions, error bars are 95\\% prediction intervals.
+"Difference" is the difference between predictions for vegetation and peat.'}
+p_veg_peat_difference
+```
+
+<!--
+# On the responsible interpretation of peat mid infrared spectra (mid-IR, MIR, FTIR)
+
+Mid infrared spectra (MIRS, also known as mid-IR, FTIR) of peat are a source of vast information on peat chemistry.
+Not only can they be used to predict other peat properties, but also to trace changes in individual molecular structures.
+However, the way from a mid infrared spectrum to a correct inference about peat chemistry is riddled by many pitfalls which often are not appropriately handled.
+
+Here, we list common pitfalls, illustrate them with examples, describe their consequences, and show how to avoid them.
+Our exposition is specifically directed to scientists who have already very basic knowledge of MIRS, but are uncertain about possible errors during the interpretation of MIRS.
+
+We do not list examples to point at other's mistakes (we have our own mistakes [---todo:ref]), but to provide accessible examples which underline how common some of the pitfalls are.
+Through the use of examples, we also want to increase awareness for the grave consequences which the misinterpretation of MIRS has for our understanding of peat chemistry and processes.
+
+## Pitfall 1: Ignoring alternative molecular structures which contribute to a peak
+
+The peak around 1620 cm$^{-1}$ is often identified as indicator of aromatics [---todo:ref].
+Even though this is true, one should be aware that proteins also contribute to this peak.
+Now, you may think that proteins have such a small abundance in peat in comparison to aromatics that their contribution to this peak can be safely ignored.
+The problem is that the contribution of a molecular structure to a MIRS peak is not only controlled by its abundance, but also by its electronegativity: more electronegative bonds absorb more infrared radiation than less electronegative bonds, and thus contribute more to a peak.
+Aromatics contribute to the 1620 cm$^{-1}$ peak via a relatively unpolarized C=C bond, proteins via the much more electronegative N-H [---todo:check] bond.
+Thus, even though proteins have a much smaller abundance in peat than aromatics, the electronegativity of their N-H bond makes them a dominant contributor to the 1620 cm$^{-1}$ peak.
+
+More generally, peaks in peat MIRS almost never can be identified with only one molecular structure: typically, more than one molecular structure absorbs infrared radiation in one region of a spectrum.
+Even more, we are mostly agnostic on the relative abundances and interrelations of these molecular structures.
+On the one hand this certainly makes it hard to provide a plausible and concrete interpretation of MIRS peaks.
+On the other hand, it makes it too easy to accidentally focus on just the molecular structures we certainly know to contribute to a peak or which have been reported in previous studies.
+
+What are potential consequences of this pitfall?
+In the worst case, you'll assume that a wrong molecular structure changed in abundance.
+In any case, inappropriate consideration of alternative molecular structures will weaken the point you want to make because there are possible alternative causes for a change in a peak which you have not actively ruled out.
+
+How common is this pitfall?
+We focus here on the introductory example of the 1620 cm$^{-1}$ peak: an erroneous interpretation is always made when this peak is attributed exclusively to proteins or aromatics, without explaining why the respective other molecular structure could not have affected the peak.
+
+How can this pitfall be avoided?
+The first best step to avoid this pitfall is to use large databases where molecular structures common in any natural organic matter are assigned to peaks for the interpretation of MIRS.
+Considering how common this pitfall is, interpretation of MIRS should never be based on just a few studies and without more general literature on MIRS, e.g.
+textbooks.
+We have compiled such a database from general and more specific literature.
+This database is accessible from the 'irpeat' R package.
+In addition, we recommend other tools with more broad focus [---todo:ref spectroscopy and fun].
+
+The second best step to avoid this pitfall can be applied when you have sufficient evidence that some molecular structure did not contribute to a peak.
+For example, consider that you made separate protein measurements and found that protein content does barely change between treatments of some experiment you analyze.
+Then (and only then), it is appropriate to label the 1620 cm$^{-1}$ peak "aromatic peak".
+Explaining and justifying why some molecular structures can  be neglected makes this decision transparent.
+
+## Pitfall 3: Confounding molecular structures with organic matter fractions
+
+## Pitfall 2: Not considering how spectral normalization affects peaks
+
+Spectral normalization is a heuristic procedure which modifies the intensity values of spectra to get rid of differences between spectra which are not caused by the analyzed sample, but for example by sensor settings, or differences in sample materials if only relative changes are interpreted.
+Common spectral normalization procedures are for example to divide all intensity values by the intensity of a specific peak (this is the basis for so-called humification indices; for peat, inensities are most often normalized to (= divided by) the intensity around 1090 cm$^{-1}$ to analyze the change in a molecular structure relative to carbohydrates [---todo:ref]), by the sum of all intensity values (e.g.
+...), or by the vector norm (vector normalization).
+Since this is a heuristic procedure, different normalizations are appropriate for different tasks.
+
+Consequences of normalization:
+
++ Intensity changes in a peak can only be interpreted relative to changes in the intensity to which the spectra was normalized.
+
+## Pitfall 3: Analyzing peat chemistry with invalid or untested spectral prediction models without being critical
+
+Because MIRS contain so much information on peat chemistry, they can be used to predict many other peat properties.
+This makes spectral prediction models both very resource efficient tools to estimate several peat properties and devices to interpret spectra (by interpreting model coefficients).
+
+Spectral prediction models typically depend on many predictor variables (independent variables), first because of the normalization which makes variables included as predictor in a model also dependent on variables not included in the model (see pitfall 2), and second because many spectral prediction models use many peaks (or other variables derived from MIRS) to make predictions.
+Using many variables increases the risk of overfitting --- the phenomenon that a model fits the training datawell, but not new data --- and therefore such models need more and more diverse testing data to be validated.
+Spectral prediction models for peat MIRS are not new [---todo:ref], but well tested spectral prediction models for peat do not exist yet.
+
+In particular, one common pitfall is to use spectral prediction models on MIRS for which the model extrapolates.
+Extrapolating with high-dimensional data is particularly risky because of the increased overfitting risk and should therefore be avoided.
+Extrapolating also includes using a model trained on foreign data not representative for peat to make predictions for peat samples.
+Using such models can cause strongly biased predictions and underestimation of prediction uncertainties.
+
+To give a concrete example: @Hodgkins.2018 developed a model to predict peat holocellulose (as a proxy for carbohydrates) and Klason lignin (as a proxy for aromatics) contents by using a diverse set of wood, grass, leaves, and paper product samples.
+The MIRS have several properties which differ from peat and cause biased predictions.
+In particular, Klason lignin contents are strongly underestimated [---todo:ref].
+As a consequence, applying these models considerably underestimates peat Klason lignin contents and in any case introduces an unknown, but certainly not negligible, amount of uncertainty in any inference drawn based on these models.
+Sadly, these models have been misapplied frequently in the past, including by us [---todo:ref] and the conclusions of all these studies which are based on the models are on shaky ground.
+
+To avoid this pitfall, you have to check that the model you want to use is representative for you data and does not extrapolate for your data.
+Whilst it is difficult to fully check whether a model is representative for your data, a good indication for this is that it was computed with MIRS from samples which are similar to your samples (i.e., if you want to make predictions for peat, use a model which was computed with MIRS from peat).
+You can check that a model does not extrapolate by computing its prediction domain.
+The prediction domain of a model is the range of the predictor variable values of the MIRS used to compute the model.
+Ideally, researchers who provide prediction models would also include code/tools to easily perform these checks.
+
+## Pitfall 4: Neglecting uncertainties of spectral prediction models
+
+Spectral prediction models are no magic and often have relatively large prediction uncertainties. For instance
+-->

### 2023-11-27 analysis/paper/002-reply-main.Rmd
Update the comment on Hodgkins et al. (2018) with comments from Klaus.

@@
 1.
-Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition and aromatics may accumulate at a higher rate with depth in tropical peatlands than estimated by @Hodgkis.2018.
+Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition and aromatics may accumulate at a higher rate with depth in tropical peatlands than estimated by @Hodgkins.2018.
 
@@
 
-Our previous study has shown that in particular the spectral prediction model for Klason lignin is biased, especially for more decomposed peat [@Teickner.2022f].
-Our reanalysis using the modified model for Klason lignin from @Teickner.2022f thus indicates that larger Klason lignin contents may be necessary to stabilize tropical near-surface peat against decomposition as well as that changes in Klason lignin with depth may be more pronounced in the tropics than estimated in @Hodgkins.2018 (supporting information S1).
+Our previous study has shown that the spectral prediction model for Klason lignin is biased, especially for more decomposed peat [@Teickner.2022f].
+Our reanalysis using the modified model for Klason lignin from @Teickner.2022f indicates that larger Klason lignin contents may be necessary to stabilize tropical near-surface peat against decomposition as well as that changes in Klason lignin with depth may be more pronounced in the tropics than estimated in @Hodgkins.2018 (supporting information S1).
 
@@
 
-We argue that we should differentiate between vague concepts such as carbohydrates and aromatics and precise concepts such as holocellulose and Klason lignin and we should always use the most precise concept possible, unless more general statements are explicitly warranted.
+We argue that we should differentiate between vague concepts such as carbohydrates and aromatics and precise concepts such as holocellulose and Klason lignin and we should always use the most precise concept possible to describe the organic matter fraction we _plan_ (or intent) to measure, unless more general statements are explicitly warranted.
 
-For example, limitations of the extraction procedure by which holocellulose and Klason lignin are defined are known (e.g., @Hatfield.2005, @Bunzel.2011, @Abu-Omar.2021) and these limitations as well as differences to other procedures get obscured by using vague words such as carbohydrates and aromatics.
+For example, @Hodgkins.2018 planned to measure Klason lignin contents because their spectral prediction model used Klason lignin data as dependent variable, however this variable is labeled as aromatics.
+Limitations of the extraction procedure by which Klason lignin are defined are known (e.g., @Hatfield.2005, @Bunzel.2011, @Abu-Omar.2021) and these limitations as well as differences to other procedures get obscured by using vague words such as carbohydrates and aromatics.
 
-Using precise concepts for organic matter fractions becomes especially important when interpreting peat chemistry based on mid infrared spectra (MIRS) or spectral prediction models because all variables derived from MIRS --- for example the widely used humification indices, as proxy for relative contents of recalcitrant organic matter fractions (e.g.
-@Broder.2012), or Klason lignin contents predicted from models --- are really only indirect estimates of organic matter fractions and there are abundant possibilities how they can be misleading if they are not sufficiently validated.
+However, using precise concepts for organic matter fractions is also important for exactly the opposite reason: to not suggest a variable has more precision than is actually the case; this becomes especially important when interpreting peat chemistry based on mid infrared spectra (MIRS) or spectral prediction models because all variables derived from MIRS --- for example the widely used humification indices, as proxy for relative contents of recalcitrant organic matter fractions (e.g.
+@Broder.2012), or Klason lignin contents predicted from models --- are in fact only indirect estimates of organic matter fractions and there are abundant possibilities how they can be misleading if they are not sufficiently validated.
+Stating these limitations requires precise concepts.
 
@@
 carbohydrates) [@Stuart.2004].
-In addition, absorption of infrared radiation (and hence peak heights) can be decoupled from abundances of molecular structures because also the electronegativity of the bonds controls absorption of mid infrared radiation [@Stuart.2004].
+In addition, absorption of infrared radiation (and hence peak heights or areas) can be decoupled from abundances of molecular structures because also the electronegativity of the bonds controls absorption of mid infrared radiation [@Stuart.2004].
+These confounding factors make it necessary to explicitly define the conditions under which a spectral prediction model or other variable derived from MIRS (e.g.
+a humification index) is a valid proxy for a specific, precisely defined, organic matter fraction.
 
-One major reason why the model for Klason lignin in @Hodgkins.2018 is biased and invalid for peat is that the only predictor variable used in this model, `arom15arom16`, is not a good indicator for Klason lignin because also proteins can absorb in the same energy range [@Stuart.2004] and therefore, estimates are biased relative to the amount of proteins in the samples [@Teickner.2022f].
-Therefore, we should interpret `arom15arom16` and similar variables, such as specific humification indices (HI$_{1630/1090}$, HI$_{1510/1090}$) only then as good proxies for aromatics if we have shown that protein contents do not differ much between the peat samples (thresholds which still needs to be established).
+One major reason why the model for Klason lignin in @Hodgkins.2018 is biased and invalid for peat is that the only predictor variable used in this model, `arom15arom16`, is not a good indicator for Klason lignin because also proteins can absorb in the same energy range [@Stuart.2004] and therefore, estimates are biased depending on the amount of proteins in the samples [@Teickner.2022f].
+Therefore, we should interpret `arom15arom16` and similar variables, such as specific humification indices (HI$_{1630/1090}$, HI$_{1510/1090}$) only then as good proxies for the relative abundance of aromatic C=C bonds if we have shown that protein contents do not differ much between the peat samples (thresholds which still need to be established).
 
-In all other cases, we should be precise in our wording by calling the variable `arom15arom16` to signal that this variable may be no good indicator for aromatics if it is confounded by proteins.
+In all other cases, where we cannot validate if a variable is a good proxy for a specific organic matter fraction, we should be precise in our wording by calling the variable `arom15arom16` (instead of Klason lignin or aromatics) to signal that this variable may be no good indicator for aromatics if it is confounded by proteins, i.e.
+we may agree on `arom15arom16` as precise name for this variable, but it must always be understood that it is first and foremost _only_ defined as sum of the area of two peaks extracted by the specific procedure proposed in @Hodgkins.2018 and nothing more.
 Obviously, this recommendation equally applies to all other variables derived from MIRS.

### 2023-11-27 analysis/paper/002-reply-main.Rmd
Update the comment on Hodgkins et al. (2018) with comments from Jeff Chanton (received via email, 2023-11-26).

@@
 That said, theoretical considerations of decomposition processes alone support the suggested gradient in peat chemistry and @Hodgkins.2018 made an important contribution in providing the first open access models and estimates for peat carbohydrate and aromatics content on a large scale.
-We hope that our comment highlights that accuracy matters and that it helps to focus research on the problems which hamper improving this accuracy.
+Increasing the accuracy of models is an important part of the scientific process and we hope that this comment will focus research on the problems which hamper improving this accuracy.
 
@@
 Consequently, in general --- and especially in (sub)tropical peatlands --- the average residual enrichment of Klason lignin due to decomposition probably has been underestimated by the original model.
-A consequence of this in light of climate change is that high latitude peat deposits may be even more prone to shifts in decomposition rates than suggested in @Hodgkins.2018.
+A consequence of this is that high latitude peat deposits may experience more decomposition under a warmer climate than suggested in @Hodgkins.2018 because a larger content of Klason lignin is necessary to stabilize peat chemically under warmer conditions.
 
@@
 
-However, using precise concepts for organic matter fractions is also important for exactly the opposite reason: to not suggest a variable has more precision than is actually the case; this becomes especially important when interpreting peat chemistry based on mid infrared spectra (MIRS) or spectral prediction models because all variables derived from MIRS --- for example the widely used humification indices, as proxy for relative contents of recalcitrant organic matter fractions (e.g.
-@Broder.2012), or Klason lignin contents predicted from models --- are in fact only indirect estimates of organic matter fractions and there are abundant possibilities how they can be misleading if they are not sufficiently validated.
+However, using precise concepts for organic matter fractions is also important for exactly the opposite reason: to make clear how much accuracy and precision a variable actually has, instead of implying that it would quantify a variable more accurate and precise than is actually the case.
+This becomes especially important when interpreting peat chemistry based on mid infrared spectra (MIRS) or spectral prediction models because all variables derived from MIRS --- for example the widely used humification indices, as proxy for relative contents of recalcitrant organic matter fractions (e.g.
+@Broder.2012), or Klason lignin contents predicted from models --- are in fact only indirect estimates of organic matter fractions.
+These indirect estimates of organic matter fractions can be misleading in abundant ways if they are not sufficiently validated.
 Stating these limitations requires precise concepts.
@@
 carbohydrates) [@Stuart.2004].
-In addition, absorption of infrared radiation (and hence peak heights or areas) can be decoupled from abundances of molecular structures because also the electronegativity of the bonds controls absorption of mid infrared radiation [@Stuart.2004].
+When molecular structures absorb infrared radiation, this causes a change in dipole moment of one or more of their bonds and the larger the change in dipole moment, the more intense is the absorption and hence the larger the peak in a MIRS.
+The same stretching results in a larger change in dipole moment for more electronegative bonds than less electronegative bonds [@Stuart.2004].
+Since C-N bonds are more electronegative than aromatic C=C bonds, this means that already a small amount of proteins in peat can contribute equally large or more to the peaks in MIRS around 1510 and 1630 cm$^{-1}$ than aromatic C=C bonds in the same region of MIRS (see Fig.
+3 in @Reuter.2020 for an example).
 These confounding factors make it necessary to explicitly define the conditions under which a spectral prediction model or other variable derived from MIRS (e.g.
@@
 
-One major reason why the model for Klason lignin in @Hodgkins.2018 is biased and invalid for peat is that the only predictor variable used in this model, `arom15arom16`, is not a good indicator for Klason lignin because also proteins can absorb in the same energy range [@Stuart.2004] and therefore, estimates are biased depending on the amount of proteins in the samples [@Teickner.2022f].
+One major reason why the model for Klason lignin in @Hodgkins.2018 is biased and invalid for peat is that the only predictor variable used in this model, `arom15arom16`, is not a good indicator for Klason lignin because, as explained above, also proteins can absorb in the same energy range [@Stuart.2004] and therefore, estimates are biased depending on the amount of proteins in the samples [@Teickner.2022f].
 Therefore, we should interpret `arom15arom16` and similar variables, such as specific humification indices (HI$_{1630/1090}$, HI$_{1510/1090}$) only then as good proxies for the relative abundance of aromatic C=C bonds if we have shown that protein contents do not differ much between the peat samples (thresholds which still need to be established).
@@
 @Cocozza.2003, @Broder.2012, @Tfaily.2014).
-Using precise words for what we have actually measured and what confounds we have considered is important to assure that this first step will actually be an improvement over the qualitative interpretation of MIRS.
+Using precise words for what we have actually measured and what confounding factors we have considered is important to assure that this first step will actually be an improvement over the qualitative interpretation of MIRS.
 

### 2023-12-06 README.Rmd
Prepare submission of the reply to EarthEarXiv, update README.

@@
 
-This repository contains the data and code for our two manuscripts (in preparation):
+This repository contains the data and code for our two manuscripts:
 
@@
 
-> `r paste(manuscript2$authors, "(in preparation):", manuscript2$title)`
+> `r paste(manuscript2$authors, "(2023):", manuscript2$title, "Not peer-reviewed preprint.")`
 
@@
     - [001-reply-main.Rmd](analysis/paper/001-reply-main.Rmd): This is the main script for manuscript 2.
-It is compiled from within `001-paper-main.Rmd` and produces the supplementary information for manuscript 2.
+It is compiled from within `001-paper-main.Rmd` and produces the supplementary information S1 for manuscript 2.
     - [002-reply-main.Rmd](analysis/paper/002-reply-main.Rmd): This script produces the document for manuscript 2. It is compiled from within `001-reply-main.Rmd`.

### 2023-12-06 analysis/paper/001-reply-main.Rmd
Prepare submission of the reply to EarthEarXiv, update README.

@@
 ---
-title: 'Supplementary Information to: "Comment on Hodgkins et al. (2018)"'
-author: "Henning Teickner, Klaus-Holger Knorr"
+title: 'Supplementary Information S1 to: "Comment on Hodgkins et al. (2018)"'
+author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
 date: ""
@@
     keep_tex: true
-    citation_package: natbib
     toc: false
-bibliography: references
-biblio-style: unsrtnat
-natbiboptions: super
+bibliography: references.bib
 header-includes:
@@
 ```
+
+$^1$ ILÖK, Ecohydrology and Biogeochemistry Group, University of Münster, Heisenbergstr. 2, 48149 Münster, Germany  
+$^*$ Corresponding author: Henning Teickner (henning.teickner@uni-muenster.de)
 
@@
 
-```{r reply-res-p-latitude-poc-shrinkage, echo=FALSE, out.width="70%", fig.height=8.6, fig.width=6.5, fig.cap='(ref:reply-res-p-latitude-poc-shrinkage-caption)'}
+```{r reply-res-p-latitude-poc-shrinkage, echo=FALSE, out.width="60%", fig.height=8.6, fig.width=6.5, fig.cap='(ref:reply-res-p-latitude-poc-shrinkage-caption)'}
 p1 <- 

### 2023-12-06 analysis/paper/002-reply-main.Rmd
Prepare submission of the reply to EarthEarXiv, update README.

@@
    - \doublespacing
+   - \usepackage{titling}
+   - \pretitle{\begin{center}
+     \bf{This manuscript is a non-peer reviewed preprint submitted to EarthArXiv. It has not been submitted for publication to a journal yet.} \LARGE\\
+     \vspace{5ex}}
+   - \posttitle{\end{center}}
 fontsize: 12pt
+abstract:  |  
+   \noindent @Hodgkins.2018 used mid infrared spectra (MIRS) to make inferences about the stability of peat against decomposition along a latitudinal gradient from the tropics to northern latitudes.
+Recently, we have shown that these spectral prediction models are biased and their predictions more uncertain than considered in @Hodgkins.2018 [@Teickner.2022f].
+Here, we show what consequences this bias and uncertainty and additional neglected uncertainty sources have for the main conclusions in @Hodgkins.2018.
+   In particular, we find that:
+
+      1.
+Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition and aromatics may accumulate at a higher rate with depth in tropical peatlands than estimated by @Hodgkins.2018.
+
+      2.
+Larger uncertainties indicate that also larger differences in aromatic contents between (sub)tropical and high latitude peat than estimated by @Hodgkins.2018 are possible, also between deeper peat.
+
+      3.
+More research should focus on how estimates of carbohydrate and aromatic contents from MIRS may be confounded by other organic matter fractions, in particular proteins.
+As a first step, this requires accurate concepts to name organic matter fractions and variables used in the interpretation of MIRS.
 ---
@@
 $^*$ Corresponding author: Henning Teickner (henning.teickner@uni-muenster.de)
+
+\newpage
+
+# Introduction {.unnumbered}
 
@@
 
+# Supporting information {.unnumbered}
+
+Supporting information S1 is available as appendix to this manuscript.
+
 # Acknowledgements {.unnumbered}
@@
 
-Data and code to reproduce our analyses are available via <https://doi.org/10.5281/zenodo.6325760.>
+Data and code to reproduce our analyses are available via <https://doi.org/10.5281/zenodo.6325760>.
 
-```{r reply-res-p-veg-peat-differnece, echo=FALSE, eval=FALSE, out.width="90%", fig.height=3.5, fig.width=7.5, fig.cap='Predicted surface ($\\le 50$ cm) peat and vegetation median holocellulose (A) and Klason lignin (B) contents for the peat cores shown in figure \\ref{fig:res-p-gaussian-beta-depth-profile-best} grouped according to \\cite{Hodgkins.2018}.
-Predictions of the (biased) original models are equivalent to the data shown in \\cite{Hodgkins.2018} (figure 5).
-Averages are median values of predictions, error bars are 95\\% prediction intervals.
-"Difference" is the difference between predictions for vegetation and peat.'}
-p_veg_peat_difference
-```
-
-<!--
-# On the responsible interpretation of peat mid infrared spectra (mid-IR, MIR, FTIR)
-
-Mid infrared spectra (MIRS, also known as mid-IR, FTIR) of peat are a source of vast information on peat chemistry.
-Not only can they be used to predict other peat properties, but also to trace changes in individual molecular structures.
-However, the way from a mid infrared spectrum to a correct inference about peat chemistry is riddled by many pitfalls which often are not appropriately handled.
-
-Here, we list common pitfalls, illustrate them with examples, describe their consequences, and show how to avoid them.
-Our exposition is specifically directed to scientists who have already very basic knowledge of MIRS, but are uncertain about possible errors during the interpretation of MIRS.
-
-We do not list examples to point at other's mistakes (we have our own mistakes [---todo:ref]), but to provide accessible examples which underline how common some of the pitfalls are.
-Through the use of examples, we also want to increase awareness for the grave consequences which the misinterpretation of MIRS has for our understanding of peat chemistry and processes.
-
-## Pitfall 1: Ignoring alternative molecular structures which contribute to a peak
-
-The peak around 1620 cm$^{-1}$ is often identified as indicator of aromatics [---todo:ref].
-Even though this is true, one should be aware that proteins also contribute to this peak.
-Now, you may think that proteins have such a small abundance in peat in comparison to aromatics that their contribution to this peak can be safely ignored.
-The problem is that the contribution of a molecular structure to a MIRS peak is not only controlled by its abundance, but also by its electronegativity: more electronegative bonds absorb more infrared radiation than less electronegative bonds, and thus contribute more to a peak.
-Aromatics contribute to the 1620 cm$^{-1}$ peak via a relatively unpolarized C=C bond, proteins via the much more electronegative N-H [---todo:check] bond.
-Thus, even though proteins have a much smaller abundance in peat than aromatics, the electronegativity of their N-H bond makes them a dominant contributor to the 1620 cm$^{-1}$ peak.
-
-More generally, peaks in peat MIRS almost never can be identified with only one molecular structure: typically, more than one molecular structure absorbs infrared radiation in one region of a spectrum.
-Even more, we are mostly agnostic on the relative abundances and interrelations of these molecular structures.
-On the one hand this certainly makes it hard to provide a plausible and concrete interpretation of MIRS peaks.
-On the other hand, it makes it too easy to accidentally focus on just the molecular structures we certainly know to contribute to a peak or which have been reported in previous studies.
-
-What are potential consequences of this pitfall?
-In the worst case, you'll assume that a wrong molecular structure changed in abundance.
-In any case, inappropriate consideration of alternative molecular structures will weaken the point you want to make because there are possible alternative causes for a change in a peak which you have not actively ruled out.
-
-How common is this pitfall?
-We focus here on the introductory example of the 1620 cm$^{-1}$ peak: an erroneous interpretation is always made when this peak is attributed exclusively to proteins or aromatics, without explaining why the respective other molecular structure could not have affected the peak.
-
-How can this pitfall be avoided?
-The first best step to avoid this pitfall is to use large databases where molecular structures common in any natural organic matter are assigned to peaks for the interpretation of MIRS.
-Considering how common this pitfall is, interpretation of MIRS should never be based on just a few studies and without more general literature on MIRS, e.g.
-textbooks.
-We have compiled such a database from general and more specific literature.
-This database is accessible from the 'irpeat' R package.
-In addition, we recommend other tools with more broad focus [---todo:ref spectroscopy and fun].
-
-The second best step to avoid this pitfall can be applied when you have sufficient evidence that some molecular structure did not contribute to a peak.
-For example, consider that you made separate protein measurements and found that protein content does barely change between treatments of some experiment you analyze.
-Then (and only then), it is appropriate to label the 1620 cm$^{-1}$ peak "aromatic peak".
-Explaining and justifying why some molecular structures can  be neglected makes this decision transparent.
-
-## Pitfall 3: Confounding molecular structures with organic matter fractions
-
-## Pitfall 2: Not considering how spectral normalization affects peaks
-
-Spectral normalization is a heuristic procedure which modifies the intensity values of spectra to get rid of differences between spectra which are not caused by the analyzed sample, but for example by sensor settings, or differences in sample materials if only relative changes are interpreted.
-Common spectral normalization procedures are for example to divide all intensity values by the intensity of a specific peak (this is the basis for so-called humification indices; for peat, inensities are most often normalized to (= divided by) the intensity around 1090 cm$^{-1}$ to analyze the change in a molecular structure relative to carbohydrates [---todo:ref]), by the sum of all intensity values (e.g.
-...), or by the vector norm (vector normalization).
-Since this is a heuristic procedure, different normalizations are appropriate for different tasks.
-
-Consequences of normalization:
-
-+ Intensity changes in a peak can only be interpreted relative to changes in the intensity to which the spectra was normalized.
-
-## Pitfall 3: Analyzing peat chemistry with invalid or untested spectral prediction models without being critical
-
-Because MIRS contain so much information on peat chemistry, they can be used to predict many other peat properties.
-This makes spectral prediction models both very resource efficient tools to estimate several peat properties and devices to interpret spectra (by interpreting model coefficients).
-
-Spectral prediction models typically depend on many predictor variables (independent variables), first because of the normalization which makes variables included as predictor in a model also dependent on variables not included in the model (see pitfall 2), and second because many spectral prediction models use many peaks (or other variables derived from MIRS) to make predictions.
-Using many variables increases the risk of overfitting --- the phenomenon that a model fits the training datawell, but not new data --- and therefore such models need more and more diverse testing data to be validated.
-Spectral prediction models for peat MIRS are not new [---todo:ref], but well tested spectral prediction models for peat do not exist yet.
-
-In particular, one common pitfall is to use spectral prediction models on MIRS for which the model extrapolates.
-Extrapolating with high-dimensional data is particularly risky because of the increased overfitting risk and should therefore be avoided.
-Extrapolating also includes using a model trained on foreign data not representative for peat to make predictions for peat samples.
-Using such models can cause strongly biased predictions and underestimation of prediction uncertainties.
-
-To give a concrete example: @Hodgkins.2018 developed a model to predict peat holocellulose (as a proxy for carbohydrates) and Klason lignin (as a proxy for aromatics) contents by using a diverse set of wood, grass, leaves, and paper product samples.
-The MIRS have several properties which differ from peat and cause biased predictions.
-In particular, Klason lignin contents are strongly underestimated [---todo:ref].
-As a consequence, applying these models considerably underestimates peat Klason lignin contents and in any case introduces an unknown, but certainly not negligible, amount of uncertainty in any inference drawn based on these models.
-Sadly, these models have been misapplied frequently in the past, including by us [---todo:ref] and the conclusions of all these studies which are based on the models are on shaky ground.
-
-To avoid this pitfall, you have to check that the model you want to use is representative for you data and does not extrapolate for your data.
-Whilst it is difficult to fully check whether a model is representative for your data, a good indication for this is that it was computed with MIRS from samples which are similar to your samples (i.e., if you want to make predictions for peat, use a model which was computed with MIRS from peat).
-You can check that a model does not extrapolate by computing its prediction domain.
-The prediction domain of a model is the range of the predictor variable values of the MIRS used to compute the model.
-Ideally, researchers who provide prediction models would also include code/tools to easily perform these checks.
-
-## Pitfall 4: Neglecting uncertainties of spectral prediction models
-
-Spectral prediction models are no magic and often have relatively large prediction uncertainties. For instance
--->
+# References {.unnumbered}

### 2023-12-06 analysis/paper/002-reply-main.Rmd
Prepare submission of the reply to EarthEarXiv.
Update link to reproducible research compendium in manuscript 2.

@@
 fontsize: 12pt
-abstract:  |  
-   \noindent @Hodgkins.2018 used mid infrared spectra (MIRS) to make inferences about the stability of peat against decomposition along a latitudinal gradient from the tropics to northern latitudes.
-Recently, we have shown that these spectral prediction models are biased and their predictions more uncertain than considered in @Hodgkins.2018 [@Teickner.2022f].
-Here, we show what consequences this bias and uncertainty and additional neglected uncertainty sources have for the main conclusions in @Hodgkins.2018.
-   In particular, we find that:
-
-      1.
-Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition and aromatics may accumulate at a higher rate with depth in tropical peatlands than estimated by @Hodgkins.2018.
-
-      2.
-Larger uncertainties indicate that also larger differences in aromatic contents between (sub)tropical and high latitude peat than estimated by @Hodgkins.2018 are possible, also between deeper peat.
-
-      3.
-More research should focus on how estimates of carbohydrate and aromatic contents from MIRS may be confounded by other organic matter fractions, in particular proteins.
-As a first step, this requires accurate concepts to name organic matter fractions and variables used in the interpretation of MIRS.
 ---
@@
 
-Data and code to reproduce our analyses are available via <https://doi.org/10.5281/zenodo.6325760>.
+Data and code to reproduce our analyses are available via <https://doi.org/10.5281/zenodo.10276230> [@Teickner.2023a].
 
